### 2022-06-08 paper.tex
init

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}

\usepackage{epsfig}
\usepackage{subcaption}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{apalike}

\usepackage[scaled=0.8]{helvet}    % Less huge \textsf{functionName}
\usepackage[misc,geometry]{ifsym} % for letter symbol
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{trees,snakes,arrows}
\usetikzlibrary{shapes,chains}
\usetikzlibrary{positioning}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}
\usepackage[nospace]{cite}

\input{macros}

\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices}
\author{
        Karl Norrman and 
        Vaishnavi Sundararajan and
        Alessandro Bruni
}

\begin{document}
\maketitle
%
\begin{abstract}
Constrained IoT devices are becoming ubiquitous in society
and there is a need for secure communication protocols that respect the
constraints under which these devices operate.
%
\mEdhoc{} is an authenticated key establishment protocol for constrained IoT
devices, currently being standardized by the Internet Engineering Task
Force (IETF).
%
A rudimentary version of \mEdhoc{} with only two key establishment methods was
formally analyzed in 2018.
%
Since then, the protocol has evolved significantly and several new key
establishment methods have been added.
%
In this paper, we present a formal analysis of all \mEdhoc{} methods in an
enhanced symbolic Dolev-Yao model using the \mTamarin{} tool.
%
We show that not all methods satisfy the authentication notion
injective agreement, but that all do satisfy a notion of implicit
authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
%
We identify other weaknesses to which we propose improvements
%
For example, a party may intend to establish a session key with a certain peer,
but ending up establishing it with another, trusted but compromised, peer.
%
We communicated our findings and proposals to the IETF, which
has incorporated some of these in newer versions of the standard.
\end{abstract}
%
%-------------------------------------------------------------------------- sec
\section{\uppercase{Introduction}}
\label{sec:introduction}

As IoT devices become more prevalent and get involved in progressively sensitive
functions in society, the need to secure their communications
becomes increasingly important.
%
Most security analysis has focused on computationally
strong devices, such as cars and web-cameras, where existing protocols like
\mDandTls{} suffice.
%
Constrained devices, on the other hand, which operate under severe
bandwidth and energy consumption restrictions, have received much less
attention.
%
These devices may be simple sensors, which only relay environment
measurements to a server every hour, but need to function autonomously without
maintenance for long periods of time.
%
The IETF standardized the Object Security for
Constrained RESTful Environments (\mOscore{}) protocol to secure communications
between constrained devices~\cite{rfc8613}.
%
However, the \mOscore{} protocol requires a pre-established security context.
%
The IETF has been discussing requirements and mechanisms for a key
exchange protocol, named Ephemeral Diffie-Hellman Over COSE (\mEdhoc), for
establishing \mOscore{} security contexts.
%
Naturally, \mEdhoc{} must work under the same constrained requirements as
\mOscore{} itself.
%
While not all use cases for \mEdhoc{} are firmly set, the overall goal is to
establish an \mOscore{} security context, under message size limitations.
%
It is therefore important to ensure that \mEdhoc{} satisfies fundamental
security properties expected from a key exchange protocol.
%The work is now being done in the IETF Lightweight Authenticated Key Exchange
%(LAKE) working group.
%

The first incarnation of \mEdhoc{} appeared in March 2016.
%
It contained two different key establishment methods, one based on a
pre-shared Diffie-Hellman (DH) cryptographic core%
\footnote{By a \emph{cryptographic core}, or simply core, we mean an academic protocol,
without encodings or application specific details required by an industrial
protocol.
%
A \emph{key establishment method} is a core with some such details added.
}
and a second based on a
variant of challenge-response signatures in the style of
\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
%
\mEdhoc{} is therefore a framework of several key establishment methods.
%
In May 2018, the core based on challenge-response signatures was replaced by
one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}.
%
Since then the protocol has undergone significant changes.
%
Three new cores, mixing challenge-response signatures and regular signature for
authentication, were added~\cite{our-analysis-selander-lake-edhoc-00}.
%

We formulate and formalize a security model covering all four key
establishment methods, which is important especially since the
\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} lacks a clear description
of the intended security model and overall security goals.
%

We perform the analysis in a symbolic Dolev-Yao model.
In this framework, we model messages as terms in an algebra, with operations such
as encryption, modelled as functions on these terms.
These functions are assumed perfect, e.g., one cannot decrypt an encrypted
message without access to the key.
The adversary, while unable to break encryption or reverse hashing, is modelled
as the network.
That is, the adversary, can block reroute, replay and modify messages at will.
A symbolic model like this, while slightly severe an abstraction, still allows
us to analyze \mEdhoc{} for logical flaws without incurring the complexity of
a computational model.
%
The standardization process is ongoing, with the authors releasing newer
versions of the \mSpec{} (see Section~\ref{sec:newdrafts} for more detail
about how these versions differ from the one analyzed here).
%

%-------------------------------------------------------------------------- sub
\subsection{Contributions}
\label{sec:contributions}
In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
establishment methods) using the \mTamarin{} tool~\cite{DBLP:conf/cav/MeierSCB13}.
%
We present a formal model we constructed of the protocol as given in the
\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00}.
%

We give an explicit adversary model for the protocol and verify
properties such as session key material and entity authentication, and perfect
forward secrecy, for all four methods.
%

The model itself is valuable as a basis for verifying further updates in the
ongoing standardization.
%
It is publicly available~\cite{edhocTamarinRepo}.
%
It took several person-months to interpret the
specification and construct the model.
%
Termination requires a hand-crafted proof oracle to guide \mTamarin{}.
%

We show that not all \mEdhoc{}'s key establishment
methods provide authentication according to the injective agreement definition
on the session key material, and none on the initiator's identity.
%
However, we show that all methods fulfill an implicit agreement property
covering the session key material and the initiator's identity.
%
We identify a number of subtleties, ambiguities and weaknesses in the
specification.
%
For example, the authentication policy requirements allows situations where a
party establishes session key material with a trusted but compromised peer, even
though the intention was to establish it with a different trusted party.
%
We provide remedies for the identified issues and have
communicated these to the IETF and the specification authors, who incorporated
some of our suggestions and currently consider how to deal with the remaining
ones.
%

%-------------------------------------------------------------------------- sub
\subsection{Comparison with Related Work}
The May 2018 version of \mEdhoc{} was formally analyzed by
\cite{DBLP:conf/secsr/BruniJPS18} using the \mProverif{}
tool~\cite{DBLP:conf/csfw/Blanchet01}.
%
Their analysis covered a pre-shared key authenticated core and one
based on \mSigma.
%
The properties checked for therein were secrecy, PFS and integrity of
application data, identity protection against an active adversary,
and strong authentication.
%

In contrast to the key establishment methods analyzed by Bruni et~al., which
were based on the well-understood pre-shared key DH and \mSigma{} protocols,
the three newly added
methods combine two unilateral authentication protocols with the goal to
constructing mutual authentication protocols.
%
Combining two protocols, which individually provide unilateral authentication,
is not guaranteed to result in a secure mutual authentication
protocol~\cite{DBLP:conf/ccs/Krawczyk16}.
%
Consequently, even though the framework is similar to the one analyzed by Bruni
et~al., the cryptographic underpinnings have significantly increased in
complexity, and is using mechanisms which have not previously been formally analyzed.
%
The set of properties we check for is also different.
%
Our analysis is further carried out using a different tool,
namely \mTamarin; different kinds of strategies to formulate and
successfully analyze the protocol are required when working with this tool.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{The \mEdhoc{} Protocol}}
\label{sec:edhoc}
\input{edhocProtocol}

%-------------------------------------------------------------------------- sec
\section{\uppercase{Formalization and Results}}
\label{sec:formalization}
\input{edhocFormalization}

%-------------------------------------------------------------------------- sec
\section{\uppercase{Discussion}}
\label{sec:discussion}
There are a few places where \mEdhoc{} can be improved,
which we found during this work and communicated to the authors.
%
We discuss them below.
%

%-------------------------------------------------------------------------- sub
\subsection{Unclear Intended Use}
\label{sec:unclearProtocolUse}
%
The \mEdhoc{} \mSpec{} lists several security goals, but they are
imprecise and difficult to interpret due to lack of context and intended usage
descriptions.
%
Without knowing how the protocol is to be used,
it is not clear whether the listed security goals are the most important ones
for constrained IoT devices.
%

The abstract goal of \mEdhoc{} is simple: establish an \mOscore{} security
context using few roundtrips and small messages.
%
From that, the design of \mEdhoc{} is mainly driven by what
can be achieved given the technical restrictions.
%
Focusing too much on what can be achieved within given restrictions, and paying
too little attention to the use cases where the
protocol is to be used and their specific goals, risks resulting in
sub-optimal trade-offs and design decisions.
%

\mEdhoc{} is intended to cover a variety of use cases, many of which are
difficult to predict today.
%
However, this does not
prevent collecting \emph{typical} use cases and user stories
to identify more specific security goals that will be important in most cases.
%

While constructing our model, we made up simple user stories to identify
security properties of interest.
%
Several of these revealed subtleties and undefined aspects of \mEdhoc{}.
%
We informed the \mEdhoc{} authors, who addressed these aspects in the
\mSpec{}.
%

\subsubsection{(Non-)Repudiation}
An access control solution for a nuclear power-plant may need to log who is
passing through a door, whereas it may be undesirable for, say, a coffee
machine to log a list of users along with their coffee preferences.
%
Via this simple thought experiment, we realized that the \mSpec{} did not
consider the concept of (non)-repudiation.
%
In response, the authors of the \mSpec{} added a paragraph discussing how
different methods relate to (non)-repudiation.

\subsubsection{Unintended Peer Authentication}
Section~3.2 of the \mSpec{} states that parties must be configured
with a policy restricting the set of peers they run \mEdhoc{} with.
%
However, the initiator is not required to verify that the \mIdcredr{} received
in the second message is the same as the one intended.
%
The following attack scenario is therefore possible.
%

Suppose someone has configured all devices in their home to be in the
allowed set of devices, but that one of the devices ($A$) is compromised.
%
If another device $B$, initiates a connection to a third device $C$, the
compromised device $A$ may interfere by responding in $C$'s place, blocking
the legitimate response from $C$.
%
Since $B$ does not verify that the identity indicated in the second message
matches the intended identity $C$, and device $A$ is part of the allowed set,
$B$ will complete and accept the \mEdhoc{} run with device $A$ instead of the
intended $C$.
%
The obvious solution is for the initiator to match \mIdcredr{} to the intended
identity indicated by the application, which we included in our model.
%
We have communicated this to the \mEdhoc{} authors and they are considering
how to resolve the issue.
%

%------------------------------------------------------------------------- sub
\subsection{Unclear Security Model}
We argue that the \mSpec{} gives too little information about what capabilities
an adversary is assumed to have, and that this leads to unclear design goals and
potentially sub-optimal design.
%

Even though \mEdhoc{} incorporates cryptographic cores from different academic
security protocols, its design does not take into account the adversary models
for which these protocols were designed.
%
For example, \mOptls{}, whose cryptographic core is essentially the same
as the \mStat{} authentication method, is designed to be secure in the CK
model~\cite{DBLP:conf/crypto/CanettiK02}.
%
The CK security model explicitly separates the secure storage of long-term
keys from storage of session state and ephemeral keys.
%
This is appropriate for modelling the use of secure modules.
%

The \mEdhoc{} authors indicated to us that it was
not necessary to consider compromised ephemeral keys separately from
compromised long-term keys.
%
The rationale is that \mSigma{} cannot protect against compromised ephemeral
keys~\cite{personalCommunication}.
%
That rationale is presumably based on the fact that the \mSigSig{} method is
closely modeled on the \mSigmaI{} variant of \mSigma{}, and that it would be
preferable to obtain a homogeneous security level among the \mEdhoc{}
methods.
%
That rationale is only true, however, if one restricts attention to session key
confidentiality of an ongoing session.
%
Secure modules provide value in other ways, for example, by allowing
constructions with Post-Compromise Security (PCS) guarantees.
%
We discussed this with the authors, and
the latest version of the \mSpec{}~\cite{latest-ietf-lake-edhoc-05} includes
recommendations on storage of long-term keys and operations on these inside a
secure module.
%

%-------------------------------------------------------------------------- sub
\subsection{Session Key Material}
\label{sec:sessionKeyMaterial}
\mEdhoc{} establishes session key material, from which session keys
can be derived using the \mEdhoc{}-Exporter.
%
The session key material is affected by \mGxy{}, and if a party uses the
\mStat{} authentication method, also by that party's secret static long-term key.
%
As shown in Section~\ref{sec:formalization}, mutual injective agreement cannot
be achieved for $P_I$.
%
If this property is not important for constrained IoT devices which cannot use
any of the other methods, then one can simply accept that the methods have
different authentication strengths.
%
Otherwise, this is a problem.
%

We identified three alternatives for resolving this.
%
One alternative is to include \mIdcredi{}, or its hash, in the first and
second messages.
%
This would, however, increase message sizes and prevent initiator identity
protection, which are grave concerns for \mEdhoc{}.
%
A second alternative is to not derive the session key material from $P_I$.
%
Doing so, however, deviates from the design of \mOptls{} (and similar protocols
from which the \mStat{}-based methods are derived), where the inclusion of
$P_I$ plays a crucial part in the security proof of resistance against
initiator ephemeral key compromise.
%
The third alternative is to include a fourth message from responder to initiator,
carrying a MAC based on a key derived from session key material including $P_I$.
%
Successful MAC verification guarantees
to the initiator that the responder injectively agrees on $P_I$.
%
We presented the options to IETF, and they decided to add a
fourth message as an option in the latest version of the
\mSpec{}~\cite{latest-ietf-lake-edhoc-05}.
%

Regardless of how this is handled, we verified that all methods
enjoy a common, but weaker, property: mutual implicit agreement
on all of $P_e, P_I$ and $P_R$, where applicable.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{Conclusions and Future Work}}
\label{sec:conclusions}
\label{sec:newdrafts}
We formally modeled all four
methods of the \mEdhoc{} \mSpec{} using \mTamarin.
%
We formulated several important security properties and identified precise
adversary models in which we verified these.
%
The properties are shown in Table~\ref{tab:props}.
%
Mutual injective agreement covers the parameters $S_P$:
responder identity, roles, session key material (except for $P_I$ when
initiator uses the \mStat{} authentication
method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
%
The responder in addition is ensured agreement on the initiators identity and
$P_I$, i.e., on the set $S_F$.
%
Implicit agreement covers all previously mentioned parameters for both peers.
%
Verification of all lemmas, including model validation lemmas, took 42 minutes
on an Intel Core i7-6500U 2.5GHz using two cores.
%
Mutual entity authentication, UKS- and KCI resistance can be inferred
from the verified properties.
%
\begin{table*}[h!]
        \centering
        \caption{Verified properties. $S_P$ contains
            roles, responder identity, session key material (excluding
            $P_I$), \mCi, \mCr, and \mSuites. $S_F$ is $S_{P}$,
            the initiator identity, and $P_I$.}
        \label{tab:props}
        \begin{tabular}{|l|c|c|c|c|}
                \hline
                & \mSigSig & \mSigStat & \mStatSig & \mStatStat \\
                \hline
                Injective agreement for I & $S_F$ & $S_F$ & $S_P$ & $S_P$\\
                Injective agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
                Implicit agreement for I & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
                Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
                PFS for session key material & \cm & \cm & \cm & \cm\\
                \hline
        \end{tabular}
\end{table*}

Further, we identified a situation where initiators may establish an \mOscore{}
security context with a different party than the application using \mEdhoc{}
intended, and proposed a simple mitigation.
%
We discussed how the IETF may extract and better define security properties to
enable easier verification.

We verified each method in isolation, and leave as future work to verify whether
the methods are secure under composition.

%\subsection{Newer Versions of the Specification}
%\label{sec:newdrafts}
In this work, we have analyzed the \mEdhoc{} version as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}.
%
There are newer versions, with the most recent version as
of February 2021~\cite{latest-ietf-lake-edhoc-05}.
%
However, the changes to the protocol over these versions are not
particularly significant for our analysis.
%

%-------------------------------------------------------------------------- ack
\section*{ACKNOWLEDGEMENTS}
This work was partially supported by
the Wallenberg AI, Autonomous Systems and Software Program (WASP) funded by
the Knut and Alice Wallenberg Foundation.
%
We are grateful to G\"oran Selander, John Mattsson and Francesca Palombini for
clarifications regarding the specification.
%

%-------------------------------------------------------------------------- bib
\bibliographystyle{apalike}
{\small
    \bibliography{refComp}
}
\end{document}

### 2022-06-30 paper.tex
[add] initial text on DoS

@@
 \input{edhocFormalization}
+
+%-------------------------------------------------------------------------- sec
+\section{\uppercase{Denial of Service}}
+\label{sec:dos}
+\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
+attacks and delegates the responsibility to handle this to the application.
+%
+\mEdhoc{} itself has limited functionality, few negotiable parameters and
+processing rules
+%
+Hence, we argue in this section that the attack surface
+is small from a DoS perspective.
+%
+However, \mEdhoc{} leaves much of the semantics of its messages undefined.
+%
+This increases the risk of applications vulnerable to DoS even if \mEdhoc{} is
+arguably not vulnerable to DoS itself.
+%
+
+Section 8.7 of the \mEdhoc{} specification explains the design philosophy w.r.t.
+DoS in two paragraphs and gives two recommendations.
+%
+The first being that mitigation of DoS based on sending proper messages should
+be handled by lower layers, e.g., by checking reachability of the sending
+party\footnote{Karl: this may be another way to perform reflection DoS attacks}.
+%
+The second being that an application may try to determine (in some way) that an
+apparently valid message is in fact probably a forgery and should be ignored or
+cause processing to stop.
+%
 

### 2022-06-30 paper.tex
[add] initial version of DoS section

@@
 \label{sec:dos}
+\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
 \mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
@@
 
+\mPoint{\mEdhoc{} gives a little info of questionable use for developers}
 Section 8.7 of the \mEdhoc{} specification explains the design philosophy w.r.t.
@@
 cause processing to stop.
+%
+
+\mPoint{Responder is more vulnerable than initiator}.
+For the same reasons as given in the IKEv2 DDoS analysis
+by Nir and Smyslov~\cite{rfc8019} the responder is more vulnerable than the
+initiator.
+%
+From a roles and message passing perspective, IKEv2 is very similar to all
+methods in \mEdhoc{}.
+%
+The RFC provides much good advice for protocols with this structure that
+\mEdhoc{} using applications can reuse.
+%
+We should look more into this.
+%
+
+\mPoint{Amplification}
+An adversary can fill a message 1 with easily generated bogus data and send this
+to a victim responder.
+%
+There is little work for the initiator, but both symmetric and asymmetric key
+operations required by the responder.
+%
+\mEdhoc{} does not protect against this, but refer to the application and its
+transport layer to resolve the problem.
+%
+This is a design decision, see Section 8.7 of \mEdhoc{} specification.
+%
+In addition to the computational resources the responder spends, it may also
+need to obtain certificate chains and verify certificate revocation servers as a
+result of the spoofed message.
+%
+The asymmetry between the adversaries effort and the responder may hence be
+significant.
+%
+Since \mEdhoc{} specifically targets constrained devices this asymmetry needs to
+be considered even more carefully than for IKEv2~\cite{rfc8019}.
+%
+
+Recommend that the role of responder is chosen carefully.
+%
+The responder is most vulnerable.
+%
+For example, is it better that a lock or one of the key cards takes on the
+responder role and becomes vulnerable?
+%
+
+\mPoint{Reflection Attacks}
+Reflection attacks are attacks where an adversary spoofs the source address of a
+victim in a message and thereby tricks the responder to send a message to the
+victim.
+%
+\mEdhoc{} leaves protection against this to the transport protocol and
+application, and suggests using a path reachability method similar to IKEv2.
+%
+In situations where the responder does not need to perform additional
+communications to verify certificates or similar, a reachability mechanism may
+be more expensive than performing the computations.
+\footnote{Find ref of that computation is often cheaper than radio
+transmission, at least energy-wise.}
+%
+However, if round-trip times are large, and half-open states hence are kept
+by the receiver for a long time, a reachability mechanism can trade
+communication for storage.
+\footnote{RFC8019 discuss some numbers as examples of tradeoffs. Should we set
+up a simple model and derive an expression for trade-offs? Could be useful for
+implementors to consider, but low academic value. This would capture also
+amplification.}
+%
+
+\mPoint{In a constrained device randomness may be easily depleted}
+Assume an adversary generate (deterministic) bogus message 1s and send these at
+a sufficiently high rate to the responder.
+%
+\mEdhoc{} specifically targets constrained devices, that presumably may not
+always have good randomness sources.
+%
+Is it possible to deplete or significantly reduce the entropy of the source and
+in this way be able to predict the responder's random input to other sessions
+(with other initiators)?
+%
+The LAKE WG have been talking about effects of poor randomness on signatures, so
+this should be an issue, no?
+%
+
+\mPoint{\mEdhoc{} can deadlock if not used correctly}
+An adversary may send forged error messages, and specifically reject proposed
+ciphersuites.
+%
+The initiator is recommended to not try that ciphersuite with the responder
+again in a new session for a while.
+%
+If the adversary rejects all options for a initiator this way, it can prevent
+communication between the initiator and that responder until the cache is
+cleared.
+%
+Potentially, an adversary could lock a party of a group out of communication all
+together this way.
+%
+This needs to be considered by applications.
+%
+
+\mPoint{Is any of the methods better than the others from a DoS perspective?}
+We considered this for quite some time, but there seems to be little difference
+in this regard.
 %

### 2022-07-01 model/oracle.py
[add] tamarin model with session key reveal

@@ -1,0 +1,279 @@
+#!/usr/bin/python3
+import sys, re
+from functools import reduce
+
+DEBUG = False
+#DEBUG = True
+
+# Put prios between 0 and 100. Above 100 is for default strategy
+MAXNPRIO = 200    # max number of prios, 0 is lowest prio
+FALLBACKPRIO = MAXNPRIO    # max number of prios, 0 is lowest prio
+prios = [(i, []) for i in range(MAXNPRIO + 1)]
+
+
+def outputPrios(goalLines, lemma):
+    rankedGoals = [str(goal) + "\n" for prioList in prios for goal in prioList[1]]
+    print("".join(rankedGoals))
+
+
+def dumpPrios(goalLines, lemma):
+    print("Prios:")
+    for pl in prios:
+        for p in pl[1]:
+            print("    > level:{}, goalNo:{}".format(pl[0], p))
+
+
+def prioritize(goalNumber, prio, goalLine):
+    prios[prio][1].insert(0, goalNumber)
+    if DEBUG:
+        goal = re.sub("\s+", " ", goalLine)
+        print("goalNo:{} prio:{} goal:{}".format(goalNumber, prio, goal))
+
+
+def genPrios(goalLines, lemma):
+    # Prioritize splitEqs over new instances
+    # splitEqs = False
+    # splitEqsLine = -1
+    # for i in range(len(goalLines)):
+    #     if re.match(".*splitEqs.*", goalLines[i]):
+    #         splitEqs = True
+    #         splitEqsLine = i
+
+    for line in goalLines:
+        goal = line.split(':')[0]
+        if "sanity" in lemma:
+            if DEBUG:
+                print("MATCHING Sanity LEMMA: {}".format(lemma))
+            if re.match(".*SKRev.*", line) or\
+               re.match(".*Completed.*", line):
+                    prioritize(goal, 90, line)
+            elif re.match(".*StR.*", line) or\
+               re.match(".*StI.*", line):
+                    prioritize(goal, 80, line)
+            elif re.match(".*KU\( 'g'\^~xx \).*", line) or\
+               re.match(".*KU\( 'g'\^~yy \).*", line) or\
+               re.match(".*KU\( ~xx.*", line) or\
+               re.match(".*KU\( ~yy.*", line) or\
+               re.match(".*~~>.*", line) or\
+               re.match(".*=.*=.*", line):
+                    prioritize(goal, 70, line)
+            elif re.match(".*LTK_.*", line):
+                    prioritize(goal, 67, line)
[…]

### 2022-07-05 paper.tex
[add] update of DoS section

--- 2022-06-08 paper.tex
@@
 \input{edhocFormalization}
+
+%-------------------------------------------------------------------------- sec
+\section{\uppercase{Denial of Service}}
+\label{sec:dos}
+%\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
+\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
+attacks and delegates the responsibility to handle this to the application.
+%
+\mEdhoc{} itself has limited functionality, few negotiable parameters and
+processing rules.
+%
+Hence, we argue in this section that its attack surface
+is limited from a DoS perspective.
+%
+However, \mEdhoc{} leaves much of the semantics of its messages undefined.
+%
+This increases the risk of applications vulnerable to DoS via \mEdhoc{}, even
+though \mEdhoc{} itself is arguably not particularly weak in this respect.
+%
+
+\paragraph{DoS mitigation is left to developers.}
+Section 8.7 of the \mEdhoc{} specification\footnote{draft-ietf-lake-edhoc-14}
+explains the design philosophy w.r.t.
+DoS in two paragraphs and gives two recommendations.
+%
+The first being that \mEdhoc{} relies on lower layers to mitigate DoS, and an
+example of DoS countermeasures is given: checking return routeability when
+receiving the first message.
+%
+The second being that an application may try to determine (in some way) that an
+apparently valid message is in fact probably a forgery and should be ignored or
+cause processing to stop.
+%
+
+No recommendation is given on how to minimize DoS effects stemming from use of
+error messages, which can be used in an arbitrary way in \mEdhoc{}.
+%
+We now elaborate on some considerations application developers needs to take
+into account.
+%
+
+\paragraph{Amplification and asymmetry.}
+For the same reasons as given in the IKEv2 DDoS analysis
+by~\cite{rfc8019} the responder is more vulnerable than the
+initiator.
+%
+The key similarity between IKEv2 and \mEdhoc{} here is that the initiator is
+anonymous in the first message and that creating the first message is
+computationally cheap and requires no storage.
+%
+No element in the first \mEdhoc{} message can be verified by the receiver as
+being legitimate as long as the format is correct and elliptic curve points
+lie on the proposed curve.
+%
+\mEdhoc{} leave it to the application and its
+transport layer to resolve any issues.
+%
+In addition to the computational resources the responder spends, communication
+patterns in the underlying protocols may require the responder to
+obtain and verify certificate chains, and verify
+certificate-revocation servers as a result of the spoofed initial message.
+%
+For example, the content of the \mADone{} information element may hint the
+responder to take further actions.
+%
+
+The asymmetry between the adversary's effort and the responder's may hence be
+significant.
+%
+Since \mEdhoc{} specifically targets constrained devices this asymmetry needs to
+be considered even more carefully than for IKEv2~\cite{rfc8019}.
+%
+
+Depending on the use case, the first \mEdhoc{} message may be sent by either
+party in a session.
+%
+For example, even though it may seem intuitive that an electronic
+key-card initiates the connection towards lock, the first \mEdhoc{} message
+could be included in the response from the lock.
+%
+This means the lock can postpone heavy verification actions until it has
+authenticated the key card after receiving the third \mEdhoc{} message.
+%
+
+Implementers are well advised to consult \cite{rfc8019}, which provides much
+good general information about DoS for protocols with this structure.
+%
+However, ultimately, a product-specific risk assessment is required to
+determine appropriate DDoS mitigations.
+%
+
+\paragraph{Reflection Attacks}
+Reflection attacks are attacks where an adversary spoofs the source address of a
+victim in a message and thereby tricks the responder to send a message to the
+victim.
+%
+\mEdhoc{} leaves protection against this to the transport protocol and
+application, and suggests using a return path reachability method similar to IKEv2.
+%
+We note that such reachability mechanisms can easily be used to make a responder
+send reachability requests to arbitrary addressable targets.
+%
+Depending on the design of such reachability mechanisms, they may require that
+unsolicited reachability requests are discarded, limiting the effects of
+anonymous DoS attacks like these.
+%
+
+In situations where the responder does not need to perform additional
+communications to verify certificates or similar, a reachability mechanism may
+be more expensive in terms of time and storage compared to continue with the
+protocol and keeping a half-open state for a period of time.
+%
+Doing so would also avoid creating secondary DoS effects on certificate storage
+nodes or certificate revocation servers.
+%
+The latter may cause a systems-level issue if many distributed
+\mEdhoc{} responders frequently requests such servers.
+%
+However, if round-trip times are large, and half-open states hence are kept
+by the receiver for a long time, a reachability mechanism can trade
+communication for storage.
+\footnote{RFC8019 discuss some numbers as examples of tradeoffs. Should we set
+up a simple model and derive an expression for trade-offs? Could be useful for
+implementors to consider, but low academic value. This would capture also
+amplification.}
+%
+
+\mPoint{In a constrained device randomness may be easily depleted}
+Assume an adversary generate (deterministic) bogus message 1s and send these at
+a sufficiently high rate to the responder.
+%
+\mEdhoc{} specifically targets constrained devices, that presumably may not
+always have good randomness sources.
+%
+Is it possible to deplete or significantly reduce the entropy of the source and
+in this way be able to predict the responder's random input to other sessions
+(with other initiators)?
+%
+The LAKE WG have been talking about effects of poor randomness on signatures, so
+this should be an issue, no?
+%
+
+\mPoint{\mEdhoc{} can deadlock if not used correctly}
+An adversary may send forged error messages, and specifically reject proposed
+ciphersuites.
+%
+The initiator is recommended to not try that ciphersuite with the responder
+again in a new session for a while.
+%
+If the adversary rejects all options for a initiator this way, it can prevent
+communication between the initiator and that responder until the cache is
+cleared.
+%
+Potentially, an adversary could lock a party of a group out of communication all
+together this way.
+%
+This needs to be considered by applications.
+%
+
+While we have not conducted a formal analysis of the fact, the structure of all
+methods are so similar that it seems plausible that none of them would be
+stronger against DDoS than the others.
+%
 

### 2022-07-07 paper.tex
[add] DoS updates

@@
 %
-\mEdhoc{} itself has limited functionality, few negotiable parameters and
-processing rules.
+The core \mEdhoc{} specification has limited functionality, few negotiable
+parameters and processing rules.
 %
@@
 %
-However, \mEdhoc{} leaves much of the semantics of its messages undefined.
+However, \mEdhoc{} leaves much of the semantics of its messages undefined or
+specified as (optional) extensions.
+%
+It is therefore difficult to draw the line between what is \mEdhoc{} and what is
+an extension.
+%
+For example, the core specification does not include a
+message transport mechanism.
 %
@@
 %
+While there is some work on DDoS aspects of \mCoap{}, and much on IoT in
+general, this does not capture the particular consequences originating with
+\mEdhoc{}.
+%
 We now elaborate on some considerations application developers needs to take
@@
 
-\paragraph{Amplification and asymmetry.}
+\paragraph{Amplification and role asymmetry.}
 For the same reasons as given in the IKEv2 DDoS analysis
@@
 %
-In addition to the computational resources the responder spends, communication
-patterns in the underlying protocols may require the responder to
+In addition to the computational resources the responder spends, \mEdhoc{}
+execution may require the responder to
 obtain and verify certificate chains, and verify
@@
 
-Implementers are well advised to consult \cite{rfc8019}, which provides much
-good general information about DoS for protocols with this structure.
-%
-However, ultimately, a product-specific risk assessment is required to
-determine appropriate DDoS mitigations.
-%
-
-\paragraph{Reflection Attacks}
+\paragraph{Reflection Attacks.}
 Reflection attacks are attacks where an adversary spoofs the source address of a
@@
 
-\mPoint{In a constrained device randomness may be easily depleted}
-Assume an adversary generate (deterministic) bogus message 1s and send these at
-a sufficiently high rate to the responder.
-%
-\mEdhoc{} specifically targets constrained devices, that presumably may not
-always have good randomness sources.
-%
-Is it possible to deplete or significantly reduce the entropy of the source and
-in this way be able to predict the responder's random input to other sessions
-(with other initiators)?
-%
-The LAKE WG have been talking about effects of poor randomness on signatures, so
-this should be an issue, no?
-%
+%\mPoint{In a constrained device randomness may be easily depleted}
+%Assume an adversary generate (deterministic) bogus message 1s and send these at
+%a sufficiently high rate to the responder.
+%%
+%\mEdhoc{} specifically targets constrained devices, that presumably may not
+%always have good randomness sources.
+%%
+%Is it possible to deplete or significantly reduce the entropy of the source and
+%in this way be able to predict the responder's random input to other sessions
+%(with other initiators)?
+%%
+%The LAKE WG have been talking about effects of poor randomness on signatures, so
+%this should be an issue, no?
+%%
 
 \mPoint{\mEdhoc{} can deadlock if not used correctly}
+
 An adversary may send forged error messages, and specifically reject proposed
@@
 
-%-------------------------------------------------------------------------- sec
+Implementers are well advised to consult \cite{rfc8019}, which provides much
+good general information about DoS for key establishment
+protocols with the 3-message structure employed by \mEdhoc{}.
+%
+However, ultimately, a product-specific risk assessment is required to
+determine appropriate DDoS mitigations.
+%
+
 \section{\uppercase{Discussion}}

### 2022-07-10 Journal Version/journal-version.tex
Added files with the new text

--- 2022-06-30 paper.tex
@@
 
-\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices}
+%\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices}
+\title{Analyzing the EDHOC Key Establishment Protocol using TAMARIN}
 \author{
@@
 \begin{abstract}
-Constrained IoT devices are becoming ubiquitous in society
-and there is a need for secure communication protocols that respect the
-constraints under which these devices operate.
+Given how common IoT devices that use constrained resources are becoming today, the need of the hour is communication protocols which can operate securely under such limitations.
 %
-\mEdhoc{} is an authenticated key establishment protocol for constrained IoT
-devices, currently being standardized by the Internet Engineering Task
-Force (IETF).
+For a few years, the Internet Engineering Task Force (IETF) has been working to standardize \mEdhoc{}, an authenticated key establishment protocol for such constrained IoT devices.
 %
-A rudimentary version of \mEdhoc{} with only two key establishment methods was
-formally analyzed in 2018.
+The first version of \mEdhoc{} was proposed in 2016.
 %
-Since then, the protocol has evolved significantly and several new key
-establishment methods have been added.
+In 2018, Bruni et al~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.
 %
-In this paper, we present a formal analysis of all \mEdhoc{} methods in an
-enhanced symbolic Dolev-Yao model using the \mTamarin{} tool.
+%By now, the protocol has been fleshed out much more, and has been augmented with multiple new key establishment methods.
+By 2021, the protocol had been fleshed out much more, with multiple new key establishment methods, and this version was formally analyzed using the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
 %
-We show that not all methods satisfy the authentication notion
-injective agreement, but that all do satisfy a notion of implicit
-authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
+Here, we build on this work, and use \mTamarin{} to formally analyze the key establishment methods in the current version of \mEdhoc{}, as well as discuss some ramifications of the choices made while designing the protocol.
+\vnote{We need to see what more we can say here, about the new stuff we're potentially adding}
 %
-We identify other weaknesses to which we propose improvements
-%
-For example, a party may intend to establish a session key with a certain peer,
-but ending up establishing it with another, trusted but compromised, peer.
-%
-We communicated our findings and proposals to the IETF, which
-has incorporated some of these in newer versions of the standard.
+%We show that not all methods satisfy the authentication notion
+%injective agreement, but that all do satisfy a notion of implicit
+%authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
+%%
+%We identify other weaknesses to which we propose improvements
+%%
+%For example, a party may intend to establish a session key with a certain peer,
+%but ending up establishing it with another, trusted but compromised, peer.
+%%
+%We communicated our findings and proposals to the IETF, which
+%has incorporated some of these in newer versions of the standard.
+%Constrained IoT devices are becoming ubiquitous in society
+%and there is a need for secure communication protocols that respect the
+%constraints under which these devices operate.
+%%
+%\mEdhoc{} is an authenticated key establishment protocol for constrained IoT
+%devices, currently being standardized by the Internet Engineering Task
+%Force (IETF).
+%%
+%A rudimentary version of \mEdhoc{} with only two key establishment methods was
+%formally analyzed in 2018.
+%%
+%Since then, the protocol has evolved significantly and several new key
+%establishment methods have been added.
+%%
+%In this paper, we present a formal analysis of all \mEdhoc{} methods in an
+%enhanced symbolic Dolev-Yao model using the \mTamarin{} tool.
+%%
+%We show that not all methods satisfy the authentication notion
+%injective agreement, but that all do satisfy a notion of implicit
+%authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
+%%
+%We identify other weaknesses to which we propose improvements
+%%
+%For example, a party may intend to establish a session key with a certain peer,
+%but ending up establishing it with another, trusted but compromised, peer.
+%%
+%We communicated our findings and proposals to the IETF, which
+%has incorporated some of these in newer versions of the standard.
 \end{abstract}
@@
 \label{sec:introduction}
+%%%%%%% START HERE %%%%%%%%%%
+\input{intro}
 
-As IoT devices become more prevalent and get involved in progressively sensitive
-functions in society, the need to secure their communications
-becomes increasingly important.
+%As IoT devices become more prevalent and get involved in progressively sensitive
+%functions in society, the need to secure their communications
+%becomes increasingly important.
+%%
+%Most security analysis has focused on computationally
+%strong devices, such as cars and web-cameras, where existing protocols like
+%\mDandTls{} suffice.
+%%
+%Constrained devices, on the other hand, which operate under severe
+%bandwidth and energy consumption restrictions, have received much less
+%attention.
+%%
+%These devices may be simple sensors, which only relay environment
+%measurements to a server every hour, but need to function autonomously without
+%maintenance for long periods of time.
+%%
+%The IETF standardized the Object Security for
+%Constrained RESTful Environments (\mOscore{}) protocol to secure communications
+%between constrained devices~\cite{rfc8613}.
+%%
+%However, the \mOscore{} protocol requires a pre-established security context.
+%%
+%The IETF has been discussing requirements and mechanisms for a key
+%exchange protocol, named Ephemeral Diffie-Hellman Over COSE (\mEdhoc), for
+%establishing \mOscore{} security contexts.
+%%
+%Naturally, \mEdhoc{} must work under the same constrained requirements as
+%\mOscore{} itself.
+%%
+%While not all use cases for \mEdhoc{} are firmly set, the overall goal is to
+%establish an \mOscore{} security context, under message size limitations.
+%%
+%It is therefore important to ensure that \mEdhoc{} satisfies fundamental
+%security properties expected from a key exchange protocol.
+%%The work is now being done in the IETF Lightweight Authenticated Key Exchange
+%%(LAKE) working group.
+%%
 %
-Most security analysis has focused on computationally
-strong devices, such as cars and web-cameras, where existing protocols like
-\mDandTls{} suffice.
+%The first incarnation of \mEdhoc{} appeared in March 2016.
+%%
+%It contained two different key establishment methods, one based on a
+%pre-shared Diffie-Hellman (DH) cryptographic core%
+%\footnote{By a \emph{cryptographic core}, or simply core, we mean an academic protocol,
+%without encodings or application specific details required by an industrial
+%protocol.
+%%
+%A \emph{key establishment method} is a core with some such details added.
+%}
+%and a second based on a
+%variant of challenge-response signatures in the style of
+%\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
+%%
+%\mEdhoc{} is therefore a framework of several key establishment methods.
+%%
+%In May 2018, the core based on challenge-response signatures was replaced by
+%one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}.
+%%
+%Since then the protocol has undergone significant changes.
+%%
+%Three new cores, mixing challenge-response signatures and regular signature for
+%authentication, were added~\cite{our-analysis-selander-lake-edhoc-00}.
+%%
 %
-Constrained devices, on the other hand, which operate under severe
-bandwidth and energy consumption restrictions, have received much less
-attention.
+%We formulate and formalize a security model covering all four key
+%establishment methods, which is important especially since the
+%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} lacks a clear description
+%of the intended security model and overall security goals.
+%%
 %
-These devices may be simple sensors, which only relay environment
-measurements to a server every hour, but need to function autonomously without
-maintenance for long periods of time.
+%We perform the analysis in a symbolic Dolev-Yao model.
+%In this framework, we model messages as terms in an algebra, with operations such
+%as encryption, modelled as functions on these terms.
+%These functions are assumed perfect, e.g., one cannot decrypt an encrypted
+%message without access to the key.
+%The adversary, while unable to break encryption or reverse hashing, is modelled
+%as the network.
+%That is, the adversary, can block reroute, replay and modify messages at will.
+%A symbolic model like this, while slightly severe an abstraction, still allows
+%us to analyze \mEdhoc{} for logical flaws without incurring the complexity of
+%a computational model.
+%%
+%The standardization process is ongoing, with the authors releasing newer
+%versions of the \mSpec{} (see Section~\ref{sec:newdrafts} for more detail
+%about how these versions differ from the one analyzed here).
+%%
 %
-The IETF standardized the Object Security for
-Constrained RESTful Environments (\mOscore{}) protocol to secure communications
-between constrained devices~\cite{rfc8613}.
+%%-------------------------------------------------------------------------- sub
+%\subsection{Contributions}
+%\label{sec:contributions}
+%In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
+%establishment methods) using the \mTamarin{} tool~\cite{DBLP:conf/cav/MeierSCB13}.
+%%
+%We present a formal model we constructed of the protocol as given in the
+%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00}.
+%%
 %
-However, the \mOscore{} protocol requires a pre-established security context.
+%We give an explicit adversary model for the protocol and verify
+%properties such as session key material and entity authentication, and perfect
+%forward secrecy, for all four methods.
+%%
 %
-The IETF has been discussing requirements and mechanisms for a key
-exchange protocol, named Ephemeral Diffie-Hellman Over COSE (\mEdhoc), for
-establishing \mOscore{} security contexts.
+%The model itself is valuable as a basis for verifying further updates in the
+%ongoing standardization.
+%%
+%It is publicly available~\cite{edhocTamarinRepo}.
+%%
+%It took several person-months to interpret the
+%specification and construct the model.
+%%
+%Termination requires a hand-crafted proof oracle to guide \mTamarin{}.
+%%
 %
-Naturally, \mEdhoc{} must work under the same constrained requirements as
-\mOscore{} itself.
+%We show that not all \mEdhoc{}'s key establishment
+%methods provide authentication according to the injective agreement definition
+%on the session key material, and none on the initiator's identity.
+%%
+%However, we show that all methods fulfill an implicit agreement property
+%covering the session key material and the initiator's identity.
+%%
+%We identify a number of subtleties, ambiguities and weaknesses in the
+%specification.
+%%
+%For example, the authentication policy requirements allows situations where a
+%party establishes session key material with a trusted but compromised peer, even
+%though the intention was to establish it with a different trusted party.
+%%
+%We provide remedies for the identified issues and have
+%communicated these to the IETF and the specification authors, who incorporated
+%some of our suggestions and currently consider how to deal with the remaining
+%ones.
+%%
 %
-While not all use cases for \mEdhoc{} are firmly set, the overall goal is to
-establish an \mOscore{} security context, under message size limitations.
+%%-------------------------------------------------------------------------- sub
+%\subsection{Comparison with Related Work}
+%The May 2018 version of \mEdhoc{} was formally analyzed by
+%\cite{DBLP:conf/secsr/BruniJPS18} using the \mProverif{}
+%tool~\cite{DBLP:conf/csfw/Blanchet01}.
+%%
+%Their analysis covered a pre-shared key authenticated core and one
+%based on \mSigma.
+%%
+%The properties checked for therein were secrecy, PFS and integrity of
+%application data, identity protection against an active adversary,
+%and strong authentication.
+%%
 %
-It is therefore important to ensure that \mEdhoc{} satisfies fundamental
-security properties expected from a key exchange protocol.
-%The work is now being done in the IETF Lightweight Authenticated Key Exchange
-%(LAKE) working group.
-%
-
-The first incarnation of \mEdhoc{} appeared in March 2016.
-%
-It contained two different key establishment methods, one based on a
-pre-shared Diffie-Hellman (DH) cryptographic core%
-\footnote{By a \emph{cryptographic core}, or simply core, we mean an academic protocol,
-without encodings or application specific details required by an industrial
-protocol.
-%
-A \emph{key establishment method} is a core with some such details added.
-}
-and a second based on a
-variant of challenge-response signatures in the style of
-\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
-%
-\mEdhoc{} is therefore a framework of several key establishment methods.
-%
-In May 2018, the core based on challenge-response signatures was replaced by
-one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}.
-%
-Since then the protocol has undergone significant changes.
-%
-Three new cores, mixing challenge-response signatures and regular signature for
-authentication, were added~\cite{our-analysis-selander-lake-edhoc-00}.
-%
-
-We formulate and formalize a security model covering all four key
-establishment methods, which is important especially since the
-\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} lacks a clear description
-of the intended security model and overall security goals.
-%
-
-We perform the analysis in a symbolic Dolev-Yao model.
-In this framework, we model messages as terms in an algebra, with operations such
-as encryption, modelled as functions on these terms.
-These functions are assumed perfect, e.g., one cannot decrypt an encrypted
-message without access to the key.
-The adversary, while unable to break encryption or reverse hashing, is modelled
-as the network.
-That is, the adversary, can block reroute, replay and modify messages at will.
-A symbolic model like this, while slightly severe an abstraction, still allows
-us to analyze \mEdhoc{} for logical flaws without incurring the complexity of
-a computational model.
-%
-The standardization process is ongoing, with the authors releasing newer
-versions of the \mSpec{} (see Section~\ref{sec:newdrafts} for more detail
-about how these versions differ from the one analyzed here).
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{Contributions}
-\label{sec:contributions}
-In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
-establishment methods) using the \mTamarin{} tool~\cite{DBLP:conf/cav/MeierSCB13}.
-%
-We present a formal model we constructed of the protocol as given in the
-\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00}.
-%
-
-We give an explicit adversary model for the protocol and verify
-properties such as session key material and entity authentication, and perfect
-forward secrecy, for all four methods.
-%
-
-The model itself is valuable as a basis for verifying further updates in the
-ongoing standardization.
-%
-It is publicly available~\cite{edhocTamarinRepo}.
-%
-It took several person-months to interpret the
-specification and construct the model.
-%
-Termination requires a hand-crafted proof oracle to guide \mTamarin{}.
-%
-
-We show that not all \mEdhoc{}'s key establishment
-methods provide authentication according to the injective agreement definition
-on the session key material, and none on the initiator's identity.
-%
-However, we show that all methods fulfill an implicit agreement property
-covering the session key material and the initiator's identity.
-%
-We identify a number of subtleties, ambiguities and weaknesses in the
-specification.
-%
-For example, the authentication policy requirements allows situations where a
-party establishes session key material with a trusted but compromised peer, even
-though the intention was to establish it with a different trusted party.
-%
-We provide remedies for the identified issues and have
-communicated these to the IETF and the specification authors, who incorporated
-some of our suggestions and currently consider how to deal with the remaining
-ones.
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{Comparison with Related Work}
-The May 2018 version of \mEdhoc{} was formally analyzed by
-\cite{DBLP:conf/secsr/BruniJPS18} using the \mProverif{}
-tool~\cite{DBLP:conf/csfw/Blanchet01}.
-%
-Their analysis covered a pre-shared key authenticated core and one
-based on \mSigma.
-%
-The properties checked for therein were secrecy, PFS and integrity of
-application data, identity protection against an active adversary,
-and strong authentication.
-%
-
-In contrast to the key establishment methods analyzed by Bruni et~al., which
-were based on the well-understood pre-shared key DH and \mSigma{} protocols,
-the three newly added
-methods combine two unilateral authentication protocols with the goal to
-constructing mutual authentication protocols.
-%
-Combining two protocols, which individually provide unilateral authentication,
-is not guaranteed to result in a secure mutual authentication
-protocol~\cite{DBLP:conf/ccs/Krawczyk16}.
-%
-Consequently, even though the framework is similar to the one analyzed by Bruni
-et~al., the cryptographic underpinnings have significantly increased in
-complexity, and is using mechanisms which have not previously been formally analyzed.
-%
-The set of properties we check for is also different.
-%
-Our analysis is further carried out using a different tool,
-namely \mTamarin; different kinds of strategies to formulate and
-successfully analyze the protocol are required when working with this tool.
+%In contrast to the key establishment methods analyzed by Bruni et~al., which
+%were based on the well-understood pre-shared key DH and \mSigma{} protocols,
+%the three newly added
+%methods combine two unilateral authentication protocols with the goal to
+%constructing mutual authentication protocols.
+%%
+%Combining two protocols, which individually provide unilateral authentication,
+%is not guaranteed to result in a secure mutual authentication
+%protocol~\cite{DBLP:conf/ccs/Krawczyk16}.
+%%
+%Consequently, even though the framework is similar to the one analyzed by Bruni
+%et~al., the cryptographic underpinnings have significantly increased in
+%complexity, and is using mechanisms which have not previously been formally analyzed.
+%%
+%The set of properties we check for is also different.
+%%
+%Our analysis is further carried out using a different tool,
+%namely \mTamarin; different kinds of strategies to formulate and
+%successfully analyze the protocol are required when working with this tool.
 %
@@
 \label{sec:edhoc}
-\input{edhocProtocol}
+\input{protocol}
 

### 2022-07-12 Journal Version/protocol.tex

% !TEX root = journal-version.tex
In this section, we describe the various key establishment methods of the \mEdhoc{} protocol.
%
Following~\cite{Norr21}, we refer to the two roles executing the protocol as the initiator $I$ and the responder $R$.
%
We annotate values with $I$ and $R$ to make explicit which role they belong to.
%

We denote by $\langle d_{t, \mathit{id}}, Q_{t, \mathit{id}}\rangle$ public-private key pairs, where $d$ is the private key, $Q$ is the public key, $t \in \{e, s\}$ indicates whether the key is ephemeral or static, and $\mathit{id}$ indicates the party to whom the key pair belongs.
%
When clear from context, we will often drop part (or all) of the subscript.
%
Static key pairs (suitable for both regular and challenge-response signatures) are long-term authentication credentials, while ephemeral key pairs are those which are generated afresh for each instantiation of the protocol.
%

Parties can authenticate using regular signatures or challenge-response signatures. 
%
In the former case, we say that they use the \emph{signature based authentication method} (\mSig{}).
%
In the latter, we say, following the terminology in the \mSpec{}, that they use the \emph{static key authentication method} (\mStat{}).
%
For more details about the challenge-response terminology, see~\cite{DBLP:conf/crypto/Krawczyk05}.

%%------------------------------------------------------------------------- sub
%

\mEdhoc{} relies on \mCose{}~\cite{rfc8152} for elliptic curve operations and
transforming points into bitstrings, and we therefore abstract those as
follows.
%
Signatures and verification thereof using party $A$'s key pair are
denoted \mSign{A}$(\cdot)$ and \mVf{A}$(\cdot)$ respectively.
%
The DH-primitive combining a private key $d$ and a point $P$ is denoted
$\mDH(d,P)$.
%
We abuse notation and let these function symbols denote operations on both
points and the corresponding bitstrings.
%

%------------------------------------------------------------------------- sub
\subsection{Framework Structure}
\label{sec:framework}
\mEdhoc{}'s goal is to establish an \mOscore{} security context,
including session key material \mbox{denoted \mSessKey{}}, and optionally transfer
application data \mADone{}, \mADtwo{} and \mADthree{}.
%
To accomplish this, the
\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} gives
a three-message protocol pattern, shown in Figure~\ref{fig:edhocFramework}.
%
We first describe this pattern and the parts that are common to all key
establishment methods.
%
Then we describe authentication and derivation of keys in more detail.
%
The latter is what differentiate the key establishment methods from each other.
%

\subsubsection{Protocol Pattern}
The first two messages negotiate authentication methods \mMethod{} and a
ciphersuite \mSuites{}.
%
In \mMethod{}, the initiator $I$ proposes which authentication method each
party should use.
%
These may differ, leading to four possible combinations:
\mSigSig{}, \mSigStat{}, \mStatSig{} and \mStatStat{}.
%
We refer to these \emph{combinations} of authentication methods simply as
\emph{methods} to align with the \mSpec{} terminology.
%
The first authentication method in a combination is the one proposed for the
initiator and the last is the one proposed for the responder.
%
The responder $R$ may reject the choice of method
or cipher suite with an error message, resulting in negotiation across
multiple \mEdhoc{} sessions.
%
Our analysis excludes error messages.
%

\mEdhoc{}'s first two messages also exchange connection identifiers \mCi{} and
\mCr{}, and public ephemeral keys, \mGx{} and \mGy{}.
%
The connection identifiers \mCi{} and \mCr{}, described in Section 3.1 of the
\mSpec{}, deserve some elaboration.
%
The \mSpec{} describes these identifiers as not serving a security purpose for
\mEdhoc{}, but only as aiding message routing to the correct \mEdhoc{} processing
entity at a party.
%
Despite this, the \mSpec{} states that they may be used by \mOscore{}, or other
protocols using the established security context, without restricting how they
are to be used.
%
Because \mEdhoc{} may need them in clear-text for routing, \mOscore{} cannot
rely on them being secret.
%
Section 7.1.1 of the \mSpec{} requires the identifiers to be unique.
%
Uniqueness is defined to mean that $\mCi{} \not = \mCr{}$ for a given session
and the \mSpec{} requires parties to verify that this is the case.
%
The same section also require that \mOscore{} must be able to retrieve the
security context based on these identifiers.
%
The intended usage of \mCi{} and \mCr{} by \mOscore{} is not made specific and
therefore it is not clear which properties should be verified.
%
We verify that the parties agree on the established values.
%

The two last messages provide identification and authentication.
%
Parties exchange identifiers for their long-term keys, \mIdcredi{} and \mIdcredr{},
as well as information elements, \mAuthi{} and \mAuthr{}, to authenticate
that the parties control the corresponding long-term keys.
%
The content of \mAuthi{} and \mAuthr{} depends on the authentication method
associated with the corresponding long-term key.
%
For example, if $\mMethod{} = \mSigStat{}$, the responder $R$ must either
reject the offer or provide an \mIdcredr{} corresponding to a key pair
$\langle\mPriv{R},\ \mPub{R}\rangle$ suitable for use with challenge-response
signatures, and
compute \mAuthr{} based on the static key authentication method \mStat{}.
%
In turn, the initiator $I$ must provide an \mIdcredi{} corresponding to a key
pair $\langle\mPriv{I},\ \mPub{I}\rangle$ suitable for a regular signature,
and compute
\mAuthi{} based on the signature based authentication method \mSig{}.
%

\subsubsection{Authentication}
Regardless of whether \mStat{} or \mSig{} is used to compute \mAuthr{}, a
MAC is first computed over \mIdcredr{}, \mCredr{}, a transcript hash of the
information exchanged so far, and \mADtwo{} if included.
%
The MAC is the result of encrypting the empty string with the Authenticated
Encryption with Additional Data (AEAD) algorithm from the ciphersuite
\mSuites{}, using the mentioned information as additional data.
%
The MAC key is derived from the ephemeral key material
\mGx{}, \mGy{}, \mX{} and \mY{}, where $I$
computes $\mDH(\mX,\ \mGy)$ and $R$ computes $\mDH(\mY,\ \mGx)$, both resulting in
the same output in the usual DH way.
%

In case $R$ uses the \mSig{} authentication method, \mAuthr{} is $R$'s
signature over the MAC itself and the data that the MAC already covers.
%
In case $R$ uses the \mStat{} authentication method, \mAuthr{} is simply the
MAC itself.
%
However, when using \mStat{}, the key for the MAC is derived, not only from the
ephemeral key material, but also from $R$'s long-term key
$\langle\mPriv{R},\ \mPub{R}\rangle$.
%
For those familiar with \mOptls, this corresponds to the 1-RTT semi-static
pattern computing the MAC key \textsf{sfk} for the \textsf{sfin}
message~\cite{DBLP:conf/eurosp/KrawczykW16}.
%
The content of \mAuthi{} is computed in the corresponding way for the initiator
$I$.
%
In Figure~\ref{fig:edhocFramework}, we denote a MAC using a key derived from
both $\langle\mPriv{R},\ \mPub{R}\rangle$ and $\langle\mX,\ \mGx\rangle$ by
$\mathit{MAC}_I$, and a MAC using a key derived from
both $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\ \mGy\rangle$ by
$\mathit{MAC}_R$.
%

Parts of the last two messages are encrypted and integrity protected, as
indicated in~\ref{fig:edhocFramework}.
%
The second message is encrypted by XORing the output of the key derivation
function HKDF (see Section~\ref{sec:keysched}) on to the plain text.
%
The third message is encrypted and integrity protected by the AEAD algorithm
determined by the ciphersuite \mSuites{}.
%

\begin{figure}
\centering
\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
\begin{tikzpicture}
    \node (ini) at (0, 0) {Initiator};
    \draw [very thick] (0, -0.25) -- (0,-2.3);
    \draw [very thick] (5.75, -0.25) -- (5.75,-2.3);
    \node (res) at (5.75,0) {Responder};
    \msg{1em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
    \msg{3em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, \{\mIdcredr, \mAuthr, \mADtwo\}};
    \msg{5em}{ini}{res}{\mMsgthree: \mCr, \{\mIdcredi, \mAuthi, \mADthree\}};
    \draw [line width=1mm] (-0.75,-2.3) -- (0.75,-2.3);
    \draw [line width=1mm] (5.75-0.75,-2.3) -- (5.75+0.75,-2.3);
    \node (padding) at (0,-2.5) {};
    \end{tikzpicture}
    \begin{tabular}{|c|c|c|}
        \hline
        \mMethod & \mAuthi & \mAuthr\\
        \hline
        \mSigSig{} & $\mSign{I}(\cdot)$ & $\mSign{R}(\cdot)$ \\
        \mSigStat{} & $\mSign{I}(\cdot)$ & $\textit{MAC}_R(\cdot)$\\
        \mStatSig{} & $\textit{MAC}_I(\cdot)$ & $\mSign{R}(\cdot)$\\
        \mStatStat{} & $\textit{MAC}_I(\cdot)$ & $\textit{MAC}_R(\cdot)$\\
        \hline
    \end{tabular}
    \caption{Structure of \mEdhoc{}: $\{t\}$ means $t$ is encrypted and integrity
protected.}
\label{fig:edhocFramework}
\end{figure}

\subsubsection{Key Schedule}
\label{sec:keysched}
At the heart of \mEdhoc{} is the key schedule depicted in
Figure~\ref{fig:kdfdiagram}.
%
\mEdhoc{} uses two functions from the \mHkdf{} interface~\cite{rfc5869} to derive keys.
%
\mHkdfExtract{} 
constructs uniformly distributed key material from random input and a salt,
while \mHkdfExpand{} generates keys from key material and a salt.
%

The key schedule is rooted in the ephemeral DH key
\mGxy{}, which is computed as $\mDH(\mX, \mGy)$ by $I$ and as $\mDH(\mY, \mGx)$
by $R$.
%
From \mGxy{}, three intermediate keys \mPRKtwo, \mPRKthree{} and
\mPRKfour{} are derived during the course of protocol execution.
%
Each of them is used for a specific message in the protocol, and from
these intermediate keys, encryption and integrity keys
(\mKtwoe, \mKtwom{}, \mKthreeae, and \mKthreem) for that message are derived.
%
The salt for generating \mPRKtwo{} is the empty string.
%

The protocol uses a running transcript hash $th$, which includes the information
transmitted so far.
%
The value of the hash, denoted $th_i$ for the $i$th message, is included in key
derivations as shown in Figure~\ref{fig:kdfdiagram}.
%

Successful protocol execution establishes the session key material \mSessKey{}
for \mOscore{}.
%
\mSessKey{} can be considered a set that always includes \mGxy{}.
%
If the initiator uses the \mStat{} authentication method, \mSessKey{} also
includes
$\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$, which we denote $P_I$.
%
If the responder uses the \mStat{} authentication method, it also includes
$\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\ \mGx)$, which we denote $P_R$.
%
From the session key material, a key exporter (\mEdhoc-Exporter) based on
\mHkdf{} is used to extract keys required for \mOscore{}.
%

\begin{figure*}[!h]
\centering
\scalebox{.785}{
\input{kdfdiagram.tex}
}
\caption{Key schedule: Light blue boxes hold DH keys ($P_e, P_I, P_R$),
orange boxes intermediate key material (\mPRKtwo, \mPRKthree, \mPRKfour), and
dark blue boxes keys for \mAead{} or \mXor{} encryption
(\mKtwoe, \mKtwom, \mKthreeae, \mKthreem).
Dashed boxes are conditionals.}
\label{fig:kdfdiagram}
\end{figure*}

As an illustrative example of the entire process, we refer to
Figure~\ref{fig:edhocsigstat}, which depicts the protocol pattern, operations
and key derivations for the \mSigStat{} method in more detail.
%
\begin{figure*}[ht]
\centering
\scalebox{.7}{
\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
\begin{tikzpicture}
    \node (ini) at (0, 0) {Initiator};
    \draw [very thick] (0, -0.5) -- (0,-14.8);
    \draw [very thick] (9, -0.5) -- (9,-14.8);
    \node[below=0.5em of ini,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{I},\ \mPub{I}\rangle,\ \mIdcredi,\ \mIdcredr,\ \mADone,\ \mADthree
    \end{array}
    $};
    \node (res) at (9,0) {Responder};
    \node[below=0.5em of res,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{R},\ \mPub{R}\rangle, \ \mIdcredr,\ \mADtwo
    \end{array}$};
    \action{3.5em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\ \mGx\rangle$};
    \msg{6.5em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
    \action{7.0em}{res}{$
      \begin{array}{c}
          \text{Generates } \mCr,\ \langle\mY{},\ \mGy\rangle\\
          \ \ P_e = \mDH(\mY,\ \mGx{})\\
          \ \ P_R = \mDH(\mPriv{R},\ \mGx{})\\
        \mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle)\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        \mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mMactwo = \mAead(\mKtwom; \langle \mIdcredr, \mTHtwo, \mCredr, \mADtwo \rangle; \textrm{``\phantom{}''}) \\
        \mKtwoe = \mHkdfExpand(\mPRKtwo, \mTHtwo)
      \end{array}$};
    \msg{21.7em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\ \mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
    \action{22.5em}{ini}{$
      \begin{array}{c}
        %\mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle) \
        \ P_e = \mDH(\mX,\ \mGy{})\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        %\mKtwoe = \mHkdfExpand(\mPRKtwo,\mTHtwo)\\
        %\mGrx = \mCredr^{x} \\
        \ \ P_R = \mDH(\mX,\ \mPub{R})\\
        \mPRKfour = \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        %\mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKfour, \mTHthree) \\
        \mMacthree = \mAead(\mKthreem; \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle; \textrm{``\phantom{}''}) \\
        \mSigthree = \mSign{I}(\langle \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle, \mMacthree \rangle)
      \end{array}$};
    \msg{35.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae; \mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
    \action{36em}{res}{$
    \begin{array}{c}
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKthree, \mTHthree) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHthree)
    \end{array}$};
    \draw [line width=1mm] (-2,-14.8) -- (2,-14.8);
    \draw [line width=1mm] (7,-14.8) -- (11,-14.8);
    \end{tikzpicture}
}
    \caption{The \mSigStat{} method.
    Tuples are denoted $\langle\cdot\rangle$, and the hash function \mHash{} is
    as determined by \mSuites{}.}
\label{fig:edhocsigstat}
\end{figure*}
%------------------------------------------------------------------------- sub

### 2022-07-12 paper.tex

@@
 
+%-------------------------------------------------------------------------- sec
+\section{\uppercase{Error handling}}
+\label{sec:errorHandling}
+\mPoint{Error handling is undefined except for algo negotiation}
+Section 6 of the \mEdhoc{} specification\footnote{draft-ietf-lake-edhoc-14}
+states that error messages can be sent at any time and by any party.
+%
+One error message type is defined (for algorithm negotiation), and the rest can
+be arbitrarily defined by the implementation.
+%
+The contents and semantics of these messages are also implementation defined.
+%
+Consequently, the error messages provide a generic message passing mechanism.
+%
+
+\textbf{Q:} Is htere any situation where it makes sense that the initiator sends
+a unsupported ciphersuite message instead of message 3?
+%
+
+\textbf{Q:} In Section H: if no message correlation support by the transport
+layer, \mCi{} and \mCr{} should be used for correlation of messages to a
+session.
+%
+Error messages do not carry \mCi{} or \mCr{}; will this create a problem?
+%
+
+\mEdhoc{} provides the error message mechanism, but gives little to no guidance
+on how it should be used safely.
+%
+Example of issue with improper use: Assume application logs error messages a
+finite log with log rotation.
+%
+If the log is used for anomaly detection or detection of sensitive events, then
+an adversary can simply inject error messages and fill the log until the
+sensitive event is overwritten.
+%
+If the filling error messages are less suspicious than the actual attack, this
+may be beneficial for the adversary.
+%
+Implementors should do proper log-separation and log management.
+%
+
+Section 6.2 gives an excample of sending an error message "method not
+supported".
+%
+This message cannot be used to negotiate method securely.
+%
+The reason is that error messages are not integrity protected, and an initator
+only proposes one method, and therefore the same tactic used for ciphersuite
+negotiation cannot be used for methods.
+%
+Because method negotiation can be viewed in the context of crypto agility, there
+is room for improvement.
+%
+
+%-------------------------------------------------------------------------- sub
 \section{\uppercase{Discussion}}

### 2022-07-12 intro.tex
Moved content out into paper.tex and added intro.tex and protocol.tex to git

--- 2022-07-10 Journal Version/journal-version.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.21 of 2022/01/12
+% !TEX root = journal-version.tex
+Traditionally, formal analysis for security protocols has been centered on devices with little to no resource constraints, where protocols like \mDandTls{} can be easily employed. 
 %
-\documentclass[runningheads]{llncs}
+However, as IoT gains popularity and is used for many applications, including potentially sensitive ones, it becomes paramount to analyze these protocols to ensure that they do not admit security flaws.
+%
+IoT protocols are often run on devices which operate under severe restrictions on resources like bandwidth and energy consumption.
+These devices are often simple in their operation, but need to communicate and function without human interference or maintenance for extended periods of time.
+%
+Other protocols have been proposed to secure communications between constrained devices.
+One such, which has been standardized by the IETF is the Object Security for Constrained RESTful Environments (\mOscore{}) protocol.
+%
+However, \mOscore{} requires the pre-establishment of a security context.
+To this end, a key exchange protocol named Ephemeral Diffie-Hellman Over COSE (\mEdhoc{}) has been under discussion at the IETF.
 
-\usepackage{epsfig}
-\usepackage{subcaption}
-\usepackage{calc}
-\usepackage{amssymb}
-\usepackage{amstext}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{multicol}
-\usepackage{pslatex}
-\usepackage{apalike}
+%
+Since \mEdhoc{} will establish security contexts for \mOscore{}, the same resource constraints (especially those pertaining to message size) apply to the former as for the latter.
+%
+While establishing security contexts for \mOscore{} is the primary goal for the \mEdhoc{} protocol, there might well be other use cases, which have not been explored in depth yet.
+%
+It is therefore, very important to ensure that most of the fundamental properties expected of key exchange protocols as established in the literature are satisfied by \mEdhoc{} as well.
+%-------------------------------------------------------------------------- sub
+\subsection{Evolution of \mEdhoc}
+\label{sec:edhocevol}
+The first \mEdhoc{} framework was introduced in March 2016. 
+%
+It allowed two different key establishment methods -- one involved a pre-shared Diffie-Hellman (DH) \emph{cryptographic core}, and the other was a variation on challenge-response signatures, {\`a} la \mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
+%
+A \emph{cryptographic core}, often just called a core, is an academic protocol, i.e. with no encodings or application-specific details as needed for an industrial protocol. 
+%
+Once these ingredients are added to a cryptographic core, we obtain a key-establishment method.
+%
+Since then, the protocol has seen multiple changes.
+%
+In May 2018, the designers replaced the challenge-response signature core with one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}, and in 2020, three new cores which mixed challenge-response signatures and ``regular'' authentication signatures were added as well~\cite{our-analysis-selander-lake-edhoc-00}.
 
-\usepackage[scaled=0.8]{helvet}    % Less huge \textsf{functionName}
-\usepackage[misc,geometry]{ifsym} % for letter symbol
-\usepackage{graphicx}
-\usepackage[dvipsnames]{xcolor}
-\usepackage{tikz}
-\usetikzlibrary{trees,snakes,arrows}
-\usetikzlibrary{shapes,chains}
-\usetikzlibrary{positioning}
-\usepackage{hyperref}
-\hypersetup{
-    colorlinks=true,
-    linkcolor=blue,
-    citecolor=blue,
-    urlcolor=blue
-}
-\usepackage[nospace]{cite}
+%-------------------------------------------------------------------------- sub
+\subsection{Related Work and Contributions}
+\label{sub:related}
+The earliest related work is by Bruni et al~\cite{DBLP:conf/secsr/BruniJPS18}, which formally analyzes the May 2018 version of \mEdhoc{} using the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01}. 
+%
+In this paper, the authors analyze two key establishment methods -- one built on a pre-shared key authentication core, and one on \mSigma{}. 
+%
+The authors check  various properties, namely secrecy, identity protection against an active adversary, strong authentication, perfect forward secrecy (PFS), and integrity of application data. 
+%
+In 2021, Norrman et al~\cite{Norr21} analyze the 2020 version of \mEdhoc{} in the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13}. 
+%
+This version of the protocol has five key establishment methods. 
+%
+In~\cite{Norr21}, the authors check for injective agreement, implicit agreement, and perfect forward secrecy for the session key material. 
+%
+The authors also discuss some fallouts of the various design choices made as part of \mEdhoc{}, and the impact of \mEdhoc{} in multiple use-case scenarios.
+%
+\vnote{Need to check this once more later. Also need to see if we can say something about Steve's SAPIC+ paper. CAREFULLY check what follows, and fill in the list of properties!!}
+%
 
-\input{macros}
+%-------------------------------------------------------------------------- sub
+\subsection{Contributions}
+\label{sec:contributions}
+In this paper, we extend the work presented in~\cite{Norr21}. 
+%
+We analyze the version of \mEdhoc{} as in the latest draft~\cite{}.
+%
+We modify the \mTamarin{} models used in that paper to obtain a significant speed-up in the verification time.
+%
+This required us to carefully modify the oracle as well.
+%
+We check for the following properties: \vnote{Fill this in}
+%
+We also consider various issues arising in relation to the use of trusted execution environments (TEEs), denial of service (DoS) attacks, sources of randomness, the negotiation of parameters (which the formal model abstracts away) for establishing the protocol, and other potential attacks and concerns.
+%
 
-%\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices}
-\title{Analyzing the EDHOC Key Establishment Protocol using TAMARIN}
-\author{
-        Karl Norrman and 
-        Vaishnavi Sundararajan and
-        Alessandro Bruni
-}
-
-\begin{document}
-\maketitle
-%
-\begin{abstract}
-Given how common IoT devices that use constrained resources are becoming today, the need of the hour is communication protocols which can operate securely under such limitations.
-%
-For a few years, the Internet Engineering Task Force (IETF) has been working to standardize \mEdhoc{}, an authenticated key establishment protocol for such constrained IoT devices.
-%
-The first version of \mEdhoc{} was proposed in 2016.
-%
-In 2018, Bruni et al~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.
-%
-%By now, the protocol has been fleshed out much more, and has been augmented with multiple new key establishment methods.
-By 2021, the protocol had been fleshed out much more, with multiple new key establishment methods, and this version was formally analyzed using the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
-%
-Here, we build on this work, and use \mTamarin{} to formally analyze the key establishment methods in the current version of \mEdhoc{}, as well as discuss some ramifications of the choices made while designing the protocol.
-\vnote{We need to see what more we can say here, about the new stuff we're potentially adding}
-%
-%We show that not all methods satisfy the authentication notion
-%injective agreement, but that all do satisfy a notion of implicit
-%authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
-%%
-%We identify other weaknesses to which we propose improvements
-%%
-%For example, a party may intend to establish a session key with a certain peer,
-%but ending up establishing it with another, trusted but compromised, peer.
-%%
-%We communicated our findings and proposals to the IETF, which
-%has incorporated some of these in newer versions of the standard.
-%Constrained IoT devices are becoming ubiquitous in society
-%and there is a need for secure communication protocols that respect the
-%constraints under which these devices operate.
-%%
-%\mEdhoc{} is an authenticated key establishment protocol for constrained IoT
-%devices, currently being standardized by the Internet Engineering Task
-%Force (IETF).
-%%
-%A rudimentary version of \mEdhoc{} with only two key establishment methods was
-%formally analyzed in 2018.
-%%
-%Since then, the protocol has evolved significantly and several new key
-%establishment methods have been added.
-%%
-%In this paper, we present a formal analysis of all \mEdhoc{} methods in an
-%enhanced symbolic Dolev-Yao model using the \mTamarin{} tool.
-%%
-%We show that not all methods satisfy the authentication notion
-%injective agreement, but that all do satisfy a notion of implicit
-%authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
-%%
-%We identify other weaknesses to which we propose improvements
-%%
-%For example, a party may intend to establish a session key with a certain peer,
-%but ending up establishing it with another, trusted but compromised, peer.
-%%
-%We communicated our findings and proposals to the IETF, which
-%has incorporated some of these in newer versions of the standard.
-\end{abstract}
-%
-%-------------------------------------------------------------------------- sec
-\section{\uppercase{Introduction}}
-\label{sec:introduction}
-%%%%%%% START HERE %%%%%%%%%%
-\input{intro}
-
-%As IoT devices become more prevalent and get involved in progressively sensitive
-%functions in society, the need to secure their communications
-%becomes increasingly important.
-%%
-%Most security analysis has focused on computationally
-%strong devices, such as cars and web-cameras, where existing protocols like
-%\mDandTls{} suffice.
-%%
-%Constrained devices, on the other hand, which operate under severe
-%bandwidth and energy consumption restrictions, have received much less
-%attention.
-%%
-%These devices may be simple sensors, which only relay environment
-%measurements to a server every hour, but need to function autonomously without
-%maintenance for long periods of time.
-%%
-%The IETF standardized the Object Security for
-%Constrained RESTful Environments (\mOscore{}) protocol to secure communications
-%between constrained devices~\cite{rfc8613}.
-%%
-%However, the \mOscore{} protocol requires a pre-established security context.
-%%
-%The IETF has been discussing requirements and mechanisms for a key
-%exchange protocol, named Ephemeral Diffie-Hellman Over COSE (\mEdhoc), for
-%establishing \mOscore{} security contexts.
-%%
-%Naturally, \mEdhoc{} must work under the same constrained requirements as
-%\mOscore{} itself.
-%%
-%While not all use cases for \mEdhoc{} are firmly set, the overall goal is to
-%establish an \mOscore{} security context, under message size limitations.
-%%
-%It is therefore important to ensure that \mEdhoc{} satisfies fundamental
-%security properties expected from a key exchange protocol.
-%%The work is now being done in the IETF Lightweight Authenticated Key Exchange
-%%(LAKE) working group.
-%%
-%
-%The first incarnation of \mEdhoc{} appeared in March 2016.
-%%
-%It contained two different key establishment methods, one based on a
-%pre-shared Diffie-Hellman (DH) cryptographic core%
-%\footnote{By a \emph{cryptographic core}, or simply core, we mean an academic protocol,
-%without encodings or application specific details required by an industrial
-%protocol.
-%%
-%A \emph{key establishment method} is a core with some such details added.
-%}
-%and a second based on a
-%variant of challenge-response signatures in the style of
-%\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
-%%
-%\mEdhoc{} is therefore a framework of several key establishment methods.
-%%
-%In May 2018, the core based on challenge-response signatures was replaced by
-%one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}.
-%%
-%Since then the protocol has undergone significant changes.
-%%
-%Three new cores, mixing challenge-response signatures and regular signature for
-%authentication, were added~\cite{our-analysis-selander-lake-edhoc-00}.
-%%
-%
-%We formulate and formalize a security model covering all four key
-%establishment methods, which is important especially since the
-%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} lacks a clear description
-%of the intended security model and overall security goals.
-%%
-%
-%We perform the analysis in a symbolic Dolev-Yao model.
-%In this framework, we model messages as terms in an algebra, with operations such
-%as encryption, modelled as functions on these terms.
-%These functions are assumed perfect, e.g., one cannot decrypt an encrypted
-%message without access to the key.
-%The adversary, while unable to break encryption or reverse hashing, is modelled
-%as the network.
-%That is, the adversary, can block reroute, replay and modify messages at will.
-%A symbolic model like this, while slightly severe an abstraction, still allows
-%us to analyze \mEdhoc{} for logical flaws without incurring the complexity of
-%a computational model.
-%%
-%The standardization process is ongoing, with the authors releasing newer
-%versions of the \mSpec{} (see Section~\ref{sec:newdrafts} for more detail
-%about how these versions differ from the one analyzed here).
-%%
-%
-%%-------------------------------------------------------------------------- sub
-%\subsection{Contributions}
-%\label{sec:contributions}
 %In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
@@
 %ones.
-%%
 %
-%%-------------------------------------------------------------------------- sub
+
+%-------------------------------------------------------------------------- sub
 %\subsection{Comparison with Related Work}
@@
 %successfully analyze the protocol are required when working with this tool.
-%
-
-%-------------------------------------------------------------------------- sec
-\section{\uppercase{The \mEdhoc{} Protocol}}
-\label{sec:edhoc}
-\input{protocol}
-
-%-------------------------------------------------------------------------- sec
-\section{\uppercase{Formalization and Results}}
-\label{sec:formalization}
-\input{edhocFormalization}
-
-%-------------------------------------------------------------------------- sec
-\section{\uppercase{Denial of Service}}
-\label{sec:dos}
-\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
-\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
-attacks and delegates the responsibility to handle this to the application.
-%
-\mEdhoc{} itself has limited functionality, few negotiable parameters and
-processing rules
-%
-Hence, we argue in this section that the attack surface
-is small from a DoS perspective.
-%
-However, \mEdhoc{} leaves much of the semantics of its messages undefined.
-%
-This increases the risk of applications vulnerable to DoS even if \mEdhoc{} is
-arguably not vulnerable to DoS itself.
-%
-
-\mPoint{\mEdhoc{} gives a little info of questionable use for developers}
-Section 8.7 of the \mEdhoc{} specification explains the design philosophy w.r.t.
-DoS in two paragraphs and gives two recommendations.
-%
-The first being that mitigation of DoS based on sending proper messages should
-be handled by lower layers, e.g., by checking reachability of the sending
-party\footnote{Karl: this may be another way to perform reflection DoS attacks}.
-%
-The second being that an application may try to determine (in some way) that an
-apparently valid message is in fact probably a forgery and should be ignored or
-cause processing to stop.
-%
-
-\mPoint{Responder is more vulnerable than initiator}.
-For the same reasons as given in the IKEv2 DDoS analysis
-by Nir and Smyslov~\cite{rfc8019} the responder is more vulnerable than the
-initiator.
-%
-From a roles and message passing perspective, IKEv2 is very similar to all
-methods in \mEdhoc{}.
-%
-The RFC provides much good advice for protocols with this structure that
-\mEdhoc{} using applications can reuse.
-%
-We should look more into this.
-%
-
-\mPoint{Amplification}
-An adversary can fill a message 1 with easily generated bogus data and send this
-to a victim responder.
-%
-There is little work for the initiator, but both symmetric and asymmetric key
-operations required by the responder.
-%
-\mEdhoc{} does not protect against this, but refer to the application and its
-transport layer to resolve the problem.
-%
-This is a design decision, see Section 8.7 of \mEdhoc{} specification.
-%
-In addition to the computational resources the responder spends, it may also
-need to obtain certificate chains and verify certificate revocation servers as a
-result of the spoofed message.
-%
-The asymmetry between the adversaries effort and the responder may hence be
-significant.
-%
-Since \mEdhoc{} specifically targets constrained devices this asymmetry needs to
-be considered even more carefully than for IKEv2~\cite{rfc8019}.
-%
-
-Recommend that the role of responder is chosen carefully.
-%
-The responder is most vulnerable.
-%
-For example, is it better that a lock or one of the key cards takes on the
-responder role and becomes vulnerable?
-%
-
-\mPoint{Reflection Attacks}
-Reflection attacks are attacks where an adversary spoofs the source address of a
-victim in a message and thereby tricks the responder to send a message to the
-victim.
-%
-\mEdhoc{} leaves protection against this to the transport protocol and
-application, and suggests using a path reachability method similar to IKEv2.
-%
-In situations where the responder does not need to perform additional
-communications to verify certificates or similar, a reachability mechanism may
-be more expensive than performing the computations.
-\footnote{Find ref of that computation is often cheaper than radio
-transmission, at least energy-wise.}
-%
-However, if round-trip times are large, and half-open states hence are kept
-by the receiver for a long time, a reachability mechanism can trade
-communication for storage.
-\footnote{RFC8019 discuss some numbers as examples of tradeoffs. Should we set
-up a simple model and derive an expression for trade-offs? Could be useful for
-implementors to consider, but low academic value. This would capture also
-amplification.}
-%
-
-\mPoint{In a constrained device randomness may be easily depleted}
-Assume an adversary generate (deterministic) bogus message 1s and send these at
-a sufficiently high rate to the responder.
-%
-\mEdhoc{} specifically targets constrained devices, that presumably may not
-always have good randomness sources.
-%
-Is it possible to deplete or significantly reduce the entropy of the source and
-in this way be able to predict the responder's random input to other sessions
-(with other initiators)?
-%
-The LAKE WG have been talking about effects of poor randomness on signatures, so
-this should be an issue, no?
-%
-
-\mPoint{\mEdhoc{} can deadlock if not used correctly}
-An adversary may send forged error messages, and specifically reject proposed
-ciphersuites.
-%
-The initiator is recommended to not try that ciphersuite with the responder
-again in a new session for a while.
-%
-If the adversary rejects all options for a initiator this way, it can prevent
-communication between the initiator and that responder until the cache is
-cleared.
-%
-Potentially, an adversary could lock a party of a group out of communication all
-together this way.
-%
-This needs to be considered by applications.
-%
-
-\mPoint{Is any of the methods better than the others from a DoS perspective?}
-We considered this for quite some time, but there seems to be little difference
-in this regard.
-%
-
-%-------------------------------------------------------------------------- sec
-\section{\uppercase{Discussion}}
-\label{sec:discussion}
-There are a few places where \mEdhoc{} can be improved,
-which we found during this work and communicated to the authors.
-%
-We discuss them below.
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{Unclear Intended Use}
-\label{sec:unclearProtocolUse}
-%
-The \mEdhoc{} \mSpec{} lists several security goals, but they are
-imprecise and difficult to interpret due to lack of context and intended usage
-descriptions.
-%
-Without knowing how the protocol is to be used,
-it is not clear whether the listed security goals are the most important ones
-for constrained IoT devices.
-%
-
-The abstract goal of \mEdhoc{} is simple: establish an \mOscore{} security
-context using few roundtrips and small messages.
-%
-From that, the design of \mEdhoc{} is mainly driven by what
-can be achieved given the technical restrictions.
-%
-Focusing too much on what can be achieved within given restrictions, and paying
-too little attention to the use cases where the
-protocol is to be used and their specific goals, risks resulting in
-sub-optimal trade-offs and design decisions.
-%
-
-\mEdhoc{} is intended to cover a variety of use cases, many of which are
-difficult to predict today.
-%
-However, this does not
-prevent collecting \emph{typical} use cases and user stories
-to identify more specific security goals that will be important in most cases.
-%
-
-While constructing our model, we made up simple user stories to identify
-security properties of interest.
-%
-Several of these revealed subtleties and undefined aspects of \mEdhoc{}.
-%
-We informed the \mEdhoc{} authors, who addressed these aspects in the
-\mSpec{}.
-%
-
-\subsubsection{(Non-)Repudiation}
-An access control solution for a nuclear power-plant may need to log who is
-passing through a door, whereas it may be undesirable for, say, a coffee
-machine to log a list of users along with their coffee preferences.
-%
-Via this simple thought experiment, we realized that the \mSpec{} did not
-consider the concept of (non)-repudiation.
-%
-In response, the authors of the \mSpec{} added a paragraph discussing how
-different methods relate to (non)-repudiation.
-
-\subsubsection{Unintended Peer Authentication}
-Section~3.2 of the \mSpec{} states that parties must be configured
-with a policy restricting the set of peers they run \mEdhoc{} with.
-%
-However, the initiator is not required to verify that the \mIdcredr{} received
-in the second message is the same as the one intended.
-%
-The following attack scenario is therefore possible.
-%
-
-Suppose someone has configured all devices in their home to be in the
-allowed set of devices, but that one of the devices ($A$) is compromised.
-%
-If another device $B$, initiates a connection to a third device $C$, the
-compromised device $A$ may interfere by responding in $C$'s place, blocking
-the legitimate response from $C$.
-%
-Since $B$ does not verify that the identity indicated in the second message
-matches the intended identity $C$, and device $A$ is part of the allowed set,
-$B$ will complete and accept the \mEdhoc{} run with device $A$ instead of the
-intended $C$.
-%
-The obvious solution is for the initiator to match \mIdcredr{} to the intended
-identity indicated by the application, which we included in our model.
-%
-We have communicated this to the \mEdhoc{} authors and they are considering
-how to resolve the issue.
-%
-
-%------------------------------------------------------------------------- sub
-\subsection{Unclear Security Model}
-We argue that the \mSpec{} gives too little information about what capabilities
-an adversary is assumed to have, and that this leads to unclear design goals and
-potentially sub-optimal design.
-%
-
-Even though \mEdhoc{} incorporates cryptographic cores from different academic
-security protocols, its design does not take into account the adversary models
-for which these protocols were designed.
-%
-For example, \mOptls{}, whose cryptographic core is essentially the same
-as the \mStat{} authentication method, is designed to be secure in the CK
-model~\cite{DBLP:conf/crypto/CanettiK02}.
-%
-The CK security model explicitly separates the secure storage of long-term
-keys from storage of session state and ephemeral keys.
-%
-This is appropriate for modelling the use of secure modules.
-%
-
-The \mEdhoc{} authors indicated to us that it was
-not necessary to consider compromised ephemeral keys separately from
-compromised long-term keys.
-%
-The rationale is that \mSigma{} cannot protect against compromised ephemeral
-keys~\cite{personalCommunication}.
-%
-That rationale is presumably based on the fact that the \mSigSig{} method is
-closely modeled on the \mSigmaI{} variant of \mSigma{}, and that it would be
-preferable to obtain a homogeneous security level among the \mEdhoc{}
-methods.
-%
-That rationale is only true, however, if one restricts attention to session key
-confidentiality of an ongoing session.
-%
-Secure modules provide value in other ways, for example, by allowing
-constructions with Post-Compromise Security (PCS) guarantees.
-%
-We discussed this with the authors, and
-the latest version of the \mSpec{}~\cite{latest-ietf-lake-edhoc-05} includes
-recommendations on storage of long-term keys and operations on these inside a
-secure module.
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{Session Key Material}
-\label{sec:sessionKeyMaterial}
-\mEdhoc{} establishes session key material, from which session keys
-can be derived using the \mEdhoc{}-Exporter.
-%
-The session key material is affected by \mGxy{}, and if a party uses the
-\mStat{} authentication method, also by that party's secret static long-term key.
-%
-As shown in Section~\ref{sec:formalization}, mutual injective agreement cannot
-be achieved for $P_I$.
-%
-If this property is not important for constrained IoT devices which cannot use
-any of the other methods, then one can simply accept that the methods have
-different authentication strengths.
-%
-Otherwise, this is a problem.
-%
-
-We identified three alternatives for resolving this.
-%
-One alternative is to include \mIdcredi{}, or its hash, in the first and
-second messages.
-%
-This would, however, increase message sizes and prevent initiator identity
-protection, which are grave concerns for \mEdhoc{}.
-%
-A second alternative is to not derive the session key material from $P_I$.
-%
-Doing so, however, deviates from the design of \mOptls{} (and similar protocols
-from which the \mStat{}-based methods are derived), where the inclusion of
-$P_I$ plays a crucial part in the security proof of resistance against
-initiator ephemeral key compromise.
-%
-The third alternative is to include a fourth message from responder to initiator,
-carrying a MAC based on a key derived from session key material including $P_I$.
-%
-Successful MAC verification guarantees
-to the initiator that the responder injectively agrees on $P_I$.
-%
-We presented the options to IETF, and they decided to add a
-fourth message as an option in the latest version of the
-\mSpec{}~\cite{latest-ietf-lake-edhoc-05}.
-%
-
-Regardless of how this is handled, we verified that all methods
-enjoy a common, but weaker, property: mutual implicit agreement
-on all of $P_e, P_I$ and $P_R$, where applicable.
-%
-
-%-------------------------------------------------------------------------- sec
-\section{\uppercase{Conclusions and Future Work}}
-\label{sec:conclusions}
-\label{sec:newdrafts}
-We formally modeled all four
-methods of the \mEdhoc{} \mSpec{} using \mTamarin.
-%
-We formulated several important security properties and identified precise
-adversary models in which we verified these.
-%
-The properties are shown in Table~\ref{tab:props}.
-%
-Mutual injective agreement covers the parameters $S_P$:
-responder identity, roles, session key material (except for $P_I$ when
-initiator uses the \mStat{} authentication
-method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
-%
-The responder in addition is ensured agreement on the initiators identity and
-$P_I$, i.e., on the set $S_F$.
-%
-Implicit agreement covers all previously mentioned parameters for both peers.
-%
-Verification of all lemmas, including model validation lemmas, took 42 minutes
-on an Intel Core i7-6500U 2.5GHz using two cores.
-%
-Mutual entity authentication, UKS- and KCI resistance can be inferred
-from the verified properties.
-%
-\begin{table*}[h!]
-        \centering
-        \caption{Verified properties. $S_P$ contains
-            roles, responder identity, session key material (excluding
-            $P_I$), \mCi, \mCr, and \mSuites. $S_F$ is $S_{P}$,
-            the initiator identity, and $P_I$.}
-        \label{tab:props}
-        \begin{tabular}{|l|c|c|c|c|}
-                \hline
-                & \mSigSig & \mSigStat & \mStatSig & \mStatStat \\
-                \hline
-                Injective agreement for I & $S_F$ & $S_F$ & $S_P$ & $S_P$\\
-                Injective agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-                Implicit agreement for I & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-                Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-                PFS for session key material & \cm & \cm & \cm & \cm\\
-                \hline
-        \end{tabular}
-\end{table*}
-
-Further, we identified a situation where initiators may establish an \mOscore{}
-security context with a different party than the application using \mEdhoc{}
-intended, and proposed a simple mitigation.
-%
-We discussed how the IETF may extract and better define security properties to
-enable easier verification.
-
-We verified each method in isolation, and leave as future work to verify whether
-the methods are secure under composition.
-
-%\subsection{Newer Versions of the Specification}
-%\label{sec:newdrafts}
-In this work, we have analyzed the \mEdhoc{} version as of July
-2020~\cite{our-analysis-selander-lake-edhoc-00}.
-%
-There are newer versions, with the most recent version as
-of February 2021~\cite{latest-ietf-lake-edhoc-05}.
-%
-However, the changes to the protocol over these versions are not
-particularly significant for our analysis.
-%
-
-%-------------------------------------------------------------------------- ack
-\section*{ACKNOWLEDGEMENTS}
-This work was partially supported by
-the Wallenberg AI, Autonomous Systems and Software Program (WASP) funded by
-the Knut and Alice Wallenberg Foundation.
-%
-We are grateful to G\"oran Selander, John Mattsson and Francesca Palombini for
-clarifications regarding the specification.
-%
-
-%-------------------------------------------------------------------------- bib
-\bibliographystyle{apalike}
-{\small
-    \bibliography{refComp}
-}
-\end{document}

### 2022-07-12 protocol.tex
Moved content out into paper.tex and added intro.tex and protocol.tex to git

--- 2022-07-12 Journal Version/protocol.tex
@@
 
+%------------------------------------------------------------------------- sub
+\subsection{Notation}
+\label{sec:notation}
 We denote by $\langle d_{t, \mathit{id}}, Q_{t, \mathit{id}}\rangle$ public-private key pairs, where $d$ is the private key, $Q$ is the public key, $t \in \{e, s\}$ indicates whether the key is ephemeral or static, and $\mathit{id}$ indicates the party to whom the key pair belongs.
@@
 For more details about the challenge-response terminology, see~\cite{DBLP:conf/crypto/Krawczyk05}.
-
-%%------------------------------------------------------------------------- sub
 %
 
-\mEdhoc{} relies on \mCose{}~\cite{rfc8152} for elliptic curve operations and
-transforming points into bitstrings, and we therefore abstract those as
-follows.
+\mEdhoc{} fundamentally utilizes elliptic curves and associated Diffie-Hellman operations. 
 %
-Signatures and verification thereof using party $A$'s key pair are
-denoted \mSign{A}$(\cdot)$ and \mVf{A}$(\cdot)$ respectively.
+Signatures using a principal $A$'s keys are denoted by \mSign{A}$(\cdot)$, while the verification thereof is denoted by \mVf{A}$(\cdot)$.
 %
-The DH-primitive combining a private key $d$ and a point $P$ is denoted
-$\mDH(d,P)$.
+A Diffie-Hellman operation which combines a private key $d$ and a point $P$ on the elliptic curve is represented as $\mDH(d, P)$.
 %
-We abuse notation and let these function symbols denote operations on both
-points and the corresponding bitstrings.
+We will often overload notation to let $P$ stand for both the point on the elliptic curve as well as the corresponding bitstring.
+
+%\mEdhoc{} relies on \mCose{}~\cite{rfc8152} for elliptic curve operations and
+%transforming points into bitstrings, and we therefore abstract those as
+%follows.
+%
+%Signatures and verification thereof using party $A$'s key pair are
+%denoted \mSign{A}$(\cdot)$ and \mVf{A}$(\cdot)$ respectively.
+%%
+%The DH-primitive combining a private key $d$ and a point $P$ is denoted
+%$\mDH(d,P)$.
+%%
+%We abuse notation and let these function symbols denote operations on both
+%points and the corresponding bitstrings.
 %
@@
 %------------------------------------------------------------------------- sub
-\subsection{Framework Structure}
-\label{sec:framework}
-\mEdhoc{}'s goal is to establish an \mOscore{} security context,
-including session key material \mbox{denoted \mSessKey{}}, and optionally transfer
-application data \mADone{}, \mADtwo{} and \mADthree{}.
+\subsection{Overall description}
+\label{sec:description}
+\mEdhoc{} is designed to establish a security context for \mOscore{}.
 %
-To accomplish this, the
-\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} gives
-a three-message protocol pattern, shown in Figure~\ref{fig:edhocFramework}.
+This context includes the session key material (we denote this by \mSessKey{}), and might also transfer application data \mADone{}, \mADtwo{}, and \mADthree{}.
 %
-We first describe this pattern and the parts that are common to all key
-establishment methods.
+The abstract protocol as given in the \mSpec{}~\cite{} consists of three messages, and is shown in Figure~\ref{fig:edhocFramework}~\cite{Norr21}.
 %
-Then we describe authentication and derivation of keys in more detail.
-%
-The latter is what differentiate the key establishment methods from each other.
+The generalized abstract protocol is the same across methods, but the authentication mechanisms and key derivation procedures differ.
 %
 
-\subsubsection{Protocol Pattern}
-The first two messages negotiate authentication methods \mMethod{} and a
-ciphersuite \mSuites{}.
+Of the three messages, the first two, among other things, serve to establish a common authentication method \mMethod{}, and a ciphersuite \mSuites{}.
 %
-In \mMethod{}, the initiator $I$ proposes which authentication method each
-party should use.
+In \mMethod{}, the party playing the Initiator role proposes an ordered list of choices for the authentication methods that each party can use. 
 %
-These may differ, leading to four possible combinations:
-\mSigSig{}, \mSigStat{}, \mStatSig{} and \mStatStat{}.
+As mentioned earlier, these may differ for the two roles, yielding four possible combinations: \mSigSig{}, \mSigStat{}, \mStatSig{}, and \mStatStat{}, where the first is the method used by the Initiator role, and the second by the Responder role.~\footnote{As in the \mSpec{}, we will call the combinations of authentication methods as methods as well.}
 %
-We refer to these \emph{combinations} of authentication methods simply as
-\emph{methods} to align with the \mSpec{} terminology.
+The party executing the Responder role may choose to reject the method or ciphersuite chosen by the Initiator by sending an error message.
 %
-The first authentication method in a combination is the one proposed for the
-initiator and the last is the one proposed for the responder.
+This results in abandoning this session and renegotiating, as the Initiator goes down their list of choices for authentication methods, and picks the next option.
 %
-The responder $R$ may reject the choice of method
-or cipher suite with an error message, resulting in negotiation across
-multiple \mEdhoc{} sessions.
+While our analysis (and indeed, \mTamarin{}) cannot naturally handle such renegotiation which maintains state and ``remembers'' the rejected method and/or ciphersuite, the formal modeling excludes this aspect of the protocol.
 %
-Our analysis excludes error messages.
+However, we will discuss the ramifications of such a renegotiation procedure and the error messages later, in Section~\ref{sec:}.
+%
+
+In addition to negotiating the method and the ciphersuite, the first two messages are also instrumental for the exchange of public ephemeral keys \mGx{} and \mGy{}, and connection identifiers \mCi{} and \mCr{}, for the initiator and responder roles respectively.
+%
+The \mSpec{} states that the connection identifiers serve only to route messages to the correct principal executing \mEdhoc{}, but then also claims that they may be used in turn by any protocols (like \mOscore{}) using the security context established by \mEdhoc{}.
+%
+Thus, no explicit security guarantees are required to be satisfied by these communication identifiers, as per the \mSpec{}.
+
+%%%%%%%% START HERE
+
+%\mEdhoc{}'s goal is to establish an \mOscore{} security context,
+%including session key material \mbox{denoted \mSessKey{}}, and optionally transfer
+%application data \mADone{}, \mADtwo{} and \mADthree{}.
+%%
+%To accomplish this, the
+%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} gives
+%a three-message protocol pattern, shown in Figure~\ref{fig:edhocFramework}.
+%
+%We first describe this pattern and the parts that are common to all key
+%establishment methods.
+%%
+%Then we describe authentication and derivation of keys in more detail.
+%%
+%The latter is what differentiate the key establishment methods from each other.
+%%
+%
+%\subsubsection{Protocol Pattern}
+%The first two messages negotiate authentication methods \mMethod{} and a
+%ciphersuite \mSuites{}.
+%%
+%In \mMethod{}, the initiator $I$ proposes which authentication method each
+%party should use.
+%%
+%These may differ, leading to four possible combinations:
+%\mSigSig{}, \mSigStat{}, \mStatSig{} and \mStatStat{}.
+%%
+%We refer to these \emph{combinations} of authentication methods simply as
+%\emph{methods} to align with the \mSpec{} terminology.
+%%
+%The first authentication method in a combination is the one proposed for the
+%initiator and the last is the one proposed for the responder.
+%%
+%The responder $R$ may reject the choice of method
+%or cipher suite with an error message, resulting in negotiation across
+%multiple \mEdhoc{} sessions.
+%%
+%Our analysis excludes error messages.
 %

### 2022-07-13 paper.tex
[add] more elaboration on DoS

@@
 \begin{abstract}
-Constrained IoT devices are becoming ubiquitous in society
-and there is a need for secure communication protocols that respect the
-constraints under which these devices operate.
+Given how common IoT devices that use constrained resources are becoming today, the need of the hour is communication protocols which can operate securely under such limitations.
 %
-\mEdhoc{} is an authenticated key establishment protocol for constrained IoT
-devices, currently being standardized by the Internet Engineering Task
-Force (IETF).
+For a few years, the Internet Engineering Task Force (IETF) has been working to standardize \mEdhoc{}, an authenticated key establishment protocol for such constrained IoT devices.
 %
-A rudimentary version of \mEdhoc{} with only two key establishment methods was
-formally analyzed in 2018.
+The first version of \mEdhoc{} was proposed in 2016.
 %
-Since then, the protocol has evolved significantly and several new key
-establishment methods have been added.
+In 2018, Bruni et al~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.
 %
-In this paper, we present a formal analysis of all \mEdhoc{} methods in an
-enhanced symbolic Dolev-Yao model using the \mTamarin{} tool.
+%By now, the protocol has been fleshed out much more, and has been augmented with multiple new key establishment methods.
+By 2021, the protocol had been fleshed out much more, with multiple new key establishment methods, and this version was formally analyzed using the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
 %
-We show that not all methods satisfy the authentication notion
-injective agreement, but that all do satisfy a notion of implicit
-authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
-%
-We identify other weaknesses to which we propose improvements
-%
-For example, a party may intend to establish a session key with a certain peer,
-but ending up establishing it with another, trusted but compromised, peer.
-%
-We communicated our findings and proposals to the IETF, which
-has incorporated some of these in newer versions of the standard.
+Here, we build on this work, and use \mTamarin{} to formally analyze the key establishment methods in the current version of \mEdhoc{}, as well as discuss some ramifications of the choices made while designing the protocol.
+\vnote{We need to see what more we can say here, about the new stuff we're potentially adding}
 \end{abstract}
@@
 \label{sec:introduction}
-
-As IoT devices become more prevalent and get involved in progressively sensitive
-functions in society, the need to secure their communications
-becomes increasingly important.
-%
-Most security analysis has focused on computationally
-strong devices, such as cars and web-cameras, where existing protocols like
-\mDandTls{} suffice.
-%
-Constrained devices, on the other hand, which operate under severe
-bandwidth and energy consumption restrictions, have received much less
-attention.
-%
-These devices may be simple sensors, which only relay environment
-measurements to a server every hour, but need to function autonomously without
-maintenance for long periods of time.
-%
-The IETF standardized the Object Security for
-Constrained RESTful Environments (\mOscore{}) protocol to secure communications
-between constrained devices~\cite{rfc8613}.
-%
-However, the \mOscore{} protocol requires a pre-established security context.
-%
-The IETF has been discussing requirements and mechanisms for a key
-exchange protocol, named Ephemeral Diffie-Hellman Over COSE (\mEdhoc), for
-establishing \mOscore{} security contexts.
-%
-Naturally, \mEdhoc{} must work under the same constrained requirements as
-\mOscore{} itself.
-%
-While not all use cases for \mEdhoc{} are firmly set, the overall goal is to
-establish an \mOscore{} security context, under message size limitations.
-%
-It is therefore important to ensure that \mEdhoc{} satisfies fundamental
-security properties expected from a key exchange protocol.
-%The work is now being done in the IETF Lightweight Authenticated Key Exchange
-%(LAKE) working group.
-%
-
-The first incarnation of \mEdhoc{} appeared in March 2016.
-%
-It contained two different key establishment methods, one based on a
-pre-shared Diffie-Hellman (DH) cryptographic core%
-\footnote{By a \emph{cryptographic core}, or simply core, we mean an academic protocol,
-without encodings or application specific details required by an industrial
-protocol.
-%
-A \emph{key establishment method} is a core with some such details added.
-}
-and a second based on a
-variant of challenge-response signatures in the style of
-\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
-%
-\mEdhoc{} is therefore a framework of several key establishment methods.
-%
-In May 2018, the core based on challenge-response signatures was replaced by
-one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}.
-%
-Since then the protocol has undergone significant changes.
-%
-Three new cores, mixing challenge-response signatures and regular signature for
-authentication, were added~\cite{our-analysis-selander-lake-edhoc-00}.
-%
-
-We formulate and formalize a security model covering all four key
-establishment methods, which is important especially since the
-\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} lacks a clear description
-of the intended security model and overall security goals.
-%
-
-We perform the analysis in a symbolic Dolev-Yao model.
-In this framework, we model messages as terms in an algebra, with operations such
-as encryption, modelled as functions on these terms.
-These functions are assumed perfect, e.g., one cannot decrypt an encrypted
-message without access to the key.
-The adversary, while unable to break encryption or reverse hashing, is modelled
-as the network.
-That is, the adversary, can block reroute, replay and modify messages at will.
-A symbolic model like this, while slightly severe an abstraction, still allows
-us to analyze \mEdhoc{} for logical flaws without incurring the complexity of
-a computational model.
-%
-The standardization process is ongoing, with the authors releasing newer
-versions of the \mSpec{} (see Section~\ref{sec:newdrafts} for more detail
-about how these versions differ from the one analyzed here).
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{Contributions}
-\label{sec:contributions}
-In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
-establishment methods) using the \mTamarin{} tool~\cite{DBLP:conf/cav/MeierSCB13}.
-%
-We present a formal model we constructed of the protocol as given in the
-\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00}.
-%
-
-We give an explicit adversary model for the protocol and verify
-properties such as session key material and entity authentication, and perfect
-forward secrecy, for all four methods.
-%
-
-The model itself is valuable as a basis for verifying further updates in the
-ongoing standardization.
-%
-It is publicly available~\cite{edhocTamarinRepo}.
-%
-It took several person-months to interpret the
-specification and construct the model.
-%
-Termination requires a hand-crafted proof oracle to guide \mTamarin{}.
-%
-
-We show that not all \mEdhoc{}'s key establishment
-methods provide authentication according to the injective agreement definition
-on the session key material, and none on the initiator's identity.
-%
-However, we show that all methods fulfill an implicit agreement property
-covering the session key material and the initiator's identity.
-%
-We identify a number of subtleties, ambiguities and weaknesses in the
-specification.
-%
-For example, the authentication policy requirements allows situations where a
-party establishes session key material with a trusted but compromised peer, even
-though the intention was to establish it with a different trusted party.
-%
-We provide remedies for the identified issues and have
-communicated these to the IETF and the specification authors, who incorporated
-some of our suggestions and currently consider how to deal with the remaining
-ones.
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{Comparison with Related Work}
-The May 2018 version of \mEdhoc{} was formally analyzed by
-\cite{DBLP:conf/secsr/BruniJPS18} using the \mProverif{}
-tool~\cite{DBLP:conf/csfw/Blanchet01}.
-%
-Their analysis covered a pre-shared key authenticated core and one
-based on \mSigma.
-%
-The properties checked for therein were secrecy, PFS and integrity of
-application data, identity protection against an active adversary,
-and strong authentication.
-%
-
-In contrast to the key establishment methods analyzed by Bruni et~al., which
-were based on the well-understood pre-shared key DH and \mSigma{} protocols,
-the three newly added
-methods combine two unilateral authentication protocols with the goal to
-constructing mutual authentication protocols.
-%
-Combining two protocols, which individually provide unilateral authentication,
-is not guaranteed to result in a secure mutual authentication
-protocol~\cite{DBLP:conf/ccs/Krawczyk16}.
-%
-Consequently, even though the framework is similar to the one analyzed by Bruni
-et~al., the cryptographic underpinnings have significantly increased in
-complexity, and is using mechanisms which have not previously been formally analyzed.
-%
-The set of properties we check for is also different.
-%
-Our analysis is further carried out using a different tool,
-namely \mTamarin; different kinds of strategies to formulate and
-successfully analyze the protocol are required when working with this tool.
-%
+\input{intro}
 
@@
 \label{sec:formalization}
-\input{edhocFormalization}
+\input{formalization}
 
@@
 \label{sec:dos}
-%\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
-\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
-attacks and delegates the responsibility to handle this to the application.
+
+% There is little about EDHOC specific DoS stuff; that is interesting in itself,
+% and motivates this section
+There is a wealth of information about DoS in general, and more specifically for
+IoT.
 %
-The core \mEdhoc{} specification has limited functionality, few negotiable
+However, there is little information regarding DoS aspects stemming
+specifically from \mEdhoc{}.
+%
+
+% Other protocols have DoS considerations, but EDHOC delegates it to developer
+In contrast to \mTls{} and IKEv2,
+the \mEdhoc{} core specification does not include any specific measures for
+countering Denial of Service (DoS) attacks.
+%
+Instead, that responsibility is delegated to the application, using
+proprietary measures or extensions to \mEdhoc{}.
+%
+
+% The core is small and probably not vulnerable to anything unexpected
+The \mEdhoc{} core specification has limited functionality, few negotiable
 parameters and processing rules.
@@
 %
+
+% However, It is difficult to define what EDHOC *is*; it s partially specified in many
+% specs
 However, \mEdhoc{} leaves much of the semantics of its messages undefined or
-specified as (optional) extensions.
+specified as extensions.
 %
-It is therefore difficult to draw the line between what is \mEdhoc{} and what is
-an extension.
+Examples of mechanisms with mostly undefined semantics include processing
+rules and content of error messages and of the additional data that is
+passed in each of the \mEdhoc{} messages.
 %
-For example, the core specification does not include a
+Another example is that the core specification does not include a
 message transport mechanism.
 %
-This increases the risk of applications vulnerable to DoS via \mEdhoc{}, even
-though \mEdhoc{} itself is arguably not particularly weak in this respect.
+The transport for \mEdhoc's primary target, i.e., to key \mOscore{},
+is \mCoap{}.
+%
+This is specified in~\cite{ietf-core-oscore-edhoc-03}, but still leaves
+many design choices to the application.
+%
+Because of this, it is difficult to draw the line between what is considered as
+\mEdhoc{} and what is an extension.
+%
+
+% Pros with undefined semantics: flexibilty; Cons: developers can shoot
+% themselves in the foot
+Undefined semantics allow the implementors much freedom
+and flexibility to design applications.
+%
+On the down-side, the few recommendations given increase the risk of
+applications being vulnerable to DoS via \mEdhoc{}, even
+though the core \mEdhoc{} specification itself is arguably not particularly
+weak in this respect.
 %
@@
 No recommendation is given on how to minimize DoS effects stemming from use of
-error messages, which can be used in an arbitrary way in \mEdhoc{}.
+error messages, which can be used in an arbitrary way in \mEdhoc{}
+(see Section~\ref{sec:errorHandling} below).
 %
@@
 authenticated the key card after receiving the third \mEdhoc{} message.
+%
+The draft IETF specification profiling \mEdhoc{} for use with
+\mCoap{} and \mOscore{} ~\cite{ietf-core-oscore-edhoc-03},
+however, defines the
+\mCoap{} client to be the \mEdhoc{} initiator, and hence reduces the flexibility
+to use that option.
 %
@@
 
-\textbf{Q:} Is htere any situation where it makes sense that the initiator sends
+\textbf{Q:} Is there any situation where it makes sense that the initiator sends
 a unsupported ciphersuite message instead of message 3?
@@
 
-Section 6.2 gives an excample of sending an error message "method not
+Section 6.2 gives an example of sending an error message "method not
 supported".

### 2022-07-14 formalization.tex
Changed through Section 3.2

% !TEX root = paper.tex
\vnote{Need to rewrite this based on the new spec}
The \mEdhoc{} \mSpec{} \cite{our-analysis-selander-lake-edhoc-00} claims
that \mEdhoc{} satisfies many security properties, but these are imprecisely
expressed and motivated.
%
In particular, there is no coherent adversary model.
%
It is therefore not clear in which context properties should be verified.
%
We resolve this by clearly specifying an adversary model, in which we can verify
properties.
%

\subsection{Formal Model}\label{sec:threat-model}
\vnote{Can we add extra references here? SAPIC+ etc?} As in~\cite{Norr21}, we verify \mEdhoc{} in the Dolev-Yao model~\cite{DY83}.
%
This is a well-established model for symbolic verification of security protocols.
%
Messages are modelled as terms in an algebra, and the various cryptographic operations are assumed to be perfect, i.e. encrypted messages can only be decrypted with the correct key, no hash collisions exist etc.
%

The adversary is assumed to be the network, and thus controls the communication channel and can see all messages being communicated as part of the protocol.
%
In addition, they can interact with an unbounded number of protocol sessions, dropping, injecting, and modifying messages at will, but cannot break cryptography.

On top of the Dolev-Yao model, we allow the adversary two extra capabilities with respect to getting access to the long-term and ephemeral keys. 
%
Long-term key reveal, denoted $\mRevLTK^{t}(A)$, stands for the adversary gaining access to a principal $A$'s long-term private key \mPriv{A} at a time point $t$.
%
Ephemeral key reveal, denoted $\mRevEph^{t}(A, k)$, stands for the adversary obtaining, at time $t$, the ephemeral private key \mPrivE{A} used by party $A$  in a session where they establish key material $k$.
%
Ephemeral keys might be stored in a less secure part of a device, but long-term keys are often stored and operate within a secure model, like a trusted execution environment (TEE). \vnote{Do we want to say more about TEEs here?}
%
Formalizing these two events allows us to have more fine-grained control over the access that an adversary has to these fundamentally different kinds of keys.
%

We now briefly describe \mTamarin{}, and how protocols are modelled in it.

%-------------------------------------------------------------------------- sub
\subsection{\mTamarin{}}
\label{sec:tamarin}
As in~\cite{Norr21}, we model and verify \mEdhoc{} in the symbolic model using \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an interactive tool for the formal verification of security protocols.
%
Protocols are modelled in \mTamarin{} as multi-set rewrite rules which encode a transition relation.
%
The elements of these multi-sets, called facts, contribute to the global system state.

Rewrite rules can be annotated with events, called actions in \mTamarin{}.
%
Communicated messages in the protocol are modelled as terms in an algebra, which specifies sets of names, variables, and allowable function symbols.
%
Facts and actions are modelled as $n$-ary predicates in the term algebra, and actions can be parametrized using terms.

An annotated multi-set rewrite rule is represented as $l \ifarrow[e] r$, where $l$ and $r$ are multi-sets, and $e$ is a multi-set of actions.
%
A sequence of actions yields a protocol execution trace.
%
Properties are defined as formulas in a first-order logic, and these formulas can be verified over execution traces.

Protocol verification in \mTamarin{} happens under an equational theory $E$.
%
For example, to represent the fact that $E$ satisfies the reversal of symmetric encryption by using a decryption operation with the key, one can write $\textit{dec}(\textit{enc}(x, y), y) =_{E} x$. 
%
The equational theory $E$ is fixed upfront to handle the functions supported by the term algebra, so we will omit the subscript for the rest of this paper.
%

Users can extend the default term algebra and equational theory in \mTamarin{} with new function symbols and unification rules for these new symbols.
%
For example, \mEdhoc{} requires authenticated encryption, which we model using the symbol \mT{aeadEncrypt}. 
%
We augment \mTamarin{} with the following rule for this operation, which represents the fact that if the adversary knows a key \mT{k}, a message \mT{m}, and authenticated data \mT{ad}, and has access to an encryption algorithm \mT{ai}, then they can obtain the message corresponding to the authenticated encryption of \mT{m} with \mT{k}~\cite{Norr21}.
\begin{verbatim}
[!KU(k), !KU(m), !KU(ad), !KU(ai)] --[]-> [!KU(aeadEncrypt(k, m, ad, ai))]
\end{verbatim}
%
Similarly, we can extend \mTamarin{} with rules for other operations, like \mXor{}.

For syntactic sugar, \mTamarin{} allows the use of let-bindings and tuples. 
%
The rules modelling a Dolev-Yao adversary and the evolution of their knowledge as the protocol executes are built into \mTamarin.
%
However, we can extend and/or modify this behaviour by adding other rules as well.

%-------------------------------------------------------------------------- sub

%%%%%%%%%%% START HERE

\subsection{Formalization of Properties}
\label{sec:desired-properties}
We use the \mTamarin{} verification
tool~\cite{DBLP:conf/cav/MeierSCB13} to encode the model and verify properties.
%
This tool uses a fragment of temporal first order logic to reason about
events and knowledge of the parties and of the adversary.
%
For conciseness we use a slightly different syntax than
that used by \mTamarin{}, but which has a direct mapping to \mTamarin{}'s logic.
%

Event types are predicates over global states of system execution.
%
Let $E$ be an event type and let $t$ be a timestamp associated with a point in a
trace.
%
Then $E^{t}(p_i)_{i\in\mathbb{N}}$ denotes an event of type $E$ associated with
a sequence of parameters $(p_i)_{i\in\mathbb{N}}$ at time $t$ in that trace.
%
In general, more than one event may have the same timestamp and hence
timestamps form a quasi order, which we denote by $t_1 \lessdot t_2$ when $t_1$
is before $t_2$ in a trace.
%
We define $\doteq$ analogously.
%
However, two events of the same type cannot have the same timestamp, so
$t_1 \doteq t_2$ implies $E^{t_1} = E^{t_2}$.
%
Two events having the same timestamp does not imply the there is a fork in the
trace, only that the two events happen simultaneously and that comparing their
timestamps results in that they are equal.
%
This notation corresponds to \mTamarin{}'s use of action facts
$E(p_i)_{i\in\mathbb{N}}@t$.
%

The event $\mK^t(p)$ denotes that the adversary knows a parameter $p$ at
time $t$.
%
Parameters are terms in a term algebra of protocol specific operations and
generic operations, e.g., tuples $\langle\cdot\rangle$.
%
Intuitively, $\mK^t(p)$ evaluates to true when $p$ is in
the closure of the
parameters the adversary observed from interacting with parties using the
protocol, under the Dolev-Yao message deduction operations and
the advanced adversary capabilities up until time $t$.
%
For a more precise definition of knowledge management, we refer to~\cite{DBLP:conf/cav/MeierSCB13}.
%
An example of a formula is
\[
    \forall t, k, k'\mLogicDot \mK^{t}(\langle k, k'\rangle)\ \rightarrow\ 
\mK^{t}(k) \land \mK^{t}(k'),
\]
expressing that if there is a time $t$ when the adversary knows the tuple
$\langle k, k'\rangle$, then the adversary knows both
$k$ and $k'$ at the same point in time.
%

An initiator $I$ considers the
protocol run started when it sends a message \mMsgone{} (event type \mIStart)
and the run completed after sending a message \mMsgthree{} (event type
\mIComplete).
%
Similarly, a responder $R$ considers the run started upon receiving
a \mMsgone{} (event type \mRStart), and completed upon receiving a \mMsgthree{}
(event type \mRComplete).
%

%----------------------------------------------------------------------- PFS
\subsubsection{Perfect Forward Secrecy (PFS)}
\label{sec:secrecy}
Informally, PFS captures the idea that session key material remains secret
even if a long-term key leaks in the future.
%
We define PFS for session key material \mSessKey{} as \mPredPfs{} in
Figure~\ref{fig:props}.
%

\begin{figure*}
\begin{align*}
    \mPredPfs \triangleq\ & \forall I, R, \mSessKey, t_2, t_3\mLogicDot
    \mK^{t_3}(\mSessKey)\  \land\ 
    (\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(I, R, \mSessKey))
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))
\end{align*}
\begin{align*}
    \mPredInjI \triangleq\ &
    \forall I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, \mSessKey, S)
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRStart^{t_1}(R, \mSessKey, S) \land t_1 \lessdot t_2)
    \land (\forall I' R' t_1' \mLogicDot \mIComplete^{t_1'}(I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1)
    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
\end{align*}
\begin{align*}
    \mPredInjR \triangleq\ &
    \forall I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(I, R, \mSessKey, S)
    \rightarrow\\
    &(\exists t_1\mLogicDot \mIStart^{t_1}(I, R, \mSessKey, S) \land t_1 \lessdot t_2)
    \land (\forall I' R' t_1' \mLogicDot \mRComplete^{t_1'}(I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1)
    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2).
\end{align*}
\begin{align*}
    \mPredImpI \triangleq\ &
    \forall I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(I, R, \mSessKey, S)
    \rightarrow\\
      &(\forall I', R', S', t_2\mLogicDot \mRComplete^{t_2}(I', R', \mSessKey, S') \rightarrow
             (I=I' \land R=R' \land S=S'))\\
      &\land (\forall I', R', S', t_1'\mLogicDot
        (\mIComplete^{t_1'}(I', R', \mSessKey, S') \rightarrow t_1' \doteq t_1
        )\\
      &\ \ \ \lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey)).
\end{align*}
%
\caption{Formalization of security properties and adversary model.}
\label{fig:props}
\end{figure*}

The first parameter $I$ of the \mIComplete{} event represents the
initiator's identity,
and the second, $R$, represents that $I$ believes $R$ to be playing
the responder role.
%
The third parameter, \mSessKey{}, is the established session key material.
%
The parameters of the \mRComplete{} event are defined analogously.
%
Specifically, the first parameter of \mRComplete{} represents the identity of
whom $R$ believes is playing the initiator role.
%
The essence of the definition is that an adversary only knows \mSessKey{} if they
compromised one of the
parties long-term keys before that party completed the run, or if the adversary
compromised any of the ephemeral keys at any time after a party starts
its protocol run.
%
One way the definition slightly differs from the corresponding \mTamarin{} lemma
is that \mTamarin{} does not allow a disjunction on the left-hand side of an
implication in a universally quantified formula.
%
In the lemma, therefore, instead of the disjunction
$\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$,
we use a single action parametrized by $I$, $R$, and \mSessKey{} to signify that
\emph{either} party has completed their role.
%

%-------------------------------------------------------------------- InjAgree
\subsubsection{Authentication}
\label{sec:authenticationDef}
We prove two different flavors of authentication, the first being classical
\emph{injective agreement} following Lowe~\cite{DBLP:conf/csfw/Lowe97a}, and
the second being an implicit agreement property.
%
Informally, injective agreement guarantees to an initiator $I$ that whenever
$I$ completes a run ostensibly with a responder $R$,
then $R$ has been engaged in the protocol as a responder,
and this run of $I$ corresponds to a unique run of $R$.
%
In addition, the property guarantees to $I$ that the two parties agree on a set
$S$ of parameters associated with the run, including, in particular, the
session key material \mSessKey{}.
%
However, we will treat \mSessKey{} separately for clarity.
%
On completion, $I$ knows that $R$ has access to the session key material.
%
The corresponding property for $R$ is analogous.
%

Traditionally, the event types used to describe injective agreement are called
\emph{Running} and \emph{Commit}, but to harmonize the presentations of
authentication and PFS in this section, we refer to these event types as
\mIStart{} and \mIComplete{} respectively for the initiator, and
\mRStart{} and \mRComplete{} for the responder.
%
For the initiator role we define injective agreement by
\mPredInjI{} in Figure~\ref{fig:props}.

%
The property captures that for an initiator $I$, either the injective agreement
property as described above holds, or the long-term key of the believed
responder $R$ has been compromised before $I$ completed its role.
%
Had the adversary compromised $R$'s long-term key, they could have generated a
message of their liking (different to what $R$ agreed on), and signed this or
computed a $\mathit{MAC}_R$ based on \mPubE{I}, \mPriv{R} and their own chosen
ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$.
%
This places no restrictions on the ephemeral
key reveals, or on the reveal of the initiator's long-term key.
%
For the responder we define the property \mPredInjR{} as in
Figure~\ref{fig:props}.
%

%------------------------------------------------------------- Implicit auth
Unlike PFS, not all \mEdhoc{} methods enjoy the injective agreement property.
%
Hence, we show for all methods a form of \emph{implicit agreement} on all the
parameters mentioned above.
%
We take inspiration from the computational model definitions of implicit
authentication, proposed by Guilhem~et~al.~\cite{DBLP:conf/csfw/GuilhemFW20}, to
modify classical injective agreement into an implicit property.
%
A small but important difference between our definition and theirs, is that
they focus on
authenticating a key and related identities, whereas we extend the more general
concept of agreeing on a set of parameters, starting from the idea of injective
agreement~\cite{DBLP:conf/csfw/Lowe97a}.
%
We use the term \emph{implicit} in this context to denote that a party $A$
assumes that any other party $B$ who knows the session key material \mSessKey{} must
be the intended party, and that $B$ (if honest) will also agree on a set
$S$ of parameters computed by the protocol, one of which is \mSessKey{}.
%
When implicit agreement holds for both roles, upon completion, $A$ is guaranteed
that $A$ has been or is engaged in exactly one protocol run with $B$ in the
opposite role, and that $B$ has been or will be able to agree on $S$.
%
The main difference to injective agreement is that $A$ concludes that if
$A$ sends the last message and this reaches $B$, then $A$ and $B$ have agreed
on $I$, $R$ and $S$.
%
While almost full explicit key authentication, as defined by
Guilhem~et~al.~\cite{DBLP:conf/csfw/GuilhemFW20}, is a similar property, our
definition does not require key confirmation, so our definition is closer to
their definition of implicit authentication.
%
In the \mTamarin{} model we split the property into one lemma for
$I$ (\mPredImpI{}) and one for $R$ (\mPredImpR{}) to save memory during
verification.
%
We show only the definition for $I$ in Figure~\ref{fig:props}, because it is
symmetric to the one for $R$.
%

For implicit agreement to hold for the initiator $I$, the ephemeral keys
can never be revealed.
%
Intuitively, the reason for this is that the implicit agreement relies on that
whomever knows the session key material is the intended responder.
%
An adversary with access to the ephemeral keys and the public keys of
both parties can compute the session key material produced by all methods.
%
However, the responder $R$'s long-term key can be revealed after $I$ completes
its run, because the adversary is still unable to compute $P_e$.
%
The initiator's long-term key can also be revealed at any time without affecting
$I$'s guarantee for the same reason.
%

%------------------------------------------------------- Agreed parameters
\subsubsection{Agreed Parameters}
\label{sec:agreedParams}
The initiator $I$ gets injective and implicit agreement guarantees on the
following partial set $S_P$ of parameters:
\begin{itemize}
    \item the roles played by itself and its peer,
    \item responder identity,
    \item session key material (which varies depending on \mEdhoc{} method),
    \item context identifiers \mCi{} and \mCr{}, and
    \item cipher suites \mSuites{}.
\end{itemize}
%
Because \mEdhoc{} aims to provide identity protection for $I$, there is no
injective agreement guarantee for $I$ that $R$ agrees on the initiator's
identity.
%
For the same reason, there is no such guarantee for $I$ with respect to
the $P_I$ part of the session key material when $I$ uses the \mStat{}
authentication method.
%
There is, however, an implicit agreement guarantee for $I$ that $R$ agrees on
$I$'s identity and the full session key material.
%
Since $R$ completes after $I$, $R$ can get injective agreement guarantees on
more parameters, namely also the initiator's identity and the full session key
material for all methods.
%
The full set of agreed parameters $S_F$ is $S_P \cup \{I, P_I\}$
when $P_I$ is
part of the session key material, and $S_P \cup \{I\}$ otherwise.
%

%------------------------------------------------------- Implied properties
\subsubsection{Inferred Properties}
From the above, other properties can be inferred to hold in our adversary model.
%
Protocols where a party does not get confirmation that their peer knows the
session key material may be susceptible to
\emph{Key-Compromise Impersonation (KCI)}
attacks~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
Attacks in this class allow an adversary in possession of a party $A$'s secret
long-term key to coerce $A$ to complete a
protocol run believing it authenticated a certain peer $B$, but where $B$ did
not engage with $A$ at all in a run.
%
Because both our notions of agreement above ensure agreement on identities,
roles and session key material, all methods passing verification of those are
also resistant to KCI attacks.
%

If a party $A$ can be coerced into believing it completed a run with $B$, but
where the session key material is actually shared with $C$ instead, the 
protocol is vulnerable to an \emph{Unknown Key-Share (UKS)}
attack~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
For the same reason as for KCI, any method for which our agreement
properties hold is also resistant to UKS attacks.
%

From the injective agreement properties it follows that each party is assured
the identity of its peer upon completion.
%
Therefore, the agreement properties also capture \emph{entity authentication}.
%

%---------------------------------------------------------------------------sub
\subsection{\mTamarin{} Encoding of \mEdhoc{}}
\label{sec:modeling}
%
We model all four methods of \mEdhoc{}, namely
\mSigSig, \mSigStat, \mStatSig{} and \mStatStat.
%
Because the methods share a lot of common structure, we derive
their \mTamarin-models from a single specification written with the aid of the
M4 macro language.
%
To keep the presentation brief, we only present the \mStatSig{} metohod, as it
illustrates the use of two different asymmetric authentication methods
simultaneously.
%
The full \mTamarin{} code for all models can be found at~\cite{edhocTamarinRepo}.
%
Variable names used in the code excerpts here are sometimes shortened compared
to the model itself to fit the paper format.
%

\subsubsection{Primitive Operations}
Our model uses the built-in theories of exclusive-or and DH operations, as
in~\cite{DBLP:conf/csfw/DreierHRS18,DBLP:conf/csfw/SchmidtMCB12}.
%
Hashing is modeled via the built-in hashing function symbol augmented
with a public constant as additional input, modelling different
hash functions.
%
The HKDF interface is represented by \mT{expa} for the
expansion operation and \mT{extr} for the extraction operation.
%
Signatures use \mTamarin's built-in theory for \mT{sign} and \mT{verify}
operations.
%
For \mAead{} operations on key \mT{k}, message \mbox{\mT{m}}, additional data \mT{ad}
and algorithm identifier \mT{ai}, we use \mT{aeadEncrypt(m, k, ad, ai)}
for encryption.
%
Decryption with verification of the integrity is defined via the equation
\begin{small}\begin{verbatim}
aeadDecrypt(aeadEncrypt(m, k, ad, ai),
    k, ad, ai) = m.
\end{verbatim}\end{small}
%
The integrity protection of AEAD covers \mT{ad}, and this equation hence requires
an adversary to know \mT{ad} even if they only wish to decrypt the data.
%
To enable the adversary to decrypt without needing to verify the integrity
we add the equation
\begin{small}\begin{verbatim}
decrypt(aeadEncrypt(m, k, ad, ai), k, ai) = m.
\end{verbatim}\end{small}
%
The latter equation is not used by honest parties.
%

\subsubsection{Protocol Environment and Adversary Model}
We model the binding between a party's identity and their long-term
key pairs using the following rules.
%
\begin{small}
\begin{verbatim}
rule registerLTK_SIG:
    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
        [!LTK_SIG($A, ~ltk),
         !PK_SIG($A, pk(~ltk)),
         Out(<$A, pk(~ltk)>)]
rule registerLTK_STAT:
    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
        [!LTK_STAT($A, ~ltk),
         !PK_STAT($A, 'g'^~ltk),
         Out(<$A, 'g'^~ltk>)]
\end{verbatim}
\end{small}
%
The two rules register long-term
key pairs for the \mSig{}- and \mStat{}-based methods respectively.
%
The fact \verb|Fr(~ltk)| creates a fresh term \mT{ltk}, representing a long-term
secret key, not known to the adversary.
%
The fact \verb|Out(<$A, pk(~ltk)>)| sends the identity of the party
owning the long-term key and the corresponding public key to the adversary.
%
The event \mT{UniqLTK} together with a corresponding restriction models the fact
that the each party is associated with exactly one long-term key.
%
Consequently, an adversary cannot register additional long-term keys for an
identity.
%
In line with the \mEdhoc{} \mSpec{}, this models an external mechanism
ensuring that long term keys are bound to correct identities, e.g.,
a certificate authority.
%

We rely on \mTamarin's{} built-in message deduction rules for a Dolev-Yao adversary.
%
To model an adversary compromising long-term keys, i.e., events of type
\mRevLTK{}, and revealing ephemeral keys, i.e., events of type
\mRevEph{}, we use standard reveal rules.
%
The timing of reveals as modelled by these events is important.
%
The long-term keys can be revealed on registration, before protocol execution.
%
The ephemeral key of a party can be revealed when the party completes,
i.e., at events of type \mIComplete{} and \mRComplete.~\footnote{A stronger, and perhaps more realistic, model would reveal ephemeral keys upon
creation at the start of the run, but we failed to get \mTamarin{} to
terminate on this.}
%

\subsubsection{Protocol Roles}
We model each method of the protocol with four rules: \mT{I1}, \mT{R2}, \mT{I3}
and \mT{R4} (with the method suffixed to the rule name).
%
Each of these represent one step of the protocol as run by the initiator $I$
or the responder $R$.
%
The rules correspond to the event types \mIStart, \mRStart, \mIComplete,  and
\mRComplete, respectively.
%
Facts prefixed with \mT{StI} carry state information between \mT{I1} and \mT{I3}.
%
A term unique to the current thread, \mT{tid}, links two rules to a given state fact.
%
Similarly, facts prefixed with \mT{StR} carry state information between the
responder role's rules.
%
Line 28 in the \mT{R2\_STAT\_SIG} rule shown below illustrate one such use of state
facts.
%

We do not model the error message that $R$ can send in response to message
\mMsgone, and hence our model does not
capture the possibility for $R$ to reject $I$'s offer.
%

We model the XOR encryption of \mT{CIPHERTEXT\_2} with the key \mT{K\_2e} using
\mTamarin{}'s built in theory for XOR, and allow each term of the encrypted
element to be attacked individually.
%
That is, we first expand \mT{K\_2e} to as many key-stream terms as there are
terms in the plaintext tuple by applying the \mHkdfExpand{} function to unique
inputs per term.
%
We then XOR each term in the plaintext with its own key-stream term.
%
This models the \mSpec{} closer than if we would have XORed \mT{K\_2e} as a
single term onto the plaintext tuple.
%
The XOR encryption can be seen in lines 19--22 in the listing of
\mT{R2\_STAT\_SIG} below.
%
\begin{small}
\begin{verbatim}
1 rule R2_STAT_SIG:
2 let
3    agreed = <CS0, CI, ~CR>
4    gx = 'g'^xx
5    data_2 = <'g'^~yy, CI, ~CR>
6    m1 = <'STAT', 'SIG', CS0, CI, gx>
7    TH_2 = h(<$H0, m1, data_2>)
8    prk_2e = extr('e', gx^~yy)
9    prk_3e2m = prk_2e
10   K_2m = expa(<$cAEAD0, TH_2, 'K_2m'>,
11               prk_3e2m)
12   protected2 = $V // ID_CRED_V
13   CRED_V = pkV
14   extAad2 = <TH_2, CRED_V>
15   assocData2 = <protected2, extAad2>
16   MAC_2 = aead('e', K_2m, assocData2,
                  $cAEAD0)
17   authV = sign(<assocData2, MAC_2>, ~ltk)
18   plainText2 = <$V, authV>
19   K_2e = expa(<$cAEAD0, TH_2,
                 'K_2e'>, prk_2e)
20   K_2e_1 = expa(<$cAEAD0, TH_2,
                   'K_2e', '1'>, prk_2e)
21   K_2e_2 = expa(<$cAEAD0, TH_2,
                   'K_2e', '2'>, prk_2e)
22   CIPHERTEXT_2 = <$V XOR K_2e_1,
                     authV XOR K_2e_2>
23   m2 = <data_2, CIPHERTEXT_2>
24   exp_sk = <gx^~yy>
25 in
26   [!LTK_SIG($V, ~ltk), !PK_SIG($V, pkV),
      In(m1), Fr(~CR), Fr(~yy), Fr(~tid)]
27   --[ExpRunningR(~tid, $V, exp_sk, agreed),
        R2(~tid, $V, m1, m2)]->
28   [StR2_STAT_SIG($V, ~ltk, ~yy, prk_3e2m,
              TH_2, CIPHERTEXT_2, gx^~yy,
              ~tid, m1, m2, agreed),
29    Out(m2)]
\end{verbatim}
\end{small}
%

To implement events and
to bind them to parameters, we use actions.
%
For example, the action \verb|ExpRunningR(~tid, $V, exp_sk, agreed)| in line 27
above implements binding of an event of type \mRStart{} to the parameters and session key
material.
%

As explained in Section~\ref{sec:agreedParams}, it is not possible to show
injective agreement on session key material when it includes
$P_I$ (not visible in the rule \mT{R2\_STAT\_SIG}).
%
Therefore, we use certain actions to implement events that include $P_I$ in the
session key material and other actions that do not.
%
Session key material which includes (resp. does not include) $P_I$ is referred
to as \mT{imp\_sk} (resp. \mT{exp\_sk}) in the
\mTamarin{} model.
%
In the case of \mSigSig{} and \mSigStat, therefore, \mT{imp\_sk} is the same as
\mT{exp\_sk}.
%

%-------------------------------------------------------------------------- sub
\subsection{\mTamarin{} Encoding of Properties}
\label{sec:propertyFormalization}
The properties and adversary model we defined in
Section~\ref{sec:desired-properties} translate directly into \mTamarin's logic,
using the straightforward mapping of events to the actions emitted from the model.
%
As an example, we show the lemma for verifying the property \mPredPfs.
%
\begin{small}
\begin{verbatim}
1 lemma secrecyPFS:
2 all-traces
3  "All u v sk #t3 #t2.
4   (K(sk)@t3 & CompletedRun(u, v, sk)@t2) ==>
5       ( (Ex #t1. LTKRev(u)@t1 & #t1 < #t2)
6       | (Ex #t1. LTKRev(v)@t1 & #t1 < #t2)
7       | (Ex #t1. EphKeyRev(sk)@t1))"
\end{verbatim}
\end{small}
%
As mentioned earlier, the action \mT{CompletedRun(u, v, sk)} in line 4 is
emitted by both the rules \mT{I3} and \mT{R4}, and corresponds
to the disjunction of events $\mIComplete^{t_2} \lor \mRComplete^{t_2}$ in the
definition of \mPredPfs{} in Section~\ref{sec:secrecy}.
%
Similarly, \mT{EphKeyRev(sk)} in line 7 models that the ephemeral
key is revealed for either $I$ or $R$, or both.
%

### 2022-07-14 paper.tex
Changed through Section 3.2

@@
 \label{sec:dos}
-
-% There is little about EDHOC specific DoS stuff; that is interesting in itself,
-% and motivates this section
-There is a wealth of information about DoS in general, and more specifically for
-IoT.
+%\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
+\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
+attacks and delegates the responsibility to handle this to the application.
 %
-However, there is little information regarding DoS aspects stemming
-specifically from \mEdhoc{}.
-%
-
-% Other protocols have DoS considerations, but EDHOC delegates it to developer
-In contrast to \mTls{} and IKEv2,
-the \mEdhoc{} core specification does not include any specific measures for
-countering Denial of Service (DoS) attacks.
-%
-Instead, that responsibility is delegated to the application, using
-proprietary measures or extensions to \mEdhoc{}.
-%
-
-% The core is small and probably not vulnerable to anything unexpected
-The \mEdhoc{} core specification has limited functionality, few negotiable
+The core \mEdhoc{} specification has limited functionality, few negotiable
 parameters and processing rules.
@@
 %
-
-% However, It is difficult to define what EDHOC *is*; it s partially specified in many
-% specs
 However, \mEdhoc{} leaves much of the semantics of its messages undefined or
-specified as extensions.
+specified as (optional) extensions.
 %
-Examples of mechanisms with mostly undefined semantics include processing
-rules and content of error messages and of the additional data that is
-passed in each of the \mEdhoc{} messages.
+It is therefore difficult to draw the line between what is \mEdhoc{} and what is
+an extension.
 %
-Another example is that the core specification does not include a
+For example, the core specification does not include a
 message transport mechanism.
 %
-The transport for \mEdhoc's primary target, i.e., to key \mOscore{},
-is \mCoap{}.
-%
-This is specified in~\cite{ietf-core-oscore-edhoc-03}, but still leaves
-many design choices to the application.
-%
-Because of this, it is difficult to draw the line between what is considered as
-\mEdhoc{} and what is an extension.
-%
-
-% Pros with undefined semantics: flexibilty; Cons: developers can shoot
-% themselves in the foot
-Undefined semantics allow the implementors much freedom
-and flexibility to design applications.
-%
-On the down-side, the few recommendations given increase the risk of
-applications being vulnerable to DoS via \mEdhoc{}, even
-though the core \mEdhoc{} specification itself is arguably not particularly
-weak in this respect.
+This increases the risk of applications vulnerable to DoS via \mEdhoc{}, even
+though \mEdhoc{} itself is arguably not particularly weak in this respect.
 %
@@
 No recommendation is given on how to minimize DoS effects stemming from use of
-error messages, which can be used in an arbitrary way in \mEdhoc{}
-(see Section~\ref{sec:errorHandling} below).
+error messages, which can be used in an arbitrary way in \mEdhoc{}.
 %
@@
 authenticated the key card after receiving the third \mEdhoc{} message.
-%
-The draft IETF specification profiling \mEdhoc{} for use with
-\mCoap{} and \mOscore{} ~\cite{ietf-core-oscore-edhoc-03},
-however, defines the
-\mCoap{} client to be the \mEdhoc{} initiator, and hence reduces the flexibility
-to use that option.
 %
@@
 
-\textbf{Q:} Is there any situation where it makes sense that the initiator sends
+\textbf{Q:} Is htere any situation where it makes sense that the initiator sends
 a unsupported ciphersuite message instead of message 3?
@@
 
-Section 6.2 gives an example of sending an error message "method not
+Section 6.2 gives an excample of sending an error message "method not
 supported".

### 2022-07-15 paper.tex
[add] some notes on error handling

--- 2022-07-13 paper.tex
@@
 
+Note: Section 5.1 says that if processing fails for some reason,
+then "typically" an error message is sent processing is terminated and protocol
+state is discarded. But: for algo negotiation (the only defined error
+message), then state is not discarded.
+
+Note: Section 5.2.2 says:
+"The Initiator MUST NOT change the supported cipher suites and
+the order of preference in SUITES_I based on previous error messages."
+But: what happens if order of preference changes for some other reason? Would it
+be good it include a flag in msg1 to indicate to the responder that the
+ciphersuite order depends on a previous run? This way, responder could react
+differently if a new cipher suite has been added without breaking the "MUST NOT"
+requirement. Need to think about this more.
+
+Note: Section 5.3.3 says about receiver processing of message 2:
+"If any processing step fails,
+the Responder MUST send an EDHOC error message back, formatted as defined in
+Section 6, and the session MUST be discontinued"
+And Section 5.4.3 about the responder processing of message 3:
+"If any processing step fails, the Responder MUST send an EDHOC error
+message back, formatted as defined in Section 6, and the session MUST
+be discontinued."
+The same is said in Section 5.5.3 about the initiator's processing of message 4.
+
 %-------------------------------------------------------------------------- sub

### 2022-07-15 paper.tex
Updates from Overleaf

--- 2022-07-14 paper.tex
@@
 \label{sec:formalization}
-\input{formalization}
+\input{edhocFormalization}
 
@@
 \label{sec:errorHandling}
-\mPoint{Error handling is undefined except for algo negotiation}
-Section 6 of the \mEdhoc{} specification\footnote{draft-ietf-lake-edhoc-14}
-states that error messages can be sent at any time and by any party.
+\mPoint{Error handling is undefined except for algo negotiation} Section 6 of the \mEdhoc{} specification\cite{draft-ietf-lake-edhoc-15} states that error messages can be sent at any time and by any party.
 %
-One error message type is defined (for algorithm negotiation), and the rest can
-be arbitrarily defined by the implementation.
+At the current version of the standard \cite{draft-ietf-lake-edhoc-15}, there are three types of error messages: type 0 is used to represent success, type 1 represents a generic error message used for debugging purposes, where the actual message is an English string, and type 2 represents failure to negotiate a common ciphersuite.
+Type 2 messages carry a list of supported ciphersuites by the responder.
 %
-The contents and semantics of these messages are also implementation defined.
+The contents and semantics of type 1 messages are dependent on the implementation.
 %
-Consequently, the error messages provide a generic message passing mechanism.
+Consequently, type 1 error messages provide a generic message passing mechanism, albeit one without a predefined semantics.
+
+%\textbf{Q:} Is there any situation where it makes sense that the initiator sends an unsupported ciphersuite message instead of message 3?
+%\textbf{Q:} In Section H: if no message correlation support by the transport layer, \mCi{} and \mCr{} should be used for correlation of messages to a session.
+
+\subsection{Lack of connection identifiers in error messages}
+Error messages do not include the optional connection identifiers \mCi{} or \mCr{} that are transmitted during the handshake.
+This is potentially problematic for connection-less transport layers, where \mCi{} and \mCr{} may be used to correlate messages.
+
+\subsection{Improper handling of error messages}
+\mEdhoc{} provides the error message mechanism, but gives little to no guidance on how it should be used safely.
 %
-
-\textbf{Q:} Is htere any situation where it makes sense that the initiator sends
-a unsupported ciphersuite message instead of message 3?
+For example, assume that an application using \mEdhoc{} logs error messages on a finite log with log rotation.
 %
-
-\textbf{Q:} In Section H: if no message correlation support by the transport
-layer, \mCi{} and \mCr{} should be used for correlation of messages to a
-session.
+If the log is used for anomaly detection or detection of sensitive events, then an adversary can simply inject error messages and fill the log until the sensitive event is overwritten.
 %
-Error messages do not carry \mCi{} or \mCr{}; will this create a problem?
-%
-
-\mEdhoc{} provides the error message mechanism, but gives little to no guidance
-on how it should be used safely.
-%
-Example of issue with improper use: Assume application logs error messages a
-finite log with log rotation.
-%
-If the log is used for anomaly detection or detection of sensitive events, then
-an adversary can simply inject error messages and fill the log until the
-sensitive event is overwritten.
-%
-If the filling error messages are less suspicious than the actual attack, this
-may be beneficial for the adversary.
+If filling error messages in the log is less suspicious than the actual attack, this may be beneficial for the adversary.
 %
 Implementors should do proper log-separation and log management.
+
+\subsection{Unspecified semantics for debugging messages}
+Section 6.2 gives an example of sending a type 1 error message with the information string "Method not supported".
 %
-
-Section 6.2 gives an excample of sending an error message "method not
-supported".
+It's important to note that this message cannot be used to negotiate the connection method securely.
 %
-This message cannot be used to negotiate method securely.
+In fact -- differently from ciphersuite negotiation -- \mEdhoc{} does not include a mechanism to support negotiating the connection method, among the four available combinations of signature keys and static Diffie-Hellman keys.
 %
-The reason is that error messages are not integrity protected, and an initator
-only proposes one method, and therefore the same tactic used for ciphersuite
-negotiation cannot be used for methods.
+The initiator selects the method on the first message, and the responder can reply with a type 1 error message along with the information string "Method not supported", but a standard compliant implementation cannot act upon this information string for negotiation purposes.
 %
-Because method negotiation can be viewed in the context of crypto agility, there
-is room for improvement.
+This is because the error message is not integrity protected and there is no secure mechanism to select a different authentication method that ensure that the preferred available authentication mechanism is always selected.
+%
+Because method negotiation can be viewed in the context of cryptographic agility, there is room for improvement.
 %

### 2022-07-15 protocol.tex
Modified root command in files

@@
-% !TEX root = journal-version.tex
+% !TEX root = paper.tex
 In this section, we describe the various key establishment methods of the \mEdhoc{} protocol.
@@
 %
+We will describe the various aspects of the protocol now.
+
+\begin{figure}[ht]
+\centering
+\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
+\begin{tikzpicture}
+    \node (ini) at (0, 0) {Initiator};
+    \draw [very thick] (0, -0.25) -- (0,-2.3);
+    \draw [very thick] (5.75, -0.25) -- (5.75,-2.3);
+    \node (res) at (5.75,0) {Responder};
+    \msg{1em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
+    \msg{3em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, \{\mIdcredr, \mAuthr, \mADtwo\}};
+    \msg{5em}{ini}{res}{\mMsgthree: \mCr, \{\mIdcredi, \mAuthi, \mADthree\}};
+    \draw [line width=1mm] (-0.75,-2.3) -- (0.75,-2.3);
+    \draw [line width=1mm] (5.75-0.75,-2.3) -- (5.75+0.75,-2.3);
+    \node (padding) at (0,-2.5) {};
+    \end{tikzpicture}
+    \caption{Structure of \mEdhoc{}: $\{t\}$ means $t$ is encrypted and integrity
+protected.~\cite{Norr21}}
+\label{fig:edhocFramework}
+\end{figure}
 
@@
 %
-As mentioned earlier, these may differ for the two roles, yielding four possible combinations: \mSigSig{}, \mSigStat{}, \mStatSig{}, and \mStatStat{}, where the first is the method used by the Initiator role, and the second by the Responder role.~\footnote{As in the \mSpec{}, we will call the combinations of authentication methods as methods as well.}
+As mentioned earlier, these may differ for the two roles, yielding four possible combinations: \mSigSig{}, \mSigStat{}, \mStatSig{}, and \mStatStat{}, where the first is the method used by the Initiator role, and the second by the Responder role.\footnote{As in the \mSpec{}, we will overload notation and refer to the combinations of authentication methods as methods as well.}
 %
@@
 %
-Thus, no explicit security guarantees are required to be satisfied by these communication identifiers, as per the \mSpec{}.
+While the \mSpec{} does not require any explicit security guarantees to be satisfied by these communication identifiers, it does, however, require that the identifiers be unique, i.e.
+in any session, $\mCi{} \neq \mCr{}$, and that the parties involved in the session can verify this uniqueness.
+%
+The \mSpec{} states that \mOscore{} should be able to use these identifiers to retrieve any particular security context.
+%
+In this work, as in~\cite{Norr21}, we verify that the parties agree on the values of \mCi{} and \mCr{}.
 
-%%%%%%%% START HERE
+Messages 2 and 3 also serve to identify and authenticate each party to the other.
+%
+These messages contain long-term key identifiers (\mIdcredi{} and \mIdcredr{}), as well as authenticating information (\mAuthi{} and \mAuthr{}) which lets each party know that the other party does indeed control the long-term key provided by them.
+%
+The authentication information changes based on the authentication method associated with the long-term key chosen by the party in question.
+%
+Consider the following example scenario.
+%
+The initiator $I$ chooses the method $\mSigStat{}$ and sends this via $\mMethod{}$ to the responder $R$.
+%
+$R$ now has an option to either reject this choice of method, or to provide an identifier \mIdcredr{} for a key pair which can be used with challenge-response signatures as well as authenticating information \mAuthr{}, as dictated by the static key authentication method \mStat{}.
+%
+$I$ will now respond with an identifier \mIdcredi{} corresponding to a key pair which is suitable for regular signatures, and provide authentication \mAuthi{} as appropriate for the signature-based method \mSig{}.
 
-%\mEdhoc{}'s goal is to establish an \mOscore{} security context,
-%including session key material \mbox{denoted \mSessKey{}}, and optionally transfer
-%application data \mADone{}, \mADtwo{} and \mADthree{}.
-%%
-%To accomplish this, the
-%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} gives
-%a three-message protocol pattern, shown in Figure~\ref{fig:edhocFramework}.
+\subsection{Key Schedule}
+\label{sec:keysched}
+As mentioned above, messages 2 and 3 of \mEdhoc{} contain authenticating information.
 %
-%We first describe this pattern and the parts that are common to all key
-%establishment methods.
-%%
-%Then we describe authentication and derivation of keys in more detail.
-%%
-%The latter is what differentiate the key establishment methods from each other.
-%%
+This contains encrypted and potentially signed MAC values, as we will describe in more detail in Section~\ref{sec:edhocauth}.
 %
-%\subsubsection{Protocol Pattern}
-%The first two messages negotiate authentication methods \mMethod{} and a
-%ciphersuite \mSuites{}.
-%%
-%In \mMethod{}, the initiator $I$ proposes which authentication method each
-%party should use.
-%%
-%These may differ, leading to four possible combinations:
-%\mSigSig{}, \mSigStat{}, \mStatSig{} and \mStatStat{}.
-%%
-%We refer to these \emph{combinations} of authentication methods simply as
-%\emph{methods} to align with the \mSpec{} terminology.
-%%
-%The first authentication method in a combination is the one proposed for the
-%initiator and the last is the one proposed for the responder.
-%%
-%The responder $R$ may reject the choice of method
-%or cipher suite with an error message, resulting in negotiation across
-%multiple \mEdhoc{} sessions.
-%%
-%Our analysis excludes error messages.
+The keys for these are generated by using a key schedule which takes at least one DH key \mGxy{} as basic input, as shown in Figure~\ref{fig:kdfdiagram}, and is intrinsic to the functioning of \mEdhoc{}.
+%
+\begin{figure}[!h]
+\scalebox{.785}{
+\input{kdfdiagram.tex}
+}
+\caption{Key schedule: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo, \mPRKthree, \mPRKfour are the intermediate key material, and \mKtwoe,\mKtwom, \mKthreeae, \mKthreem are the encryption keys for \mAead{} or \mXor{}.
+Dashed boxes are conditionals.~\cite{Norr21}}
+\label{fig:kdfdiagram}
+\end{figure}
 %
 
-\mEdhoc{}'s first two messages also exchange connection identifiers \mCi{} and
-\mCr{}, and public ephemeral keys, \mGx{} and \mGy{}.
+To derive keys, \mEdhoc{} uses two functions from the \mHkdf{} interface~\cite{rfc5869}, \mHkdfExtract{} and \mHkdfExpand{}.
 %
-The connection identifiers \mCi{} and \mCr{}, described in Section 3.1 of the
-\mSpec{}, deserve some elaboration.
+Both functions take as argument two values -- a salt, and an input. 
 %
-The \mSpec{} describes these identifiers as not serving a security purpose for
-\mEdhoc{}, but only as aiding message routing to the correct \mEdhoc{} processing
-entity at a party.
-%
-Despite this, the \mSpec{} states that they may be used by \mOscore{}, or other
-protocols using the established security context, without restricting how they
-are to be used.
-%
-Because \mEdhoc{} may need them in clear-text for routing, \mOscore{} cannot
-rely on them being secret.
-%
-Section 7.1.1 of the \mSpec{} requires the identifiers to be unique.
-%
-Uniqueness is defined to mean that $\mCi{} \not = \mCr{}$ for a given session
-and the \mSpec{} requires parties to verify that this is the case.
-%
-The same section also require that \mOscore{} must be able to retrieve the
-security context based on these identifiers.
-%
-The intended usage of \mCi{} and \mCr{} by \mOscore{} is not made specific and
-therefore it is not clear which properties should be verified.
-%
-We verify that the parties agree on the established values.
+For \mHkdfExtract{}, the input is a DH key, while for \mHkdfExpand{}, it is intermediate key material.
 %
 
-The two last messages provide identification and authentication.
+As mentioned earlier, the fundamental building block for the key schedule is the ephemeral DH key \mGxy{}, which  is computed in two different ways by $I$ (as $\mDH(\mX, \mGy)$) and $R$ (as $\mDH(\mY, \mGx)$).
 %
-Parties exchange identifiers for their long-term keys, \mIdcredi{} and \mIdcredr{},
-as well as information elements, \mAuthi{} and \mAuthr{}, to authenticate
-that the parties control the corresponding long-term keys.
+This key gives rise to intermediate keys \mPRKtwo{}, \mPRKthree{} and
+\mPRKfour{}, which can be derived as part of protocol execution.
 %
-The content of \mAuthi{} and \mAuthr{} depends on the authentication method
-associated with the corresponding long-term key.
-%
-For example, if $\mMethod{} = \mSigStat{}$, the responder $R$ must either
-reject the offer or provide an \mIdcredr{} corresponding to a key pair
-$\langle\mPriv{R},\ \mPub{R}\rangle$ suitable for use with challenge-response
-signatures, and
-compute \mAuthr{} based on the static key authentication method \mStat{}.
-%
-In turn, the initiator $I$ must provide an \mIdcredi{} corresponding to a key
-pair $\langle\mPriv{I},\ \mPub{I}\rangle$ suitable for a regular signature,
-and compute
-\mAuthi{} based on the signature based authentication method \mSig{}.
+Each intermediate key gives rise to encryption and integrity keys (\mKtwoe, \mKtwom{}, \mKthreeae, and \mKthreem) for a particular message in the protocol.
 %
 
-\subsubsection{Authentication}
-Regardless of whether \mStat{} or \mSig{} is used to compute \mAuthr{}, a
-MAC is first computed over \mIdcredr{}, \mCredr{}, a transcript hash of the
-information exchanged so far, and \mADtwo{} if included.
+In order to generate the final keys, the two \mHkdf{} algorithms use various values for salt.
 %
-The MAC is the result of encrypting the empty string with the Authenticated
-Encryption with Additional Data (AEAD) algorithm from the ciphersuite
-\mSuites{}, using the mentioned information as additional data.
+\mPRKtwo{} is generated by the \mHkdfExtract{} algorithm while using the empty string as the salt.
 %
-The MAC key is derived from the ephemeral key material
-\mGx{}, \mGy{}, \mX{} and \mY{}, where $I$
-computes $\mDH(\mX,\ \mGy)$ and $R$ computes $\mDH(\mY,\ \mGx)$, both resulting in
-the same output in the usual DH way.
+\mPRKthree{} and \mPRKfour{} are separately generated if $R$ or $I$ uses the \mStat{} method, using the corresponding DH key as input and the previous intermediate key as salt.
+%
+The key $P_{R}$ is used if the responder uses the \mStat{} authentication method.
+%
+$P_{R}$ is the key computed as $\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\ \mGx)$.
+%
+On the other hand, the key $P_{I}$ is used if the initiator uses \mStat{}, and is computed as $\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$.
+
+These intermediate keys are fed into an \mHkdfExpand{} algorithm, as shown in Figure~\ref{fig:kdfdiagram}. 
+%
+This algorithm uses as the salt a value $\textit{th}$, which is a running hash of the information transmitted thus far as part of the protocol.
+%
+$\textit{th}_{i}$ denotes the hash corresponding to the $i^{\rm{th}}$ message.
+
+At the end of a successful run of the protocol, the session key material for \mOscore{} is established as \mSessKey{}, which is a set of various keys.
+%
+This set always includes \mGxy{}, and if the initiator (resp. the responder) uses the \mStat{} authentication method, then it also includes $P_{I}$ (resp. $P_{R}$).
+%
+Once \mSessKey{} has been established, an \mHkdf{}-based key exporter named \mEdhoc-Exporter extracts the keys required for \mOscore{}.
 %
 
-In case $R$ uses the \mSig{} authentication method, \mAuthr{} is $R$'s
-signature over the MAC itself and the data that the MAC already covers.
+\subsection{About authentication in \mEdhoc{}}
+\label{sec:edhocauth}
+We now describe how the authentication information is constructed, depending on which method is used.
 %
-In case $R$ uses the \mStat{} authentication method, \mAuthr{} is simply the
-MAC itself.
+For both \mStat{} and \mSig{}, the following information is used to compute \mAuthr{}: \mIdcredr{}, \mCredr{}, a transcript hash of all the communicated information thus far in the protocol, and \mADtwo{}, if included.
 %
-However, when using \mStat{}, the key for the MAC is derived, not only from the
-ephemeral key material, but also from $R$'s long-term key
-$\langle\mPriv{R},\ \mPub{R}\rangle$.
+\mAuthi{} uses the same pieces of information corresponding to the initiator role.
 %
-For those familiar with \mOptls, this corresponds to the 1-RTT semi-static
-pattern computing the MAC key \textsf{sfk} for the \textsf{sfin}
-message~\cite{DBLP:conf/eurosp/KrawczykW16}.
+A MAC is obtained by feeding this material as additional data and the empty string as input to the Authenticated Encryption with Additional Data (AEAD) encryption algorithm as indicated in the established ciphersuite \mSuites{}.
 %
-The content of \mAuthi{} is computed in the corresponding way for the initiator
-$I$.
+The encryption key for the AEAD algorithm is constructed, for both roles, using the ephemeral key material \mGx{}, \mGy{}, \mX{}, and \mY{}. 
 %
-In Figure~\ref{fig:edhocFramework}, we denote a MAC using a key derived from
-both $\langle\mPriv{R},\ \mPub{R}\rangle$ and $\langle\mX,\ \mGx\rangle$ by
-$\mathit{MAC}_I$, and a MAC using a key derived from
-both $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\ \mGy\rangle$ by
-$\mathit{MAC}_R$.
+The initiator computes $\mDH(\mX, \mGy)$ while the responder computes $\mDH(\mY, \mGx)$, and DH operations give rise to the same key under both these computations.
+
+When $I$ uses the \mSig{} authentication method, \mAuthi{} is $I$'s signature over the MAC along with the data covered by the MAC.
 %
-
-Parts of the last two messages are encrypted and integrity protected, as
-indicated in~\ref{fig:edhocFramework}.
+However, when $I$ uses the \mStat{} method, \mAuthi{} is just the MAC, with one difference: the MAC key is derived using both the ephemeral key material as well as the long-term key for the initiator $\langle\mPriv{I},\ \mPub{I}\rangle$.
 %
-The second message is encrypted by XORing the output of the key derivation
-function HKDF (see Section~\ref{sec:keysched}) on to the plain text.
+This is similar to the 1-RTT semi-static pattern in \mOptls which computes the MAC key \textsf{sfk} for the message \textsf{sfin}~\cite{DBLP:conf/eurosp/KrawczykW16}.
 %
-The third message is encrypted and integrity protected by the AEAD algorithm
-determined by the ciphersuite \mSuites{}.
+The same procedure works for $R$ as well (with the values corresponding to $R$). 
 %
-
-\begin{figure}
+The abstract calculation for the authentication values is as shown in Table~\ref{tab:authvalues}~\cite{Norr21}.
+%
+We use $\mathit{MAC}_{I}$ (resp.
+$\mathit{MAC}_{R}$) to denote a MAC which uses an encryption key constructed using $\langle\mPriv{R},\ \mPub{R}\rangle$ and $\langle\mX,\ \mGx\rangle$ (resp.
+using $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\ \mGy\rangle$).
+\begin{table}[ht]
 \centering
-\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
-\begin{tikzpicture}
-    \node (ini) at (0, 0) {Initiator};
-    \draw [very thick] (0, -0.25) -- (0,-2.3);
-    \draw [very thick] (5.75, -0.25) -- (5.75,-2.3);
-    \node (res) at (5.75,0) {Responder};
-    \msg{1em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
-    \msg{3em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, \{\mIdcredr, \mAuthr, \mADtwo\}};
-    \msg{5em}{ini}{res}{\mMsgthree: \mCr, \{\mIdcredi, \mAuthi, \mADthree\}};
-    \draw [line width=1mm] (-0.75,-2.3) -- (0.75,-2.3);
-    \draw [line width=1mm] (5.75-0.75,-2.3) -- (5.75+0.75,-2.3);
-    \node (padding) at (0,-2.5) {};
-    \end{tikzpicture}
 \begin{tabular}{|c|c|c|}
@@
 \end{tabular}
-    \caption{Structure of \mEdhoc{}: $\{t\}$ means $t$ is encrypted and integrity
-protected.}
-\label{fig:edhocFramework}
-\end{figure}
-
-\subsubsection{Key Schedule}
-\label{sec:keysched}
-At the heart of \mEdhoc{} is the key schedule depicted in
-Figure~\ref{fig:kdfdiagram}.
-%
-\mEdhoc{} uses two functions from the \mHkdf{} interface~\cite{rfc5869} to derive keys.
-%
-\mHkdfExtract{} 
-constructs uniformly distributed key material from random input and a salt,
-while \mHkdfExpand{} generates keys from key material and a salt.
+\caption{The outer functions for each method \mMethod{}~\cite{Norr21}}
+\label{tab:authvalues}
+\end{table}
 %
 
-The key schedule is rooted in the ephemeral DH key
-\mGxy{}, which is computed as $\mDH(\mX, \mGy)$ by $I$ and as $\mDH(\mY, \mGx)$
-by $R$.
+In addition to this, communications 2 and 3 contain messages which are encrypted and integrity protected, as can be seen in~\ref{fig:edhocFramework}.
 %
-From \mGxy{}, three intermediate keys \mPRKtwo, \mPRKthree{} and
-\mPRKfour{} are derived during the course of protocol execution.
+The second communication is encrypted by performing an \mXor{} operation between the plaintext and the output of the key derivation function \mHkdf{}, as in Section~\ref{sec:keysched}.
 %
-Each of them is used for a specific message in the protocol, and from
-these intermediate keys, encryption and integrity keys
-(\mKtwoe, \mKtwom{}, \mKthreeae, and \mKthreem) for that message are derived.
+For the third message, encryption and integrity protection is assured by the \mAead{} algorithm which is part of the established ciphersuite \mSuites{}.
+
+In Figure~\ref{fig:edhocsigstat} we show an example of protocol execution under the \mSigStat{} method. 
 %
-The salt for generating \mPRKtwo{} is the empty string.
-%
+The figure describes in detail the various message patterns, and operations and key derivations used to construct these messages.
+\vnote{Do we still have the other method diagrams that we wrote out for the original paper but did not include?}
 
-The protocol uses a running transcript hash $th$, which includes the information
-transmitted so far.
-%
-The value of the hash, denoted $th_i$ for the $i$th message, is included in key
-derivations as shown in Figure~\ref{fig:kdfdiagram}.
-%
-
-Successful protocol execution establishes the session key material \mSessKey{}
-for \mOscore{}.
-%
-\mSessKey{} can be considered a set that always includes \mGxy{}.
-%
-If the initiator uses the \mStat{} authentication method, \mSessKey{} also
-includes
-$\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$, which we denote $P_I$.
-%
-If the responder uses the \mStat{} authentication method, it also includes
-$\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\ \mGx)$, which we denote $P_R$.
-%
-From the session key material, a key exporter (\mEdhoc-Exporter) based on
-\mHkdf{} is used to extract keys required for \mOscore{}.
-%
-
-\begin{figure*}[!h]
-\centering
-\scalebox{.785}{
-\input{kdfdiagram.tex}
-}
-\caption{Key schedule: Light blue boxes hold DH keys ($P_e, P_I, P_R$),
-orange boxes intermediate key material (\mPRKtwo, \mPRKthree, \mPRKfour), and
-dark blue boxes keys for \mAead{} or \mXor{} encryption
-(\mKtwoe, \mKtwom, \mKthreeae, \mKthreem).
-Dashed boxes are conditionals.}
-\label{fig:kdfdiagram}
-\end{figure*}
-
-As an illustrative example of the entire process, we refer to
-Figure~\ref{fig:edhocsigstat}, which depicts the protocol pattern, operations
-and key derivations for the \mSigStat{} method in more detail.
-%
-\begin{figure*}[ht]
+\begin{figure}[ht]
 \centering
@@
     \end{array}$};
-    \action{3.5em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\ \mGx\rangle$};
-    \msg{6.5em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
-    \action{7.0em}{res}{$
+    \action{3.7em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\ \mGx\rangle$};
+    \msg{7em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
+    \action{8em}{res}{$
       \begin{array}{c}
@@
       \end{array}$};
-    \msg{21.7em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\ \mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
-    \action{22.5em}{ini}{$
+    \msg{24em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\ \mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
+    \action{25em}{ini}{$
       \begin{array}{c}
@@
       \end{array}$};
-    \msg{35.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae; \mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
-    \action{36em}{res}{$
+    \msg{38.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae; \mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
+    \action{40em}{res}{$
     \begin{array}{c}
@@
 }
-    \caption{The \mSigStat{} method.
-    Tuples are denoted $\langle\cdot\rangle$, and the hash function \mHash{} is
-    as determined by \mSuites{}.}
+    \caption{The \mSigStat{} method for \mEdhoc{}. $\langle\cdot\rangle$ denotes a tuple, and the hash function \mHash{} is as established in the ciphersuite \mSuites{}.~\cite{Norr21}}
 \label{fig:edhocsigstat}
-\end{figure*}
+\end{figure}
 %------------------------------------------------------------------------- sub

### 2022-07-15 kdfdiagram.tex

% !TEX root =  protocol.tex

\begin{tikzpicture}[%
    >=latex,              % Nice arrows; your taste may be different
    start chain=going below,    % General flow is top-to-bottom
    node distance=4mm and 60mm, % Global setup of box spacing
    every join/.style={norm},   % Default linetype for connecting boxes
    ]
% ------------------------------------------------- 
% A few box styles 
% <on chain> *and* <on grid> reduce the need for manual relative
% positioning of nodes
\tikzset{
terminput/.style={rounded corners},
term/.style={rounded corners},
  base/.style={draw, thick, on chain, on grid, align=center, minimum height=4ex},
  dhkbox/.style={draw=cbgreen, fill=cbgreen!25, rectangle},
  dhk/.style={base, dhkbox},
  prkbox/.style={draw=cborange, fill=cborange!25, rectangle},
  prk/.style={base, prkbox},
  %hkdfext/.style={base, draw=Green3, fill=Green3!25, isosceles triangle, isosceles triangle apex angle=60, anchor=base, shape border rotate=-90, text width=6em},
  hkdfext/.style={base, draw=black, fill=none, rectangle},
  %hkdfexp/.style={base, draw=orange, fill=orange!50, isosceles triangle, isosceles triangle apex angle=60, anchor=base, shape border rotate=-90, text width=6em},
  hkdfexp/.style={base, draw=black, fill=none, rectangle},
  keybbox/.style={draw=cbnavy, fill=cbnavy!25, rectangle},
  keyb/.style={base, keybbox, text width=4em},
  % -------------------------------------------------
  norm/.style={->, draw, black},
  cond/.style={base, draw=black, dashed, fill=none, rectangle},
  txt/.style={base, draw=none, fill=none}
  }
% -------------------------------------------------
% Start by placing the nodes
%\node [terminput] (u1) {Salt};
%\node [hkdfext, join] (h1) {\mHkdfExtract};
\node [prk, join] (p2) {\mPRKtwo};
\node [cond, join] (c1) {R uses \mStat?};

\node [prk, below=6mm of c1.south] (p3) {\mPRKthree};
\draw [->, norm] (c1.south) -- (p3.north) node[midway, right] {N};

\node [cond, join, below=10mm of p3.south] (c2) {I uses \mStat?};
\node [prk, below=8mm of c2.south] (p4) {\mPRKfour};
\draw [->, norm] (c2.south) -- (p4.north) node[midway, right] {N};

\node [hkdfext, right=3cm of p3] (h3) {\mHkdfExtract};
\node [hkdfext, right=3cm of p4] (h5) {\mHkdfExtract};

\node [hkdfexp, shape border rotate=180, left= 2.5cm of p4] (h6) {\mHkdfExpand};
\node [keyb, join, left=3cm of h6] (k3) {\mKthreem};
\node [hkdfexp, shape border rotate=180, below= 0.8cm of h6] (h9) {\mHkdfExpand};
\node [txt, join, left=1cm of h9.west] (t4) {EDHOC-Exporter};

\node [hkdfexp, shape border rotate=180, left= 2.5cm of p3] (h4) {\mHkdfExpand};
\node [keyb, join, left=3cm of h4] (k2) {\mKtwom};

\node [hkdfexp, shape border rotate=180, left= 2.5cm of p2] (h2) {\mHkdfExpand};
\node [keyb, join, left=3cm of h2] (k1) {\mKtwoe};

\node [hkdfexp, shape border rotate=180, below= 0.8cm of h4] (h8) {\mHkdfExpand};
\node [keyb, below=0.8cm of k2] (k2b) {\mKthreeae};

\node [txt, left=1cm of k1.west] (t1) {Enc (XOR) \\ in m2};
\node [txt, left=1cm of k2.west] (t2) {\mMactwo~(signed if \\ R uses \mSig)};
\node [txt, left=1cm of k2b.west] (t2b) {\mAead\ in m3};
\node [txt, left=1cm of k3.west] (t3) {\mMacthree~(signed if \\ I uses \mSig)};

\draw [->, norm] (k1.west) -- (t1.east);
\draw [->, norm] (k2.west) -- (t2.east);
\draw [->, norm] (k2b.west) -- (t2b.east);
\draw [->, norm] (k3.west) -- (t3.east);

\draw [->, norm] (p3.south) ++(0,-0.5) -- (h8);
\draw [->, norm] (h8) -- (k2b);
\draw [->, norm] (p2) -- (h2); 
\draw [->, norm] (c1.east) -- ++(1.85,0) -- (h3.north) node[midway,above left] {Y};
%\draw [->, norm] (h3.south) -- ++(0,-1) -- ++(-3,0);
\draw [->, norm] (h3.west) -- (p3.east);
\draw [->, norm] (p3) -- (h4); 
\draw [->, norm] (c2.east) -- ++(1.91,0) -- (h5.north) node[midway,above left] {Y};
%\draw [->, norm] (h5.south) -- ++(0,-1) -- ++(-3,0);
\draw [->, norm] (h5.west) -- (p4.east);
\draw [->, norm] (p4) -- (h6);
\draw [->, norm] (p4.west) ++(-0.25,-0) -- ++(0,-0.8) -- (h9.east);

\node [hkdfext, right=3cm of p2] (h1) {\mHkdfExtract};
\node [dhk, right=2.7cm of h1] (p0) {$\mGxy$};
\node [terminput, text width=2em, below = 0.2cm of p0] (u1) {Salt};
\draw [->] (h1.west) -- (p2.east);
\draw [->] (u1.west) -- ++(-1.52,0) -- (h1.south);
\draw [->] (p0.west) -- (h1.east);

\node [dhk, right = 2.7cm of h3] (u2) {$P_R$};
\draw [->, norm] (u2.west) -- (h3.east);

\node [dhk, right = 2.7cm of h5] (u3) {$P_I$};
\draw [->, norm] (u3.west) -- (h5.east);

\node [term, above = 0.55cm of h4] (u5) {\mTHtwo};
\draw [->, dotted, shorten >=1mm] (u5) -- (h4);
\draw [->, dotted, shorten >=1mm] (u5) -- (h2);

\node [term, above = 0.5cm of h6] (u6) {\mTHthree};
\draw [->, dotted, shorten >=1mm] (u6) -- (h6);
\draw [->, dotted, shorten >=1mm] (u6) -- (h8);

%\node [term, below= 1cm of h9] (u7) {\mTHfour};
\node [term, below=0.4cm of p4] (u7) {\mTHfour};
%\draw [->, dotted, shorten >=1mm] (u7) -- (h9.south);
\draw [-> , dotted ] (u7.west) -- ([yshift=-0.4em] h9.east);

%\matrix [draw, ultra thick, double, below=2em of t1.east] {
%  \node [dhkbox, semithick, label=right:DH key] {}; \\
%  \node [prkbox, semithick, label=right:Intermediate key material] {}; \\
%  \node [keybbox, semithick, label=right:Key for \mAead{} or \mXor] {}; \\
%};

%
% ------------------------------------------------- 
% 
%\path (h2.east) to node [near start, yshift=1em] {$n$} (c3); 
%  \draw [o->,lccong] (h2.east) -- (p8);
%\path (p3.east) to node [yshift=-1em] {$k \leq 0$} (c4r); 
%  \draw [o->,lcnorm] (p3.east) -- (p9);
% -------------------------------------------------
% A last flourish which breaks all the rules
%\draw [->,MediumPurple4, dotted, thick, shorten >=1mm]
%  (p9.south) -- ++(5mm,-3mm)  -- ++(27mm,0) 
%  |- node [black, near end, yshift=0.75em, it]
%    {(When message + resources available)} (p0);
% -------------------------------------------------
\end{tikzpicture}

### 2022-07-15 paper.tex

--- 2022-07-15 paper.tex
@@
 \label{sec:errorHandling}
-\mPoint{Error handling is undefined except for algo negotiation}
-Section 6 of the \mEdhoc{} specification\footnote{draft-ietf-lake-edhoc-14}
-states that error messages can be sent at any time and by any party.
+\mPoint{Error handling is undefined except for algo negotiation} Section 6 of the \mEdhoc{} specification\cite{draft-ietf-lake-edhoc-15} states that error messages can be sent at any time and by any party.
 %
-One error message type is defined (for algorithm negotiation), and the rest can
-be arbitrarily defined by the implementation.
+At the current version of the standard \cite{draft-ietf-lake-edhoc-15}, there are three types of error messages: type 0 is used to represent success, type 1 represents a generic error message used for debugging purposes, where the actual message is an English string, and type 2 represents failure to negotiate a common ciphersuite.
+Type 2 messages carry a list of supported ciphersuites by the responder.
 %
-The contents and semantics of these messages are also implementation defined.
+The contents and semantics of type 1 messages are dependent on the implementation.
 %
-Consequently, the error messages provide a generic message passing mechanism.
+Consequently, type 1 error messages provide a generic message passing mechanism, albeit one without a predefined semantics.
+
+%\textbf{Q:} Is there any situation where it makes sense that the initiator sends an unsupported ciphersuite message instead of message 3?
+%\textbf{Q:} In Section H: if no message correlation support by the transport layer, \mCi{} and \mCr{} should be used for correlation of messages to a session.
+
+\subsection{Lack of connection identifiers in error messages}
+Error messages do not include the optional connection identifiers \mCi{} or \mCr{} that are transmitted during the handshake.
+This is potentially problematic for connection-less transport layers, where \mCi{} and \mCr{} may be used to correlate messages.
+
+\subsection{Improper handling of error messages}
+\mEdhoc{} provides the error message mechanism, but gives little to no guidance on how it should be used safely.
 %
-
-\textbf{Q:} Is there any situation where it makes sense that the initiator sends
-a unsupported ciphersuite message instead of message 3?
+For example, assume that an application using \mEdhoc{} logs error messages on a finite log with log rotation.
 %
-
-\textbf{Q:} In Section H: if no message correlation support by the transport
-layer, \mCi{} and \mCr{} should be used for correlation of messages to a
-session.
+If the log is used for anomaly detection or detection of sensitive events, then an adversary can simply inject error messages and fill the log until the sensitive event is overwritten.
 %
-Error messages do not carry \mCi{} or \mCr{}; will this create a problem?
-%
-
-\mEdhoc{} provides the error message mechanism, but gives little to no guidance
-on how it should be used safely.
-%
-Example of issue with improper use: Assume application logs error messages a
-finite log with log rotation.
-%
-If the log is used for anomaly detection or detection of sensitive events, then
-an adversary can simply inject error messages and fill the log until the
-sensitive event is overwritten.
-%
-If the filling error messages are less suspicious than the actual attack, this
-may be beneficial for the adversary.
+If filling error messages in the log is less suspicious than the actual attack, this may be beneficial for the adversary.
 %
 Implementors should do proper log-separation and log management.
+
+\subsection{Unspecified semantics for debugging messages}
+Section 6.2 gives an example of sending a type 1 error message with the information string "Method not supported".
 %
-
-Section 6.2 gives an example of sending an error message "method not
-supported".
+It's important to note that this message cannot be used to negotiate the connection method securely.
 %
-This message cannot be used to negotiate method securely.
+In fact -- differently from ciphersuite negotiation -- \mEdhoc{} does not include a mechanism to support negotiating the connection method, among the four available combinations of signature keys and static Diffie-Hellman keys.
 %
-The reason is that error messages are not integrity protected, and an initator
-only proposes one method, and therefore the same tactic used for ciphersuite
-negotiation cannot be used for methods.
+The initiator selects the method on the first message, and the responder can reply with a type 1 error message along with the information string "Method not supported", but a standard compliant implementation cannot act upon this information string for negotiation purposes.
 %
-Because method negotiation can be viewed in the context of crypto agility, there
-is room for improvement.
+This is because the error message is not integrity protected and there is no secure mechanism to select a different authentication method that ensure that the preferred available authentication mechanism is always selected.
+%
+Because method negotiation can be viewed in the context of cryptographic agility, there is room for improvement.
 %

### 2022-07-18 formalization.tex
Changes to section 3

@@
 \subsection{Formal Model}\label{sec:threat-model}
-\vnote{Can we add extra references here? SAPIC+ etc?} As in~\cite{Norr21}, we verify \mEdhoc{} in the Dolev-Yao model~\cite{DY83}.
+\vnote{Can we add extra references here? SAPIC+ etc? Can leak session keys as well.}
+
+As in~\cite{Norr21}, we verify \mEdhoc{} in the Dolev-Yao model~\cite{DY83}.
 %
@@
 
-We now briefly describe \mTamarin{}, and how protocols are modelled in it.
+We now briefly describe \mTamarin{}, and how we model \mEdhoc{} in it here.
 
@@
 The elements of these multi-sets, called facts, contribute to the global system state.
+%
+For syntactic sugar, \mTamarin{} also allows the use of let-bindings and tuples. 
+%
+For ease of presentation, we will present the model and properties in a slightly different syntax in this paper, but this syntax can be directly mapped to that of \mTamarin.
 
@@
 %
-Properties are defined as formulas in a first-order logic, and these formulas can be verified over execution traces.
+Event types are predicates over global states generated during protocol execution.
+%
+Consider an event type $E$ and a timestamp $t$ as part of a trace.
+%
+By $E^{t}(p_i)_{i\in\mathbb{N}}$, we represent an event of type $E$ occurring at time $t$ in a trace, parametrized by the sequence of values $(p_i)_{i\in\mathbb{N}}$ (corresponding to the action fact $E(p_i)_{i\in\mathbb{N}}@t$ in \mTamarin).
+%
+Thus, the time points form a quasi order, and we denote the fact that $t_{1}$ comes before $t_{2}$ in a protocol trace by $t_{1} \lessdot t_{2}$, and that $t_{1}$ and $t_{2}$ stand for the same time point in a trace by $t_{1} \doteq t_{2}$.
+%
+\mTamarin{} allows events to occur at the same time point, with one restriction: multiple events of the same type cannot occur simultaneously, so if $t_{1} \doteq t_{2}$, then $E^{t_{1}} = E^{t_{2}}$.
+
+Properties are defined as formulas in a fragment of temporal first order logic, and these formulas can be verified over execution traces.
 
@@
 
-For syntactic sugar, \mTamarin{} allows the use of let-bindings and tuples. 
 %
@@
 %
-However, we can extend and/or modify this behaviour by adding other rules as well.
-
-%-------------------------------------------------------------------------- sub
-
-%%%%%%%%%%% START HERE
-
-\subsection{Formalization of Properties}
-\label{sec:desired-properties}
-We use the \mTamarin{} verification
-tool~\cite{DBLP:conf/cav/MeierSCB13} to encode the model and verify properties.
+We can extend and/or modify this behaviour by adding other rules as well.
 %
-This tool uses a fragment of temporal first order logic to reason about
-events and knowledge of the parties and of the adversary.
+To denote that the adversary has access to a message $p$ at time $t$, we use $\mK^{t}(p)$.
 %
-For conciseness we use a slightly different syntax than
-that used by \mTamarin{}, but which has a direct mapping to \mTamarin{}'s logic.
+This stands for the fact that as the adversary interacts with the parties executing the protocol, at time point $t$, the adversary can derive the term $p$ using the Dolev-Yao deduction rules and any additional adversary capabilities specified in \mTamarin.
 %
-
-Event types are predicates over global states of system execution.
-%
-Let $E$ be an event type and let $t$ be a timestamp associated with a point in a
-trace.
-%
-Then $E^{t}(p_i)_{i\in\mathbb{N}}$ denotes an event of type $E$ associated with
-a sequence of parameters $(p_i)_{i\in\mathbb{N}}$ at time $t$ in that trace.
-%
-In general, more than one event may have the same timestamp and hence
-timestamps form a quasi order, which we denote by $t_1 \lessdot t_2$ when $t_1$
-is before $t_2$ in a trace.
-%
-We define $\doteq$ analogously.
-%
-However, two events of the same type cannot have the same timestamp, so
-$t_1 \doteq t_2$ implies $E^{t_1} = E^{t_2}$.
-%
-Two events having the same timestamp does not imply the there is a fork in the
-trace, only that the two events happen simultaneously and that comparing their
-timestamps results in that they are equal.
-%
-This notation corresponds to \mTamarin{}'s use of action facts
-$E(p_i)_{i\in\mathbb{N}}@t$.
-%
-
-The event $\mK^t(p)$ denotes that the adversary knows a parameter $p$ at
-time $t$.
-%
-Parameters are terms in a term algebra of protocol specific operations and
-generic operations, e.g., tuples $\langle\cdot\rangle$.
-%
-Intuitively, $\mK^t(p)$ evaluates to true when $p$ is in
-the closure of the
-parameters the adversary observed from interacting with parties using the
-protocol, under the Dolev-Yao message deduction operations and
-the advanced adversary capabilities up until time $t$.
-%
-For a more precise definition of knowledge management, we refer to~\cite{DBLP:conf/cav/MeierSCB13}.
-%
-An example of a formula is
+As an example, consider the following implication 
 \[
@@
 \]
-expressing that if there is a time $t$ when the adversary knows the tuple
-$\langle k, k'\rangle$, then the adversary knows both
-$k$ and $k'$ at the same point in time.
+which represents the fact that if the adversary gets to know the pair of keys $\langle k, k' \rangle$ at a time point $t$, then the adversary also knows each of those keys $k$ and $k'$ at time point $t$ as well.
+%
+For more details about how \mTamarin{} manages adversary knowledge, see~\cite{DBLP:conf/cav/MeierSCB13}.
 %
 
-An initiator $I$ considers the
-protocol run started when it sends a message \mMsgone{} (event type \mIStart)
-and the run completed after sending a message \mMsgthree{} (event type
-\mIComplete).
-%
-Similarly, a responder $R$ considers the run started upon receiving
-a \mMsgone{} (event type \mRStart), and completed upon receiving a \mMsgthree{}
-(event type \mRComplete).
+%-------------------------------------------------------------------------- sub
+
+\subsection{Model and Desired Properties}
+\label{sec:desired-properties}
+In this section, we describe how we model \mEdhoc{} and the various properties we want it to satisfy.
 %
 
-%----------------------------------------------------------------------- PFS
-\subsubsection{Perfect Forward Secrecy (PFS)}
-\label{sec:secrecy}
-Informally, PFS captures the idea that session key material remains secret
-even if a long-term key leaks in the future.
+The party $I$ executing the initiator role considers a run of the protocol begun as soon as it sends the first message \mMsgone{} with event type \mIStart, and considers the run ended once it has sent the third message \mMsgthree{} with event type \mIComplete.
 %
-We define PFS for session key material \mSessKey{} as \mPredPfs{} in
-Figure~\ref{fig:props}.
+Similarly, the responder $R$ considers a run started upon receiving \mMsgone{} with event type \mRStart, and finished upon receiving \mMsgthree{} with type \mRComplete.
 %
 
-\begin{figure*}
+%
+In this work, we consider the following properties: perfect forward secrecy, injective agreement for both the initiator and the responder, and implicit agreement. 
+%
+Agreement is considered on a set of parameters $S$ which also contains the session key material \mSessKey{}.
+%
+We will first describe in detail both properties, and then describe the contents of the set $S$.
+%
+We formalize these properties as shown in Figure~\ref{fig:props}, which is a modified version of a figure in~\cite{Norr21}.
+\vnote{The figure changes because of TEEs}
+
+\begin{figure*}[ht]
 \begin{align*}
@@
 
-The first parameter $I$ of the \mIComplete{} event represents the
-initiator's identity,
-and the second, $R$, represents that $I$ believes $R$ to be playing
-the responder role.
+%----------------------------------------------------------------------- PFS
+\subsubsection{Perfect Forward Secrecy (PFS): }
+\label{sec:secrecy}
+Informally, by perfect forward secrecy, we mean that the adversary cannot gain access to the session key material \mSessKey{}, even if a long-term key is leaked after the establishment of \mSessKey{}.
 %
-The third parameter, \mSessKey{}, is the established session key material.
+PFS for session key material \mSessKey{} is formalized as \mPredPfs{} in Figure~\ref{fig:props}.
+
+In the \mIComplete{} event, $I$ represents the identity of the initiator, and $R$ represents the identity of the party who the initiator believes is playing the responder role, while \mSessKey{} stands for the established session key material.
 %
-The parameters of the \mRComplete{} event are defined analogously.
+\mRComplete{} has analogous parameters; in particular, the responder $R$ believes $I$ is the party playing the initiator role.
 %
-Specifically, the first parameter of \mRComplete{} represents the identity of
-whom $R$ believes is playing the initiator role.
+Essentially, the property states that an adversary obtains \mSessKey{} only if a party's long-term key is compromised before their run ends, or if an ephemeral key is compromised at any time after a party's run began.
+\vnote{This has to change once we add TEEs}
 %
-The essence of the definition is that an adversary only knows \mSessKey{} if they
-compromised one of the
-parties long-term keys before that party completed the run, or if the adversary
-compromised any of the ephemeral keys at any time after a party starts
-its protocol run.
+The property in Figure~\ref{fig:props} is unlike the actual \mTamarin{} lemma in one minor way: \mTamarin's syntax does not allow disjunctions to appear on the left-hand side of an implication inside a universally-quantified formula.
 %
-One way the definition slightly differs from the corresponding \mTamarin{} lemma
-is that \mTamarin{} does not allow a disjunction on the left-hand side of an
-implication in a universally quantified formula.
-%
-In the lemma, therefore, instead of the disjunction
-$\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$,
-we use a single action parametrized by $I$, $R$, and \mSessKey{} to signify that
-\emph{either} party has completed their role.
-%
+Therefore, in the \mTamarin{} code, instead of using the disjunction $\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$ to model the fact that either party can have completed their execution, we use a single action parametrized by the terms $I$, $R$, and \mSessKey.
 
 %-------------------------------------------------------------------- InjAgree
-\subsubsection{Authentication}
+\subsubsection{Authentication Properties:}
 \label{sec:authenticationDef}
-We prove two different flavors of authentication, the first being classical
-\emph{injective agreement} following Lowe~\cite{DBLP:conf/csfw/Lowe97a}, and
-the second being an implicit agreement property.
+Following~\cite{Norr21}, we prove two different kinds of authentication properties, namely \emph{injective agreement} in the style of~\cite{DBLP:conf/csfw/Lowe97a}, and implicit agreement.
 %
-Informally, injective agreement guarantees to an initiator $I$ that whenever
-$I$ completes a run ostensibly with a responder $R$,
-then $R$ has been engaged in the protocol as a responder,
-and this run of $I$ corresponds to a unique run of $R$.
+Injective agreement can be guaranteed to either party running the protocol.
 %
-In addition, the property guarantees to $I$ that the two parties agree on a set
-$S$ of parameters associated with the run, including, in particular, the
-session key material \mSessKey{}.
+For the initiator $I$, it guarantees to $I$ that whenever $I$ believes that they have completed a run with $R$ as responder, then the party $R$ has indeed executed the protocol in the role of a responder, and that this run of $I$ uniquely corresponds to one of $R$ where the set of parameters is $S$ and includes, in particular, the session key material \mSessKey{}.
 %
-However, we will treat \mSessKey{} separately for clarity.
-%
-On completion, $I$ knows that $R$ has access to the session key material.
-%
-The corresponding property for $R$ is analogous.
-%
-
-Traditionally, the event types used to describe injective agreement are called
-\emph{Running} and \emph{Commit}, but to harmonize the presentations of
-authentication and PFS in this section, we refer to these event types as
-\mIStart{} and \mIComplete{} respectively for the initiator, and
-\mRStart{} and \mRComplete{} for the responder.
-%
-For the initiator role we define injective agreement by
-\mPredInjI{} in Figure~\ref{fig:props}.
+It can be defined for $R$ in a similar manner.
 
 %
-The property captures that for an initiator $I$, either the injective agreement
-property as described above holds, or the long-term key of the believed
-responder $R$ has been compromised before $I$ completed its role.
+We formalize injective agreement for the initiator role as \mPredInjI{} and for the responder role as \mPredInjR{} in Figure~\ref{fig:props}.
 %
-Had the adversary compromised $R$'s long-term key, they could have generated a
-message of their liking (different to what $R$ agreed on), and signed this or
-computed a $\mathit{MAC}_R$ based on \mPubE{I}, \mPriv{R} and their own chosen
-ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$.
+For the initiator $I$, this property represents the fact that either injective agreement (as described above) holds, or the long-term key of the party $R$ assumed to be playing the responder role has been compromised before $I$'s role finished.
 %
-This places no restrictions on the ephemeral
-key reveals, or on the reveal of the initiator's long-term key.
+If the adversary has managed to compromise $R$'s long-term key, they can generate any message of their liking and signed it using said key, or computed a $\mathit{MAC}_{R}$ using the long-term keys \mPubE{I}, \mPriv{R}, and an ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$ of their choice.
 %
-For the responder we define the property \mPredInjR{} as in
-Figure~\ref{fig:props}.
-%
+The analogous definition holds for the responder $R$.
+\vnote{Check this, and modify to deal with TEEs}
 
 %------------------------------------------------------------- Implicit auth
-Unlike PFS, not all \mEdhoc{} methods enjoy the injective agreement property.
+As part of our analysis, we found that all the \mEdhoc{} methods satisfy PFS.
 %
-Hence, we show for all methods a form of \emph{implicit agreement} on all the
-parameters mentioned above.
+However, this is not the case for the injective agreement property as stated above.
 %
-We take inspiration from the computational model definitions of implicit
-authentication, proposed by Guilhem~et~al.~\cite{DBLP:conf/csfw/GuilhemFW20}, to
-modify classical injective agreement into an implicit property.
+Thus, we show a different property, a form of implicit agreement on the same set of parameters, which is guaranteed for all methods.
 %
-A small but important difference between our definition and theirs, is that
-they focus on
-authenticating a key and related identities, whereas we extend the more general
-concept of agreeing on a set of parameters, starting from the idea of injective
-agreement~\cite{DBLP:conf/csfw/Lowe97a}.
+This modification is inspired by the definitions of implicit authentication in the computational model~\cite{DBLP:conf/csfw/GuilhemFW20}.
 %
-We use the term \emph{implicit} in this context to denote that a party $A$
-assumes that any other party $B$ who knows the session key material \mSessKey{} must
-be the intended party, and that $B$ (if honest) will also agree on a set
-$S$ of parameters computed by the protocol, one of which is \mSessKey{}.
+While that paper focuses on authenticating just a key and related identities, our definition encompasses a general set of parameters, as in the notion of injective agreement proposed by Lowe~\cite{DBLP:conf/csfw/Lowe97a}.
+
+The ``implicit'' in the name of the property stands for the fact that a party $A$ assumes that any party $B$ who has access to the session key material \mSessKey{} must, in fact, be the intended party, and that if $B$ is honest, $B$ will agree on a set $S$ of parameters which includes \mSessKey.
 %
-When implicit agreement holds for both roles, upon completion, $A$ is guaranteed
-that $A$ has been or is engaged in exactly one protocol run with $B$ in the
-opposite role, and that $B$ has been or will be able to agree on $S$.
+Implicit agreement for both roles guarantees to $A$ that $A$ is or has been involved in exactly one protocol execution with $B$, and that $B$ agrees or will agree with $A$ on $S$.
 %
-The main difference to injective agreement is that $A$ concludes that if
-$A$ sends the last message and this reaches $B$, then $A$ and $B$ have agreed
-on $I$, $R$ and $S$.
+This property diverges from injective agreement in that upon sending the last message, $A$ concludes that if this message reaches $B$, then $A$ and $B$ agree on each other's identities and roles, and the set $S$.
+
+Note that for implicit agreement to hold for $I$, the ephemeral keys must not be revealed, since the property relies on the fact that the intended responder is the only one who knows the session key material.
 %
-While almost full explicit key authentication, as defined by
-Guilhem~et~al.~\cite{DBLP:conf/csfw/GuilhemFW20}, is a similar property, our
-definition does not require key confirmation, so our definition is closer to
-their definition of implicit authentication.
+If the adversary has access to the ephemeral keys, they can use it along with the public keys of $I$ and $R$ to compute the session key material.
 %
-In the \mTamarin{} model we split the property into one lemma for
-$I$ (\mPredImpI{}) and one for $R$ (\mPredImpR{}) to save memory during
-verification.
+However, either party's long-term key can be revealed after the other party has finished their execution, since this still leaves the adversary unable to compute \mGxy{}.
+
+Since \mTamarin{} runs out of memory to verify this property as is, we split it into two lemmas -- \mPredImpI{} for $I$ for one \mPredImpR{} for $R$.
 %
-We show only the definition for $I$ in Figure~\ref{fig:props}, because it is
-symmetric to the one for $R$.
+Figure~\ref{fig:props} contains the definition for \mPredImpI{}.
 %
-
-For implicit agreement to hold for the initiator $I$, the ephemeral keys
-can never be revealed.
-%
-Intuitively, the reason for this is that the implicit agreement relies on that
-whomever knows the session key material is the intended responder.
-%
-An adversary with access to the ephemeral keys and the public keys of
-both parties can compute the session key material produced by all methods.
-%
-However, the responder $R$'s long-term key can be revealed after $I$ completes
-its run, because the adversary is still unable to compute $P_e$.
-%
-The initiator's long-term key can also be revealed at any time without affecting
-$I$'s guarantee for the same reason.
-%
+\mPredImpR{} is formalized similarly, so we omit it.
 
 %------------------------------------------------------- Agreed parameters
-\subsubsection{Agreed Parameters}
+\subsubsection{Parameters in set $S$:}
 \label{sec:agreedParams}
+We now describe the set $S$ of parameters upon which the two parties obtain guarantees via the above properties.
+%
 The initiator $I$ gets injective and implicit agreement guarantees on the
-following partial set $S_P$ of parameters:
+following partial set $S_P$ of parameters~\cite{Norr21}:
 \begin{itemize}
@@
     \item responder identity,
-    \item session key material (which varies depending on \mEdhoc{} method),
+    \item session key material (which varies depending on the \mEdhoc{} method),
     \item context identifiers \mCi{} and \mCr{}, and
@@
 %
-Because \mEdhoc{} aims to provide identity protection for $I$, there is no
-injective agreement guarantee for $I$ that $R$ agrees on the initiator's
-identity.
+
+Due to the initiator being guaranteed identity protection under \mEdhoc{}, $I$ cannot get explicit agreement with $R$ on the initiator's identity.
 %
-For the same reason, there is no such guarantee for $I$ with respect to
-the $P_I$ part of the session key material when $I$ uses the \mStat{}
-authentication method.
+Similarly, when using the \mStat{} authentication method, $I$ does not get any such guarantees about $P_{I}$.
 %
-There is, however, an implicit agreement guarantee for $I$ that $R$ agrees on
-$I$'s identity and the full session key material.
+However, $I$ does get implicit agreement with $R$ about $I$'s identity and the full set $S_{F}$ of agreed parameters.
 %
-Since $R$ completes after $I$, $R$ can get injective agreement guarantees on
-more parameters, namely also the initiator's identity and the full session key
-material for all methods.
+In contrast, since $R$'s run finishes after that of $I$, $R$ can get explicit injective agreement assurances on the full set $S_{F}$ of agreed parameters.
 %
-The full set of agreed parameters $S_F$ is $S_P \cup \{I, P_I\}$
-when $P_I$ is
-part of the session key material, and $S_P \cup \{I\}$ otherwise.
+The full set of agreed parameters $S_F$ is $S_P \cup \{I, P_I\}$ when $P_I$ is part of the session key material, and $S_P \cup \{I\}$ otherwise.
 %
 
-%------------------------------------------------------- Implied properties
-\subsubsection{Inferred Properties}
-From the above, other properties can be inferred to hold in our adversary model.
+In addition to these properties, a couple of properties can be inferred without being explicitly modelled and verified. 
 %
-Protocols where a party does not get confirmation that their peer knows the
-session key material may be susceptible to
-\emph{Key-Compromise Impersonation (KCI)}
-attacks~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
+One might be interested in checking is called Key-Compromise Impersonation (KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
 %
-Attacks in this class allow an adversary in possession of a party $A$'s secret
-long-term key to coerce $A$ to complete a
-protocol run believing it authenticated a certain peer $B$, but where $B$ did
-not engage with $A$ at all in a run.
+Key-Compromise Impersonation (KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97} often occurs in protocols where a party cannot be assured that the peer they are involved in a run with actually knows the session key material.
 %
-Because both our notions of agreement above ensure agreement on identities,
-roles and session key material, all methods passing verification of those are
-also resistant to KCI attacks.
+A KCI attack occurs when an adversary who has access to $A$'s long-term private key to make $A$ believe that they completed an execution with a peer $B$, while $B$ did not participate in said execution at all.
 %
+Our above notions of agreement ensure that both parties agree on each other's identity, role, and session key material.
+%
+Therefore, all \mEdhoc{} methods that satisfy these agreement properties also avoid KCI attacks.
 
-If a party $A$ can be coerced into believing it completed a run with $B$, but
-where the session key material is actually shared with $C$ instead, the 
-protocol is vulnerable to an \emph{Unknown Key-Share (UKS)}
-attack~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
+Another kind of attack is an Unknown Key Share (UKS)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
 %
-For the same reason as for KCI, any method for which our agreement
-properties hold is also resistant to UKS attacks.
+As part of a UKS attack, a party $A$ can be made to believe that it finished an execution with a party $B$, but where the session key material is actually shared between $A$ and $C$ instead.
 %
+Again, due to the agreement on identities and session key material, any method that satisfies the above agreement properties also resists UKS attacks.
+%
+Overall, the agreement properties capture entity authentication, and satisfy any properties based on that notion.
 
-From the injective agreement properties it follows that each party is assured
-the identity of its peer upon completion.
-%
-Therefore, the agreement properties also capture \emph{entity authentication}.
-%
+\vnote{Is there anything else we can add thanks to the TEEs?}
 
 %---------------------------------------------------------------------------sub
-\subsection{\mTamarin{} Encoding of \mEdhoc{}}
+\subsection{Encoding \mEdhoc{} in \mTamarin}
 \label{sec:modeling}
 %
-We model all four methods of \mEdhoc{}, namely
-\mSigSig, \mSigStat, \mStatSig{} and \mStatStat.
+In this section, we describe how we model the \mSigSig, \mSigStat, \mStatSig, and \mStatStat{} methods of \mEdhoc{} in \mTamarin.
 %
-Because the methods share a lot of common structure, we derive
-their \mTamarin-models from a single specification written with the aid of the
-M4 macro language.
+As in~\cite{Norr21}, we construct the \mTamarin{} model by utilizing the fact that all methods of \mEdhoc{} share a common underlying structure (as shown in Figure~\ref{fig:edhocFramework}).
 %
-To keep the presentation brief, we only present the \mStatSig{} metohod, as it
-illustrates the use of two different asymmetric authentication methods
-simultaneously.
+We do so by writing a single specification file in the M4 macro language, which generates all the methods.
+%
+As in~\cite{Norr21}, we only present the \mStatSig{} method which illustrates the use of two different authentication methods.
 %
@@
 %
-Variable names used in the code excerpts here are sometimes shortened compared
-to the model itself to fit the paper format.
+
+As mentioned in Section~\ref{sec:tamarin}, we extend the default equational theory of \mTamarin{} to handle various operations used in \mEdhoc.
+%
+Of the built-in theories, we use the ones for exclusive-or (\mXor), Diffie-Hellman operations, signatures (\mT{sign} and \mT{verify} operations), and hashing~\cite{DBLP:conf/csfw/DreierHRS18,DBLP:conf/csfw/SchmidtMCB12}.
+%
+We augment the hashing function symbol to add an extra input, which we use to model different hash functions.
 %
 
-\subsubsection{Primitive Operations}
-Our model uses the built-in theories of exclusive-or and DH operations, as
-in~\cite{DBLP:conf/csfw/DreierHRS18,DBLP:conf/csfw/SchmidtMCB12}.
+In addition to these default operations, \mEdhoc{} is built over \mHkdfExpand, \mHkdfExtract, and the \mAead{} functions.
 %
-Hashing is modeled via the built-in hashing function symbol augmented
-with a public constant as additional input, modelling different
-hash functions.
+We represent \mHkdfExpand by \mT{expa} and \mHkdfExtract by \mT{extr}.
 %
-The HKDF interface is represented by \mT{expa} for the
-expansion operation and \mT{extr} for the extraction operation.
+\mAead{} operations need us to add extra equations to the underlying theory.
 %
-Signatures use \mTamarin's built-in theory for \mT{sign} and \mT{verify}
-operations.
+A term encrypted using \mAead{} is represented by \mT{aeadEncrypt(m, k, ad, ai)}, where \mT{m} is the underlying message, \mT{k} is the encrypting key, \mT{ad} is the additional data, and \mT{ai} is the identifier for the encryption algorithm.
 %
-For \mAead{} operations on key \mT{k}, message \mbox{\mT{m}}, additional data \mT{ad}
-and algorithm identifier \mT{ai}, we use \mT{aeadEncrypt(m, k, ad, ai)}
-for encryption.
+Decryption of such a term is defined via two equations that we add to \mTamarin's theory.
 %
-Decryption with verification of the integrity is defined via the equation
-\begin{small}\begin{verbatim}
-aeadDecrypt(aeadEncrypt(m, k, ad, ai),
-    k, ad, ai) = m.
-\end{verbatim}\end{small}
+The following equation requires the decrypting party to know the additional data \mT{ad} to decrypt this encrypted term with verification of its integrity. 
+\begin{small}
+\begin{verbatim}
+aeadDecrypt(aeadEncrypt(m, k, ad, ai), k, ad, ai) = m.
+\end{verbatim}
+\end{small}
 %
-The integrity protection of AEAD covers \mT{ad}, and this equation hence requires
-an adversary to know \mT{ad} even if they only wish to decrypt the data.
+Only the above equation is used by honest parties, but the adversary should also be able to decrypt without having to go through the additional step of identity verification.
 %
-To enable the adversary to decrypt without needing to verify the integrity
-we add the equation
+To this end, we also add the following equation, where the adversary does not need access to the additional data \mT{ad} in order to decrypt.
 \begin{small}\begin{verbatim}
@@
 %
-The latter equation is not used by honest parties.
+
+Having described how we adapt the equational theory to model \mEdhoc, we now move on to the modelling of the adversary model and the environment in which the protocol is executed.
+%
+We extend the built-in Dolev-Yao adversary rules which are part of \mTamarin.
+%
+We use the following rules to capture the link between a party's identity and their long-term key pairs, in the \mSig{}- and in the \mStat{}-based methods respectively.
+\begin{center}
+\begin{minipage}{0.48\textwidth}
+\begin{small}
+\begin{verbatim}
+1 rule registerLTK_SIG:
+2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
+3        [!LTK_SIG($A, ~ltk),
+4         !PK_SIG($A, pk(~ltk)),
+5         Out(<$A, pk(~ltk)>)]
+\end{verbatim}
+\end{small}
+\end{minipage}
+\hfill\vline\hfill
+\begin{minipage}{0.48\textwidth}
+\begin{small}
+\begin{verbatim}
+1 rule registerLTK_STAT:
+2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
+3        [!LTK_STAT($A, ~ltk),
+4         !PK_STAT($A, 'g'^~ltk),
+5         Out(<$A, 'g'^~ltk>)]
+\end{verbatim}
+\end{small}
+\end{minipage}
+\end{center}
+
+Using the fact \verb|Fr(~ltk)|, \mTamarin{} creates a new term \mT{ltk} and uses it to represent a secret long-term key.
+%
+Via the \verb|Out(<$A, pk(~ltk)>)| fact, \mTamarin{} puts out onto the communication channel the identity of the party to whom this long-term key belongs, along with their public key. 
+%
+Since the adversary has access to the communication channel, they can pick up all of this information.
+%
+The event \mT{UniqLTK} parametrized by a party's identity and their long-term key models the unique correspondence between those two values.
+%
+As a result, this rules out a party owning multiple long-term keys -- in particular, it keeps the adversary from registering long-term keys in some honest party's name.
+%
+This aligns well with an external mechanism such as a certificate authority ensuring that long-term keys are uniquely assigned to the corresponding identities, which is ensured by the \mSpec.
+
+To model the reveal of long-term keys and ephemeral keys to an adversary, we use standard reveal rules and events of type \mRevLTK{} and \mRevEph, respectively. 
+%
+It is also important to keep track of the time points at which these events occur.
+%
+Long-term keys can be revealed on registration, even before the protocol begins.
+%
+However, ephemeral keys in our model can only be revealed when a principal completes their role, i.e. simultaneous with events of type \mIComplete{} and \mRComplete.
+%
+While we did try to reveal ephemeral keys at the start of protocol execution, we could not get the verification to terminate in such a model.
+\vnote{Do we still not get \mTamarin{} to terminate if we reveal ephemeral keys at the start of the protocol? With the TEEs etc?}
+%
+Having set out the capabilities of the adversary, we now model the execution of the honest agents' roles.
+
+%%%%%%%%%%% START HERE
+For each protocol method, we use two rules apiece for the initiator and responder -- \mT{I1}, \mT{R2}, \mT{I3}, \mT{R4}.
+%
+Each of these stand for one step of the protocol as executed by either party.
+%
+To disambiguate, we will attach the method to the rule name.
+%
+These four rules directly map to the event types \mIStart, \mRStart, \mIComplete, and \mRComplete, respectively.
+%
+We show the \mT{R2\_STAT\_SIG} rule below to illustrate the various aspects of the \mTamarin{} modelling we are describing here.
+
+In order to keep track of the initiator's state, we use facts prefixed with \mT{StI}, which carry information between the \mT{I1} and \mT{I3} rules.
+%
+Similarly, for the responder's state, we have \mT{StR} to carry state data between \mT{R2} and \mT{R4}.
+%
+In order to link two rules to a state fact, we use \mT{tid}, which is unique to the current session.
+%
+The use of these state facts can be seen in line 28 in the \mT{R2\_STAT\_SIG} rule.
+
+Note that we do not model any error message that $R$ might send in response to message \mMsgone rejecting $I$'s choice of ciphersuite and/or method.
 %
 
-\subsubsection{Protocol Environment and Adversary Model}
-We model the binding between a party's identity and their long-term
-key pairs using the following rules.
+As in~\cite{Norr21}, we use \mTamarin's in-built equational theory for \mXor{}, and model the \mXor{} encryption of \mT{CIPHERTEXT\_2} with the key \mT{K\_2e} by allowing each part of the encrypted term to be ``separately attacked''.
 %
-\begin{small}
-\begin{verbatim}
-rule registerLTK_SIG:
-    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
-        [!LTK_SIG($A, ~ltk),
-         !PK_SIG($A, pk(~ltk)),
-         Out(<$A, pk(~ltk)>)]
-rule registerLTK_STAT:
-    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
-        [!LTK_STAT($A, ~ltk),
-         !PK_STAT($A, 'g'^~ltk),
-         Out(<$A, 'g'^~ltk>)]
-\end{verbatim}
-\end{small}
+This means that we first expand \mT{K\_2e} to the same number of key terms as subterms in the plaintext tuple.
 %
-The two rules register long-term
-key pairs for the \mSig{}- and \mStat{}-based methods respectively.
+This is done by applying \mHkdfExpand{} to unique inputs per subterm.
 %
-The fact \verb|Fr(~ltk)| creates a fresh term \mT{ltk}, representing a long-term
-secret key, not known to the adversary.
+After this, we \mXor{} each subterm with its own key term.
 %
-The fact \verb|Out(<$A, pk(~ltk)>)| sends the identity of the party
-owning the long-term key and the corresponding public key to the adversary.
+This is more faithful to the \mSpec{} than \mXor-ing \mT{K\_2e} on its own with the plaintext tuple.
 %
-The event \mT{UniqLTK} together with a corresponding restriction models the fact
-that the each party is associated with exactly one long-term key.
-%
-Consequently, an adversary cannot register additional long-term keys for an
-identity.
-%
-In line with the \mEdhoc{} \mSpec{}, this models an external mechanism
-ensuring that long term keys are bound to correct identities, e.g.,
-a certificate authority.
-%
+This can be seen in lines 19--22 in the code for \mT{R2\_STAT\_SIG}.
+\vnote{Add stuff here that changes with the TEE}
 
-We rely on \mTamarin's{} built-in message deduction rules for a Dolev-Yao adversary.
-%
-To model an adversary compromising long-term keys, i.e., events of type
-\mRevLTK{}, and revealing ephemeral keys, i.e., events of type
-\mRevEph{}, we use standard reveal rules.
-%
-The timing of reveals as modelled by these events is important.
-%
-The long-term keys can be revealed on registration, before protocol execution.
-%
-The ephemeral key of a party can be revealed when the party completes,
-i.e., at events of type \mIComplete{} and \mRComplete.~\footnote{A stronger, and perhaps more realistic, model would reveal ephemeral keys upon
-creation at the start of the run, but we failed to get \mTamarin{} to
-terminate on this.}
-%
-
-\subsubsection{Protocol Roles}
-We model each method of the protocol with four rules: \mT{I1}, \mT{R2}, \mT{I3}
-and \mT{R4} (with the method suffixed to the rule name).
-%
-Each of these represent one step of the protocol as run by the initiator $I$
-or the responder $R$.
-%
-The rules correspond to the event types \mIStart, \mRStart, \mIComplete,  and
-\mRComplete, respectively.
-%
-Facts prefixed with \mT{StI} carry state information between \mT{I1} and \mT{I3}.
-%
-A term unique to the current thread, \mT{tid}, links two rules to a given state fact.
-%
-Similarly, facts prefixed with \mT{StR} carry state information between the
-responder role's rules.
-%
-Line 28 in the \mT{R2\_STAT\_SIG} rule shown below illustrate one such use of state
-facts.
-%
-
-We do not model the error message that $R$ can send in response to message
-\mMsgone, and hence our model does not
-capture the possibility for $R$ to reject $I$'s offer.
-%
-
-We model the XOR encryption of \mT{CIPHERTEXT\_2} with the key \mT{K\_2e} using
-\mTamarin{}'s built in theory for XOR, and allow each term of the encrypted
-element to be attacked individually.
-%
-That is, we first expand \mT{K\_2e} to as many key-stream terms as there are
-terms in the plaintext tuple by applying the \mHkdfExpand{} function to unique
-inputs per term.
-%
-We then XOR each term in the plaintext with its own key-stream term.
-%
-This models the \mSpec{} closer than if we would have XORed \mT{K\_2e} as a
-single term onto the plaintext tuple.
-%
-The XOR encryption can be seen in lines 19--22 in the listing of
-\mT{R2\_STAT\_SIG} below.
-%
-\begin{small}
+{
+\bgroup
+\parindent 0pt
+\begin{minipage}{0.49\textwidth}
+\bgroup
+\parindent 0pt
+\begin{scriptsize}
 \begin{verbatim}
@@
 18   plainText2 = <$V, authV>
+\end{verbatim}
+\end{scriptsize}
+\egroup
+\end{minipage}
+\begin{minipage}{0.02\textwidth}
+\end{minipage}
+\begin{minipage}{0.48\textwidth}
+\bgroup
+\parindent 0pt
+\begin{scriptsize}
+\begin{verbatim}
 19   K_2e = expa(<$cAEAD0, TH_2,
@@
 \end{verbatim}
-\end{small}
+\end{scriptsize}
+\egroup
+\end{minipage}
+\\
+\egroup
+%
+}
+
+As mentioned earlier, we use actions to represent parametrized events.
+%
+For example, in line 27 above, the action \verb|ExpRunningR(~tid, $V, exp_sk, agreed)| represents an event of type \mRStart{} parametrized by the session id, the responder's identity, and the session key material \mT{exp\_sk}.
+%
+The \mT{exp} in the name of the variable for session key material represents the fact that the agreement property satisfied by this key is explicit, i.e., it includes $P_{I}$, as in Section~\ref{sec:agreedParams}.
+%
+We use \mT{imp\_sk} for the corresponding session key material which does not include $P_{I}$.
+%
+For the \mSigSig{} and \mSigStat{} methods, therefore, the two values are the same.
+
+%%-------------------------------------------------------------------------- sub
+%\subsection{Encoding the Desired Properties in \mTamarin}
+%\label{sec:propertyFormalization}
+%The properties and adversary model we defined in
+%Section~\ref{sec:desired-properties} translate directly into \mTamarin's logic,
+%using the straightforward mapping of events to the actions emitted from the model.
 %
 
-To implement events and
-to bind them to parameters, we use actions.
+The properties we listed in Section~\ref{sec:desired-properties} translate directly into \mTamarin's logic.
 %
-For example, the action \verb|ExpRunningR(~tid, $V, exp_sk, agreed)| in line 27
-above implements binding of an event of type \mRStart{} to the parameters and session key
-material.
+We show the \mTamarin{} lemma which encodes the \mPredPfs{} property. 
 %
+Other properties are formalized similarly.
 
-As explained in Section~\ref{sec:agreedParams}, it is not possible to show
-injective agreement on session key material when it includes
-$P_I$ (not visible in the rule \mT{R2\_STAT\_SIG}).
-%
-Therefore, we use certain actions to implement events that include $P_I$ in the
-session key material and other actions that do not.
-%
-Session key material which includes (resp. does not include) $P_I$ is referred
-to as \mT{imp\_sk} (resp. \mT{exp\_sk}) in the
-\mTamarin{} model.
-%
-In the case of \mSigSig{} and \mSigStat, therefore, \mT{imp\_sk} is the same as
-\mT{exp\_sk}.
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{\mTamarin{} Encoding of Properties}
-\label{sec:propertyFormalization}
-The properties and adversary model we defined in
-Section~\ref{sec:desired-properties} translate directly into \mTamarin's logic,
-using the straightforward mapping of events to the actions emitted from the model.
-%
-As an example, we show the lemma for verifying the property \mPredPfs.
-%
 \begin{small}
@@
 2 all-traces
-3  "All u v sk #t3 #t2.
+3  ``All u v sk #t3 #t2.
 4   (K(sk)@t3 & CompletedRun(u, v, sk)@t2) ==>
@@
 %
-As mentioned earlier, the action \mT{CompletedRun(u, v, sk)} in line 4 is
-emitted by both the rules \mT{I3} and \mT{R4}, and corresponds
-to the disjunction of events $\mIComplete^{t_2} \lor \mRComplete^{t_2}$ in the
-definition of \mPredPfs{} in Section~\ref{sec:secrecy}.
+
+In this formalization, we use the action \mT{CompletedRun(u, v, sk)} (in line 4) to represent the disjunction of the events $\mIComplete^{t_{2}}$ and $\mRComplete^{t_{2}}$.
 %
-Similarly, \mT{EphKeyRev(sk)} in line 7 models that the ephemeral
-key is revealed for either $I$ or $R$, or both.
+As expected, this action is emitted by both \mT{I3} and \mT{R4}.
 %
+Similarly, the action \mT{EphKeyRev(sk)} in line 7 stands for the reveal of the ephemeral key for $I$ or $R$ or both.

### 2022-07-18 paper.tex
Changes to section 3

@@
 "The Initiator MUST NOT change the supported cipher suites and
-the order of preference in SUITES_I based on previous error messages."
+the order of preference in SUITES\_I based on previous error messages."
 But: what happens if order of preference changes for some other reason? Would it

### 2022-07-18 paper.tex
[add] initial section on TEE

@@
 
+%-------------------------------------------------------------------------- sec
+\section{\uppercase{Trusted Execution Environments}}
+\label{sec:TEE}
+The main purpose of Trusted Execution Environments (TEE) is to prevent an
+adversary with physical access to a device from accessing the computational
+circuits.
+%
+They come in various forms of differing complexity.
+%
+Some, like ARM Trust Zone and Intel SGX are general purpose execution
+environments, which are flexibly programmable.
+%
+Other, like the Universal Subscriber Identity Modules (USIM) used for
+authentication to 3GPP mobile networks, have application specific interfaces for
+authentication and key agreement protocols etc.
+%
+A key aspect is how much of the application is placed in the TEE and how much is
+outside.
+%
+
+For larger devices that include general purpose processors with Trust Zone or
+SGX, entire \mEdhoc{} and \mOscore{} may reside inside the TEE.
+%
+For constrained IoT devices in the lower end of the scale, a TEE may have to be
+implemented using a special-purpose integrated circuit.
+%
+In the latter case, it may be beneficial to follow a minimalistic approach, and
+contain the long-term key and only
+the operations that need access to it in the TEE for cost reasons.
+%
+
+It may at first appear more secure to implement as much as possible inside the
+TEE, but there is a security trade-off: the more code inside the TEE,
+the higher the risk of implementation errors in the security-critical code.
+%
+Because the security critical code runs in the area where the long-term keys
+reside, an implementation error here risks leaking information of the key to the
+adversary.
+%
+From this perspective, it may be beneficial to follow the minimalistic approach
+even when having access to Trust Zone or SGX.
+%
+
+A slightly more secure division of functionality is to also keep the
+session key inside
+the TEE and extend the interface to accept messages and return the
+(en/de)crypted counter part.
+%
+That is, the interface exposes \mAead{} functions in the interface.
+%
+
 %-------------------------------------------------------------------------- sub
+\subsection{Practical Aspects and Use Case}
+\label{sec:TEE:useCase}
+Authentication of a specific IoT device assumes that this devices is the only
+entity with access to the long-term key associated with the corresponding
+identity.
+%
+Since IoT devices may be accessible to adversaries, e.g., an insider cloning a
+key card, the long-term keys must be appropriately protected.
+%
+A state of the art approach is to use a Trusted Execution Environment (TEE),
+which holds the key and provides an API for operations using the key.
+%
+This is the approach taken by the
+Trust-Zone based \mMuEdhoc{}~\cite{DBLP:conf/codaspy/Hristozov0XFLS21}
+implementation for example.
+%
+Typical operations include signatures using the long-term private key of a
+party.
+%
+
+%-------------------------------------------------------------------------- sub
+\subsection{System and Adversary Model Extensions Supporting TEE}
+\label{sec:TEE:advModel}
+%
+We extend the adversary model of~\cite{Norr21} by allowing the adversary
+access to a TEE interface on a party of their choosing before and after the
+\mEdhoc{} session.
+%
+This corresponds to a so-called weak post-compromise security
+(PCS)~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
+%
+
+We split the \mEdhoc{} functionality as follows.
+%
+The TEE contains the long-term key and allows the non-TEE parts of the
+application to perform the operations on it via its interface.
+%
+More precisely, parties using the \mSig{} authentication method use a TEE with
+an interface which accepts a message and returns the signature of that message
+using the party's private long-term key.
+%
+Parties using the \mStat{} authentication method use a TEE with an interface
+accepting $g^x$ and returning $g^x$ raised to the party's private long-term key.
+\knote{It should accept any value. Adversary may not stick to elements of the
+    form $g^x$. Or we explicitly state that the TEE checks that the received
+input lies on the curve (but this may not be sufficient: are the curves cyclic
+groups? Need to check).}
+%
+The non-TEE parts implement \mEdhoc{} using this interface.
+%
+This interface requires the least functionality from the TEE, reducing the TEE's
+complexity and possibly cost of its implementation.
+%
+The functional split is suitable even for the situation where a
+constrained device has implemented only the storage of the long-term key in a
+special purpose circuit with minimal processing functionality.
+%
+Since \mEdhoc{} focuses on constrained IoT devices, it seems appropriate to
+cater for this setting.
+%
+
+%-------------------------------------------------------------------------- sub
+\subsection{Extended Formalism and Security Properties}
+\label{sec:TEE:fmAndProps}
+Formally, we model the TEE interface by adding two new rewrite rules:
+%
+\begin{small}
+\begin{verbatim}
+rule forge_SIG:
+   [!LTK_SIG($A, ~ltk), In(xx)] --[TEE($A)]-> [Out(sign(xx, ~ltk))]!>
+
+rule exp_STAT:
+   [!LTK_STAT($A, ~ltk), In('g'^~xx)] --[TEE($A)]-> [Out(('g'^~xx)^~ltk)]
+\end{verbatim}
+\end{small}
+%
+These rules allow the adversary to obtain terms representing signatures on a
+value of their choice to forge signatures (\verb|forge_SIG|), or to obtain terms
+representing a curve point of their choice raised to the power of the
+long-term key (\verb|exp_STAT|).
+%
+
+Because it is a trivial attack when the adversary access these rules with values
+from the test session, we must disqualify those rule applications.
+%
+We do so by creating an action fact \verb|TEE($A)|, where \verb|$A$| is the
+identity corresponding to the private key used, and then augmenting the
+properties with a condition that no such action fact exists from the start of
+the protocol execution and its end.
+%
+Care must be taken when specifying the start and the end.
+%
+Specifically, ...
+\knote{Need to check this. Should we not allow access to both parties keys
+    before and after session? How could we otherwise claim KCI?
+}
+\knote{Need to check whether auth properties are correct. $\forall$ constructs
+    should be like the $\exists$ constructs of PFS. Not only for symmetry, but
+    they look incorrect now (should not be $\lor \forall \ldots$).
+}
+
+%-------------------------------------------------------------------------- sec
 \section{\uppercase{Discussion}}

### 2022-07-18 paper.tex
[add] text describing the simplified XOR encryption

@@
 %
-We do so by creating an action fact \verb|TEE($A)|, where \verb|$A$| is the
+We do so by creating an action fact \verb|TEE($A)|, where \verb|$A| is the
 identity corresponding to the private key used, and then augmenting the
@@
 }
+
+\paragraph{XOR expressiveness}
+We did not manage to get Tamarin to terminate for all lemmas and all methods
+with the added TEE extension.
+%
+When the responder uses the \mStat{} authentication method Tamarin did not
+complete the in a reasonable time for the authentication lemmas.
+%
+To combat this, we simplified the modeling of XOR encryption of message\_2 for
+these lemmas.
+%
+More precisely, the second message contains the two values \verb|V| and
+\verb|authV|, and we generally model the XOR encryption by xoring each of them
+with their own key-stream terms.
+%
+For the problematic lemmas however, we resorted to xoring the entire tuple
+\verb|<V, auth| with single key-stream term.
+%
+Given that no attacks were identified using our original modeling, this
+simplification is intuitively not a severe restriction.
+%
+However, it is possible that we now miss attacks on these methods 
+combining properties of the two-term XOR encryption with TEE capabilities.
+%
 
@@
 \end{table*}
+\knote{This should be extended with verification times and with TEE.
+    It should also show that when responder uses \mStat{} and \mSig{}, a
+    different XOR encryption is used.
+}
 

### 2022-07-18 paper.tex
[fix] spelling and clarifications

@@
 %
-To combat this, we simplified the modeling of XOR encryption of message\_2 for
-these lemmas.
+To combat this, we simplified the modeling of XOR encryption of the second
+message for these lemmas when the responder uses \mStat{} authentication.
 %
@@
 For the problematic lemmas however, we resorted to xoring the entire tuple
-\verb|<V, auth| with single key-stream term.
+\verb|<V, auth>| with a single key-stream term.
 %

### 2022-07-19 paper.tex
[fix] remove m4-macros from code snippets
[add] relations to related work on TEE

@@
 %
-This corresponds to a so-called weak post-compromise security
+This allows us to distinguish between a full compromise and a weak one, using
+the terminology of~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
+%
+A weak compromise is when the adversary have temporary access to the TEE.
+%
+Seen in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add adversary
+capabilities corresponding to a server adversary, i.e., an adversary that
+compromises a party, learns its ephemeral keys and may temporarily access its
+TEE, but does not learn the party's long-term key.
+%
+The model of~\cite{Norr21} only captures server adversary capabilities without
+access to the TEE, via the leakage of ephemeral keys, and local adversary
+capabilities, or more specifically, a secure-world adversary with read-write
+access to the
+secure world where the long-term keys reside via the leakage of long-term keys.
+%
+Our extension allows analysis of so-called weak post-compromise security
 (PCS)~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
@@
 rule forge_SIG:
-   [!LTK_SIG($A, ~ltk), In(xx)] --[TEE($A)]-> [Out(sign(xx, ~ltk))]!>
+   [!LTK_SIG($A, ~ltk), In(xx)] --[TEE($A)]-> [Out(sign(xx, ~ltk))]
 

### 2022-07-19 paper.tex
[fix] clarifying start/end of run

@@
 %
-Parties using the \mStat{} authentication method use a TEE with an interface
-accepting $g^x$ and returning $g^x$ raised to the party's private long-term key.
+A party $U$ using the \mStat{} authentication method use a TEE with an interface
+accepting a point $p$ on the curve and returning $\mDH(p,\ \mGu{})$.
+%
 \knote{It should accept any value. Adversary may not stick to elements of the
@@
 %
-Specifically, ...
-\knote{Need to check this. Should we not allow access to both parties keys
-    before and after session? How could we otherwise claim KCI?
-}
-\knote{Need to check whether auth properties are correct. $\forall$ constructs
-    should be like the $\exists$ constructs of PFS. Not only for symmetry, but
-    they look incorrect now (should not be $\lor \forall \ldots$).
-}
+Specifically, the start and end of the protocol run must be viewed w.r.t. the
+role.
+%
+For example, the injective agreement property for the initiator in
+the \mSigSig{} method requires that the adversary does not have access to the
+TEE of the
+responder from the time of transmission of the first message until the second
+message is received by the initiator.
+%
+Because reception of one message and transmission of the next one at a party
+is an atomic operation, these timepoints represent the start and end of the
+protocol run from the perspective of the initiator.
+%
 
-\paragraph{XOR expressiveness}
 We did not manage to get Tamarin to terminate for all lemmas and all methods
@@
 %
-More precisely, the second message contains the two values \verb|V| and
-\verb|authV|, and we generally model the XOR encryption by xoring each of them
-with their own key-stream terms.
+More precisely, the second message contains the two values \verb|R| (the
+responder's identity) and
+\verb|authR| (the authentication information), and we generally model the
+XOR encryption by xoring each of them with their own key-stream terms.
 %
 For the problematic lemmas however, we resorted to xoring the entire tuple
-\verb|<V, auth>| with a single key-stream term.
+\verb|<R, authR>| with a single key-stream term.
 %
@@
 from the verified properties.
+\knote{In Section 3.3 we say that injective agreement does \emph{not} hold for
+    all methods. It is a little contradictory to the table below where we
+    exclude $P_I$. It would be better to show it as not holding when $P_i$ is
+    included (both for consistency and for clarity).
+}
 %

### 2022-07-19 intro.tex
Modified root command in files
[fix] review and update of intro

@@
-% !TEX root = journal-version.tex
-Traditionally, formal analysis for security protocols has been centered on devices with little to no resource constraints, where protocols like \mDandTls{} can be easily employed. 
+% !TEX root = paper.tex
+IoT protocols are often run on devices which operate under severe restrictions on resources like bandwidth and energy consumption.
 %
-However, as IoT gains popularity and is used for many applications, including potentially sensitive ones, it becomes paramount to analyze these protocols to ensure that they do not admit security flaws.
+These constrained devices are often simple in their operation, but need to communicate and function without human interference or maintenance for extended periods of time.
 %
-IoT protocols are often run on devices which operate under severe restrictions on resources like bandwidth and energy consumption.
-These devices are often simple in their operation, but need to communicate and function without human interference or maintenance for extended periods of time.
+IETF is standardizing new protocols to secure communications between constrained devices.
 %
-Other protocols have been proposed to secure communications between constrained devices.
-One such, which has been standardized by the IETF is the Object Security for Constrained RESTful Environments (\mOscore{}) protocol.
+One such is the Object Security for Constrained RESTful Environments (\mOscore{}) protocol. 
 %
 However, \mOscore{} requires the pre-establishment of a security context.
-To this end, a key exchange protocol named Ephemeral Diffie-Hellman Over COSE (\mEdhoc{}) has been under discussion at the IETF.
-
+%
+To this end, a key exchange protocol named Ephemeral Diffie-Hellman Over COSE
+(\mEdhoc{}) is under discussion in the IETF. 
 %
@@
 %
-It is therefore, very important to ensure that most of the fundamental properties expected of key exchange protocols as established in the literature are satisfied by \mEdhoc{} as well.
+It is therefore important to ensure that most of the fundamental properties expected of key exchange protocols as established in the literature are satisfied by \mEdhoc{} as well.
+%
+
 %-------------------------------------------------------------------------- sub
@@
 %
-A \emph{cryptographic core}, often just called a core, is an academic protocol, i.e. with no encodings or application-specific details as needed for an industrial protocol. 
+A \emph{cryptographic core}, often just called a core, is an academic protocol,
+i.e., with no encodings or application-specific details as needed for an industrial protocol. 
 %
@@
 %
-In May 2018, the designers replaced the challenge-response signature core with one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}, and in 2020, three new cores which mixed challenge-response signatures and ``regular'' authentication signatures were added as well~\cite{our-analysis-selander-lake-edhoc-00}.
+In May 2018, the designers replaced the challenge-response signature core with
+one based on \mSigma{}
+(SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}, and in
+2020, three new cores, which mixed challenge-response signatures and regular signatures were added as well~\cite{our-analysis-selander-lake-edhoc-00}.
 
@@
 \label{sub:related}
-The earliest related work is by Bruni et al~\cite{DBLP:conf/secsr/BruniJPS18}, which formally analyzes the May 2018 version of \mEdhoc{} using the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01}. 
+The earliest related work is~\cite{DBLP:conf/secsr/BruniJPS18}, which formally analyzes the May 2018 version of \mEdhoc{} using the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01}. 
 %
-In this paper, the authors analyze two key establishment methods -- one built on a pre-shared key authentication core, and one on \mSigma{}. 
+In this paper, the authors analyze two key establishment methods -- one built on
+a pre-shared key authentication core, and one based on \mSigma{}. 
 %
@@
 %
-In 2021, Norrman et al~\cite{Norr21} analyze the 2020 version of \mEdhoc{} in the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13}. 
+Later work~\cite{Norr21} analyze the May 2020 version of \mEdhoc{} in the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13}. 
 %
@@
 We analyze the version of \mEdhoc{} as in the latest draft~\cite{}.
+\knote{possibly}
 %
-We modify the \mTamarin{} models used in that paper to obtain a significant speed-up in the verification time.
-%
-This required us to carefully modify the oracle as well.
+We extend the adversary model and \mTamarin{} models to capture weak
+post-compromise security (PCS) and modeling of Trusted Execution Environments
+(TEE).
 %

### 2022-07-19 oldSections/edhocProtocol.tex
[fix] moved old sources to separate dir to avoid confusion

--- 2022-07-12 Journal Version/protocol.tex
@@
 % !TEX root = journal-version.tex
-In this section, we describe the various key establishment methods of the \mEdhoc{} protocol.
+
+We now present the structure of the protocol using notation for key material,
+elliptic curve operations and
+identities mostly adopted from NIST SP 800-56A Rev. 3~\cite{sp800-56a-rev3}.
 %
-Following~\cite{Norr21}, we refer to the two roles executing the protocol as the initiator $I$ and the responder $R$.
+One notable difference is that we refer to the two roles executing the protocol
+as the initiator $I$ and the responder $R$.
 %
-We annotate values with $I$ and $R$ to make explicit which role they belong to.
+We do this to avoid confusing roles with the parties taking them on.
+%
+Values in the analysis are subscripted with $I$ and $R$ when necessary to
+distinguish which role is associated with the values.
 %
 
-We denote by $\langle d_{t, \mathit{id}}, Q_{t, \mathit{id}}\rangle$ public-private key pairs, where $d$ is the private key, $Q$ is the public key, $t \in \{e, s\}$ indicates whether the key is ephemeral or static, and $\mathit{id}$ indicates the party to whom the key pair belongs.
+%------------------------------------------------------------------------- sub
+\subsection{Preliminaries}
+\label{sec:preliminaries}
+Private/public key pairs are written
+$\langle d_{t,\mathit{id}},\,Q_{t,\mathit{id}}\rangle$,
+where $d$ is the private key, $Q$ the public key, $t \in \{e, s\}$ denotes
+whether the key is ephemeral or static, and $\mathit{id}$ is the role or
+party controlling the key pair.
 %
-When clear from context, we will often drop part (or all) of the subscript.
+When irrelevant, we drop the subscript or parts thereof.
 %
-Static key pairs (suitable for both regular and challenge-response signatures) are long-term authentication credentials, while ephemeral key pairs are those which are generated afresh for each instantiation of the protocol.
+Ephemeral key pairs are generated fresh for each instantiation of the protocol
+and static key pairs are long-term keys used for authentication.
 %
-
-Parties can authenticate using regular signatures or challenge-response signatures. 
+Static key pairs are suitable for either regular signatures or
+challenge-response signatures.
 %
-In the former case, we say that they use the \emph{signature based authentication method} (\mSig{}).
+When a party uses regular signatures for authentication, we say that they use
+the \emph{signature based authentication method}, or \mSig{} for short.
 %
-In the latter, we say, following the terminology in the \mSpec{}, that they use the \emph{static key authentication method} (\mStat{}).
+When a party uses challenge-response signatures for authentication, we say that
+they use the \emph{static key authentication method}, or \mStat{} for short.
 %
-For more details about the challenge-response terminology, see~\cite{DBLP:conf/crypto/Krawczyk05}.
-
-%%------------------------------------------------------------------------- sub
+The latter naming may appear confusing since signature keys are equally static,
+but is chosen to make the connection to the \mSpec{} clear.
+%
+We adopt the challenge-response terminology for this style of authentication
+from~\cite{DBLP:conf/crypto/Krawczyk05}.
 %

### 2022-07-19 formalization.tex
[add] update of protocol description

@@
 %
-Long-term key reveal, denoted $\mRevLTK^{t}(A)$, stands for the adversary gaining access to a principal $A$'s long-term private key \mPriv{A} at a time point $t$.
+Long-term key reveal, denoted $\mRevLTK^{t}(A)$, stands for the adversary
+gaining access to a party $A$'s long-term private key \mPriv{A} at a time point $t$.
 %
@@
 %
-However, ephemeral keys in our model can only be revealed when a principal completes their role, i.e. simultaneous with events of type \mIComplete{} and \mRComplete.
+However, ephemeral keys in our model can only be revealed when a party completes their role, i.e. simultaneous with events of type \mIComplete{} and \mRComplete.
 %

### 2022-07-19 protocol.tex
[fix] clarifying start/end of run
[add] update of protocol description

@@
 % !TEX root = paper.tex
-In this section, we describe the various key establishment methods of the \mEdhoc{} protocol.
+In this section, we describe the various key establishment methods of the
+\mEdhoc{} protocol.
 %
-Following~\cite{Norr21}, we refer to the two roles executing the protocol as the initiator $I$ and the responder $R$.
+Following~\cite{Norr21}, we refer to the two roles executing the protocol as
+the initiator $I$ and the responder $R$.
 %
@@
 \label{sec:notation}
-We denote by $\langle d_{t, \mathit{id}}, Q_{t, \mathit{id}}\rangle$ public-private key pairs, where $d$ is the private key, $Q$ is the public key, $t \in \{e, s\}$ indicates whether the key is ephemeral or static, and $\mathit{id}$ indicates the party to whom the key pair belongs.
+We denote by $\langle d_{t, \mathit{id}}, Q_{t, \mathit{id}}\rangle$
+public-private key pairs, where $d$ is the private key, $Q$ is the public key,
+$t \in \{e, s\}$ indicates whether the key is ephemeral or static,
+and $\mathit{id}$ indicates the party to whom the key pair belongs.
 %
-When clear from context, we will often drop part (or all) of the subscript.
+When clear from context, we will often drop part (or all) of the subscripts.
 %
-Static key pairs (suitable for both regular and challenge-response signatures) are long-term authentication credentials, while ephemeral key pairs are those which are generated afresh for each instantiation of the protocol.
+Static key pairs (suitable for regular or challenge-response signatures)
+are long-term authentication credentials, whereas ephemeral key pairs are those
+which are generated afresh for each execution of the protocol.
 %
 
-Parties can authenticate using regular signatures or challenge-response signatures. 
+Parties can authenticate using regular signatures or challenge-response
+signatures.
 %
-In the former case, we say that they use the \emph{signature based authentication method} (\mSig{}).
+In the former case, we say that they use the
+\emph{signature based authentication method} (\mSig{}).
 %
-In the latter, we say, following the terminology in the \mSpec{}, that they use the \emph{static key authentication method} (\mStat{}).
+In the latter, we say, following the terminology in the \mSpec{},
+that they use the \emph{static key authentication method} (\mStat{}).
 %
-For more details about the challenge-response terminology, see~\cite{DBLP:conf/crypto/Krawczyk05}.
+We adopt the challenge-response terminology
+from~\cite{DBLP:conf/crypto/Krawczyk05}.
 %
 
-\mEdhoc{} fundamentally utilizes elliptic curves and associated Diffie-Hellman operations. 
+\mEdhoc{} fundamentally use elliptic curves and associated Diffie-Hellman
+operations.
 %
-Signatures using a principal $A$'s keys are denoted by \mSign{A}$(\cdot)$, while the verification thereof is denoted by \mVf{A}$(\cdot)$.
+Signatures using a party $A$'s keys are denoted by \mSign{A}$(\cdot)$, while
+the verification thereof is denoted by \mVf{A}$(\cdot)$.
 %
-A Diffie-Hellman operation which combines a private key $d$ and a point $P$ on the elliptic curve is represented as $\mDH(d, P)$.
+A Diffie-Hellman operation which combines a private key $d$ and a point $P$
+on the elliptic curve is represented as $\mDH(d, P)$.
 %
-We will often overload notation to let $P$ stand for both the point on the elliptic curve as well as the corresponding bitstring.
-
+We will often overload notation to let $P$ stand for both the point on the
+elliptic curve as well as the corresponding bitstring encoding.
+%
 %\mEdhoc{} relies on \mCose{}~\cite{rfc8152} for elliptic curve operations and
@@
 %------------------------------------------------------------------------- sub
-\subsection{Overall description}
+\subsection{Overall Description}
 \label{sec:description}
@@
 %
-This context includes the session key material (we denote this by \mSessKey{}), and might also transfer application data \mADone{}, \mADtwo{}, and \mADthree{}.
+This context includes the session key material (we denote this by \mSessKey{}).
 %
-The abstract protocol as given in the \mSpec{}~\cite{} consists of three messages, and is shown in Figure~\ref{fig:edhocFramework}~\cite{Norr21}.
+The abstract protocol as given in the \mSpec{}~\cite{} consists of three
+messages, and is shown in Figure~\ref{fig:edhocFramework}~\cite{Norr21}.
 %
-The generalized abstract protocol is the same across methods, but the authentication mechanisms and key derivation procedures differ. 
+The generalized abstract protocol is the same across methods,
+and may transfer application data \mADone{}, \mADtwo{}, and \mADthree{} in
+addition to establishing the context.
+%
+The authentication mechanisms and key derivation procedures differ between
+methods.
 %
 We will describe the various aspects of the protocol now.
-
+%
 \begin{figure}[ht]
@@
 \end{figure}
-
-Of the three messages, the first two, among other things, serve to establish a common authentication method \mMethod{}, and a ciphersuite \mSuites{}.
 %
-In \mMethod{}, the party playing the Initiator role proposes an ordered list of choices for the authentication methods that each party can use. 
+Of the three messages, the first two, among other things, serve to establish a
+common authentication method \mMethod{}, and a ciphersuite \mSuites{}.
 %
-As mentioned earlier, these may differ for the two roles, yielding four possible combinations: \mSigSig{}, \mSigStat{}, \mStatSig{}, and \mStatStat{}, where the first is the method used by the Initiator role, and the second by the Responder role.\footnote{As in the \mSpec{}, we will overload notation and refer to the combinations of authentication methods as methods as well.}
+In \mMethod{}, the party playing the initiator role proposes which
+authentication methods the two parties shall use, and an ordered list
+of choices for the ciphersuite.
 %
-The party executing the Responder role may choose to reject the method or ciphersuite chosen by the Initiator by sending an error message.
+As mentioned earlier, the authentication methods may differ for the two roles,
+yielding four possible
+combinations: \mSigSig{}, \mSigStat{}, \mStatSig{}, and \mStatStat{},
+where the first authentication method in each combination is used by the
+initiator role, and the second by the responder role.
+\footnote{As in the \mSpec{}, we will from now on overload
+notation and refer to the combinations of authentication methods as methods as
+well.}
 %
-This results in abandoning this session and renegotiating, as the Initiator goes down their list of choices for authentication methods, and picks the next option.
+The party executing the responder role may choose to reject the method or
+ciphersuite chosen by the initiator by sending an error message.
 %
-While our analysis (and indeed, \mTamarin{}) cannot naturally handle such renegotiation which maintains state and ``remembers'' the rejected method and/or ciphersuite, the formal modeling excludes this aspect of the protocol.
+This results in abandoning this session and renegotiating, as the initiator
+goes down their list of choices for ciphersuites, and picks the next option for
+a next execution of the protocol.
 %
-However, we will discuss the ramifications of such a renegotiation procedure and the error messages later, in Section~\ref{sec:}.
+Our analysis does not cover such renegotiation which requires maintaining state
+between executions to remember the rejected ciphersuites.
+%
+However, we will discuss the ramifications of such a renegotiation procedure
+and the error messages later, in Section~\ref{sec:}.
 %
 
-In addition to negotiating the method and the ciphersuite, the first two messages are also instrumental for the exchange of public ephemeral keys \mGx{} and \mGy{}, and connection identifiers \mCi{} and \mCr{}, for the initiator and responder roles respectively.
+In addition to negotiating the ciphersuite, the first two messages are
+also instrumental for the exchange of public ephemeral keys \mGx{} and \mGy{},
+and connection identifiers \mCi{} and \mCr{}, for the initiator and responder
+roles respectively.
 %
-The \mSpec{} states that the connection identifiers serve only to route messages to the correct principal executing \mEdhoc{}, but then also claims that they may be used in turn by any protocols (like \mOscore{}) using the security context established by \mEdhoc{}.
+The \mSpec{} states that the connection identifiers serve only to route messages
+to the correct party executing \mEdhoc{}, but also claims that they may be
+used in turn by protocols (like \mOscore{}) using the security context
+established by \mEdhoc{}.
 %
-While the \mSpec{} does not require any explicit security guarantees to be satisfied by these communication identifiers, it does, however, require that the identifiers be unique, i.e.
-in any session, $\mCi{} \neq \mCr{}$, and that the parties involved in the session can verify this uniqueness.
+While the \mSpec{} does not require any explicit security guarantees to be
+satisfied by these connection identifiers, it does, however, require that the
+identifiers be unique, i.e., in any session, $\mCi{} \neq \mCr{}$, and that the
+parties involved in the session can verify this uniqueness.
 %
-The \mSpec{} states that \mOscore{} should be able to use these identifiers to retrieve any particular security context.
+More precisely, the \mSpec{} states that \mOscore{} should be able to use these
+identifiers to retrieve any particular security context.
 %
-In this work, as in~\cite{Norr21}, we verify that the parties agree on the values of \mCi{} and \mCr{}.
-
-Messages 2 and 3 also serve to identify and authenticate each party to the other.
+In this work, as in~\cite{Norr21}, we verify that the parties agree on the
+values of \mCi{} and \mCr{}.
 %
-These messages contain long-term key identifiers (\mIdcredi{} and \mIdcredr{}), as well as authenticating information (\mAuthi{} and \mAuthr{}) which lets each party know that the other party does indeed control the long-term key provided by them.
+The second and third messages also serve to identify and authenticate each
+party to the other.
 %
-The authentication information changes based on the authentication method associated with the long-term key chosen by the party in question.
+These messages contain long-term key identifiers (\mIdcredi{} and \mIdcredr{}).
+%
+Additionally, the messages contain authenticating information
+(\mAuthi{} and \mAuthr{}), which lets each
+party know that the other party does indeed control the long-term key
+associated with these identifiers.
+%
+The authentication information is different for each authentication method.
 %
@@
 %
-The initiator $I$ chooses the method $\mSigStat{}$ and sends this via $\mMethod{}$ to the responder $R$.
+The initiator $I$ chooses the method $\mSigStat{}$ and sends this via
+$\mMethod{}$ to the responder $R$.
 %
-$R$ now has an option to either reject this choice of method, or to provide an identifier \mIdcredr{} for a key pair which can be used with challenge-response signatures as well as authenticating information \mAuthr{}, as dictated by the static key authentication method \mStat{}.
+$R$ now has an option to either reject this choice of method, or to provide an
+identifier \mIdcredr{} for a key pair which can be used with challenge-response
+signatures as well as authenticating information \mAuthr{}, as dictated by
+the static key authentication method \mStat{}.
 %
-$I$ will now respond with an identifier \mIdcredi{} corresponding to a key pair which is suitable for regular signatures, and provide authentication \mAuthi{} as appropriate for the signature-based method \mSig{}.
+$I$ will then respond with an identifier \mIdcredi{} corresponding to a key
+pair, which is suitable for regular signatures, and provide authentication
+\mAuthi{} as appropriate for the signature-based method \mSig{}.
 
+%-------------------------------------------------------------------------- sub
 \subsection{Key Schedule}
 \label{sec:keysched}
-As mentioned above, messages 2 and 3 of \mEdhoc{} contain authenticating information.
+As mentioned above, the second and third messages of \mEdhoc{} contain
+authenticating information.
 %
-This contains encrypted and potentially signed MAC values, as we will describe in more detail in Section~\ref{sec:edhocauth}.
+This information is either a signature or a message authentication code (MAC),
+as we will describe in more detail in Section~\ref{sec:edhocauth}.
 %
-The keys for these are generated by using a key schedule which takes at least one DH key \mGxy{} as basic input, as shown in Figure~\ref{fig:kdfdiagram}, and is intrinsic to the functioning of \mEdhoc{}.
+The keys for these are generated by using a key schedule which takes at least
+one DH key \mGxy{} as basic input, as shown in Figure~\ref{fig:kdfdiagram},
+and is intrinsic to the functioning of \mEdhoc{}.
 %
@@
 }
-\caption{Key schedule: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo, \mPRKthree, \mPRKfour are the intermediate key material, and \mKtwoe,\mKtwom, \mKthreeae, \mKthreem are the encryption keys for \mAead{} or \mXor{}.
+\caption{Key schedule: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo, \mPRKthree,
+\mPRKfour{} is the intermediate key material, and
+\mKtwoe,\mKtwom, \mKthreeae, \mKthreem{} are the encryption keys for \mAead{}
+or \mXor{}.
 Dashed boxes are conditionals.~\cite{Norr21}}
@@
 
-To derive keys, \mEdhoc{} uses two functions from the \mHkdf{} interface~\cite{rfc5869}, \mHkdfExtract{} and \mHkdfExpand{}.
+To derive keys, \mEdhoc{} uses two functions from the \mHkdf{}
+interface~\cite{rfc5869}, \mHkdfExtract{} and \mHkdfExpand{}.
 %
@@
 %
-For \mHkdfExtract{}, the input is a DH key, while for \mHkdfExpand{}, it is intermediate key material.
+For \mHkdfExtract{}, the input is a DH key, while for \mHkdfExpand{},
+it is intermediate key material.
 %
 
-As mentioned earlier, the fundamental building block for the key schedule is the ephemeral DH key \mGxy{}, which  is computed in two different ways by $I$ (as $\mDH(\mX, \mGy)$) and $R$ (as $\mDH(\mY, \mGx)$).
+As mentioned earlier, the fundamental building block for the key schedule is
+the ephemeral DH key \mGxy{}, which is computed in two different ways by
+$I$ (as $\mDH(\mX, \mGy)$) and $R$ (as $\mDH(\mY, \mGx)$).
 %
@@
 %
-Each intermediate key gives rise to encryption and integrity keys (\mKtwoe, \mKtwom{}, \mKthreeae, and \mKthreem) for a particular message in the protocol.
+Each intermediate key gives rise to encryption and integrity keys
+(\mKtwoe, \mKtwom{}, \mKthreeae, and \mKthreem)
+for a particular message in the protocol.
 %
 
-In order to generate the final keys, the two \mHkdf{} algorithms use various values for salt.
+In order to generate the final keys, the two \mHkdf{} algorithms use various
+values for salt.
 %
-\mPRKtwo{} is generated by the \mHkdfExtract{} algorithm while using the empty string as the salt.
+\mPRKtwo{} is generated by the \mHkdfExtract{} algorithm while using the empty
+string as the salt.
 %
-\mPRKthree{} and \mPRKfour{} are separately generated if $R$ or $I$ uses the \mStat{} method, using the corresponding DH key as input and the previous intermediate key as salt.
+\mPRKthree{} and \mPRKfour{} are separately generated if $R$ or $I$ uses the
+\mStat{} method, using the corresponding DH key as input and the previous
+intermediate key as salt.
 %
@@
 %
-On the other hand, the key $P_{I}$ is used if the initiator uses \mStat{}, and is computed as $\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$.
+Similarly, the key $P_{I}$ is used if the initiator uses \mStat{}, and is computed as $\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$.
+%
 
-These intermediate keys are fed into an \mHkdfExpand{} algorithm, as shown in Figure~\ref{fig:kdfdiagram}. 
+These intermediate keys are fed into \mHkdfExpand{}, as shown in Figure~\ref{fig:kdfdiagram}.
 %
-This algorithm uses as the salt a value $\textit{th}$, which is a running hash of the information transmitted thus far as part of the protocol.
+\mHkdfExpand{} here uses as salt a value $\textit{th}$, which is a running hash of the information transmitted thus far as part of the protocol.
 %
 $\textit{th}_{i}$ denotes the hash corresponding to the $i^{\rm{th}}$ message.
+%
 
-At the end of a successful run of the protocol, the session key material for \mOscore{} is established as \mSessKey{}, which is a set of various keys.
+At the end of a successful run of the protocol, the session key material is
+established as \mSessKey{}, which we define as a set of various keys.
 %
@@
 %
-Once \mSessKey{} has been established, an \mHkdf{}-based key exporter named \mEdhoc-Exporter extracts the keys required for \mOscore{}.
+We discuss which material should be included in \mSessKey{} in more detail in
+the formal model in Section~\ref{sec:formalization} and the consequences of
+various choices in Section~\ref{sec:sessionKeyMaterial}.
+%
+Once \mSessKey{} has been established, an \mHkdf{}-based key exporter named
+\mEdhoc-Exporter extracts the keys required by the security protocol.
 %
 
-\subsection{About authentication in \mEdhoc{}}
+%-------------------------------------------------------------------------- sub
+\subsection{About Authentication in \mEdhoc{}}
 \label{sec:edhocauth}
@@
 %
-For both \mStat{} and \mSig{}, the following information is used to compute \mAuthr{}: \mIdcredr{}, \mCredr{}, a transcript hash of all the communicated information thus far in the protocol, and \mADtwo{}, if included.
+For both \mStat{} and \mSig{}, the following information is used to compute
+\mAuthr{}: \mIdcredr{}, \mCredr{}, a transcript hash of all the communicated
+information thus far in the protocol, and \mADtwo{}, if included.
 %
-\mAuthi{} uses the same pieces of information corresponding to the initiator role.
+\mAuthi{} uses the same pieces of information, but corresponding to the
+initiator role.
 %
-A MAC is obtained by feeding this material as additional data and the empty string as input to the Authenticated Encryption with Additional Data (AEAD) encryption algorithm as indicated in the established ciphersuite \mSuites{}.
+A MAC is obtained by feeding this material as additional data and the empty
+string as input to the Authenticated Encryption with Additional Data (AEAD)
+encryption algorithm as indicated in the established ciphersuite \mSuites{}.
 %
-The encryption key for the AEAD algorithm is constructed, for both roles, using the ephemeral key material \mGx{}, \mGy{}, \mX{}, and \mY{}. 
+The encryption key for the AEAD algorithm is constructed, for both roles,
+using the ephemeral key material \mGx{}, \mGy{}, \mX{}, and \mY{}.
 %
-The initiator computes $\mDH(\mX, \mGy)$ while the responder computes $\mDH(\mY, \mGx)$, and DH operations give rise to the same key under both these computations.
+The initiator computes $\mDH(\mX, \mGy)$ while the responder computes
+$\mDH(\mY, \mGx)$, and DH operations give rise to the same key under both
+these computations.
+%
 
-When $I$ uses the \mSig{} authentication method, \mAuthi{} is $I$'s signature over the MAC along with the data covered by the MAC.
+When $I$ uses the \mSig{} authentication method, \mAuthi{} is $I$'s signature
+over the MAC along with the data covered by the MAC.
 %
-However, when $I$ uses the \mStat{} method, \mAuthi{} is just the MAC, with one difference: the MAC key is derived using both the ephemeral key material as well as the long-term key for the initiator $\langle\mPriv{I},\ \mPub{I}\rangle$.
+However, when $I$ uses the \mStat{} method, \mAuthi{} is just the MAC, with
+one difference: the MAC key is derived using both the ephemeral key material
+\mGxy{} as well as the long-term key for the initiator
+$\langle\mPriv{I},\ \mPub{I}\rangle$.
 %
-This is similar to the 1-RTT semi-static pattern in \mOptls which computes the MAC key \textsf{sfk} for the message \textsf{sfin}~\cite{DBLP:conf/eurosp/KrawczykW16}.
+This is similar to the 1-RTT semi-static pattern in \mOptls which computes the
+MAC key \textsf{sfk} for the message
+\textsf{sfin}~\cite{DBLP:conf/eurosp/KrawczykW16}.
 %
-The same procedure works for $R$ as well (with the values corresponding to $R$). 
+The same procedures works for $R$ as well (with the values corresponding to $R$).
 %
-The abstract calculation for the authentication values is as shown in Table~\ref{tab:authvalues}~\cite{Norr21}.
+The abstract calculation for the authentication information values is as shown
+in Table~\ref{tab:authvalues}~\cite{Norr21}.
 %
-We use $\mathit{MAC}_{I}$ (resp.
-$\mathit{MAC}_{R}$) to denote a MAC which uses an encryption key constructed using $\langle\mPriv{R},\ \mPub{R}\rangle$ and $\langle\mX,\ \mGx\rangle$ (resp.
-using $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\ \mGy\rangle$).
+We use $\mathit{MAC}_{I}$ (resp. $\mathit{MAC}_{R}$) to denote a MAC which uses
+an encryption key constructed using
+$\langle\mPriv{R},\ \mPub{R}\rangle$ and $\langle\mX,\ \mGx\rangle$
+(resp. using $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\ \mGy\rangle$).
+%
 \begin{table}[ht]
@@
 
-In addition to this, communications 2 and 3 contain messages which are encrypted and integrity protected, as can be seen in~\ref{fig:edhocFramework}.
+In addition to this, the second and third messages have
+provisions for transferring application messages which are encrypted and
+integrity protected, as can be seen in~\ref{fig:edhocFramework}.
 %
-The second communication is encrypted by performing an \mXor{} operation between the plaintext and the output of the key derivation function \mHkdf{}, as in Section~\ref{sec:keysched}.
+The second message is encrypted by performing a bit-wise \mXor{} between
+the plaintext and the output of the key derivation function \mHkdf{}, as
+in Section~\ref{sec:keysched}.
 %
-For the third message, encryption and integrity protection is assured by the \mAead{} algorithm which is part of the established ciphersuite \mSuites{}.
+For the third message, encryption and integrity protection is assured by the
+\mAead{} algorithm which is part of the established ciphersuite \mSuites{}.
+%
 
-In Figure~\ref{fig:edhocsigstat} we show an example of protocol execution under the \mSigStat{} method. 
+In Figure~\ref{fig:edhocsigstat} we show an example of protocol execution under
+the \mSigStat{} method.
 %
-The figure describes in detail the various message patterns, and operations and key derivations used to construct these messages.
+The figure describes in detail the various message patterns, operations and
+key derivations used to construct these messages.
 \vnote{Do we still have the other method diagrams that we wrote out for the original paper but did not include?}
+\knote{I don't think we have space for that.}
 
@@
 \end{figure}
-%------------------------------------------------------------------------- sub

### 2022-07-20 formalization.tex
Changes to discussion sections, minor typo fixes elsewhere

@@
 3    agreed = <CS0, CI, ~CR>
-4    gx = 'g'^xx
-5    data_2 = <'g'^~yy, CI, ~CR>
-6    m1 = <'STAT', 'SIG', CS0, CI, gx>
+4    gx = `g'^xx
+5    data_2 = <`g'^~yy, CI, ~CR>
+6    m1 = <`STAT', `SIG', CS0, CI, gx>
 7    TH_2 = h(<$H0, m1, data_2>)
-8    prk_2e = extr('e', gx^~yy)
+8    prk_2e = extr(`e', gx^~yy)
 9    prk_3e2m = prk_2e
-10   K_2m = expa(<$cAEAD0, TH_2, 'K_2m'>,
+10   K_2m = expa(<$cAEAD0, TH_2, `K_2m'>,
 11               prk_3e2m)
@@
 15   assocData2 = <protected2, extAad2>
-16   MAC_2 = aead('e', K_2m, assocData2,
+16   MAC_2 = aead(`e', K_2m, assocData2,
                   $cAEAD0)
@@
 19   K_2e = expa(<$cAEAD0, TH_2,
-                 'K_2e'>, prk_2e)
+                 `K_2e'>, prk_2e)
 20   K_2e_1 = expa(<$cAEAD0, TH_2,
-                   'K_2e', '1'>, prk_2e)
+                   `K_2e', `1'>, prk_2e)
 21   K_2e_2 = expa(<$cAEAD0, TH_2,
-                   'K_2e', '2'>, prk_2e)
+                   `K_2e', `2'>, prk_2e)
 22   CIPHERTEXT_2 = <$V XOR K_2e_1,

### 2022-07-20 paper.tex
[add] update of protocol description
Changes to discussion sections, minor typo fixes elsewhere

@@
 \input{formalization}
-
-%-------------------------------------------------------------------------- sec
-\section{\uppercase{Denial of Service}}
-\label{sec:dos}
-
-% There is little about EDHOC specific DoS stuff; that is interesting in itself,
-% and motivates this section
-There is a wealth of information about DoS in general, and more specifically for
-IoT.
-%
-However, there is little information regarding DoS aspects stemming
-specifically from \mEdhoc{}.
-%
-
-% Other protocols have DoS considerations, but EDHOC delegates it to developer
-In contrast to \mTls{} and IKEv2,
-the \mEdhoc{} core specification does not include any specific measures for
-countering Denial of Service (DoS) attacks.
-%
-Instead, that responsibility is delegated to the application, using
-proprietary measures or extensions to \mEdhoc{}.
-%
-
-% The core is small and probably not vulnerable to anything unexpected
-The \mEdhoc{} core specification has limited functionality, few negotiable
-parameters and processing rules.
-%
-Hence, we argue in this section that its attack surface
-is limited from a DoS perspective.
-%
-
-% However, It is difficult to define what EDHOC *is*; it s partially specified in many
-% specs
-However, \mEdhoc{} leaves much of the semantics of its messages undefined or
-specified as extensions.
-%
-Examples of mechanisms with mostly undefined semantics include processing
-rules and content of error messages and of the additional data that is
-passed in each of the \mEdhoc{} messages.
-%
-Another example is that the core specification does not include a
-message transport mechanism.
-%
-The transport for \mEdhoc's primary target, i.e., to key \mOscore{},
-is \mCoap{}.
-%
-This is specified in~\cite{ietf-core-oscore-edhoc-03}, but still leaves
-many design choices to the application.
-%
-Because of this, it is difficult to draw the line between what is considered as
-\mEdhoc{} and what is an extension.
-%
-
-% Pros with undefined semantics: flexibilty; Cons: developers can shoot
-% themselves in the foot
-Undefined semantics allow the implementors much freedom
-and flexibility to design applications.
-%
-On the down-side, the few recommendations given increase the risk of
-applications being vulnerable to DoS via \mEdhoc{}, even
-though the core \mEdhoc{} specification itself is arguably not particularly
-weak in this respect.
-%
-
-\paragraph{DoS mitigation is left to developers.}
-Section 8.7 of the \mEdhoc{} specification\footnote{draft-ietf-lake-edhoc-14}
-explains the design philosophy w.r.t.
-DoS in two paragraphs and gives two recommendations.
-%
-The first being that \mEdhoc{} relies on lower layers to mitigate DoS, and an
-example of DoS countermeasures is given: checking return routeability when
-receiving the first message.
-%
-The second being that an application may try to determine (in some way) that an
-apparently valid message is in fact probably a forgery and should be ignored or
-cause processing to stop.
-%
-
-No recommendation is given on how to minimize DoS effects stemming from use of
-error messages, which can be used in an arbitrary way in \mEdhoc{}
-(see Section~\ref{sec:errorHandling} below).
-%
-While there is some work on DDoS aspects of \mCoap{}, and much on IoT in
-general, this does not capture the particular consequences originating with
-\mEdhoc{}.
-%
-We now elaborate on some considerations application developers needs to take
-into account.
-%
-
-\paragraph{Amplification and role asymmetry.}
-For the same reasons as given in the IKEv2 DDoS analysis
-by~\cite{rfc8019} the responder is more vulnerable than the
-initiator.
-%
-The key similarity between IKEv2 and \mEdhoc{} here is that the initiator is
-anonymous in the first message and that creating the first message is
-computationally cheap and requires no storage.
-%
-No element in the first \mEdhoc{} message can be verified by the receiver as
-being legitimate as long as the format is correct and elliptic curve points
-lie on the proposed curve.
-%
-\mEdhoc{} leave it to the application and its
-transport layer to resolve any issues.
-%
-In addition to the computational resources the responder spends, \mEdhoc{}
-execution may require the responder to
-obtain and verify certificate chains, and verify
-certificate-revocation servers as a result of the spoofed initial message.
-%
-For example, the content of the \mADone{} information element may hint the
-responder to take further actions.
-%
-
-The asymmetry between the adversary's effort and the responder's may hence be
-significant.
-%
-Since \mEdhoc{} specifically targets constrained devices this asymmetry needs to
-be considered even more carefully than for IKEv2~\cite{rfc8019}.
-%
-
-Depending on the use case, the first \mEdhoc{} message may be sent by either
-party in a session.
-%
-For example, even though it may seem intuitive that an electronic
-key-card initiates the connection towards lock, the first \mEdhoc{} message
-could be included in the response from the lock.
-%
-This means the lock can postpone heavy verification actions until it has
-authenticated the key card after receiving the third \mEdhoc{} message.
-%
-The draft IETF specification profiling \mEdhoc{} for use with
-\mCoap{} and \mOscore{} ~\cite{ietf-core-oscore-edhoc-03},
-however, defines the
-\mCoap{} client to be the \mEdhoc{} initiator, and hence reduces the flexibility
-to use that option.
-%
-
-\paragraph{Reflection Attacks.}
-Reflection attacks are attacks where an adversary spoofs the source address of a
-victim in a message and thereby tricks the responder to send a message to the
-victim.
-%
-\mEdhoc{} leaves protection against this to the transport protocol and
-application, and suggests using a return path reachability method similar to IKEv2.
-%
-We note that such reachability mechanisms can easily be used to make a responder
-send reachability requests to arbitrary addressable targets.
-%
-Depending on the design of such reachability mechanisms, they may require that
-unsolicited reachability requests are discarded, limiting the effects of
-anonymous DoS attacks like these.
-%
-
-In situations where the responder does not need to perform additional
-communications to verify certificates or similar, a reachability mechanism may
-be more expensive in terms of time and storage compared to continue with the
-protocol and keeping a half-open state for a period of time.
-%
-Doing so would also avoid creating secondary DoS effects on certificate storage
-nodes or certificate revocation servers.
-%
-The latter may cause a systems-level issue if many distributed
-\mEdhoc{} responders frequently requests such servers.
-%
-However, if round-trip times are large, and half-open states hence are kept
-by the receiver for a long time, a reachability mechanism can trade
-communication for storage.
-\footnote{RFC8019 discuss some numbers as examples of tradeoffs. Should we set
-up a simple model and derive an expression for trade-offs? Could be useful for
-implementors to consider, but low academic value. This would capture also
-amplification.}
-%
-
-%\mPoint{In a constrained device randomness may be easily depleted}
-%Assume an adversary generate (deterministic) bogus message 1s and send these at
-%a sufficiently high rate to the responder.
-%%
-%\mEdhoc{} specifically targets constrained devices, that presumably may not
-%always have good randomness sources.
-%%
-%Is it possible to deplete or significantly reduce the entropy of the source and
-%in this way be able to predict the responder's random input to other sessions
-%(with other initiators)?
-%%
-%The LAKE WG have been talking about effects of poor randomness on signatures, so
-%this should be an issue, no?
-%%
-
-\mPoint{\mEdhoc{} can deadlock if not used correctly}
-
-An adversary may send forged error messages, and specifically reject proposed
-ciphersuites.
-%
-The initiator is recommended to not try that ciphersuite with the responder
-again in a new session for a while.
-%
-If the adversary rejects all options for a initiator this way, it can prevent
-communication between the initiator and that responder until the cache is
-cleared.
-%
-Potentially, an adversary could lock a party of a group out of communication all
-together this way.
-%
-This needs to be considered by applications.
-%
-
-While we have not conducted a formal analysis of the fact, the structure of all
-methods are so similar that it seems plausible that none of them would be
-stronger against DDoS than the others.
-%
-
-Implementers are well advised to consult \cite{rfc8019}, which provides much
-good general information about DoS for key establishment
-protocols with the 3-message structure employed by \mEdhoc{}.
-%
-However, ultimately, a product-specific risk assessment is required to
-determine appropriate DDoS mitigations.
-%
-
-%-------------------------------------------------------------------------- sec
-\section{\uppercase{Error handling}}
-\label{sec:errorHandling}
-\mPoint{Error handling is undefined except for algo negotiation} Section 6 of the \mEdhoc{} specification\cite{draft-ietf-lake-edhoc-15} states that error messages can be sent at any time and by any party.
-%
-At the current version of the standard \cite{draft-ietf-lake-edhoc-15}, there are three types of error messages: type 0 is used to represent success, type 1 represents a generic error message used for debugging purposes, where the actual message is an English string, and type 2 represents failure to negotiate a common ciphersuite.
-Type 2 messages carry a list of supported ciphersuites by the responder.
-%
-The contents and semantics of type 1 messages are dependent on the implementation.
-%
-Consequently, type 1 error messages provide a generic message passing mechanism, albeit one without a predefined semantics.
-
-%\textbf{Q:} Is there any situation where it makes sense that the initiator sends an unsupported ciphersuite message instead of message 3?
-%\textbf{Q:} In Section H: if no message correlation support by the transport layer, \mCi{} and \mCr{} should be used for correlation of messages to a session.
-
-\subsection{Lack of connection identifiers in error messages}
-Error messages do not include the optional connection identifiers \mCi{} or \mCr{} that are transmitted during the handshake.
-This is potentially problematic for connection-less transport layers, where \mCi{} and \mCr{} may be used to correlate messages.
-
-\subsection{Improper handling of error messages}
-\mEdhoc{} provides the error message mechanism, but gives little to no guidance on how it should be used safely.
-%
-For example, assume that an application using \mEdhoc{} logs error messages on a finite log with log rotation.
-%
-If the log is used for anomaly detection or detection of sensitive events, then an adversary can simply inject error messages and fill the log until the sensitive event is overwritten.
-%
-If filling error messages in the log is less suspicious than the actual attack, this may be beneficial for the adversary.
-%
-Implementors should do proper log-separation and log management.
-
-\subsection{Unspecified semantics for debugging messages}
-Section 6.2 gives an example of sending a type 1 error message with the information string "Method not supported".
-%
-It's important to note that this message cannot be used to negotiate the connection method securely.
-%
-In fact -- differently from ciphersuite negotiation -- \mEdhoc{} does not include a mechanism to support negotiating the connection method, among the four available combinations of signature keys and static Diffie-Hellman keys.
-%
-The initiator selects the method on the first message, and the responder can reply with a type 1 error message along with the information string "Method not supported", but a standard compliant implementation cannot act upon this information string for negotiation purposes.
-%
-This is because the error message is not integrity protected and there is no secure mechanism to select a different authentication method that ensure that the preferred available authentication mechanism is always selected.
-%
-Because method negotiation can be viewed in the context of cryptographic agility, there is room for improvement.
-%
-
-Note: Section 5.1 says that if processing fails for some reason,
-then "typically" an error message is sent processing is terminated and protocol
-state is discarded. But: for algo negotiation (the only defined error
-message), then state is not discarded.
-
-Note: Section 5.2.2 says:
-"The Initiator MUST NOT change the supported cipher suites and
-the order of preference in SUITES\_I based on previous error messages."
-But: what happens if order of preference changes for some other reason? Would it
-be good it include a flag in msg1 to indicate to the responder that the
-ciphersuite order depends on a previous run? This way, responder could react
-differently if a new cipher suite has been added without breaking the "MUST NOT"
-requirement. Need to think about this more.
-
-Note: Section 5.3.3 says about receiver processing of message 2:
-"If any processing step fails,
-the Responder MUST send an EDHOC error message back, formatted as defined in
-Section 6, and the session MUST be discontinued"
-And Section 5.4.3 about the responder processing of message 3:
-"If any processing step fails, the Responder MUST send an EDHOC error
-message back, formatted as defined in Section 6, and the session MUST
-be discontinued."
-The same is said in Section 5.5.3 about the initiator's processing of message 4.
 
@@
 A party $U$ using the \mStat{} authentication method use a TEE with an interface
-accepting a point $p$ on the curve and returning $\mDH(p,\ \mGu{})$.
+accepting a point $P$ on the curve and returning $\mDH(P,\ \mGu{})$.
 %
@@
 %-------------------------------------------------------------------------- sec
-\section{\uppercase{Discussion}}
+\section{\uppercase{Error handling}}
+\label{sec:errorHandling}
+Section 6 of the current \mEdhoc{} specification~\cite{draft-ietf-lake-edhoc-15} states that error messages can be sent at any time and by any party.
+%
+There are three types of error messages: type 0 is used to represent success, type 1 represents a generic error message used for debugging purposes, and type 2 represents the failure to negotiate a common ciphersuite.
+%
+Type 2 messages carry a list of supported ciphersuites by the responder.
+%
+The contents and semantics of type 1 messages are dependent on the implementation.
+%
+Consequently, type 1 error messages provide a generic message passing mechanism, albeit without a predefined semantics.
+%
+We now discuss some issues with this specification of error messages.
+
+\subsection{Lack of connection identifiers}
+Error messages do not include the optional connection identifiers \mCi{} or \mCr{}.
+%
+This is potentially problematic for connection-less transport layers, where these identifiers may be used to correlate messages.
+
+\subsection{Improper handling}
+\mEdhoc{} provides an error message mechanism, but little to no guidance on how it should be used safely.
+%
+For example, assume that an application using \mEdhoc{} logs error messages on a finite log with rotation.
+%
+If the log is used for anomaly detection or detection of sensitive events, then an adversary can simply inject error messages and fill the log until the sensitive event is overwritten.
+%
+In such situations, implementors should do proper log-separation and log management.
+
+\subsection{Unspecified semantics for debugging messages}
+Section 6.2 gives an example of sending a type 1 error message with the information string "Method not supported".
+%
+Note that \mEdhoc{} does not include a mechanism to support the negotiating of the connection method.
+%
+The initiator selects the method as part of the first message, and if the responder replies with a type 1 error message along with the information string "Method not supported", the standard does not tell an implementation how to act upon this information string for negotiation purposes.
+%
+An adversary can fake an error message, and there is no algorithm that is specified to ensure that a common available authentication mechanism is always selected.
+%
+In particular, note that Section 5.1 in the specification states that if processing fails for some reason, then "typically" an error message is sent, processing is terminated and protocol state is discarded.
+%
+But negotiation needs us to remember state.
+
+With respect to ciphersuite negotiation, Section 5.2.2 of the specification also states ``The Initiator MUST NOT change the supported cipher suites and
+the order of preference in SUITES\_I based on previous error messages.'' 
+%
+However, \mSuites is sent in the clear, and therefore can be modified by the adversary.
+%
+It is important to consider what happens if the order of preference in \mSuites is changed either by the adversary, or by the initiator themselves for some other reason.
+%
+It might be helpful to include a flag in \mMsgone to indicate to the responder that a (modified) ciphersuite ordering depends on that from a previous execution.
+
+%-------------------------------------------------------------------------- sec
+\section{\uppercase{Improvements to the specification}}
 \label{sec:discussion}
-There are a few places where \mEdhoc{} can be improved,
-which we found during this work and communicated to the authors.
+The \mEdhoc{} \mSpec{} has many points where it can be improved. 
 %
-We discuss them below.
+We discovered a few of these issues in~\cite{Norr21}, but in this work, we also discuss a couple of new issues.
 %
+We communicated the points discussed in~\cite{Norr21} to the authors of the \mSpec{}.
+%
+Here we also list the protocol developers' response to being told of these ambiguities in the \mSpec{}.
 
@@
 %
-The \mEdhoc{} \mSpec{} lists several security goals, but they are
-imprecise and difficult to interpret due to lack of context and intended usage
-descriptions.
+There are several security goals listed in the \mSpec{}.
 %
-Without knowing how the protocol is to be used,
-it is not clear whether the listed security goals are the most important ones
-for constrained IoT devices.
+However, these goals are informal, and imprecisely specified, and the lack of intended usage makes them difficult to interpret as well.
+%
+Since we have no indication about the use cases in which \mEdhoc{} can be employed, it becomes tough to evaluate if the listed security goals are the ones that we need the protocol to satisfy.
+
+The overarching goal of \mEdhoc{} is to establish a security context for \mOscore{} using as few roundtrips and as small messages as possible, which dictates the design of the protocol.
+%
+However, \mEdhoc{} might be intended for multiple use cases, some of which might not even be covered in the \mSpec{}, which might need different guarantees from the ones mentioned therein.
+%
+In order to identify security goals that might apply to multiple such cases, we made up some user stories corresponding to typical use cases, which helped us identify some subtle points of concern in \mEdhoc.
+%
+We communicated these issues to the authors of \mEdhoc{}.
+%
+\vnote{Add responses from the developers to these questions}
+
+\subsubsection{Non-Repudiation:}
+Access control mechanisms can vary depending on what the underlying application is.
+%
+A nuclear power-plant might need to keep track of who enters and leaves, but a coffee machine that logs every user and their coffee preference might lead to a privacy concern.
+%
+This thought experiment allowed us to perceive that the \mEdhoc{} \mSpec{} did not even consider non-repudiation.
+%
+When we pointed this out to the authors of the \mSpec, they recognized this concern, and added a discussion about how the various \mEdhoc{} methods satisfy (or not) non-repudiation.
+
+\subsubsection{Unintended Peer Authentication: }
+In Section~3.2 in the \mSpec{}, the authors state that parties executing \mEdhoc{} must be configured such that they only run \mEdhoc{} with a restricted (according to some reasonable policy) set of peers.
+%
+However, \mIdcredr{} in the second message is sent in the clear, and can therefore lead to some masquerade attacks by the adversary.
 %
 
-The abstract goal of \mEdhoc{} is simple: establish an \mOscore{} security
-context using few roundtrips and small messages.
+Suppose a person configures the restriction policy such that every device in their home is allowed to set up an \mEdhoc{} session with any other device in the home.
 %
-From that, the design of \mEdhoc{} is mainly driven by what
-can be achieved given the technical restrictions.
+However, the adversary has managed to gain control of one of these devices (say $A$).
 %
-Focusing too much on what can be achieved within given restrictions, and paying
-too little attention to the use cases where the
-protocol is to be used and their specific goals, risks resulting in
-sub-optimal trade-offs and design decisions.
+Device $A$ is part of the allowed group of devices, so $B$ will accept an \mEdhoc{} run with $A$, even though $A$ is now compromised.
+%
+In addition, since $B$ does not explicitly check to see if the \mIdcredr{} received in the second message corresponds to the intended identity, it leads to attacks where $B$ wants to talk to another non-compromised home device $C$, but $A$ responds in place of $C$ and $B$ continues with the run anyway.
+%
+This requires the initiator to match \mIdcredr{} to the identity intended by them when they send out the first message.
+%
+We included this check in our model.
+%
+We communicated this to the authors of \mEdhoc{}.
+\vnote{And what did they do about it?}
+
+\subsubsection{Denial of Service: }
+Denial of Service (DoS) attacks for IoT protocols are very well-studied.
+%
+But DoS is not well documented for \mEdhoc{} in particular, since it delegates DoS handling to the application.
+%
+The \mSpec{} does not include any specific measures for countering such attacks.
+
+Section 8.7 of the \mEdhoc{} specification~\footnote{draft-ietf-lake-edhoc-14} gives two recommendations for countering DoS attacks.
+\vnote{Is this the latest spec? The spec we’re analyzing? Which one?}
+%
+First, \mEdhoc{} relies on lower layers to mitigate DoS, and an example of DoS countermeasures is given: checking the return address upon receipt of the first message.
+%
+Second, an application may try to determine (in some way) that a seemingly valid message is in fact probably a forgery and should be ignored or cause the protocol to abort.
+
+No recommendation is given on how to minimize DoS effects stemming from the use of error messages, which can be used in a relatively arbitrary way in \mEdhoc{} (see Section~\ref{sec:errorHandling}).
+%
+We now elaborate on some considerations that the application developers needs to take into account.
 %
 
-\mEdhoc{} is intended to cover a variety of use cases, many of which are
-difficult to predict today.
+As in the IKEv2 DDoS analysis by~\cite{rfc8019}, the responder is more vulnerable than the initiator.
 %
-However, this does not
-prevent collecting \emph{typical} use cases and user stories
-to identify more specific security goals that will be important in most cases.
+The key similarity is that the initiator is allowed identity protection in the first message.
+%
+None of the elements in the first \mEdhoc{} message can be proved by the responder to be illegal, as long as the format is correct and the chosen elliptic curve points lie on the pre-chosen curve.
+%
+Additionally, execution may require the responder to obtain and verify certificate chains.
+%
+For example, the content of the \mADone{} information element may need the responder to take further action.
 %
 
-While constructing our model, we made up simple user stories to identify
-security properties of interest.
+The asymmetry between the adversary's effort towards the first message and that of the responder may hence be significant.
 %
-Several of these revealed subtleties and undefined aspects of \mEdhoc{}.
+Note that since \mEdhoc{} is to be used in conjunction with \mCoap{} and \mOscore{}, either party executing the \mCoap{} protocol might be the one playing the initiator role in \mEdhoc.
 %
-We informed the \mEdhoc{} authors, who addressed these aspects in the
-\mSpec{}.
+Since \mEdhoc{} is designed for constrained devices, this asymmetry needs to be considered even more carefully than for IKEv2~\cite{rfc8019}.
+
+Reflection attacks are another possibility.
 %
+These are attacks where an adversary spoofs the source address of a victim in a message and thereby tricks the responder into sending a message to the victim.
+%
+\mEdhoc{} suggests using a return path reachability method similar to IKEv2.
+%
+Note that even such reachability mechanisms can easily be used to make a responder send reachability requests to arbitrary addressable targets.
+%
+However, such mechanisms might require that unsolicited reachability requests be discarded, limiting the effects of
+anonymous DoS attacks.
+%
+In situations where the responder does not need to perform additional communications to verify certificates etc, a reachability mechanism may be more expensive in terms of time and storage compared to continuing with the protocol while keeping a ``half-open’’ state for a period of time.
+%
+The latter may cause a systems-level issue if many distributed \mEdhoc{} responders frequently request such servers.
+%
+However, if round-trip times are large, and half-open states hence are kept by the receiver for a long time, a reachability mechanism can trade communication for storage.
 
-\subsubsection{(Non-)Repudiation}
-An access control solution for a nuclear power-plant may need to log who is
-passing through a door, whereas it may be undesirable for, say, a coffee
-machine to log a list of users along with their coffee preferences.
+An adversary may also send forged error messages (see Section~\ref{sec:errorHandling}), and specifically reject proposed ciphersuites.
 %
-Via this simple thought experiment, we realized that the \mSpec{} did not
-consider the concept of (non)-repudiation.
+It is recommended for the initiator to not try that ciphersuite with the responder again in a new session.
 %
-In response, the authors of the \mSpec{} added a paragraph discussing how
-different methods relate to (non)-repudiation.
-
-\subsubsection{Unintended Peer Authentication}
-Section~3.2 of the \mSpec{} states that parties must be configured
-with a policy restricting the set of peers they run \mEdhoc{} with.
+If the adversary rejects all options for a initiator this way, it can prevent
+communication between the initiator and that responder until the cache is
+cleared.
 %
-However, the initiator is not required to verify that the \mIdcredr{} received
-in the second message is the same as the one intended.
-%
-The following attack scenario is therefore possible.
-%
-
-Suppose someone has configured all devices in their home to be in the
-allowed set of devices, but that one of the devices ($A$) is compromised.
-%
-If another device $B$, initiates a connection to a third device $C$, the
-compromised device $A$ may interfere by responding in $C$'s place, blocking
-the legitimate response from $C$.
-%
-Since $B$ does not verify that the identity indicated in the second message
-matches the intended identity $C$, and device $A$ is part of the allowed set,
-$B$ will complete and accept the \mEdhoc{} run with device $A$ instead of the
-intended $C$.
-%
-The obvious solution is for the initiator to match \mIdcredr{} to the intended
-identity indicated by the application, which we included in our model.
-%
-We have communicated this to the \mEdhoc{} authors and they are considering
-how to resolve the issue.
+Potentially, an adversary could lock a party of a group out of communication altogether this way.
 %
@@
 \subsection{Unclear Security Model}
-We argue that the \mSpec{} gives too little information about what capabilities
-an adversary is assumed to have, and that this leads to unclear design goals and
-potentially sub-optimal design.
+In addition to unclear contexts in which \mEdhoc{} can be used, the \mSpec{} also does not provide enough detail about adversary capabilities. 
 %
-
-Even though \mEdhoc{} incorporates cryptographic cores from different academic
-security protocols, its design does not take into account the adversary models
-for which these protocols were designed.
+The four different cryptographic cores included in \mEdhoc{} are based on academic protocols which are designed to work with highly specific but potentially different adversary models.
 %
-For example, \mOptls{}, whose cryptographic core is essentially the same
-as the \mStat{} authentication method, is designed to be secure in the CK
-model~\cite{DBLP:conf/crypto/CanettiK02}.
+%For example, \mOptls{}, which essentially has the same core as \mStat{}, is shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, which explicitly separates long-term key storage from that of the ephemeral keys and session state.
+%%
+%This fits in well with our use of TEEs.
+Since \mSigma{} cannot protect against compromised ephemeral keys, the authors of \mEdhoc{} felt that considering the compromise of ephemeral keys separately from that of long-term keys was not required~\cite{personalCommunication}.
 %
-The CK security model explicitly separates the secure storage of long-term
-keys from storage of session state and ephemeral keys.
+The reason was presumably based on the fact that the \mSigSig{} method was modeled closely on the \mSigmaI{} variant of \mSigma{}, and that it would be preferable to obtain a homogeneous security level among the \mEdhoc{}
+methods.~\cite{Norr21}
 %
-This is appropriate for modelling the use of secure modules.
+However, this only holds true for situations where one is only interested in the session key confidentiality for an ongoing session.
 %
-
-The \mEdhoc{} authors indicated to us that it was
-not necessary to consider compromised ephemeral keys separately from
-compromised long-term keys.
+Secure modules, in addition to being able to store long-term keys, are useful in other ways too. 
 %
-The rationale is that \mSigma{} cannot protect against compromised ephemeral
-keys~\cite{personalCommunication}.
+They also provide Post-Compromise Security (PCS) guarantees, for example.
 %
-That rationale is presumably based on the fact that the \mSigSig{} method is
-closely modeled on the \mSigmaI{} variant of \mSigma{}, and that it would be
-preferable to obtain a homogeneous security level among the \mEdhoc{}
-methods.
-%
-That rationale is only true, however, if one restricts attention to session key
-confidentiality of an ongoing session.
-%
-Secure modules provide value in other ways, for example, by allowing
-constructions with Post-Compromise Security (PCS) guarantees.
-%
-We discussed this with the authors, and
-the latest version of the \mSpec{}~\cite{latest-ietf-lake-edhoc-05} includes
-recommendations on storage of long-term keys and operations on these inside a
-secure module.
-%
-
+Upon discussion with the authors, they included recommendations on how to store long-term keys, and how to perform operations on these keys inside a secure module.
+\vnote{This needs more description?}
 %-------------------------------------------------------------------------- sub
@@
 \label{sec:sessionKeyMaterial}
-\mEdhoc{} establishes session key material, from which session keys
-can be derived using the \mEdhoc{}-Exporter.
+As we have seen, \mEdhoc{} establishes some session key material, which can be fed into the \mEdhoc{}-Exporter to derive session keys.
 %
-The session key material is affected by \mGxy{}, and if a party uses the
-\mStat{} authentication method, also by that party's secret static long-term key.
+This key material is directly influenced by \mGxy{} and a party's secret static long-term key, if they use \mStat{}.
 %
-As shown in Section~\ref{sec:formalization}, mutual injective agreement cannot
-be achieved for $P_I$.
+As mentioned earlier, mutual explicit injective agreement cannot be obtained for $P_{I}$.
 %
-If this property is not important for constrained IoT devices which cannot use
-any of the other methods, then one can simply accept that the methods have
-different authentication strengths.
+This might prove to be a problem, and we propose three potential alternatives for addressing this.
 %
-Otherwise, this is a problem.
+\begin{itemize}
+\item Include \mIdcredi{} or a hash thereof in the first and second messages: 
 %
-
-We identified three alternatives for resolving this.
+This increases message size and also leaks the initiator's identity.
+\item Derive session key material without using $P_{I}$: 
 %
-One alternative is to include \mIdcredi{}, or its hash, in the first and
-second messages.
+This deviates from the structure of protocols like \mOptls{} from which the \mStat{}-based methods draw heavy inspiration.
 %
-This would, however, increase message sizes and prevent initiator identity
-protection, which are grave concerns for \mEdhoc{}.
+In those protocols, $P_{I}$'s presence in the computation of session key material is crucial to obtaining resistance against the compromise of the initiator's ephemeral key.
+\item Include a fourth message from the responder to the initiator which contains a MAC based on information obtained using $P_{I}$:
 %
-A second alternative is to not derive the session key material from $P_I$.
-%
-Doing so, however, deviates from the design of \mOptls{} (and similar protocols
-from which the \mStat{}-based methods are derived), where the inclusion of
-$P_I$ plays a crucial part in the security proof of resistance against
-initiator ephemeral key compromise.
-%
-The third alternative is to include a fourth message from responder to initiator,
-carrying a MAC based on a key derived from session key material including $P_I$.
-%
-Successful MAC verification guarantees
-to the initiator that the responder injectively agrees on $P_I$.
-%
-We presented the options to IETF, and they decided to add a
-fourth message as an option in the latest version of the
-\mSpec{}~\cite{latest-ietf-lake-edhoc-05}.
-%
-
-Regardless of how this is handled, we verified that all methods
-enjoy a common, but weaker, property: mutual implicit agreement
-on all of $P_e, P_I$ and $P_R$, where applicable.
-%
+While this option increases the number of messages in the protocol, the authors decided to go with this option for situations where it is necessary to provide explicit injective agreement on $P_{I}$.
+\end{itemize}
 

### 2022-07-20 protocol.tex
Changes to discussion sections, minor typo fixes elsewhere

@@
 }
-\caption{Key schedule: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo, \mPRKthree,
-\mPRKfour{} is the intermediate key material, and
-\mKtwoe,\mKtwom, \mKthreeae, \mKthreem{} are the encryption keys for \mAead{}
-or \mXor{}.
+\caption{Key schedule: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo, \mPRKthree, \mPRKfour{} are the intermediate key material, and \mKtwoe,\mKtwom, \mKthreeae, \mKthreem{} are the encryption keys for \mAead{} or \mXor{}.
 Dashed boxes are conditionals.~\cite{Norr21}}

### 2022-07-20 protocol.tex
[fix] separate the non-formal aspects of TEEs into their own section

@@
 \end{figure}
+
+%------------------------------------------------------------------------- sec
+\section{Implementation Aspects and Key Protection}
+\label{sec:TEE}
+Adversaries may have physical access to implementations, this is perhaps an
+even higher risk in an unsupervised IoT setting.
+%
+Implementations can make use of Trusted Execution Environments (TEE) hardened
+against physical attacks to mitigate adversaries with physical access to the
+cryptographic keys.
+%
+
+TEEs come in various forms of differing complexity.
+%
+Some, like ARM Trust Zone and Intel SGX are general purpose execution
+environments, which are flexibly programmable.
+%
+Other, like the Universal Subscriber Identity Modules (USIM) used for
+authentication to 3GPP mobile networks, have application specific interfaces for
+authentication and key agreement protocols etc.
+%
+A key aspect is how much of the application is placed in the TEE and how much is
+outside.
+%
+
+For larger devices that include general purpose processors with Trust Zone or
+SGX, entire \mEdhoc{} and \mOscore{} may reside inside the TEE.
+%
+For constrained IoT devices in the lower end of the scale, a TEE may have to be
+implemented using a special-purpose integrated circuit.
+%
+In the latter case, it may be beneficial to follow a minimalistic approach, and
+contain the long-term key and only
+the operations that need access to it in the TEE for cost reasons.
+%
+
+It may at first appear more secure to implement as much as possible inside the
+TEE, but there is a security trade-off: the more code inside the TEE,
+the higher the risk of implementation errors in the security-critical code.
+%
+Because the security critical code runs in the area where the long-term keys
+reside, an implementation error here risks leaking information of the key to the
+adversary.
+%
+From this perspective, it may be beneficial to follow the minimalistic approach
+even when having access to Trust Zone or SGX.
+%
+
+A slightly more secure division of functionality is to also keep the
+session key inside
+the TEE and extend the interface to accept messages and return the
+(en/de)crypted counter part.
+%
+That is, the interface exposes \mAead{} functions in the interface.
+%

### 2022-07-21 formalization.tex

@@
 
-As in~\cite{Norr21}, we verify \mEdhoc{} in the Dolev-Yao model~\cite{DY83}.
+As in~\cite{Norr21}, we verify \mEdhoc{} in an extended
+Dolev-Yao model~\cite{DY83}.
 %
-This is a well-established model for symbolic verification of security protocols.
+Dolev-Yao is a well-established model for symbolic verification of security
+protocols.
 %
-Messages are modelled as terms in an algebra, and the various cryptographic operations are assumed to be perfect, i.e.
-encrypted messages can only be decrypted with the correct key, no hash collisions exist etc.
+Messages are modelled as terms in an algebra, and the various cryptographic
+operations are assumed to be perfect, e.g., encrypted messages can only be
+decrypted with the correct key, and there are no hash collisions exist.
 %
 
-The adversary is assumed to be the network, and thus controls the communication channel and can see all messages being communicated as part of the protocol.
+The adversary is assumed to be the network, and thus controls the communication
+channel and can see all messages being communicated as part of the protocol.
 %
-In addition, they can interact with an unbounded number of protocol sessions, dropping, injecting, and modifying messages at will, but cannot break cryptography.
+In addition, they can interact with an unbounded number of protocol sessions,
+drop, inject and modify messages at will.
+%
 
-On top of the Dolev-Yao model, we allow the adversary two extra capabilities with respect to getting access to the long-term and ephemeral keys. 
+On top of the Dolev-Yao model,~\cite{Norr21} allow the adversary two extra
+capabilities with respect to getting access to the long-term and ephemeral keys.
 %
@@
 %
-Ephemeral key reveal, denoted $\mRevEph^{t}(A, k)$, stands for the adversary obtaining, at time $t$, the ephemeral private key \mPrivE{A} used by party $A$  in a session where they establish key material $k$.
+Ephemeral key reveal, denoted $\mRevEph^{t}(A, k)$, stands for the adversary
+obtaining, at time $t$, the ephemeral private key \mPrivE{A} used by party $A$
+in a session where they establish key material $k$.
 %
-Ephemeral keys might be stored in a less secure part of a device, but long-term keys are often stored and operate within a secure model, like a trusted execution environment (TEE).
-\vnote{Do we want to say more about TEEs here?}
+% Ephemeral keys might be stored in a less secure part of a device,
+% but long-term keys are often stored and operate within a secure model,
+% like a trusted execution environment (TEE). \vnote{Do we want to say more about TEEs here?}
 %
-Formalizing these two events allows us to have more fine-grained control over the access that an adversary has to these fundamentally different kinds of keys.
+Formalizing these two capabilities allows more fine-grained control
+over the access that an adversary has to these fundamentally different kinds of
+keys.
+%
+
+%---------------------------------------------------------------------- subsub
+\subsubsection{System and Adversary Model Extensions Supporting TEE.}
+\label{sec:TEE:advModel}
+%
+We extend the adversary model of~\cite{Norr21} by allowing the adversary
+access to a TEE interface on a party of their choosing before and after the
+\mEdhoc{} session.
+%
+This allows us to distinguish between a full compromise and a weak one, using
+the terminology of~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
+%
+A weak compromise is when the adversary have temporary access to the TEE.
+%
+Seen in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add adversary
+capabilities corresponding to a server adversary, i.e., an adversary that
+compromises a party, learns its ephemeral keys and may temporarily access its
+TEE, but does not learn the party's long-term key.
+%
+The model of~\cite{Norr21} only captures server adversary capabilities without
+access to the TEE, via the leakage of ephemeral keys, and local adversary
+capabilities, or more specifically, a secure-world adversary with read-write
+access to the
+secure world where the long-term keys reside via the leakage of long-term keys.
+%
+Our extension allows analysis of so-called weak post-compromise security
+(PCS)~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
+%
+
+We split the \mEdhoc{} functionality as follows.
+%
+The TEE contains the long-term key and allows the non-TEE parts of the
+application to perform the operations on it via its interface.
+%
+More precisely, parties using the \mSig{} authentication method use a TEE with
+an interface which accepts a message and returns the signature of that message
+using the party's private long-term key.
+%
+A party $U$ using the \mStat{} authentication method use a TEE with an interface
+accepting a point $P$ on the curve and returning $\mDH(P,\ \mGu{})$.
+%
+
+The non-TEE parts implement \mEdhoc{} using this interface.
+%
+This interface requires the least functionality from the TEE, reducing the TEE's
+complexity and possibly cost of its implementation.
+%
+The functional split is suitable even for the situation where a
+constrained device has implemented only the storage of the long-term key in a
+special purpose circuit with minimal processing functionality.
+%
+Since \mEdhoc{} focuses on constrained IoT devices, it seems appropriate to
+cater for this setting.
+%
+
+%---------------------------------------------------------------------- subsub
+\subsubsection{Extended Formalism and Security Properties.}
+\label{sec:TEE:fmAndProps}
+Formally, we model the TEE interface by adding two new rewrite rules:
+%
+\begin{small}
+\begin{verbatim}
+rule forge_SIG:
+   [!LTK_SIG($A, ~ltk), In(xx)] --[TEE($A)]-> [Out(sign(xx, ~ltk))]
+
+rule exp_STAT:
+   [!LTK_STAT($A, ~ltk), In('g'^~xx)] --[TEE($A)]-> [Out(('g'^~xx)^~ltk)]
+\end{verbatim}
+\end{small}
+%
+These rules allow the adversary to obtain terms representing signatures on a
+value of their choice to forge signatures (\verb|forge_SIG|), or to obtain terms
+representing a curve point of their choice raised to the power of the
+long-term key (\verb|exp_STAT|).
+%
+
+Because it is a trivial attack when the adversary access these rules with values
+from the test session, we must disqualify those rule applications.
+%
+We do so by creating an action fact \verb|TEE($A)|, where \verb|$A| is the
+identity corresponding to the private key used, and then augmenting the
+properties with a condition that no such action fact exists from the start of
+the protocol execution and its end.
+%
+Care must be taken when specifying the start and the end.
+%
+Specifically, the start and end of the protocol run must be viewed w.r.t. the
+role.
+%
+For example, the injective agreement property for the initiator in
+the \mSigSig{} method requires that the adversary does not have access to the
+TEE of the
+responder from the time of transmission of the first message until the second
+message is received by the initiator.
+%
+Because reception of one message and transmission of the next one at a party
+is an atomic operation, these timepoints represent the start and end of the
+protocol run from the perspective of the initiator.
+%
+
+We did not manage to get Tamarin to terminate for all lemmas and all methods
+with the added TEE extension.
+%
+When the responder uses the \mStat{} authentication method Tamarin did not
+complete the in a reasonable time for the authentication lemmas.
+%
+To combat this, we simplified the modeling of XOR encryption of the second
+message for these lemmas when the responder uses \mStat{} authentication.
+%
+More precisely, the second message contains the two values \verb|R| (the
+responder's identity) and
+\verb|authR| (the authentication information), and we generally model the
+XOR encryption by xoring each of them with their own key-stream terms.
+%
+For the problematic lemmas however, we resorted to xoring the entire tuple
+\verb|<R, authR>| with a single key-stream term.
+%
+Given that no attacks were identified using our original modeling, this
+simplification is intuitively not a severe restriction.
+%
+However, it is possible that we now miss attacks on these methods
+combining properties of the two-term XOR encryption with TEE capabilities.
 %
@@
 We now briefly describe \mTamarin{}, and how we model \mEdhoc{} in it here.
+%
 
@@
 \label{sec:tamarin}
-As in~\cite{Norr21}, we model and verify \mEdhoc{} in the symbolic model using \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an interactive tool for the formal verification of security protocols.
+We extend the formal \mEdhoc{} model of~\cite{Norr21}, using the symbolic
+model tool \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an interactive
+tool for the formal verification of security protocols.
 %
-Protocols are modelled in \mTamarin{} as multi-set rewrite rules which encode a transition relation.
+Protocols are modelled in \mTamarin{} as multiset rewrite rules which encode a
+transition relation.
 %
-The elements of these multi-sets, called facts, contribute to the global system state.
+The elements of these multisets, called facts, contribute to the global
+system state.
 %
@@
 %
-For ease of presentation, we will present the model and properties in a slightly different syntax in this paper, but this syntax can be directly mapped to that of \mTamarin.
+For ease of presentation, we will present the model and properties in a
+slightly different syntax in this paper, but this syntax can be directly
+mapped to that of \mTamarin.
+%
 
@@
 %
-Communicated messages in the protocol are modelled as terms in an algebra, which specifies sets of names, variables, and allowable function symbols.
+Communicated messages in the protocol are modelled as terms in an algebra,
+which specifies sets of names, variables, and allowable function symbols.
 %
-Facts and actions are modelled as $n$-ary predicates in the term algebra, and actions can be parametrized using terms.
+Facts and actions are modelled as $n$-ary predicates in the term algebra,
+and actions can be parametrized using terms.
+%
 
-An annotated multi-set rewrite rule is represented as $l \ifarrow[e] r$, where $l$ and $r$ are multi-sets, and $e$ is a multi-set of actions.
+An annotated multiset rewrite-rule is represented as
+$l \ifarrow[e] r$, where $l$ and $r$ are multisets, and $e$ is
+a multiset of actions.
 %
@@
 %
-By $E^{t}(p_i)_{i\in\mathbb{N}}$, we represent an event of type $E$ occurring at time $t$ in a trace, parametrized by the sequence of values $(p_i)_{i\in\mathbb{N}}$ (corresponding to the action fact $E(p_i)_{i\in\mathbb{N}}@t$ in \mTamarin).
+By $E^{t}(p_i)_{i\in\mathbb{N}}$, we represent an event of type $E$ occurring
+at time $t$ in a trace, parametrized by the sequence of values
+$(p_i)_{i\in\mathbb{N}}$
+(corresponding to the action fact $E(p_i)_{i\in\mathbb{N}}@t$ in \mTamarin).
 %
-Thus, the time points form a quasi order, and we denote the fact that $t_{1}$ comes before $t_{2}$ in a protocol trace by $t_{1} \lessdot t_{2}$, and that $t_{1}$ and $t_{2}$ stand for the same time point in a trace by $t_{1} \doteq t_{2}$.
+Thus, the time points form a quasi order, and we denote the fact that
+$t_{1}$ comes before $t_{2}$ in a protocol trace by
+$t_{1} \lessdot t_{2}$, and that $t_{1}$ and $t_{2}$ stand for the same
+time point in a trace by $t_{1} \doteq t_{2}$.
 %
-\mTamarin{} allows events to occur at the same time point, with one restriction: multiple events of the same type cannot occur simultaneously, so if $t_{1} \doteq t_{2}$, then $E^{t_{1}} = E^{t_{2}}$.
+\mTamarin{} allows events to occur at the same time point, with one
+restriction: multiple events of the same type cannot occur simultaneously,
+so if $t_{1} \doteq t_{2}$, then $E^{t_{1}} = E^{t_{2}}$.
+%
 
-Properties are defined as formulas in a fragment of temporal first order logic, and these formulas can be verified over execution traces.
+Properties are defined as formulas in a fragment of temporal first order logic,
+and these formulas can be verified over execution traces.
+%
 
@@
 %
-For example, to represent the fact that $E$ satisfies the reversal of symmetric encryption by using a decryption operation with the key, one can write $\textit{dec}(\textit{enc}(x, y), y) =_{E} x$. 
+For example, to represent the fact that $E$ satisfies the reversal of
+symmetric encryption by using a decryption operation with the key,
+one can write $\textit{dec}(\textit{enc}(x, y), y) =_{E} x$.
 %
-The equational theory $E$ is fixed upfront to handle the functions supported by the term algebra, so we will omit the subscript for the rest of this paper.
+The equational theory $E$ is fixed upfront to handle the functions supported by
+the term algebra, so we will omit the subscript for the rest of this paper.
 %
 
-Users can extend the default term algebra and equational theory in \mTamarin{} with new function symbols and unification rules for these new symbols.
+Users can extend the default term algebra and equational theory in
+\mTamarin{} with new function symbols and unification rules for these new symbols.
 %
-For example, \mEdhoc{} requires authenticated encryption, which we model using the symbol \mT{aeadEncrypt}. 
+For example, \mEdhoc{} requires authenticated encryption, which we model using
+the symbol \mT{aeadEncrypt}.
 %
-We augment \mTamarin{} with the following rule for this operation, which represents the fact that if the adversary knows a key \mT{k}, a message \mT{m}, and authenticated data \mT{ad}, and has access to an encryption algorithm \mT{ai}, then they can obtain the message corresponding to the authenticated encryption of \mT{m} with \mT{k}~\cite{Norr21}.
+We augment \mTamarin{} with the following rule for this operation,
+which represents the fact that if the adversary knows a key \mT{k},
+a message \mT{m}, and authenticated data \mT{ad},
+and has access to an encryption algorithm \mT{ai},
+then they can obtain the message corresponding to the authenticated encryption
+of \mT{m} with \mT{k}~\cite{Norr21}.
+%
 \begin{verbatim}
@@
 %
-Similarly, we can extend \mTamarin{} with rules for other operations, like \mXor{}.
+We make used of \mTamarin{}'s builtin rules for other operations, like \mXor{}.
+%
 
+The rules modeling a Dolev-Yao adversary and the evolution of their knowledge
+as the protocol executes are built into \mTamarin.
 %
-The rules modelling a Dolev-Yao adversary and the evolution of their knowledge as the protocol executes are built into \mTamarin.
-%
-We can extend and/or modify this behaviour by adding other rules as well.
+We can extend and/or modify this behavior by adding other rules as well.
 %
@@
 %
-This stands for the fact that as the adversary interacts with the parties executing the protocol, at time point $t$, the adversary can derive the term $p$ using the Dolev-Yao deduction rules and any additional adversary capabilities specified in \mTamarin.
+This stands for the fact that as the adversary interacts with the parties
+executing the protocol, at time point $t$, the adversary can derive the
+term $p$ using the Dolev-Yao deduction rules and any additional adversary
+capabilities specified in \mTamarin.
 %
@@
 \]
-which represents the fact that if the adversary gets to know the pair of keys $\langle k, k' \rangle$ at a time point $t$, then the adversary also knows each of those keys $k$ and $k'$ at time point $t$ as well.
+which represents the fact that if the adversary gets to know the pair of
+keys $\langle k, k' \rangle$ at a time point $t$, then the adversary also
+knows each of those keys $k$ and $k'$ at time point $t$ as well.
 %
-For more details about how \mTamarin{} manages adversary knowledge, see~\cite{DBLP:conf/cav/MeierSCB13}.
+For more details about how \mTamarin{} manages adversary knowledge,
+see~\cite{DBLP:conf/cav/MeierSCB13}.
 %
@@
 %-------------------------------------------------------------------------- sub
-
 \subsection{Model and Desired Properties}
 \label{sec:desired-properties}
-In this section, we describe how we model \mEdhoc{} and the various properties we want it to satisfy.
+In this section, we describe how we model \mEdhoc{} and the various properties
+we want it to satisfy.
 %
 
-The party $I$ executing the initiator role considers a run of the protocol begun as soon as it sends the first message \mMsgone{} with event type \mIStart, and considers the run ended once it has sent the third message \mMsgthree{} with event type \mIComplete.
+The party $I$ executing the initiator role considers a run of the protocol
+begun as soon as it sends the first message \mMsgone{} with event type
+\mIStart, and considers the run ended once it has sent the third message
+\mMsgthree{} with event type \mIComplete.
 %
-Similarly, the responder $R$ considers a run started upon receiving \mMsgone{} with event type \mRStart, and finished upon receiving \mMsgthree{} with type \mRComplete.
+Similarly, the responder $R$ considers a run started upon receiving \mMsgone{}
+with event type \mRStart, and finished upon receiving \mMsgthree{} with type
+\mRComplete.
 %
 
+In this work, we consider the following properties:
+perfect forward secrecy, injective agreement for both the initiator and the
+responder, and implicit agreement.
 %
-In this work, we consider the following properties: perfect forward secrecy, injective agreement for both the initiator and the responder, and implicit agreement. 
+Agreement is considered on a set of parameters $S$ which also contains the
+session key material \mSessKey{}.
 %
-Agreement is considered on a set of parameters $S$ which also contains the session key material \mSessKey{}.
+We will first describe in detail both properties, and then describe the
+contents of the set $S$.
 %
-We will first describe in detail both properties, and then describe the contents of the set $S$.
-%
-We formalize these properties as shown in Figure~\ref{fig:props}, which is a modified version of a figure in~\cite{Norr21}.
+We formalize these properties as shown in Figure~\ref{fig:props}, which is
+a modified version of a figure in~\cite{Norr21}.
 \vnote{The figure changes because of TEEs}
@@
 %----------------------------------------------------------------------- PFS
-\subsubsection{Perfect Forward Secrecy (PFS): }
+\subsubsection{Perfect Forward Secrecy (PFS).}
 \label{sec:secrecy}
-Informally, by perfect forward secrecy, we mean that the adversary cannot gain access to the session key material \mSessKey{}, even if a long-term key is leaked after the establishment of \mSessKey{}.
+Informally, by perfect forward secrecy, we mean that the adversary cannot
+gain access to the session key material \mSessKey{}, even if a long-term key is
+leaked after the establishment of \mSessKey{}.
 %
-PFS for session key material \mSessKey{} is formalized as \mPredPfs{} in Figure~\ref{fig:props}.
+PFS for session key material \mSessKey{} is formalized as \mPredPfs{}
+in Figure~\ref{fig:props}.
+%
 
-In the \mIComplete{} event, $I$ represents the identity of the initiator, and $R$ represents the identity of the party who the initiator believes is playing the responder role, while \mSessKey{} stands for the established session key material.
+In the \mIComplete{} event, $I$ represents the identity of the initiator,
+and $R$ represents the identity of the party who the initiator believes is
+playing the responder role, while \mSessKey{} stands for the established
+session key material.
 %
-\mRComplete{} has analogous parameters; in particular, the responder $R$ believes $I$ is the party playing the initiator role.
+\mRComplete{} has analogous parameters; in particular, the responder $R$
+believes $I$ is the party playing the initiator role.
 %
-Essentially, the property states that an adversary obtains \mSessKey{} only if a party's long-term key is compromised before their run ends, or if an ephemeral key is compromised at any time after a party's run began.
-\vnote{This has to change once we add TEEs}
+Intuitively, the PFS property states that an adversary obtains \mSessKey{}
+only if a party's long-term key is compromised before their run ends, if an
+ephemeral key is compromised at any time after a party's run began.
 %
-The property in Figure~\ref{fig:props} is unlike the actual \mTamarin{} lemma in one minor way: \mTamarin's syntax does not allow disjunctions to appear on the left-hand side of an implication inside a universally-quantified formula.
+This is to hold even if the adversary has access to the TEE of either party
+before and after the protocol execution.
 %
-Therefore, in the \mTamarin{} code, instead of using the disjunction $\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$ to model the fact that either party can have completed their execution, we use a single action parametrized by the terms $I$, $R$, and \mSessKey.
+The property in Figure~\ref{fig:props} is unlike the actual \mTamarin{} lemma
+in one minor way: \mTamarin's logic does not allow disjunctions to appear on
+the left-hand side of an implication inside a universally-quantified formula.
+%
+Therefore, in the \mTamarin{} code, instead of using the disjunction
+$\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$
+to model the fact that either party can have completed their execution, we use
+a single action parametrized by the terms $I$, $R$, and \mSessKey.
 
 %-------------------------------------------------------------------- InjAgree
-\subsubsection{Authentication Properties:}
+\subsubsection{Authentication Properties.}
 \label{sec:authenticationDef}
-Following~\cite{Norr21}, we prove two different kinds of authentication properties, namely \emph{injective agreement} in the style of~\cite{DBLP:conf/csfw/Lowe97a}, and implicit agreement.
+Following~\cite{Norr21}, we prove two different kinds of authentication
+properties, namely \emph{injective agreement} in the style
+of~\cite{DBLP:conf/csfw/Lowe97a}, and implicit agreement.
 %
@@
 %
-For the initiator $I$, it guarantees to $I$ that whenever $I$ believes that they have completed a run with $R$ as responder, then the party $R$ has indeed executed the protocol in the role of a responder, and that this run of $I$ uniquely corresponds to one of $R$ where the set of parameters is $S$ and includes, in particular, the session key material \mSessKey{}.
+For the initiator $I$, it guarantees to $I$ that whenever $I$ believes that
+they have completed a run with $R$ as responder, then the party $R$ has indeed
+executed the protocol in the role of a responder, and that this run of $I$
+uniquely corresponds to one of $R$ where the set of parameters is $S$ and
+includes, in particular, the session key material \mSessKey{}.
 %
@@
 %
-We formalize injective agreement for the initiator role as \mPredInjI{} and for the responder role as \mPredInjR{} in Figure~\ref{fig:props}.
+We formalize injective agreement for the initiator role as \mPredInjI{} and
+for the responder role as \mPredInjR{} in Figure~\ref{fig:props}.
 %
-For the initiator $I$, this property represents the fact that either injective agreement (as described above) holds, or the long-term key of the party $R$ assumed to be playing the responder role has been compromised before $I$'s role finished.
+For the initiator $I$, this property represents the fact that either
+injective agreement (as described above) holds, or the long-term key of
+the party $R$ assumed to be playing the responder role has been
+compromised before $I$'s role finished.
 %
-If the adversary has managed to compromise $R$'s long-term key, they can generate any message of their liking and signed it using said key, or computed a $\mathit{MAC}_{R}$ using the long-term keys \mPubE{I}, \mPriv{R}, and an ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$ of their choice.
+The property is to hold even if the adversary has access to the TEE of party $R$
+before and after the protocol execution.
 %
+%If the adversary has managed to compromise $R$'s long-term key, they can
+%generate any message of their liking and sign it using said key,
+%or compute a $\mathit{MAC}_{R}$ using the long-term keys \mPubE{I},
+%\mPriv{R}, and an ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$
+%of their choice.
+%%
 The analogous definition holds for the responder $R$.
-\vnote{Check this, and modify to deal with TEEs}
+%
 
@@
 %
-Thus, we show a different property, a form of implicit agreement on the same set of parameters, which is guaranteed for all methods.
+Thus, we show a different property, a form of implicit agreement on the same
+set of parameters, which is guaranteed for all methods.
 %
-This modification is inspired by the definitions of implicit authentication in the computational model~\cite{DBLP:conf/csfw/GuilhemFW20}.
+This modification is inspired by the definitions of implicit authentication in
+the computational model~\cite{DBLP:conf/csfw/GuilhemFW20}.
 %
-While that paper focuses on authenticating just a key and related identities, our definition encompasses a general set of parameters, as in the notion of injective agreement proposed by Lowe~\cite{DBLP:conf/csfw/Lowe97a}.
+While that paper focuses on authenticating just a key and related identities,
+our definition encompasses a general set of parameters, as in the notion of
+injective agreement proposed by Lowe~\cite{DBLP:conf/csfw/Lowe97a}.
 
-The ``implicit'' in the name of the property stands for the fact that a party $A$ assumes that any party $B$ who has access to the session key material \mSessKey{} must, in fact, be the intended party, and that if $B$ is honest, $B$ will agree on a set $S$ of parameters which includes \mSessKey.
+The ``implicit'' in the name of the property stands for the fact that a party
+$A$ assumes that any party $B$ who has access to the session key material
+\mSessKey{} must, in fact, be the intended party, and that if $B$ is honest,
+$B$ will agree on a set $S$ of parameters which includes \mSessKey.
 %
-Implicit agreement for both roles guarantees to $A$ that $A$ is or has been involved in exactly one protocol execution with $B$, and that $B$ agrees or will agree with $A$ on $S$.
+Implicit agreement for both roles guarantees to $A$ that $A$ is or has been
+involved in exactly one protocol execution with $B$, and that $B$ agrees or
+will agree with $A$ on $S$.
 %
-This property diverges from injective agreement in that upon sending the last message, $A$ concludes that if this message reaches $B$, then $A$ and $B$ agree on each other's identities and roles, and the set $S$.
+This property diverges from injective agreement in that upon sending
+the last message, $A$ concludes that if this message reaches $B$, then $A$
+and $B$ agree on each other's identities and roles, and the set $S$.
+%
 
-Note that for implicit agreement to hold for $I$, the ephemeral keys must not be revealed, since the property relies on the fact that the intended responder is the only one who knows the session key material.
+Note that for implicit agreement to hold for $I$, the ephemeral keys must not
+be revealed since the property relies on the fact that the intended responder
+is the only one who knows the session key material.
 %
-If the adversary has access to the ephemeral keys, they can use it along with the public keys of $I$ and $R$ to compute the session key material.
+If the adversary has access to the ephemeral keys, they can use them along with
+the public keys of $I$ and $R$ to compute the session key material.
 %
-However, either party's long-term key can be revealed after the other party has finished their execution, since this still leaves the adversary unable to compute \mGxy{}.
+However, either party's long-term key can be revealed after the other
+party has finished their execution, since this still leaves the adversary
+unable to compute \mGxy{}.
+%
 
-Since \mTamarin{} runs out of memory to verify this property as is, we split it into two lemmas -- \mPredImpI{} for $I$ for one \mPredImpR{} for $R$.
+Since \mTamarin{} runs out of memory to verify this property as is,
+we split it into two lemmas -- \mPredImpI{} for $I$ for one \mPredImpR{} for $R$.
 %
@@
 \mPredImpR{} is formalized similarly, so we omit it.
+%
 
 %------------------------------------------------------- Agreed parameters
-\subsubsection{Parameters in set $S$:}
+\subsubsection{Parameters in set $S$.}
 \label{sec:agreedParams}
-We now describe the set $S$ of parameters upon which the two parties obtain guarantees via the above properties.
+We now describe the set $S$ of parameters upon which the two parties obtain
+guarantees via the above properties.
 %
@@
 
-Due to the initiator being guaranteed identity protection under \mEdhoc{}, $I$ cannot get explicit agreement with $R$ on the initiator's identity.
+Due to the initiator being guaranteed identity protection under \mEdhoc{}, $I$
+cannot get explicit agreement with $R$ on the initiator's identity.
 %
-Similarly, when using the \mStat{} authentication method, $I$ does not get any such guarantees about $P_{I}$.
+Similarly, when using the \mStat{} authentication method, $I$ does not get any
+such guarantees about $P_{I}$.
 %
-However, $I$ does get implicit agreement with $R$ about $I$'s identity and the full set $S_{F}$ of agreed parameters.
+However, $I$ does get implicit agreement with $R$ about $I$'s identity and the
+full set $S_{F}$ of agreed parameters.
 %
-In contrast, since $R$'s run finishes after that of $I$, $R$ can get explicit injective agreement assurances on the full set $S_{F}$ of agreed parameters.
+In contrast, since $R$'s run finishes after that of $I$, $R$ can get explicit
+injective agreement assurances on the full set $S_{F}$ of agreed parameters.
 %
-The full set of agreed parameters $S_F$ is $S_P \cup \{I, P_I\}$ when $P_I$ is part of the session key material, and $S_P \cup \{I\}$ otherwise.
+The full set of agreed parameters $S_F$ is $S_P \cup \{I, P_I\}$ when $P_I$
+is part of the session key material, and $S_P \cup \{I\}$ otherwise.
 %
 
-In addition to these properties, a couple of properties can be inferred without being explicitly modelled and verified. 
+In addition to these properties, a couple of properties can be inferred
+without being explicitly modelled and verified.
 %
-One might be interested in checking is called Key-Compromise Impersonation (KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
+One such property is Key-Compromise Impersonation
+(KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
 %
-Key-Compromise Impersonation (KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97} often occurs in protocols where a party cannot be assured that the peer they are involved in a run with actually knows the session key material.
+A KCI attack occurs when an adversary who has access to $A$'s long-term private
+key to make $A$ believe that they completed an execution with a peer $B$,
+while $B$ did not participate in said execution at all.
 %
-A KCI attack occurs when an adversary who has access to $A$'s long-term private key to make $A$ believe that they completed an execution with a peer $B$, while $B$ did not participate in said execution at all.
+This is in particular relevant when \mStat{} authentication methods are used.
 %
-Our above notions of agreement ensure that both parties agree on each other's identity, role, and session key material.
+Our above notions of agreement ensure that both parties agree on each
+other's identity, role, and session key material.
 %
-Therefore, all \mEdhoc{} methods that satisfy these agreement properties also avoid KCI attacks.
+Therefore, all \mEdhoc{} methods that satisfy these agreement properties also
+avoid KCI attacks.
+%
 
-Another kind of attack is an Unknown Key Share (UKS)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
+Another kind of attack is Unknown Key Share attacks
+(UKS)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
 %
-As part of a UKS attack, a party $A$ can be made to believe that it finished an execution with a party $B$, but where the session key material is actually shared between $A$ and $C$ instead.
+As part of a UKS attack, a party $A$ can be made to believe that it finished
+an execution with a party $B$, but where the session key material is actually
+shared between $A$ and $C$ instead.
 %
-Again, due to the agreement on identities and session key material, any method that satisfies the above agreement properties also resists UKS attacks.
+Again, due to the agreement on identities and session key material, any method
+that satisfies the above agreement properties also resists UKS attacks.
 %
-Overall, the agreement properties capture entity authentication, and satisfy any properties based on that notion.
-
-\vnote{Is there anything else we can add thanks to the TEEs?}
+Overall, the agreement properties capture entity authentication,
+and satisfy any properties based on that notion.
+%
+However, see Section~\ref{sec:unintendedPeerAuth} for a discussion on the
+interaction between \mEdhoc{} and the application leading to similar issues.
+%
 
@@
 %
-In this section, we describe how we model the \mSigSig, \mSigStat, \mStatSig, and \mStatStat{} methods of \mEdhoc{} in \mTamarin.
+In this section, we describe how we model the \mSigSig, \mSigStat, \mStatSig,
+and \mStatStat{} methods of \mEdhoc{} in \mTamarin.
 %
-As in~\cite{Norr21}, we construct the \mTamarin{} model by utilizing the fact that all methods of \mEdhoc{} share a common underlying structure (as shown in Figure~\ref{fig:edhocFramework}).
+As in~\cite{Norr21}, we construct the \mTamarin{} model by utilizing the fact
+that all methods of \mEdhoc{} share a common underlying structure
+(as shown in Figure~\ref{fig:edhocFramework}).
 %
-We do so by writing a single specification file in the M4 macro language, which generates all the methods.
+We do so by using a single specification file in the M4 macro language,
+which generates all the methods.
 %
-As in~\cite{Norr21}, we only present the \mStatSig{} method which illustrates the use of two different authentication methods.
+As in~\cite{Norr21}, we only present the \mStatSig{} method which illustrates
+the use of two different authentication methods.
 %
@@
 
-As mentioned in Section~\ref{sec:tamarin}, we extend the default equational theory of \mTamarin{} to handle various operations used in \mEdhoc.
+As mentioned in Section~\ref{sec:tamarin}, we extend the default
+equational theory of \mTamarin{} to handle various operations used in \mEdhoc.
 %
-Of the built-in theories, we use the ones for exclusive-or (\mXor), Diffie-Hellman operations, signatures (\mT{sign} and \mT{verify} operations), and hashing~\cite{DBLP:conf/csfw/DreierHRS18,DBLP:conf/csfw/SchmidtMCB12}.
+Of the built-in theories, we use the ones for exclusive-or (\mXor),
+Diffie-Hellman operations, signatures (\mT{sign} and \mT{verify} operations),
+and hashing~\cite{DBLP:conf/csfw/DreierHRS18,DBLP:conf/csfw/SchmidtMCB12}.
 %
-We augment the hashing function symbol to add an extra input, which we use to model different hash functions.
+We augment the hashing function symbol to add an extra input, which we use to
+model different hash functions.
 %
 
-In addition to these default operations, \mEdhoc{} is built over \mHkdfExpand, \mHkdfExtract, and the \mAead{} functions.
+In addition to these default operations, \mEdhoc{} is built over
+\mHkdfExpand, \mHkdfExtract, and the \mAead{} functions.
 %
-We represent \mHkdfExpand by \mT{expa} and \mHkdfExtract by \mT{extr}.
+We represent \mHkdfExpand{} by \mT{expa} and \mHkdfExtract{} by \mT{extr}.
 %
@@
 %
-A term encrypted using \mAead{} is represented by \mT{aeadEncrypt(m, k, ad, ai)}, where \mT{m} is the underlying message, \mT{k} is the encrypting key, \mT{ad} is the additional data, and \mT{ai} is the identifier for the encryption algorithm.
+A term encrypted using \mAead{} is represented by \mT{aeadEncrypt(m, k, ad, ai)},
+where \mT{m} is the underlying message, \mT{k} is the encrypting key,
+\mT{ad} is the additional data, and \mT{ai} is the identifier for the
+encryption algorithm.
 %
-Decryption of such a term is defined via two equations that we add to \mTamarin's theory.
+Decryption of such a term is defined via two equations that we add to
+\mTamarin's theory.
 %
-The following equation requires the decrypting party to know the additional data \mT{ad} to decrypt this encrypted term with verification of its integrity. 
+The following equation requires the decrypting party to know the additional
+data \mT{ad} to decrypt this encrypted term with verification of its integrity.
 \begin{small}
@@
 %
-Only the above equation is used by honest parties, but the adversary should also be able to decrypt without having to go through the additional step of identity verification.
+Only the above equation is used by honest parties, but the adversary should
+also be able to decrypt without having to go through the additional step of
+identity verification.
 %
-To this end, we also add the following equation, where the adversary does not need access to the additional data \mT{ad} in order to decrypt.
+To this end, we also add the following equation, where the adversary does not
+need access to the additional data \mT{ad} in order to decrypt.
+%
 \begin{small}\begin{verbatim}
@@
 
-Having described how we adapt the equational theory to model \mEdhoc, we now move on to the modelling of the adversary model and the environment in which the protocol is executed.
+Having described how we adapt the equational theory to model \mEdhoc,
+we now move on to the modelling of the adversary model and the environment in
+which the protocol is executed.
 %
@@
 %
-We use the following rules to capture the link between a party's identity and their long-term key pairs, in the \mSig{}- and in the \mStat{}-based methods respectively.
+We use the following rules to capture the link between a party's identity and
+their long-term key pairs, in the \mSig{}- and in the \mStat{}-based methods
+respectively.
 \begin{center}
@@
 
-Using the fact \verb|Fr(~ltk)|, \mTamarin{} creates a new term \mT{ltk} and uses it to represent a secret long-term key.
+Using the fact \verb|Fr(~ltk)|, \mTamarin{} creates a new term \mT{ltk} and uses
+it to represent a secret long-term key.
 %
-Via the \verb|Out(<$A, pk(~ltk)>)| fact, \mTamarin{} puts out onto the communication channel the identity of the party to whom this long-term key belongs, along with their public key. 
+Via the \verb|Out(<$A, pk(~ltk)>)| fact, \mTamarin{} puts out onto the
+communication channel the identity of the party to whom this long-term
+key belongs, along with their public key.
 %
-Since the adversary has access to the communication channel, they can pick up all of this information.
+Since the adversary has access to the communication channel, they can pick up
+all of this information.
 %
-The event \mT{UniqLTK} parametrized by a party's identity and their long-term key models the unique correspondence between those two values.
+The event \mT{UniqLTK} parametrized by a party's identity and their long-term
+key models the unique correspondence between those two values.
 %
-As a result, this rules out a party owning multiple long-term keys -- in particular, it keeps the adversary from registering long-term keys in some honest party's name.
+As a result, this rules out a party owning multiple long-term keys
+-- in particular, it keeps the adversary from registering long-term keys in
+some honest party's name.
 %
-This aligns well with an external mechanism such as a certificate authority ensuring that long-term keys are uniquely assigned to the corresponding identities, which is ensured by the \mSpec.
+This aligns well with an external mechanism such as a certificate authority
+ensuring that long-term keys are uniquely assigned to the corresponding
+identities, which is ensured by the \mSpec.
 
-To model the reveal of long-term keys and ephemeral keys to an adversary, we use standard reveal rules and events of type \mRevLTK{} and \mRevEph, respectively. 
+To model the reveal of long-term keys and ephemeral keys to an adversary,
+we use standard reveal rules and events of type \mRevLTK{} and \mRevEph,
+respectively.
 %
@@
 %
-However, ephemeral keys in our model can only be revealed when a party completes their role, i.e. simultaneous with events of type \mIComplete{} and \mRComplete.
+Ephemeral keys in our model can only be revealed when a party
+completes their role, i.e., simultaneously with events of type \mIComplete{}
+and \mRComplete.
 %
-While we did try to reveal ephemeral keys at the start of protocol execution, we could not get the verification to terminate in such a model.
-\vnote{Do we still not get \mTamarin{} to terminate if we reveal ephemeral keys at the start of the protocol? With the TEEs etc?}
+Having set out the capabilities of the adversary, we now model the execution
+of the honest agents' roles.
 %
-Having set out the capabilities of the adversary, we now model the execution of the honest agents' roles.
 
-%%%%%%%%%%% START HERE
-For each protocol method, we use two rules apiece for the initiator and responder -- \mT{I1}, \mT{R2}, \mT{I3}, \mT{R4}.
+For each protocol method, we use two rules apiece for the initiator and
+responder -- \mT{I1}, \mT{R2}, \mT{I3}, \mT{R4}.
 %
@@
 %
-These four rules directly map to the event types \mIStart, \mRStart, \mIComplete, and \mRComplete, respectively.
+These four rules directly map to the event types
+\mIStart, \mRStart, \mIComplete, and \mRComplete, respectively.
 %
-We show the \mT{R2\_STAT\_SIG} rule below to illustrate the various aspects of the \mTamarin{} modelling we are describing here.
+We show the \mT{R2\_STAT\_SIG} rule below to illustrate the various aspects
+of the \mTamarin{} modeling we are describing here.
+%
 
-In order to keep track of the initiator's state, we use facts prefixed with \mT{StI}, which carry information between the \mT{I1} and \mT{I3} rules.
+In order to keep track of the initiator's state, we use facts prefixed with
+\mT{StI}, which carry information between the \mT{I1} and \mT{I3} rules.
 %
-Similarly, for the responder's state, we have \mT{StR} to carry state data between \mT{R2} and \mT{R4}.
+Similarly, for the responder's state, we have \mT{StR} to carry state data
+between \mT{R2} and \mT{R4}.
 %
-In order to link two rules to a state fact, we use \mT{tid}, which is unique to the current session.
+In order to link two rules to a state fact, we use \mT{tid}, which
+is unique to the current session.
 %
 The use of these state facts can be seen in line 28 in the \mT{R2\_STAT\_SIG} rule.
-
-Note that we do not model any error message that $R$ might send in response to message \mMsgone rejecting $I$'s choice of ciphersuite and/or method.
 %
 
-As in~\cite{Norr21}, we use \mTamarin's in-built equational theory for \mXor{}, and model the \mXor{} encryption of \mT{CIPHERTEXT\_2} with the key \mT{K\_2e} by allowing each part of the encrypted term to be ``separately attacked''.
+Note that we do not model any error message that $R$ might send in response
+to message \mMsgone rejecting $I$'s choice of ciphersuite and/or method.
 %
-This means that we first expand \mT{K\_2e} to the same number of key terms as subterms in the plaintext tuple.
+
+As in~\cite{Norr21}, we model the \mXor{} encryption of \mT{CIPHERTEXT\_2} with
+the key \mT{K\_2e} by allowing each part of the encrypted term to be
+separately attacked.
+%
+This means that we first expand \mT{K\_2e} to the same number of key terms as
+subterms in the plaintext tuple.
 %
@@
 %
-This is more faithful to the \mSpec{} than \mXor-ing \mT{K\_2e} on its own with the plaintext tuple.
+This is more faithful to the \mSpec{} than \mXor-ing \mT{K\_2e} on its own
+with the plaintext tuple.
 %
 This can be seen in lines 19--22 in the code for \mT{R2\_STAT\_SIG}.
-\vnote{Add stuff here that changes with the TEE}
+%
+
+As we extended the model with TEEs and augmented the adversary's capabilities
+with access to them, \mTamarin{} failed to complete in a reasonable time for
+some combination of authentication methods and security properties (see XYZ)
+\knote{add reference to discussion section on this}.
+%
+To circumvent the problem, we simplified the \mXor{} encryption to \mXor-ing a
+single term on the entire tuple for yjese cases.
+%
 
@@
 %
-For example, in line 27 above, the action \verb|ExpRunningR(~tid, $V, exp_sk, agreed)| represents an event of type \mRStart{} parametrized by the session id, the responder's identity, and the session key material \mT{exp\_sk}.
+For example, in line 27 above, the action
+\verb|ExpRunningR(~tid, $V, exp_sk, agreed)| represents an event of type
+\mRStart{} parametrized by the session id, the responder's identity, and the
+session key material \mT{exp\_sk}.
 %
-The \mT{exp} in the name of the variable for session key material represents the fact that the agreement property satisfied by this key is explicit, i.e., it includes $P_{I}$, as in Section~\ref{sec:agreedParams}.
+The \mT{exp} in the name of the variable for session key material represents
+the fact that the agreement property satisfied by this key is explicit, i.e.,
+it includes $P_{I}$, as in Section~\ref{sec:agreedParams}.
 %
-We use \mT{imp\_sk} for the corresponding session key material which does not include $P_{I}$.
+We use \mT{imp\_sk} for the corresponding session key material which does not
+include $P_{I}$.
 %
 For the \mSigSig{} and \mSigStat{} methods, therefore, the two values are the same.
+%
 
-%%-------------------------------------------------------------------------- sub
+%-------------------------------------------------------------------------- sub
 %\subsection{Encoding the Desired Properties in \mTamarin}
@@
 
-The properties we listed in Section~\ref{sec:desired-properties} translate directly into \mTamarin's logic.
+The properties we listed in Section~\ref{sec:desired-properties}
+translate directly into \mTamarin's logic.
 %
@@
 Other properties are formalized similarly.
+%
 
@@
 2    all-traces
-3  ``All u v sk #t3 #t2.
-4   (K(sk)@t3 & CompletedRun(u, v, sk)@t2) ==>
+3    "All u v sk tid #t3 #t2.
+4        (K(sk)@t3 & CompletedRun(u, v, sk, tid)@t2) ==>
 5        ( (Ex #t1. LTKRev(u)@t1 & #t1 < #t2)
 6        | (Ex #t1. LTKRev(v)@t1 & #t1 < #t2)
-7       | (Ex #t1. EphKeyRev(sk)@t1))"
+7        | (Ex #t1. EphKeyRev(sk)@t1)
+8        | (Ex m1 #t0 #k. I1(tid, u, v, m1)@t0 & TEE(v)@k & (t0 < k | k < t2))
+9        | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & (t0 < k | k < t2))
+10       )"
 \end{verbatim}
@@
 
-In this formalization, we use the action \mT{CompletedRun(u, v, sk)} (in line 4) to represent the disjunction of the events $\mIComplete^{t_{2}}$ and $\mRComplete^{t_{2}}$.
+In this formalization, we use the action \mT{CompletedRun(u, v, sk)}
+(in line 4) to represent the disjunction of the events $\mIComplete^{t_{2}}$
+and $\mRComplete^{t_{2}}$.
 %
@@
 %
-Similarly, the action \mT{EphKeyRev(sk)} in line 7 stands for the reveal of the ephemeral key for $I$ or $R$ or both.
+Similarly, the action \mT{EphKeyRev(sk)} in line 7 stands for the reveal of
+the ephemeral key for $I$ or $R$ or both.
+%
+Lines 8 and 9 captures that the parties TEEs must be inaccessible to the
+adversary between the start and end of the execution as seen by each party
+respectively.
+%

### 2022-07-21 paper.tex

@@
 \input{formalization}
-
-%-------------------------------------------------------------------------- sec
-\section{\uppercase{Trusted Execution Environments}}
-\label{sec:TEE}
-The main purpose of Trusted Execution Environments (TEE) is to prevent an
-adversary with physical access to a device from accessing the computational
-circuits.
-%
-They come in various forms of differing complexity.
-%
-Some, like ARM Trust Zone and Intel SGX are general purpose execution
-environments, which are flexibly programmable.
-%
-Other, like the Universal Subscriber Identity Modules (USIM) used for
-authentication to 3GPP mobile networks, have application specific interfaces for
-authentication and key agreement protocols etc.
-%
-A key aspect is how much of the application is placed in the TEE and how much is
-outside.
-%
-
-For larger devices that include general purpose processors with Trust Zone or
-SGX, entire \mEdhoc{} and \mOscore{} may reside inside the TEE.
-%
-For constrained IoT devices in the lower end of the scale, a TEE may have to be
-implemented using a special-purpose integrated circuit.
-%
-In the latter case, it may be beneficial to follow a minimalistic approach, and
-contain the long-term key and only
-the operations that need access to it in the TEE for cost reasons.
-%
-
-It may at first appear more secure to implement as much as possible inside the
-TEE, but there is a security trade-off: the more code inside the TEE,
-the higher the risk of implementation errors in the security-critical code.
-%
-Because the security critical code runs in the area where the long-term keys
-reside, an implementation error here risks leaking information of the key to the
-adversary.
-%
-From this perspective, it may be beneficial to follow the minimalistic approach
-even when having access to Trust Zone or SGX.
-%
-
-A slightly more secure division of functionality is to also keep the
-session key inside
-the TEE and extend the interface to accept messages and return the
-(en/de)crypted counter part.
-%
-That is, the interface exposes \mAead{} functions in the interface.
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{Practical Aspects and Use Case}
-\label{sec:TEE:useCase}
-Authentication of a specific IoT device assumes that this devices is the only
-entity with access to the long-term key associated with the corresponding
-identity.
-%
-Since IoT devices may be accessible to adversaries, e.g., an insider cloning a
-key card, the long-term keys must be appropriately protected.
-%
-A state of the art approach is to use a Trusted Execution Environment (TEE),
-which holds the key and provides an API for operations using the key.
-%
-This is the approach taken by the
-Trust-Zone based \mMuEdhoc{}~\cite{DBLP:conf/codaspy/Hristozov0XFLS21}
-implementation for example.
-%
-Typical operations include signatures using the long-term private key of a
-party.
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{System and Adversary Model Extensions Supporting TEE}
-\label{sec:TEE:advModel}
-%
-We extend the adversary model of~\cite{Norr21} by allowing the adversary
-access to a TEE interface on a party of their choosing before and after the
-\mEdhoc{} session.
-%
-This allows us to distinguish between a full compromise and a weak one, using
-the terminology of~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
-%
-A weak compromise is when the adversary have temporary access to the TEE.
-%
-Seen in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add adversary
-capabilities corresponding to a server adversary, i.e., an adversary that
-compromises a party, learns its ephemeral keys and may temporarily access its
-TEE, but does not learn the party's long-term key.
-%
-The model of~\cite{Norr21} only captures server adversary capabilities without
-access to the TEE, via the leakage of ephemeral keys, and local adversary
-capabilities, or more specifically, a secure-world adversary with read-write
-access to the
-secure world where the long-term keys reside via the leakage of long-term keys.
-%
-Our extension allows analysis of so-called weak post-compromise security
-(PCS)~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
-%
-
-We split the \mEdhoc{} functionality as follows.
-%
-The TEE contains the long-term key and allows the non-TEE parts of the
-application to perform the operations on it via its interface.
-%
-More precisely, parties using the \mSig{} authentication method use a TEE with
-an interface which accepts a message and returns the signature of that message
-using the party's private long-term key.
-%
-A party $U$ using the \mStat{} authentication method use a TEE with an interface
-accepting a point $P$ on the curve and returning $\mDH(P,\ \mGu{})$.
-%
-\knote{It should accept any value. Adversary may not stick to elements of the
-    form $g^x$. Or we explicitly state that the TEE checks that the received
-input lies on the curve (but this may not be sufficient: are the curves cyclic
-groups? Need to check).}
-%
-The non-TEE parts implement \mEdhoc{} using this interface.
-%
-This interface requires the least functionality from the TEE, reducing the TEE's
-complexity and possibly cost of its implementation.
-%
-The functional split is suitable even for the situation where a
-constrained device has implemented only the storage of the long-term key in a
-special purpose circuit with minimal processing functionality.
-%
-Since \mEdhoc{} focuses on constrained IoT devices, it seems appropriate to
-cater for this setting.
-%
-
-%-------------------------------------------------------------------------- sub
-\subsection{Extended Formalism and Security Properties}
-\label{sec:TEE:fmAndProps}
-Formally, we model the TEE interface by adding two new rewrite rules:
-%
-\begin{small}
-\begin{verbatim}
-rule forge_SIG:
-   [!LTK_SIG($A, ~ltk), In(xx)] --[TEE($A)]-> [Out(sign(xx, ~ltk))]
-
-rule exp_STAT:
-   [!LTK_STAT($A, ~ltk), In('g'^~xx)] --[TEE($A)]-> [Out(('g'^~xx)^~ltk)]
-\end{verbatim}
-\end{small}
-%
-These rules allow the adversary to obtain terms representing signatures on a
-value of their choice to forge signatures (\verb|forge_SIG|), or to obtain terms
-representing a curve point of their choice raised to the power of the
-long-term key (\verb|exp_STAT|).
-%
-
-Because it is a trivial attack when the adversary access these rules with values
-from the test session, we must disqualify those rule applications.
-%
-We do so by creating an action fact \verb|TEE($A)|, where \verb|$A| is the
-identity corresponding to the private key used, and then augmenting the
-properties with a condition that no such action fact exists from the start of
-the protocol execution and its end.
-%
-Care must be taken when specifying the start and the end.
-%
-Specifically, the start and end of the protocol run must be viewed w.r.t. the
-role.
-%
-For example, the injective agreement property for the initiator in
-the \mSigSig{} method requires that the adversary does not have access to the
-TEE of the
-responder from the time of transmission of the first message until the second
-message is received by the initiator.
-%
-Because reception of one message and transmission of the next one at a party
-is an atomic operation, these timepoints represent the start and end of the
-protocol run from the perspective of the initiator.
-%
-
-We did not manage to get Tamarin to terminate for all lemmas and all methods
-with the added TEE extension.
-%
-When the responder uses the \mStat{} authentication method Tamarin did not
-complete the in a reasonable time for the authentication lemmas.
-%
-To combat this, we simplified the modeling of XOR encryption of the second
-message for these lemmas when the responder uses \mStat{} authentication.
-%
-More precisely, the second message contains the two values \verb|R| (the
-responder's identity) and
-\verb|authR| (the authentication information), and we generally model the
-XOR encryption by xoring each of them with their own key-stream terms.
-%
-For the problematic lemmas however, we resorted to xoring the entire tuple
-\verb|<R, authR>| with a single key-stream term.
-%
-Given that no attacks were identified using our original modeling, this
-simplification is intuitively not a severe restriction.
-%
-However, it is possible that we now miss attacks on these methods
-combining properties of the two-term XOR encryption with TEE capabilities.
-%
 
@@
 
-\subsubsection{Non-Repudiation:}
+\subsubsection{Non-Repudiation.}
 Access control mechanisms can vary depending on what the underlying application is.
@@
 
-\subsubsection{Unintended Peer Authentication: }
+\subsubsection{Unintended Peer Authentication.}
+\label{sec:unintendedPeerAuth}
 In Section~3.2 in the \mSpec{}, the authors state that parties executing \mEdhoc{} must be configured such that they only run \mEdhoc{} with a restricted (according to some reasonable policy) set of peers.
@@
 
-\subsubsection{Denial of Service: }
+\subsubsection{Denial of Service.}
 Denial of Service (DoS) attacks for IoT protocols are very well-studied.

### 2022-07-21 protocol.tex

@@
 \label{sec:TEE}
-Adversaries may have physical access to implementations, this is perhaps an
-even higher risk in an unsupervised IoT setting.
+Authentication of a specific IoT device assumes that this devices is the only
+entity with access to the long-term key associated with the corresponding
+identity.
 %
-Implementations can make use of Trusted Execution Environments (TEE) hardened
-against physical attacks to mitigate adversaries with physical access to the
-cryptographic keys.
+Since IoT devices may be accessible to adversaries, e.g., an insider cloning a
+key card, the long-term keys must be appropriately protected.
+%
+A state of the art approach is to use a Trusted Execution Environment (TEE),
+which holds the key and provides an API for operations using the key.
+%
+This is the approach taken by the
+Trust-Zone based \mMuEdhoc{}~\cite{DBLP:conf/codaspy/Hristozov0XFLS21}
+implementation for example.
+%
+Typical operations include signatures using the long-term private key of a
+party.
 %

### 2022-07-21 paper.tex
[add] noted method type on model [add] table with verification times

@@
 %
-Verification of all lemmas, including model validation lemmas, took 42 minutes
-on an Intel Core i7-6500U 2.5GHz using two cores.
+All measurements were taken on a server with 2 Intel(R) Xeon(R) Gold
+6242 CPU @ 2.80GHz with a total of 32 cores / 64 threads and reserved
+256 GiB of memory for the scheduled job (the machine has a total of
+384 GiB installed).
+%
+Time measurements are rounded to the nearest second and we report both
+real and user times (i.e. time to complete the task and total CPU time
+spent for the task).
 %
@@
 %
-\begin{table*}[h!]
+\begin{table*}[t]
         \centering
-        \caption{Verified properties. $S_P$ contains
-            roles, responder identity, session key material (excluding
+          \label{tab:props}
+          \resizebox{\textwidth}{!}{%
+            \begin{tabular}{l|c|c|c|c}
+              \textbf{Lemma $\backslash$ Method} & \textbf{SIG-SIG} & \textbf{SIG-STAT} & \textbf{STAT-SIG} & \textbf{STAT-STAT}\\
+              \hline
+              \textbf{Injective agreement / I} & $S_F$ & $S_F$ & $S_P\ddagger$ & $S_P\ddagger$\\
+              \textbf{Injective agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
+              \textbf{Time (real/user)} & 0m27s / 11m17s & 9m27s / 239m49s & 1m21s / 35m36s & 18m16s / 499m2s\\
+              \textbf{Implicit agreement / I} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
+              \textbf{Implicit agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
+              \textbf{Time (real/user)} & 2m37s / 65m40s & 5m28s / 131m12s $\dagger$ & 8m54s / 212m21s & 7m37s / 188m46s $\dagger$\\
+              \textbf{Session key secrecy} & \cm & \cm & \cm & \cm\\
+              \textbf{Time (real/user)} & 0m41s / 16m4s & 1m52s / 45m27s & 6m31s / 156m26s & 9m47s / 232m2s\\
+              \textbf{Secrecy of AD3} & \cm & \cm & \cm & \cm\\
+              \textbf{Integrity of AD3} & \cm & \cm & \cm & \cm\\
+              \textbf{Time (real/user)} & 1m40s / 38m57s & 5m53s / 138m43s & 3m35s / 85m12s & 45m6s / 1029m1s\\
+            \end{tabular}}\\[1em]
+
+          \caption{Verified properties. $S_P$ contains roles,
+            responder identity, session key material (excluding
             $P_I$), \mCi, \mCr, and \mSuites. $S_F$ is $S_{P}$,
-            the initiator identity, and $P_I$.}
-        \label{tab:props}
-        \begin{tabular}{|l|c|c|c|c|}
-                \hline
-                & \mSigSig & \mSigStat & \mStatSig & \mStatStat \\
-                \hline
-                Injective agreement for I & $S_F$ & $S_F$ & $S_P$ & $S_P$\\
-                Injective agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-                Implicit agreement for I & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-                Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-                PFS for session key material & \cm & \cm & \cm & \cm\\
-                \hline
-        \end{tabular}
+            the initiator identity, and $P_I$.\\
+            $\dagger$: For SIG-STAT and STAT-STAT implicit
+            authentication using the full XOR model the verification
+            was killed at 14246m43s and 2057m47s of computation time,
+            respectively. The time reported is for verification under
+            the simplified XOR modeling of message 2.\\
+            $\ddagger$: When the initiator is using STAT mode we
+            cannot get explicit agreement on the intiator's own key
+            material, therefore in these two cases we check injective
+            agreement for all key material except $P_I$.\\
+          }
+        % \begin{tabular}{|l|c|c|c|c|}
+        %         \hline
+        %         & \mSigSig & \mSigStat & \mStatSig & \mStatStat \\
+        %         \hline
+        %         Injective agreement for I & $S_F$ & $S_F$ & $S_P$ & $S_P$\\
+        %         Injective agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
+        %         Implicit agreement for I & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
+        %         Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
+        %         PFS for session key material & \cm & \cm & \cm & \cm\\
+        %         \hline
+        % \end{tabular}
 \end{table*}
-\knote{This should be extended with verification times and with TEE.
-    It should also show that when responder uses \mStat{} and \mSig{}, a
-    different XOR encryption is used.
-}
 

### 2022-07-21 paper.tex
[add] updates and fixes to section 6

@@
 %
-In 2018, Bruni et al~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.
+In 2018, Bruni et al.~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.
 %
@@
 %
-It might be helpful to include a flag in \mMsgone to indicate to the responder that a (modified) ciphersuite ordering depends on that from a previous execution.
+It might be helpful to include a flag in \mMsgone to indicate to the responder
+that a (modified) ciphersuite ordering depends on that from a previous
+execution.
 
@@
 %
-We discovered a few of these issues in~\cite{Norr21}, but in this work, we also discuss a couple of new issues.
+We discovered a few of these issues in~\cite{Norr21}, but in this work,
+we also discuss a couple of new issues.
 %
-We communicated the points discussed in~\cite{Norr21} to the authors of the \mSpec{}.
+We communicated the points discussed in~\cite{Norr21} to the authors of the
+\mSpec{}.
 %
-Here we also list the protocol developers' response to being told of these ambiguities in the \mSpec{}.
+Here we also list the protocol developers' response to being told of these
+ambiguities in the \mSpec{}.
 
@@
 %
-However, these goals are informal, and imprecisely specified, and the lack of intended usage makes them difficult to interpret as well.
+However, these goals are informal and imprecisely specified, and the lack of
+intended usage makes them difficult to interpret.
 %
-Since we have no indication about the use cases in which \mEdhoc{} can be employed, it becomes tough to evaluate if the listed security goals are the ones that we need the protocol to satisfy.
+Since we have little indication about use case restrictions, it is also hard
+to evaluate whether the security goals are the most relevant ones.
+%
 
-The overarching goal of \mEdhoc{} is to establish a security context for \mOscore{} using as few roundtrips and as small messages as possible, which dictates the design of the protocol.
+In order to identify security goals that might be relevant
+we made up some user stories corresponding to typical use cases,
+which helped us identify some subtle points of concern in \mEdhoc.
 %
-However, \mEdhoc{} might be intended for multiple use cases, some of which might not even be covered in the \mSpec{}, which might need different guarantees from the ones mentioned therein.
+We communicated these points to the authors of \mEdhoc{}.
 %
-In order to identify security goals that might apply to multiple such cases, we made up some user stories corresponding to typical use cases, which helped us identify some subtle points of concern in \mEdhoc.
+
+%----------------------------------------------------------------------- subsub
+\subsubsection{Non-Repudiation.}
+Access control mechanisms can vary depending on what the underlying application
+is.
 %
-We communicated these issues to the authors of \mEdhoc{}.
+A nuclear power-plant might need to keep track of who enters and leaves, but a
+coffee machine that logs every user and their coffee preference might lead to a
+privacy concern.
 %
-\vnote{Add responses from the developers to these questions}
+This simple thought experiment allowed us to identify that the \mEdhoc{}
+\mSpec{} did not even consider non-repudiation.
+%
+When we pointed this out to the authors of the \mSpec, they recognized this
+concern, and added a discussion about how the various \mEdhoc{} methods
+satisfy (or not) non-repudiation.
+%
 
-\subsubsection{Non-Repudiation.}
-Access control mechanisms can vary depending on what the underlying application is.
-%
-A nuclear power-plant might need to keep track of who enters and leaves, but a coffee machine that logs every user and their coffee preference might lead to a privacy concern.
-%
-This thought experiment allowed us to perceive that the \mEdhoc{} \mSpec{} did not even consider non-repudiation.
-%
-When we pointed this out to the authors of the \mSpec, they recognized this concern, and added a discussion about how the various \mEdhoc{} methods satisfy (or not) non-repudiation.
-
+%----------------------------------------------------------------------- subsub
 \subsubsection{Unintended Peer Authentication.}
 \label{sec:unintendedPeerAuth}
-In Section~3.2 in the \mSpec{}, the authors state that parties executing \mEdhoc{} must be configured such that they only run \mEdhoc{} with a restricted (according to some reasonable policy) set of peers.
+In Section~3.2 in the \mSpec{}, the authors state that parties executing
+\mEdhoc{} must be configured such that they only run \mEdhoc{} with a
+restricted (according to some reasonable policy) set of peers.
 %
-However, \mIdcredr{} in the second message is sent in the clear, and can therefore lead to some masquerade attacks by the adversary.
+However, the responsibilities of verifying which identity was authenticated is
+not clearly split between the application and \mEdhoc{} and therefore an attack
+similar in effect to UKS is possible.
 %
 
-Suppose a person configures the restriction policy such that every device in their home is allowed to set up an \mEdhoc{} session with any other device in the home.
+Suppose a person configures the restriction policy such that every device in
+their home is allowed to set up an \mEdhoc{} session with any other device in
+the home.
 %
-However, the adversary has managed to gain control of one of these devices (say $A$).
+However, the adversary has managed to gain control of one of these devices (say
+$A$).
 %
-Device $A$ is part of the allowed group of devices, so $B$ will accept an \mEdhoc{} run with $A$, even though $A$ is now compromised.
+Device $A$ is part of the allowed group of devices, so $B$ will accept an
+\mEdhoc{} run with $A$, even though $A$ is now compromised.
 %
-In addition, since $B$ does not explicitly check to see if the \mIdcredr{} received in the second message corresponds to the intended identity, it leads to attacks where $B$ wants to talk to another non-compromised home device $C$, but $A$ responds in place of $C$ and $B$ continues with the run anyway.
+If the application in $B$ tasks the \mEdhoc{} implementation to establish a
+security context with $C$, $B$ will send an initial message to $C$.
 %
-This requires the initiator to match \mIdcredr{} to the identity intended by them when they send out the first message.
+However, if $A$ responds to the initial message before $C$ can, the \mSpec{}
+does not require $B$ to verify that \mIdcredr{} matches what is expected for
+$C$.
+%
+A valid \mEdhoc{} implementation may then look up the appropriate credentials
+for $A$, complete the execution and deliver the security context to the
+application.
+%
+The application gets no indication of that the security context is shared with
+$A$ rather that $C$ as intended.
+%
+
+A straight forward check that the received \mIdcredr{} matches the expected
+identity thwarts the problem.
 %
@@
 %
-We communicated this to the authors of \mEdhoc{}.
-\vnote{And what did they do about it?}
-
-\subsubsection{Denial of Service.}
-Denial of Service (DoS) attacks for IoT protocols are very well-studied.
-%
-But DoS is not well documented for \mEdhoc{} in particular, since it delegates DoS handling to the application.
-%
-The \mSpec{} does not include any specific measures for countering such attacks.
-
-Section 8.7 of the \mEdhoc{} specification~\footnote{draft-ietf-lake-edhoc-14} gives two recommendations for countering DoS attacks.
-\vnote{Is this the latest spec? The spec we’re analyzing? Which one?}
-%
-First, \mEdhoc{} relies on lower layers to mitigate DoS, and an example of DoS countermeasures is given: checking the return address upon receipt of the first message.
-%
-Second, an application may try to determine (in some way) that a seemingly valid message is in fact probably a forgery and should be ignored or cause the protocol to abort.
-
-No recommendation is given on how to minimize DoS effects stemming from the use of error messages, which can be used in a relatively arbitrary way in \mEdhoc{} (see Section~\ref{sec:errorHandling}).
-%
-We now elaborate on some considerations that the application developers needs to take into account.
+We communicated this to the authors of \mEdhoc{}, who in the latest version of
+the \mSpec{}, added that \mEdhoc{} makes \mIdcredr{} available to the
+application, and added a new appendix D describing credential validations that
+the application must take care of.
 %
 
-As in the IKEv2 DDoS analysis by~\cite{rfc8019}, the responder is more vulnerable than the initiator.
+%----------------------------------------------------------------------- subsub
+\subsubsection{Denial of Service.}
+Denial of Service (DoS) attacks for IoT protocols are well-studied,
+but DoS aspects particular to \mEdhoc{} is not.
 %
-The key similarity is that the initiator is allowed identity protection in the first message.
+The \mSpec{} does not include any specific measures for countering such attacks,
+but assumes the application takes care of it.
+
+Section 8.7 of the latest version of the \mEdhoc{}
+\mSpec{}~\footnote{draft-ietf-lake-edhoc-15}
+gives two recommendations for countering DoS attacks.
 %
-None of the elements in the first \mEdhoc{} message can be proved by the responder to be illegal, as long as the format is correct and the chosen elliptic curve points lie on the pre-chosen curve.
+First, \mEdhoc{} relies on lower layers to mitigate DoS, and an example of DoS
+countermeasures is given: checking the return address upon receipt of the first
+message.
 %
-Additionally, execution may require the responder to obtain and verify certificate chains.
-%
-For example, the content of the \mADone{} information element may need the responder to take further action.
+Second, a recommendation is given that applications may try to determine (in
+some way) that a seemingly valid message is in fact probably a forgery and
+should be ignored or cause the protocol to abort.
 %
 
-The asymmetry between the adversary's effort towards the first message and that of the responder may hence be significant.
+No recommendation is given on how to minimize DoS effects stemming from the use
+of error messages, which can be used in a relatively arbitrary way in \mEdhoc{}
+(see Section~\ref{sec:errorHandling}).
 %
-Note that since \mEdhoc{} is to be used in conjunction with \mCoap{} and \mOscore{}, either party executing the \mCoap{} protocol might be the one playing the initiator role in \mEdhoc.
+We now elaborate on some considerations that the application developers need to
+take into account.
 %
-Since \mEdhoc{} is designed for constrained devices, this asymmetry needs to be considered even more carefully than for IKEv2~\cite{rfc8019}.
 
-Reflection attacks are another possibility.
+As in the IKEv2 DDoS analysis by~\cite{rfc8019}, the responder in \mEdhoc{}
+is more vulnerable than the initiator.
 %
-These are attacks where an adversary spoofs the source address of a victim in a message and thereby tricks the responder into sending a message to the victim.
+The key similarity is that the initiator enjoys identity protection in the
+first message.
+%
+None of the elements in the first \mEdhoc{} message can be proved by the
+responder to be invalid, as long as the format is correct and the chosen
+elliptic curve points lie on the chosen curve.
+%
+Additionally, execution may require the responder to obtain and verify
+certificate chains.
+%
+For example, the content of the \mADone{} information element may need the
+responder to take further action.
+%
+Since \mEdhoc{} is designed for constrained devices, this asymmetry needs to be
+considered even more carefully than for IKEv2~\cite{rfc8019}.
+%
+
+The asymmetry between the adversary's effort towards the first message and that
+of the responder may hence be significant and the possibility for applications
+to select which entity acts as initiator and which acts as responder may
+important.
+%
+\mEdhoc{} is profiled to work in conjunction with \mCoap{} and
+\mOscore{}~\cite{ietf-core-oscore-edhoc-03}.
+%
+The profiling locks the \mEdhoc{} initiator role to the \mCoap{} client role and
+hence reduces the possibilities to change the \mEdhoc{} roles according to DoS
+vulnerability preferences.
+%
+
+Reflection attacks are another risk.
+%
+These are attacks where an adversary spoofs the source address of a victim in a
+message and thereby tricks the responder into sending a message to the victim.
 %
@@
 %
-Note that even such reachability mechanisms can easily be used to make a responder send reachability requests to arbitrary addressable targets.
+Note that even such reachability mechanisms can easily be used to make a
+responder send reachability requests to arbitrary addressable targets.
 %
-However, such mechanisms might require that unsolicited reachability requests be discarded, limiting the effects of
-anonymous DoS attacks.
+However, such mechanisms might require that unsolicited reachability requests be
+discarded, limiting the effects of anonymous DoS attacks.
 %
-In situations where the responder does not need to perform additional communications to verify certificates etc, a reachability mechanism may be more expensive in terms of time and storage compared to continuing with the protocol while keeping a ``half-open’’ state for a period of time.
+In situations where the responder does not need to perform additional
+communications to verify certificates etc, a reachability mechanism may be more
+expensive in terms of time and storage compared to continuing with the protocol
+while keeping a half-open state for a period of time.
 %
-The latter may cause a systems-level issue if many distributed \mEdhoc{} responders frequently request such servers.
+However, if round-trip times are large, and half-open states hence are kept by
+the receiver for a long time, a reachability mechanism can trade communication
+for storage.
 %
-However, if round-trip times are large, and half-open states hence are kept by the receiver for a long time, a reachability mechanism can trade communication for storage.
 
-An adversary may also send forged error messages (see Section~\ref{sec:errorHandling}), and specifically reject proposed ciphersuites.
+\mEdhoc{} responders querying external servers for certificate revocation checks
+or lookups may casue systems-level issues if many distributed \mEdhoc{}
+responders frequently queries such servers when DoSed by clients.
 %
-It is recommended for the initiator to not try that ciphersuite with the responder again in a new session.
+
+An adversary may also send forged error messages (see
+Section~\ref{sec:errorHandling}), and specifically reject proposed ciphersuites.
 %
-If the adversary rejects all options for a initiator this way, it can prevent
-communication between the initiator and that responder until the cache is
-cleared.
+The \mEdhoc{} \mSpec{} recommends the initiator to not try that ciphersuite with
+the responder again in a new session.
 %
-Potentially, an adversary could lock a party of a group out of communication altogether this way.
+If the adversary rejects all options for an initiator this way, it can prevent
+communication between the initiator and that responder until the initiator's
+cache is cleared.
+%
+Potentially, an adversary could lock a party of a group out of communication
+altogether this way.
 %
@@
 \subsection{Unclear Security Model}
-In addition to unclear contexts in which \mEdhoc{} can be used, the \mSpec{} also does not provide enough detail about adversary capabilities. 
+In addition to unclear contexts in which
+\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail about
+adversary capabilities.
 %
-The four different cryptographic cores included in \mEdhoc{} are based on academic protocols which are designed to work with highly specific but potentially different adversary models.
+The four different cryptographic cores included in \mEdhoc{} are based on
+academic protocols which are designed to work with highly specific but
+potentially different adversary models.
 %
-%For example, \mOptls{}, which essentially has the same core as \mStat{}, is shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, which explicitly separates long-term key storage from that of the ephemeral keys and session state.
+%For example, \mOptls{}, which essentially has the same core as \mStat{}, is
+%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, which
+%explicitly separates long-term key storage from that of the ephemeral keys and
+%session state.
 %%
 %This fits in well with our use of TEEs.
-Since \mSigma{} cannot protect against compromised ephemeral keys, the authors of \mEdhoc{} felt that considering the compromise of ephemeral keys separately from that of long-term keys was not required~\cite{personalCommunication}.
+Since \mSigma{} cannot protect against compromised ephemeral keys, the authors
+of \mEdhoc{} felt that considering the compromise of ephemeral keys separately
+from that of long-term keys was not required~\cite{personalCommunication}.
 %
-The reason was presumably based on the fact that the \mSigSig{} method was modeled closely on the \mSigmaI{} variant of \mSigma{}, and that it would be preferable to obtain a homogeneous security level among the \mEdhoc{}
+The reason was presumably based on the fact that the \mSigSig{} method was
+modeled closely on the \mSigmaI{} variant of \mSigma{}, and that it would be
+preferable to obtain a homogeneous security level among the \mEdhoc{}
 methods.~\cite{Norr21}
 %
-However, this only holds true for situations where one is only interested in the session key confidentiality for an ongoing session.
+However, this only holds true for situations where one is only interested in the
+session key confidentiality for an ongoing session.
 %
-Secure modules, in addition to being able to store long-term keys, are useful in other ways too. 
+Secure modules, in addition to being able to store long-term keys, are useful in
+other ways too.
 %
-They also provide Post-Compromise Security (PCS) guarantees, for example.
+They also provide weak Post-Compromise Security (PCS) guarantees as discussed
+above.
 %
-Upon discussion with the authors, they included recommendations on how to store long-term keys, and how to perform operations on these keys inside a secure module.
-\vnote{This needs more description?}
+Upon discussion with the authors, they included recommendations on how to store
+long-term keys, and how to perform operations on these keys inside a secure
+module.
+%
+
 %-------------------------------------------------------------------------- sub
@@
 \label{sec:sessionKeyMaterial}
-As we have seen, \mEdhoc{} establishes some session key material, which can be fed into the \mEdhoc{}-Exporter to derive session keys.
+As we have seen, \mEdhoc{} establishes some
+session key material, which can be fed into the \mEdhoc{}-Exporter to derive
+session keys.
 %
-This key material is directly influenced by \mGxy{} and a party's secret static long-term key, if they use \mStat{}.
+This key material is directly influenced by \mGxy{} and a party's secret static
+long-term key, if they use \mStat{}.
 %
-As mentioned earlier, mutual explicit injective agreement cannot be obtained for $P_{I}$.
+As mentioned earlier, mutual explicit injective agreement cannot be obtained for
+$P_{I}$.
 %
-This might prove to be a problem, and we propose three potential alternatives for addressing this.
+This might prove to be a problem, and we proposed three potential alternatives
+for addressing this to the \mEdhoc{} authors.
 %
 \begin{itemize}
-\item Include \mIdcredi{} or a hash thereof in the first and second messages: 
+    \item Include \mIdcredi{} or a hash thereof in the first and
+    second messages:
 %
     This increases message size and also leaks the initiator's identity.
+%
     \item Derive session key material without using $P_{I}$:
 %
-This deviates from the structure of protocols like \mOptls{} from which the \mStat{}-based methods draw heavy inspiration.
+    This deviates from the structure of protocols like \mOptls{} from which the
+    \mStat{}-based methods draw heavy inspiration.
 %
-In those protocols, $P_{I}$'s presence in the computation of session key material is crucial to obtaining resistance against the compromise of the initiator's ephemeral key.
-\item Include a fourth message from the responder to the initiator which contains a MAC based on information obtained using $P_{I}$:
+    In those protocols, $P_{I}$'s presence in the computation of session key
+    material is crucial to obtaining resistance against the compromise of the
+    initiator's ephemeral key.
 %
-While this option increases the number of messages in the protocol, the authors decided to go with this option for situations where it is necessary to provide explicit injective agreement on $P_{I}$.
+    \item Include a fourth message from the
+    responder to the initiator which contains a MAC based on information
+    obtained using $P_{I}$:
 \end{itemize}
+%
+While the last option increases the number of messages in the protocol, the
+\mEdhoc{} authors decided to go with this option for situations where it is
+necessary to provide explicit injective agreement on $P_{I}$.
+%
 
@@
 %
-There are newer versions, with the most recent version as
-of February 2021~\cite{latest-ietf-lake-edhoc-05}.
+There are newer versions and our analysis applies to at least the version from
+February 2021~\cite{latest-ietf-lake-edhoc-05}.
 %
-However, the changes to the protocol over these versions are not
-particularly significant for our analysis.
+The latest version is \emph{draft-ietf-lake-edhoc-15}
+%
+\knote{we are already citing too many different versions, so I don't think we
+    should add this to the bib. Potentially, we should tomorrow instead remove
+    and just mention some of the later versions.}
 %

### 2022-07-21 formalization.tex
Minor changes, fixing typos etc

@@
 % !TEX root = paper.tex
-\vnote{Need to rewrite this based on the new spec}
 The \mEdhoc{} \mSpec{} \cite{our-analysis-selander-lake-edhoc-00} claims
@@
 \subsection{Formal Model}\label{sec:threat-model}
-\vnote{Can we add extra references here? SAPIC+ etc? Can leak session keys as well.}
-
 As in~\cite{Norr21}, we verify \mEdhoc{} in an extended
@@
 a modified version of a figure in~\cite{Norr21}.
-\vnote{The figure changes because of TEEs}
-
 \begin{figure*}[ht]
@@
 3        [!LTK_STAT($A, ~ltk),
-4         !PK_STAT($A, 'g'^~ltk),
-5         Out(<$A, 'g'^~ltk>)]
+4         !PK_STAT($A, `g'^~ltk),
+5         Out(<$A, `g'^~ltk>)]
 \end{verbatim}
@@
 with access to them, \mTamarin{} failed to complete in a reasonable time for
-some combination of authentication methods and security properties (see XYZ)
-\knote{add reference to discussion section on this}.
+some combination of authentication methods and security properties (see Section~\ref{sec:conclusion} for a detailed discussion).
 %
 To circumvent the problem, we simplified the \mXor{} encryption to \mXor-ing a
-single term on the entire tuple for yjese cases.
+single term on the entire tuple for these cases. \\
 %
@@
 }
+\vspace{5mm}
 
@@
 %
-Other properties are formalized similarly.
+Other properties are formalized similarly. \pagebreak
 %

### 2022-07-21 protocol.tex
[add] updates and fixes to section 6
[fix] typos, wrong parenthesis placement in lemmas [add] tee events in lemmas

@@
 %
-This is similar to the 1-RTT semi-static pattern in \mOptls which computes the
+This is similar to the 1-RTT semi-static pattern in \mOptls{} which computes the
 MAC key \textsf{sfk} for the message
@@
 key derivations used to construct these messages.
-\vnote{Do we still have the other method diagrams that we wrote out for the original paper but did not include?}
-\knote{I don't think we have space for that.}
-
+%\vnote{Do we still have the other method diagrams that we wrote out for the original paper but did not include?}
+%\knote{I don't think we have space for that.}
+%
 \begin{figure}[ht]

### 2022-07-21 paper.tex

@@
 Time measurements are rounded to the nearest second and we report both
-real and user times (i.e. time to complete the task and total CPU time
+the real time (time to complete the task) and user time (total CPU time
 spent for the task).
 %
-Mutual entity authentication, UKS- and KCI resistance can be inferred
+Mutual entity authentication, UKS- and KCI-resistance can be inferred
 from the verified properties.
-\knote{In Section 3.3 we say that injective agreement does \emph{not} hold for
-    all methods. It is a little contradictory to the table below where we
-    exclude $P_I$. It would be better to show it as not holding when $P_i$ is
-    included (both for consistency and for clarity).
-}
 %
@@
         \centering
-          \label{tab:props}
           \resizebox{\textwidth}{!}{%
@@
             \end{tabular}}\\[1em]
-
           \caption{Verified properties. $S_P$ contains roles,
@@
           }
+                    \label{tab:props}
         % \begin{tabular}{|l|c|c|c|c|}
@@
 %
-The latest version is \emph{draft-ietf-lake-edhoc-15}
+The latest version is \emph{draft-ietf-lake-edhoc-15}.
 %
+
 \knote{we are already citing too many different versions, so I don't think we

### 2022-07-22 formalization.tex
[fix] typos, wrong parenthesis placement in lemmas [add] tee events in lemmas
Minor changes. Added hlines to Table 2

@@
 %
-We resolve this by clearly specifying an adversary model, in which we can verify
+We resolve this by clearly specifying an adversary model, in which we can 
+verify
 properties.
@@
 
-The adversary is assumed to be the network, and thus controls the communication
+The adversary is assumed to be the network, and thus controls the 
+communication
 channel and can see all messages being communicated as part of the protocol.
@@
 On top of the Dolev-Yao model,~\cite{Norr21} allow the adversary two extra
-capabilities with respect to getting access to the long-term and ephemeral keys.
+capabilities with respect to getting access to the long-term and ephemeral 
+keys.
 %
 Long-term key reveal, denoted $\mRevLTK^{t}(A)$, stands for the adversary
-gaining access to a party $A$'s long-term private key \mPriv{A} at a time point $t$.
+gaining access to a party $A$'s long-term private key \mPriv{A} at a time point 
+$t$.
 %
@@
 % but long-term keys are often stored and operate within a secure model,
-% like a trusted execution environment (TEE). \vnote{Do we want to say more about TEEs here?}
+% like a trusted execution environment (TEE). \vnote{Do we want to say more 
+% about TEEs here?}
 %
@@
 %
-This allows us to distinguish between a full compromise and a weak one, using
+This allows us to distinguish between a full compromise and a weak one, 
+using
 the terminology of~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
@@
 %
-Seen in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add adversary
+Seen in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add 
+adversary
 capabilities corresponding to a server adversary, i.e., an adversary that
@@
 access to the
-secure world where the long-term keys reside via the leakage of long-term keys.
+secure world where the long-term keys reside via the leakage of long-term 
+keys.
 %
@@
 %
-More precisely, parties using the \mSig{} authentication method use a TEE with
-an interface which accepts a message and returns the signature of that message
+More precisely, parties using the \mSig{} authentication method use a TEE 
+with
+an interface which accepts a message and returns the signature of that 
+message
 using the party's private long-term key.
 %
-A party $U$ using the \mStat{} authentication method use a TEE with an interface
+A party $U$ using the \mStat{} authentication method use a TEE with an 
+interface
 accepting a point $P$ on the curve and returning $\mDH(P,\ \mGu{})$.
@@
 %
-This interface requires the least functionality from the TEE, reducing the TEE's
+This interface requires the least functionality from the TEE, reducing the 
+TEE's
 complexity and possibly cost of its implementation.
@@
 The functional split is suitable even for the situation where a
-constrained device has implemented only the storage of the long-term key in a
+constrained device has implemented only the storage of the long-term key in 
+a
 special purpose circuit with minimal processing functionality.
@@
 
-We did not manage to get Tamarin to terminate for all lemmas and all methods
+We did not manage to get Tamarin to terminate for all lemmas and all 
+methods
 with the added TEE extension.
@@
 
-We now briefly describe \mTamarin{}, and how we model \mEdhoc{} in it here.
+We now briefly describe \mTamarin{}, and how we model \mEdhoc{} in it 
+here.
 %
@@
 We extend the formal \mEdhoc{} model of~\cite{Norr21}, using the symbolic
-model tool \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an interactive
+model tool \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an 
+interactive
 tool for the formal verification of security protocols.
 %
-Protocols are modelled in \mTamarin{} as multiset rewrite rules which encode a
+Protocols are modelled in \mTamarin{} as multiset rewrite rules which encode 
+a
 transition relation.
@@
 %
-Event types are predicates over global states generated during protocol execution.
+Event types are predicates over global states generated during protocol 
+execution.
 %
@@
 %
-The equational theory $E$ is fixed upfront to handle the functions supported by
+The equational theory $E$ is fixed upfront to handle the functions supported 
+by
 the term algebra, so we will omit the subscript for the rest of this paper.
@@
 Users can extend the default term algebra and equational theory in
-\mTamarin{} with new function symbols and unification rules for these new symbols.
+\mTamarin{} with new function symbols and unification rules for these new 
+symbols.
 %
-For example, \mEdhoc{} requires authenticated encryption, which we model using
+For example, \mEdhoc{} requires authenticated encryption, which we model 
+using
 the symbol \mT{aeadEncrypt}.
@@
 and has access to an encryption algorithm \mT{ai},
-then they can obtain the message corresponding to the authenticated encryption
+then they can obtain the message corresponding to the authenticated 
+encryption
 of \mT{m} with \mT{k}~\cite{Norr21}.
@@
 %
-We make used of \mTamarin{}'s builtin rules for other operations, like \mXor{}.
+We make used of \mTamarin{}'s builtin rules for other operations, like 
+\mXor{}.
 %
 
-The rules modeling a Dolev-Yao adversary and the evolution of their knowledge
+The rules modeling a Dolev-Yao adversary and the evolution of their 
+knowledge
 as the protocol executes are built into \mTamarin.
@@
 %
-To denote that the adversary has access to a message $p$ at time $t$, we use $\mK^{t}(p)$.
+To denote that the adversary has access to a message $p$ at time $t$, we use 
+$\mK^{t}(p)$.
 %
-This stands for the fact that as the adversary interacts with the parties
+This 
+%stands for the fact that as the adversary interacts with the parties
 executing the protocol, at time point $t$, the adversary can derive the
@@
 \label{sec:desired-properties}
-In this section, we describe how we model \mEdhoc{} and the various properties
+In this section, we describe how we model \mEdhoc{} and the various 
+properties
 we want it to satisfy.
@@
 %
-Similarly, the responder $R$ considers a run started upon receiving \mMsgone{}
-with event type \mRStart, and finished upon receiving \mMsgthree{} with type
+Similarly, the responder $R$ considers a run started upon receiving 
+\mMsgone{}
+with event type \mRStart, and finished upon receiving \mMsgthree{} with 
+type
 \mRComplete.
@@
     \mK^{t_3}(\mSessKey)\  \land\ 
-    (\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(I, R, \mSessKey))
+    (\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(I, R, 
+\mSessKey))
     \rightarrow\\
@@
     \mPredInjI \triangleq\ &
-    \forall I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, \mSessKey, S)
+    \forall I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, 
+\mSessKey, S)
     \rightarrow\\
-    &(\exists t_1\mLogicDot \mRStart^{t_1}(R, \mSessKey, S) \land t_1 \lessdot t_2)
+    &(\exists t_1\mLogicDot \mRStart^{t_1}(R, \mSessKey, S) \land t_1 \lessdot 
+t_2)
     \land (\forall I' R' t_1' \mLogicDot \mIComplete^{t_1'}(I' , R', \mSessKey, S)
@@
     \mPredInjR \triangleq\ &
-    \forall I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(I, R, \mSessKey, S)
+    \forall I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(I, R, 
+\mSessKey, S)
     \rightarrow\\
-    &(\exists t_1\mLogicDot \mIStart^{t_1}(I, R, \mSessKey, S) \land t_1 \lessdot t_2)
+    &(\exists t_1\mLogicDot \mIStart^{t_1}(I, R, \mSessKey, S) \land t_1 \lessdot 
+t_2)
     \land (\forall I' R' t_1' \mLogicDot \mRComplete^{t_1'}(I' , R', \mSessKey, S)
@@
     \mPredImpI \triangleq\ &
-    \forall I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(I, R, \mSessKey, S)
+    \forall I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(I, R, \mSessKey, 
+S)
     \rightarrow\\
-      &(\forall I', R', S', t_2\mLogicDot \mRComplete^{t_2}(I', R', \mSessKey, S') \rightarrow
+      &(\forall I', R', S', t_2\mLogicDot \mRComplete^{t_2}(I', R', \mSessKey, S') 
+\rightarrow
              (I=I' \land R=R' \land S=S'))\\
@@
 Therefore, in the \mTamarin{} code, instead of using the disjunction
-$\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$
+$\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, 
+\mSessKey)$
 to model the fact that either party can have completed their execution, we use
@@
 For the initiator $I$, it guarantees to $I$ that whenever $I$ believes that
-they have completed a run with $R$ as responder, then the party $R$ has indeed
+they have completed a run with $R$ as responder, then the party $R$ has 
+indeed
 executed the protocol in the role of a responder, and that this run of $I$
@@
 %or compute a $\mathit{MAC}_{R}$ using the long-term keys \mPubE{I},
-%\mPriv{R}, and an ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$
-%of their choice.
+%\mPriv{R}, and an ephemeral key pair $\langle\mPrivE{R},\ 
+%\mPubE{R}\rangle$
+%of 
+%their choice.
 %%
@@
 %
-However, this is not the case for the injective agreement property as stated above.
+However, this is not the case for the injective agreement property as stated 
+above.
 %
@@
 %
-If the adversary has access to the ephemeral keys, they can use them along with
+If the adversary has access to the ephemeral keys, they can use them along 
+with
 the public keys of $I$ and $R$ to compute the session key material.
@@
 Since \mTamarin{} runs out of memory to verify this property as is,
-we split it into two lemmas -- \mPredImpI{} for $I$ for one \mPredImpR{} for $R$.
+we split it into two lemmas -- \mPredImpI{} for $I$ for one \mPredImpR{} for 
+$R$.
 %
@@
     \item responder identity,
-    \item session key material (which varies depending on the \mEdhoc{} method),
+    \item session key material (which varies depending on the \mEdhoc{} 
+method),
     \item context identifiers \mCi{} and \mCr{}, and
@@
 %
-Similarly, when using the \mStat{} authentication method, $I$ does not get any
+Similarly, when using the \mStat{} authentication method, $I$ does not get 
+any
 such guarantees about $P_{I}$.
@@
 %
-A KCI attack occurs when an adversary who has access to $A$'s long-term private
+A KCI attack occurs when an adversary who has access to $A$'s long-term 
+private
 key to make $A$ believe that they completed an execution with a peer $B$,
@@
 %
-Again, due to the agreement on identities and session key material, any method
+Again, due to the agreement on identities and session key material, any 
+method
 that satisfies the above agreement properties also resists UKS attacks.
@@
 %
-The full \mTamarin{} code for all models can be found at~\cite{edhocTamarinRepo}.
+The full \mTamarin{} code for all models can be found 
+at~\cite{edhocTamarinRepo}.
 %
@@
 As mentioned in Section~\ref{sec:tamarin}, we extend the default
-equational theory of \mTamarin{} to handle various operations used in \mEdhoc.
+equational theory of \mTamarin{} to handle various operations used in 
+\mEdhoc.
 %
@@
 Diffie-Hellman operations, signatures (\mT{sign} and \mT{verify} operations),
-and hashing~\cite{DBLP:conf/csfw/DreierHRS18,DBLP:conf/csfw/SchmidtMCB12}.
+and 
+hashing~\cite{DBLP:conf/csfw/DreierHRS18,DBLP:conf/csfw/SchmidtMCB12}.
 %
-We augment the hashing function symbol to add an extra input, which we use to
+We 
+%augment the hashing function symbol to add an extra input, which we use 
+to
 model different hash functions.
@@
 %
-We represent \mHkdfExpand{} by \mT{expa} and \mHkdfExtract{} by \mT{extr}.
+We represent \mHkdfExpand{} by \mT{expa} and \mHkdfExtract{} by 
+\mT{extr}.
 %
@@
 %
-A term encrypted using \mAead{} is represented by \mT{aeadEncrypt(m, k, ad, ai)},
+A term encrypted using \mAead{} is represented by \mT{aeadEncrypt(m, k, ad, 
+ai)},
 where \mT{m} is the underlying message, \mT{k} is the encrypting key,
@@
 Having described how we adapt the equational theory to model \mEdhoc,
-we now move on to the modelling of the adversary model and the environment in
+we now move on to the modelling of the adversary model and the 
+environment in
 which the protocol is executed.
 %
-We extend the built-in Dolev-Yao adversary rules which are part of \mTamarin.
+We extend the built-in Dolev-Yao adversary rules which are part of 
+\mTamarin.
 %
@@
 
-Using the fact \verb|Fr(~ltk)|, \mTamarin{} creates a new term \mT{ltk} and uses
+Using the fact \verb|Fr(~ltk)|, \mTamarin{} creates a new term \mT{ltk} and 
+uses
 it to represent a secret long-term key.
@@
 %
-Since the adversary has access to the communication channel, they can pick up
+Since the adversary has access to the communication channel, they can pick 
+up
 all of this information.
 %
-The event \mT{UniqLTK} parametrized by a party's identity and their long-term
-key models the unique correspondence between those two values.
+The event \mT{UniqLTK} parametrized by a party's identity and their 
+long-term
+key 
+models the unique correspondence between those two values.
 %
@@
 %
-It is also important to keep track of the time points at which these events occur.
+It is also important to keep track of the time points at which these events 
+occur.
 %
-Long-term keys can be revealed on registration, even before the protocol begins.
+Long-term keys can be revealed on registration, even before the protocol 
+begins.
 %
@@
 %
-The use of these state facts can be seen in line 28 in the \mT{R2\_STAT\_SIG} rule.
+The use of these state facts can be seen in line 28 in the \mT{R2\_STAT\_SIG} 
+rule.
 %
@@
 
-As in~\cite{Norr21}, we model the \mXor{} encryption of \mT{CIPHERTEXT\_2} with
+As in~\cite{Norr21}, we model the \mXor{} encryption of 
+\mT{CIPHERTEXT\_2} with
 the key \mT{K\_2e} by allowing each part of the encrypted term to be
@@
 %
-This means that we first expand \mT{K\_2e} to the same number of key terms as
+This means that we first expand \mT{K\_2e} to the same number of key terms 
+as
 subterms in the plaintext tuple.
@@
 
-As we extended the model with TEEs and augmented the adversary's capabilities
+As we extended the model with TEEs and augmented the adversary's 
+capabilities
 with access to them, \mTamarin{} failed to complete in a reasonable time for
-some combination of authentication methods and security properties (see Section~\ref{sec:conclusion} for a detailed discussion).
+some combination of authentication methods and security properties (see 
+Section~\ref{sec:conclusion} for a detailed discussion).
 %
-To circumvent the problem, we simplified the \mXor{} encryption to \mXor-ing a
+To circumvent the problem, we simplified the \mXor{} encryption to 
+\mXor-ing 
+a
 single term on the entire tuple for these cases. \\
@@
 %
-We use \mT{imp\_sk} for the corresponding session key material which does not
+We use \mT{imp\_sk} for the corresponding session key material which does 
+not
 include $P_{I}$.
 %
-For the \mSigSig{} and \mSigStat{} methods, therefore, the two values are the same.
+For the \mSigSig{} and \mSigStat{} methods, therefore, the two values are the 
+same.
 %
@@
 %Section~\ref{sec:desired-properties} translate directly into \mTamarin's logic,
-%using the straightforward mapping of events to the actions emitted from the model.
+%using the straightforward mapping of events to the actions emitted from the 
+%model.
 %
@@
 %
-Other properties are formalized similarly. \pagebreak
+Other properties are formalized similarly. 
 %
@@
 2    all-traces
-3    "All u v sk tid #t3 #t2.
+3    ``All u v sk tid #t3 #t2.
 4        (K(sk)@t3 & CompletedRun(u, v, sk, tid)@t2) ==>

### 2022-07-22 intro.tex
Minor changes. Added hlines to Table 2

@@
 % !TEX root = paper.tex
-IoT protocols are often run on devices which operate under severe restrictions on resources like bandwidth and energy consumption.
+IoT protocols are often run on devices which operate under severe restrictions 
+on resources like bandwidth and energy consumption.
 %
-These constrained devices are often simple in their operation, but need to communicate and function without human interference or maintenance for extended periods of time.
+These constrained devices are often simple in their operation, but need to 
+communicate and function without human interference or maintenance for 
+extended periods of time.
 %
-IETF is standardizing new protocols to secure communications between constrained devices.
+IETF is standardizing new protocols to secure communications between 
+constrained devices.
 %
-One such is the Object Security for Constrained RESTful Environments (\mOscore{}) protocol. 
+One such is the Object Security for Constrained RESTful Environments 
+(\mOscore{}) protocol. 
 %
@@
 %
-To this end, a key exchange protocol named Ephemeral Diffie-Hellman Over COSE
+To this end, a key exchange protocol named Ephemeral Diffie-Hellman Over 
+COSE
 (\mEdhoc{}) is under discussion in the IETF. 
 %
-Since \mEdhoc{} will establish security contexts for \mOscore{}, the same resource constraints (especially those pertaining to message size) apply to the former as for the latter.
+Since \mEdhoc{} will establish security contexts for \mOscore{}, the same 
+resource constraints (especially those pertaining to message size) apply to the 
+former as for the latter.
 %
-While establishing security contexts for \mOscore{} is the primary goal for the \mEdhoc{} protocol, there might well be other use cases, which have not been explored in depth yet.
+While establishing security contexts for \mOscore{} is the primary goal for the 
+\mEdhoc{} protocol, there might well be other use cases, which have not been 
+explored in depth yet.
 %
-It is therefore important to ensure that most of the fundamental properties expected of key exchange protocols as established in the literature are satisfied by \mEdhoc{} as well.
+It is therefore important to ensure that most of the fundamental properties 
+expected of key exchange protocols as established in the literature are satisfied 
+by \mEdhoc{} as well.
 %
@@
 %
-It allowed two different key establishment methods -- one involved a pre-shared Diffie-Hellman (DH) \emph{cryptographic core}, and the other was a variation on challenge-response signatures, {\`a} la \mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
+It allowed two different key establishment methods -- one involved a 
+pre-shared 
+Diffie-Hellman (DH) \emph{cryptographic core}, and the other was a 
+variation on challenge-response signatures, {\`a} la 
+\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
+
 %
 A \emph{cryptographic core}, often just called a core, is an academic protocol,
-i.e., with no encodings or application-specific details as needed for an industrial protocol. 
+i.e., with no encodings or application-specific details as needed for an 
+industrial protocol. 
 %
-Once these ingredients are added to a cryptographic core, we obtain a key-establishment method.
+Once these ingredients are added to a cryptographic core, we obtain a 
+key-establishment 
+method.
 %
@@
 one based on \mSigma{}
-(SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}, and in
-2020, three new cores, which mixed challenge-response signatures and regular signatures were added as well~\cite{our-analysis-selander-lake-edhoc-00}.
+(SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}, and 
+in
+2020, three new cores, which mixed challenge-response signatures and regular 
+signatures were added as well~\cite{our-analysis-selander-lake-edhoc-00}.
 
@@
 \label{sub:related}
-The earliest related work is~\cite{DBLP:conf/secsr/BruniJPS18}, which formally analyzes the May 2018 version of \mEdhoc{} using the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01}. 
+The earliest related work is~\cite{DBLP:conf/secsr/BruniJPS18}, which formally 
+analyzes the May 2018 version of \mEdhoc{} using the \mProverif{} 
+tool~\cite{DBLP:conf/csfw/Blanchet01}. 
 %
-In this paper, the authors analyze two key establishment methods -- one built on
+In this paper, the authors analyze two key establishment methods -- one built 
+on
 a pre-shared key authentication core, and one based on \mSigma{}. 
 %
-The authors check  various properties, namely secrecy, identity protection against an active adversary, strong authentication, perfect forward secrecy (PFS), and integrity of application data. 
+The authors check  various properties, namely secrecy, identity protection 
+against an active adversary, strong authentication, perfect forward secrecy 
+(PFS), and integrity of application data. 
 %
-Later work~\cite{Norr21} analyze the May 2020 version of \mEdhoc{} in the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13}. 
+Later work~\cite{Norr21} analyze the May 2020 version of \mEdhoc{} in the 
+\mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13}. 
 %
@@
 %
-In~\cite{Norr21}, the authors check for injective agreement, implicit agreement, and perfect forward secrecy for the session key material. 
+In~\cite{Norr21}, the authors check for injective agreement, implicit 
+agreement, and perfect forward secrecy for the session key material. 
 %
-The authors also discuss some fallouts of the various design choices made as part of \mEdhoc{}, and the impact of \mEdhoc{} in multiple use-case scenarios.
-%
-\vnote{Need to check this once more later. Also need to see if we can say something about Steve's SAPIC+ paper. CAREFULLY check what follows, and fill in the list of properties!!}
-%
+The authors also discuss some fallouts of the various design choices made as 
+part of \mEdhoc{}, and the impact of \mEdhoc{} in multiple use-case scenarios.
+%%
 
@@
 %
-We analyze the version of \mEdhoc{} as in the latest draft~\cite{}.
-\knote{possibly}
+We analyze the version of \mEdhoc{} as in~\cite{}.
 %
 We extend the adversary model and \mTamarin{} models to capture weak
-post-compromise security (PCS) and modeling of Trusted Execution Environments
+post-compromise security (PCS) and model Trusted Execution Environments
 (TEE).
 %
-We check for the following properties: \vnote{Fill this in}
+We also modify the model to consider a stronger version of the \mXor{} encryption whenever possible. 
 %
-We also consider various issues arising in relation to the use of trusted execution environments (TEEs), denial of service (DoS) attacks, sources of randomness, the negotiation of parameters (which the formal model abstracts away) for establishing the protocol, and other potential attacks and concerns.
+We check for the following properties: 
+\begin{itemize}
+\item Perfect forward secrecy for the session key material
+\item Injective agreement
+\item Implicit agreement
+\item Weak post-compromise security
+\item Secrecy and integrity of \mADthree
+\end{itemize}
+\vnote{Check this!}
 %
+We also discuss various issues arising in relation to the use of trusted 
+execution environments (TEEs), denial of service (DoS) attacks, sources of 
+randomness, the negotiation of parameters (which the formal model abstracts 
+away) for establishing the protocol, and other potential attacks and concerns.
+%
+We have communicated these issues to the developers of the protocol.
 
 %In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
-%establishment methods) using the \mTamarin{} tool~\cite{DBLP:conf/cav/MeierSCB13}.
+%establishment methods) using the \mTamarin{} 
+%tool~\cite{DBLP:conf/cav/MeierSCB13}.
 %%
-%We present a formal model we constructed of the protocol as given in the
+%We 
+%present a formal model we constructed of the protocol as given in the
 %\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00}.
@@
 %We give an explicit adversary model for the protocol and verify
-%properties such as session key material and entity authentication, and perfect
+%properties such as session key material and entity authentication, and 
+%perfect
 %forward secrecy, for all four methods.
@@
 %We show that not all \mEdhoc{}'s key establishment
-%methods provide authentication according to the injective agreement definition
+%methods provide authentication according to the injective agreement 
+%definition
 %on the session key material, and none on the initiator's identity.
@@
 %%
-%For example, the authentication policy requirements allows situations where a
-%party establishes session key material with a trusted but compromised peer, even
+%For example, the authentication policy requirements allows situations where 
+%a
+%party establishes session key material with a trusted but compromised peer, 
+%even
 %though the intention was to establish it with a different trusted party.
@@
 %We provide remedies for the identified issues and have
-%communicated these to the IETF and the specification authors, who incorporated
-%some of our suggestions and currently consider how to deal with the remaining
+%communicated these to the IETF and the specification authors, who 
+%incorporated
+%some of our suggestions and currently consider how to deal with the 
+%remaining
 %ones.
@@
 %
-%In contrast to the key establishment methods analyzed by Bruni et~al., which
-%were based on the well-understood pre-shared key DH and \mSigma{} protocols,
+%In contrast to the key establishment methods analyzed by Bruni et~al., 
+%which
+%were based on the well-understood pre-shared key DH and \mSigma{} 
+%protocols,
 %the three newly added
@@
 %%
-%Combining two protocols, which individually provide unilateral authentication,
+%Combining two protocols, which individually provide unilateral 
+%authentication,
 %is not guaranteed to result in a secure mutual authentication
@@
 %%
-%Consequently, even though the framework is similar to the one analyzed by Bruni
+%Consequently, even though the framework is similar to the one analyzed by 
+%Bruni
 %et~al., the cryptographic underpinnings have significantly increased in
-%complexity, and is using mechanisms which have not previously been formally analyzed.
+%complexity, and is using mechanisms which have not previously been 
+%formally analyzed.
 %%

### 2022-07-22 paper.tex
Minor changes. Added hlines to Table 2

--- 2022-07-21 paper.tex
@@
 
-\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices}
+\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT 
+Devices}
 \author{
@@
 \begin{abstract}
-Given how common IoT devices that use constrained resources are becoming today, the need of the hour is communication protocols which can operate securely under such limitations.
+Given how common IoT devices that use constrained resources are becoming 
+today, the need of the hour is communication protocols which can operate 
+securely under such limitations.
 %
-For a few years, the Internet Engineering Task Force (IETF) has been working to standardize \mEdhoc{}, an authenticated key establishment protocol for such constrained IoT devices.
+For a few years, the Internet Engineering Task Force (IETF) has been working 
+to standardize \mEdhoc{}, an authenticated key establishment protocol for 
+such constrained IoT devices.
 %
@@
 %
-In 2018, Bruni et al.~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.
+In 2018, Bruni et al.~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} 
+tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of 
+\mEdhoc{}, which had only two key establishment methods.  
 %
-%By now, the protocol has been fleshed out much more, and has been augmented with multiple new key establishment methods.
-By 2021, the protocol had been fleshed out much more, with multiple new key establishment methods, and this version was formally analyzed using the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
+%By now, the protocol has been fleshed out much more, and has been 
+%augmented with multiple new key establishment methods.
+By 2021, the protocol had been fleshed out much more, with multiple new key 
+establishment methods, and this version was formally analyzed using the 
+\mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
 %
-Here, we build on this work, and use \mTamarin{} to formally analyze the key establishment methods in the current version of \mEdhoc{}, as well as discuss some ramifications of the choices made while designing the protocol.
-\vnote{We need to see what more we can say here, about the new stuff we're potentially adding}
+Here, we build on this work, by modifying the model, adding new properties 
+that we analyze, and discussing some of the aspects of the latest \mEdhoc{} 
+specification. In particular, we extend the modelling with trusted execution 
+environments (TEEs), modify the way we model \mXor{} encryption, and 
+verify weak post-compromise security (PCS) as well as the secrecy and 
+integrity of some additional data used as part of the protocol.
 \end{abstract}
@@
 \label{sec:errorHandling}
-Section 6 of the current \mEdhoc{} specification~\cite{draft-ietf-lake-edhoc-15} states that error messages can be sent at any time and by any party.
+Section 6 of the current \mEdhoc{} 
+specification~\cite{draft-ietf-lake-edhoc-15} 
+states that error messages can be sent at any time and by any party.
 %
-There are three types of error messages: type 0 is used to represent success, type 1 represents a generic error message used for debugging purposes, and type 2 represents the failure to negotiate a common ciphersuite.
+There are three types of error messages: type 0 is used to represent success, 
+type 1 represents a generic error message used for debugging purposes, and 
+type 2 represents the failure to negotiate a common ciphersuite. 
 %
@@
 %
-The contents and semantics of type 1 messages are dependent on the implementation.
+The contents and semantics of type 1 messages are dependent on the 
+implementation.
 %
-Consequently, type 1 error messages provide a generic message passing mechanism, albeit without a predefined semantics.
+Consequently, type 1 error messages provide a generic message passing 
+mechanism, albeit without a predefined semantics.
 %
 We now discuss some issues with this specification of error messages.
+%
+In this section, we use ``specification'' to mean the latest specification, as 
+in~\cite{draft-ietf-lake-edhoc-15}.
 
 \subsection{Lack of connection identifiers}
-Error messages do not include the optional connection identifiers \mCi{} or \mCr{}.
+Error messages do not include the optional connection identifiers \mCi{} or 
+\mCr{}.
 %
-This is potentially problematic for connection-less transport layers, where these identifiers may be used to correlate messages.
+This is potentially problematic for connection-less transport layers, where 
+these identifiers may be used to correlate messages.
 
 \subsection{Improper handling}
-\mEdhoc{} provides an error message mechanism, but little to no guidance on how it should be used safely.
+\mEdhoc{} provides an error message mechanism, but little to no guidance on 
+how it should be used safely.
 %
-For example, assume that an application using \mEdhoc{} logs error messages on a finite log with rotation.
+For example, assume that an application using \mEdhoc{} logs error messages 
+on a finite log with rotation.
 %
-If the log is used for anomaly detection or detection of sensitive events, then an adversary can simply inject error messages and fill the log until the sensitive event is overwritten.
+If the log is used for anomaly detection or detection of sensitive events, then 
+an adversary can simply inject error messages and fill the log until the 
+sensitive event is overwritten.
 %
-In such situations, implementors should do proper log-separation and log management.
+In such situations, implementors should do proper log-separation and log 
+management.
 
 \subsection{Unspecified semantics for debugging messages}
-Section 6.2 gives an example of sending a type 1 error message with the information string "Method not supported".
+Section 6.2 gives an example of sending a type 1 error message with the 
+information string ``Method not supported".
 %
-Note that \mEdhoc{} does not include a mechanism to support the negotiating of the connection method.
+Note that \mEdhoc{} does not include a mechanism to support the 
+negotiating of the connection method.
 %
-The initiator selects the method as part of the first message, and if the responder replies with a type 1 error message along with the information string "Method not supported", the standard does not tell an implementation how to act upon this information string for negotiation purposes.
+The initiator selects the method as part of the first message, and if the 
+responder replies with a type 1 error message along with the information 
+string ``Method not supported", the standard does not tell an 
+implementation how to act upon this information string for negotiation 
+purposes.
 %
-An adversary can fake an error message, and there is no algorithm that is specified to ensure that a common available authentication mechanism is always selected.
+An adversary can fake an error message, and there is no algorithm that is 
+specified to ensure that a common available authentication mechanism is 
+always selected.
 %
-In particular, note that Section 5.1 in the specification states that if processing fails for some reason, then "typically" an error message is sent, processing is terminated and protocol state is discarded.
+In particular, note that Section 5.1 in the specification states that if processing 
+fails for some reason, then ``typically" an error message is sent, processing is 
+terminated and protocol state is discarded. 
 %
@@
 
-With respect to ciphersuite negotiation, Section 5.2.2 of the specification also states ``The Initiator MUST NOT change the supported cipher suites and
+With respect to ciphersuite negotiation, Section 5.2.2 of the specification also 
+states ``The Initiator MUST NOT change the supported cipher suites and
 the order of preference in SUITES\_I based on previous error messages.'' 
 %
-However, \mSuites is sent in the clear, and therefore can be modified by the adversary.
+However, \mSuites is sent in the clear, and therefore can be modified by the 
+adversary.
 %
-It is important to consider what happens if the order of preference in \mSuites is changed either by the adversary, or by the initiator themselves for some other reason.
+It is important to consider what happens if the order of preference in 
+\mSuites{} is changed either by the adversary, or by the initiator themselves 
+for some other reason.
 %
-It might be helpful to include a flag in \mMsgone to indicate to the responder
+It might be helpful to include a flag in \mMsgone{} to indicate to the 
+responder
 that a (modified) ciphersuite ordering depends on that from a previous
@@
 %
-Here we also list the protocol developers' response to being told of these
+Here, we also list the protocol developers' response (if any) to being told of these
 ambiguities in the \mSpec{}.
@@
 \subsubsection{Non-Repudiation.}
-Access control mechanisms can vary depending on what the underlying application
+Access control mechanisms can vary depending on what the underlying 
+application
 is.
 %
-A nuclear power-plant might need to keep track of who enters and leaves, but a
+A nuclear power-plant might need to keep track of who enters and leaves, but 
+a
 coffee machine that logs every user and their coffee preference might lead to a
@@
 However, the responsibilities of verifying which identity was authenticated is
-not clearly split between the application and \mEdhoc{} and therefore an attack
+not clearly split between the application and \mEdhoc{} and therefore an 
+attack
 similar in effect to UKS is possible.
@@
 %
-However, the adversary has managed to gain control of one of these devices (say
+However, the adversary has managed to gain control of one of these devices 
+(say
 $A$).
@@
 %
-A valid \mEdhoc{} implementation may then look up the appropriate credentials
+A valid \mEdhoc{} implementation may then look up the appropriate 
+credentials
 for $A$, complete the execution and deliver the security context to the
@@
 %
-We communicated this to the authors of \mEdhoc{}, who in the latest version of
+We communicated this to the authors of \mEdhoc{}, who in the latest version 
+of
 the \mSpec{}, added that \mEdhoc{} makes \mIdcredr{} available to the
-application, and added a new appendix D describing credential validations that
+application, and added an appendix D listing credential validations 
+that
 the application must take care of.
@@
 %
-The \mSpec{} does not include any specific measures for countering such attacks,
+The \mSpec{} does not include any specific measures for countering such 
+attacks,
 but assumes the application takes care of it.
@@
 Section 8.7 of the latest version of the \mEdhoc{}
-\mSpec{}~\footnote{draft-ietf-lake-edhoc-15}
+\mSpec{}~\cite{draft-ietf-lake-edhoc-15}
 gives two recommendations for countering DoS attacks.
@@
 First, \mEdhoc{} relies on lower layers to mitigate DoS, and an example of DoS
-countermeasures is given: checking the return address upon receipt of the first
+countermeasures is given: checking the return address upon receipt of the 
+first
 message.
@@
 
-No recommendation is given on how to minimize DoS effects stemming from the use
-of error messages, which can be used in a relatively arbitrary way in \mEdhoc{}
+No recommendation is given on how to minimize DoS effects stemming from 
+the use
+of error messages, which can be used in a relatively arbitrary way in 
+\mEdhoc{}
 (see Section~\ref{sec:errorHandling}).
 %
-We now elaborate on some considerations that the application developers need to
+We now elaborate on some considerations that the application developers 
+need to
 take into account.
@@
 %
-For example, the content of the \mADone{} information element may need the
+For example, the content of the \mADone{} information element may need 
+the
 responder to take further action.
 %
-Since \mEdhoc{} is designed for constrained devices, this asymmetry needs to be
+Since \mEdhoc{} is designed for constrained devices, this asymmetry needs to 
+be
 considered even more carefully than for IKEv2~\cite{rfc8019}.
@@
 
-The asymmetry between the adversary's effort towards the first message and that
+The asymmetry between the adversary's effort towards the first message and 
+that
 of the responder may hence be significant and the possibility for applications
@@
 %
-The profiling locks the \mEdhoc{} initiator role to the \mCoap{} client role and
-hence reduces the possibilities to change the \mEdhoc{} roles according to DoS
+The profiling locks the \mEdhoc{} initiator role to the \mCoap{} client role 
+and
+hence reduces the possibilities to change the \mEdhoc{} roles according to 
+DoS
 vulnerability preferences.
@@
 These are attacks where an adversary spoofs the source address of a victim in a
-message and thereby tricks the responder into sending a message to the victim.
+message and thereby tricks the responder into sending a message to the 
+victim.
 %
@@
 %
-However, such mechanisms might require that unsolicited reachability requests be
+However, such mechanisms might require that unsolicited reachability 
+requests be
 discarded, limiting the effects of anonymous DoS attacks.
@@
 In situations where the responder does not need to perform additional
-communications to verify certificates etc, a reachability mechanism may be more
-expensive in terms of time and storage compared to continuing with the protocol
+communications to verify certificates etc, a reachability mechanism may be 
+more
+expensive in terms of time and storage compared to continuing with the 
+protocol
 while keeping a half-open state for a period of time.
@@
 However, if round-trip times are large, and half-open states hence are kept by
-the receiver for a long time, a reachability mechanism can trade communication
+the receiver for a long time, a reachability mechanism can trade 
+communication
 for storage.
@@
 
-\mEdhoc{} responders querying external servers for certificate revocation checks
+\mEdhoc{} responders querying external servers for certificate revocation 
+checks
 or lookups may casue systems-level issues if many distributed \mEdhoc{}
@@
 %
-The \mEdhoc{} \mSpec{} recommends the initiator to not try that ciphersuite with
+The \mEdhoc{} \mSpec{} recommends the initiator to not try that ciphersuite 
+with
 the responder again in a new session.
@@
 In addition to unclear contexts in which
-\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail about
+\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail 
+about
 adversary capabilities.
@@
 %For example, \mOptls{}, which essentially has the same core as \mStat{}, is
-%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, which
-%explicitly separates long-term key storage from that of the ephemeral keys and
+%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, 
+%which
+%explicitly separates long-term key storage from that of the ephemeral keys 
+%and
 %session state.
@@
 %This fits in well with our use of TEEs.
-Since \mSigma{} cannot protect against compromised ephemeral keys, the authors
-of \mEdhoc{} felt that considering the compromise of ephemeral keys separately
+Since \mSigma{} cannot protect against compromised ephemeral keys, the 
+authors
+of \mEdhoc{} felt that considering the compromise of ephemeral keys 
+separately
 from that of long-term keys was not required~\cite{personalCommunication}.
@@
 %
-However, this only holds true for situations where one is only interested in the
+However, this only holds true for situations where one is only interested in 
+the
 session key confidentiality for an ongoing session.
 %
-Secure modules, in addition to being able to store long-term keys, are useful in
+Secure modules, in addition to being able to store long-term keys, are useful 
+in
 other ways too.
 %
-They also provide weak Post-Compromise Security (PCS) guarantees as discussed
+They also provide weak Post-Compromise Security (PCS) guarantees as 
+discussed
 above.
 %
-Upon discussion with the authors, they included recommendations on how to store
+Upon discussion with the authors, they included recommendations on how to 
+store
 long-term keys, and how to perform operations on these keys inside a secure
@@
 %
-As mentioned earlier, mutual explicit injective agreement cannot be obtained for
+As mentioned earlier, mutual explicit injective agreement cannot be obtained 
+for
 $P_{I}$.
 %
-This might prove to be a problem, and we proposed three potential alternatives
+This might prove to be a problem, and we proposed three potential 
+alternatives
 for addressing this to the \mEdhoc{} authors.
@@
 \section{\uppercase{Conclusions and Future Work}}
-\label{sec:conclusions}
-\label{sec:newdrafts}
+\label{sec:conclusion}
+%\label{sec:newdrafts}
 We formally modeled all four
@@
 %
-Implicit agreement covers all previously mentioned parameters for both peers.
+Implicit agreement covers all previously mentioned parameters for both 
+peers.
 %
@@
 Time measurements are rounded to the nearest second and we report both
-real and user times (i.e. time to complete the task and total CPU time
+the real time (time to complete the task) and user time (total CPU time
 spent for the task).
 %
-Mutual entity authentication, UKS- and KCI resistance can be inferred
+Mutual entity authentication, UKS- and KCI-resistance can be inferred
 from the verified properties.
-\knote{In Section 3.3 we say that injective agreement does \emph{not} hold for
-    all methods. It is a little contradictory to the table below where we
-    exclude $P_I$. It would be better to show it as not holding when $P_i$ is
-    included (both for consistency and for clarity).
-}
 %
+
 \begin{table*}[t]
         \centering
-          \label{tab:props}
           \resizebox{\textwidth}{!}{%
-            \begin{tabular}{l|c|c|c|c}
-              \textbf{Lemma $\backslash$ Method} & \textbf{SIG-SIG} & \textbf{SIG-STAT} & \textbf{STAT-SIG} & \textbf{STAT-STAT}\\
+            \begin{tabular}{|l|c|c|c|c|}
             \hline
-              \textbf{Injective agreement / I} & $S_F$ & $S_F$ & $S_P\ddagger$ & $S_P\ddagger$\\
+              \textbf{Lemma $\backslash$ Method} & \textbf{SIG-SIG} & 
+\textbf{SIG-STAT} & \textbf{STAT-SIG} & \textbf{STAT-STAT}\\
+              \hline
+              \textbf{Injective agreement / I} & $S_F$ & $S_F$ & $S_P\ddagger$ & 
+$S_P\ddagger$\\
               \textbf{Injective agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-              \textbf{Time (real/user)} & 0m27s / 11m17s & 9m27s / 239m49s & 1m21s / 35m36s & 18m16s / 499m2s\\
+              \textbf{Time (real/user)} & 0m27s / 11m17s & 9m27s / 239m49s & 
+1m21s / 35m36s & 18m16s / 499m2s\\
               \textbf{Implicit agreement / I} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
               \textbf{Implicit agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-              \textbf{Time (real/user)} & 2m37s / 65m40s & 5m28s / 131m12s $\dagger$ & 8m54s / 212m21s & 7m37s / 188m46s $\dagger$\\
+              \textbf{Time (real/user)} & 2m37s / 65m40s & 5m28s / 131m12s 
+$\dagger$ 
+& 8m54s / 212m21s & 7m37s / 188m46s $\dagger$\\
               \textbf{Session key secrecy} & \cm & \cm & \cm & \cm\\
-              \textbf{Time (real/user)} & 0m41s / 16m4s & 1m52s / 45m27s & 6m31s / 156m26s & 9m47s / 232m2s\\
+              \textbf{Time (real/user)} & 0m41s / 16m4s & 1m52s / 45m27s & 6m31s / 
+156m26s & 9m47s / 232m2s\\
               \textbf{Secrecy of AD3} & \cm & \cm & \cm & \cm\\
               \textbf{Integrity of AD3} & \cm & \cm & \cm & \cm\\
-              \textbf{Time (real/user)} & 1m40s / 38m57s & 5m53s / 138m43s & 3m35s / 85m12s & 45m6s / 1029m1s\\
+              \textbf{Time (real/user)} & 1m40s / 38m57s & 5m53s / 138m43s & 
+3m35s / 85m12s & 45m6s / 1029m1s\\
+			\hline
             \end{tabular}}\\[1em]
-
           \caption{Verified properties. $S_P$ contains roles,
@@
             authentication using the full XOR model the verification
-            was killed at 14246m43s and 2057m47s of computation time,
+            had to be abandoned at 14246m43s and 2057m47s of computation time,
             respectively. The time reported is for verification under
@@
 
-Further, we identified a situation where initiators may establish an \mOscore{}
+Further, we identified a situation where initiators may establish an 
+\mOscore{}
 security context with a different party than the application using \mEdhoc{}
@@
 %
-We discussed how the IETF may extract and better define security properties to
+We discussed how the IETF may extract and better define security properties 
+to
 enable easier verification.
 
-We verified each method in isolation, and leave as future work to verify whether
+We verified each method in isolation, and leave as future work to verify 
+whether
 the methods are secure under composition.
 
-%\subsection{Newer Versions of the Specification}
-%\label{sec:newdrafts}
-In this work, we have analyzed the \mEdhoc{} version as of July
+\subsection{A Note About the Version of \mEdhoc}
+\label{sec:newdrafts}
+In this work, we have analyzed the \mEdhoc{} specification as of July
 2020~\cite{our-analysis-selander-lake-edhoc-00}.
 %
-There are newer versions and our analysis applies to at least the version from
-February 2021~\cite{latest-ietf-lake-edhoc-05}.
+While our formal analysis applies to at least the version from February 2021, there are newer versions.
 %
-The latest version is \emph{draft-ietf-lake-edhoc-15}
+The latest version is from July 2022~\cite{draft-ietf-lake-edhoc-15}, which, among other things, changes the key derivation function.
 %
-\knote{we are already citing too many different versions, so I don't think we
+However, we do refer to various aspects (error handling, denial of service etc) of this latest version, in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
+
+%
+
+\knote{we 
+%are already citing too many different versions, so I don't think we
     should add this to the bib. Potentially, we should tomorrow instead remove
     and just mention some of the later versions.}
+\vnote{We do need to cite the latest one, just so people can look it up. It would 
+be more sensible to remove the citation for the one from February 2021, and 
+instead be more concrete about which version of EDHOC is the latest one that 
+our formal analysis applies to.}
 %
@@
 This work was partially supported by
-the Wallenberg AI, Autonomous Systems and Software Program (WASP) funded by
+the Wallenberg AI, Autonomous Systems and Software Program (WASP) 
+funded by
 the Knut and Alice Wallenberg Foundation.
 %
-We are grateful to G\"oran Selander, John Mattsson and Francesca Palombini for
+We are grateful to G\"oran Selander, John Mattsson and Francesca Palombini 
+for
 clarifications regarding the specification.

### 2022-07-22 protocol.tex
Minor changes. Added hlines to Table 2

--- 2022-07-21 protocol.tex
@@
 %
-We annotate values with $I$ and $R$ to make explicit which role they belong to.
+We annotate values with $I$ and $R$ to make explicit which role they belong 
+to.
 %
@@
 Static key pairs (suitable for regular or challenge-response signatures)
-are long-term authentication credentials, whereas ephemeral key pairs are those
+are long-term authentication credentials, whereas ephemeral key pairs are 
+those
 which are generated afresh for each execution of the protocol.
@@
 %
-This context includes the session key material (we denote this by \mSessKey{}).
+This context includes the session key material (we denote this by 
+\mSessKey{}).
 %
@@
 \centering
-\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
+\tikzset{>=latex, every msg/.style={draw=thick}, every 
+node/.style={fill=none,text=black}}
 \begin{tikzpicture}
@@
     \node (res) at (5.75,0) {Responder};
-    \msg{1em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
-    \msg{3em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, \{\mIdcredr, \mAuthr, \mADtwo\}};
+    \msg{1em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, 
+\mADone};
+    \msg{3em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, \{\mIdcredr, \mAuthr, 
+\mADtwo\}};
     \msg{5em}{ini}{res}{\mMsgthree: \mCr, \{\mIdcredi, \mAuthi, \mADthree\}};
@@
 \footnote{As in the \mSpec{}, we will from now on overload
-notation and refer to the combinations of authentication methods as methods as
+notation and refer to the combinations of authentication methods as methods 
+as
 well.}
@@
 %
-Our analysis does not cover such renegotiation which requires maintaining state
+Our analysis does not cover such renegotiation which requires maintaining 
+state
 between executions to remember the rejected ciphersuites.
@@
 In addition to negotiating the ciphersuite, the first two messages are
-also instrumental for the exchange of public ephemeral keys \mGx{} and \mGy{},
+also instrumental for the exchange of public ephemeral keys \mGx{} and 
+\mGy{},
 and connection identifiers \mCi{} and \mCr{}, for the initiator and responder
@@
 %
-The \mSpec{} states that the connection identifiers serve only to route messages
+The \mSpec{} states that the connection identifiers serve only to route 
+messages
 to the correct party executing \mEdhoc{}, but also claims that they may be
@@
 %
-These messages contain long-term key identifiers (\mIdcredi{} and \mIdcredr{}).
+These messages contain long-term key identifiers (\mIdcredi{} and 
+\mIdcredr{}).
 %
@@
 $R$ now has an option to either reject this choice of method, or to provide an
-identifier \mIdcredr{} for a key pair which can be used with challenge-response
-signatures as well as authenticating information \mAuthr{}, as dictated by
+identifier \mIdcredr{} for a key pair which can be used with 
+challenge-response
+signatures 
+as well as authenticating information \mAuthr{}, as dictated by
 the static key authentication method \mStat{}.
@@
 }
-\caption{Key schedule: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo, \mPRKthree, \mPRKfour{} are the intermediate key material, and \mKtwoe,\mKtwom, \mKthreeae, \mKthreem{} are the encryption keys for \mAead{} or \mXor{}.
-Dashed boxes are conditionals.~\cite{Norr21}}
+\caption{Key schedule: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo, 
+\mPRKthree, \mPRKfour{} are the intermediate key material, and 
+\mKtwoe,\mKtwom, 
+\mKthreeae, \mKthreem{} are the encryption keys for \mAead{} or 
+\mXor{}. Dashed boxes are conditionals.~\cite{Norr21}}
 \label{fig:kdfdiagram}
@@
 %
-\mPRKtwo{} is generated by the \mHkdfExtract{} algorithm while using the empty
+\mPRKtwo{} is generated by the \mHkdfExtract{} algorithm while using the 
+empty
 string as the salt.
@@
 %
-The key $P_{R}$ is used if the responder uses the \mStat{} authentication method.
+The key $P_{R}$ is used if the responder uses the \mStat{} authentication 
+method.
 %
-$P_{R}$ is the key computed as $\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\ \mGx)$.
+$P_{R}$ is the key computed as $\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\ 
+\mGx)$.
 %
-Similarly, the key $P_{I}$ is used if the initiator uses \mStat{}, and is computed as $\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$.
+Similarly, the key $P_{I}$ is used if the initiator uses \mStat{}, and is 
+computed as $\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$.
 %
 
-These intermediate keys are fed into \mHkdfExpand{}, as shown in Figure~\ref{fig:kdfdiagram}.
+These intermediate keys are fed into \mHkdfExpand{}, as shown in 
+Figure~\ref{fig:kdfdiagram}.
 %
-\mHkdfExpand{} here uses as salt a value $\textit{th}$, which is a running hash of the information transmitted thus far as part of the protocol.
+\mHkdfExpand{} here uses as salt a value $\textit{th}$, which is a running 
+hash of the information transmitted thus far as part of the protocol.
 %
@@
 %
-This set always includes \mGxy{}, and if the initiator (resp. the responder) uses the \mStat{} authentication method, then it also includes $P_{I}$ (resp. $P_{R}$).
+This set always includes \mGxy{}, and if the initiator (resp. the responder) uses 
+the \mStat{} authentication method, then it also includes $P_{I}$ (resp. 
+$P_{R}$).
 %
-We discuss which material should be included in \mSessKey{} in more detail in
+We 
+%discuss which material should be included in \mSessKey{} in more detail 
+in
 the formal model in Section~\ref{sec:formalization} and the consequences of
@@
 %
-Once \mSessKey{} has been established, an \mHkdf{}-based key exporter named
+Once \mSessKey{} has been established, an \mHkdf{}-based key exporter 
+named
 \mEdhoc-Exporter extracts the keys required by the security protocol.
@@
 \label{sec:edhocauth}
-We now describe how the authentication information is constructed, depending on which method is used.
+We now describe how the authentication information is constructed, 
+depending on which method is used.
 %
@@
 %
-This is similar to the 1-RTT semi-static pattern in \mOptls which computes the
+This is similar to the 1-RTT semi-static pattern in \mOptls which computes 
+the
 MAC key \textsf{sfk} for the message
@@
 %
-The same procedures works for $R$ as well (with the values corresponding to $R$).
+The same procedures works for $R$ as well (with the values corresponding to 
+$R$).
 %
@@
 %
-We use $\mathit{MAC}_{I}$ (resp. $\mathit{MAC}_{R}$) to denote a MAC which uses
+We use $\mathit{MAC}_{I}$ (resp. $\mathit{MAC}_{R}$) to denote a MAC 
+which uses
 an encryption key constructed using
 $\langle\mPriv{R},\ \mPub{R}\rangle$ and $\langle\mX,\ \mGx\rangle$
-(resp. using $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\ \mGy\rangle$).
+(resp. using $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\ 
+\mGy\rangle$).
 %
@@
 
-In Figure~\ref{fig:edhocsigstat} we show an example of protocol execution under
+In Figure~\ref{fig:edhocsigstat} we show an example of protocol execution 
+under
 the \mSigStat{} method.
@@
 key derivations used to construct these messages.
-\vnote{Do we still have the other method diagrams that we wrote out for the original paper but did not include?}
-\knote{I don't think we have space for that.}
-
+%\vnote{Do we still have the other method diagrams that we wrote out for the 
+%original paper but did not include?}
+%\knote{I don't think we have space for that.}
+%
 \begin{figure}[ht]
@@
 \scalebox{.7}{
-\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
+\tikzset{>=latex, every msg/.style={draw=thick}, every 
+node/.style={fill=none,text=black}}
 \begin{tikzpicture}
@@
     \begin{array}{c}
-        \text{Knows}\ \langle\mPriv{I},\ \mPub{I}\rangle,\ \mIdcredi,\ \mIdcredr,\ \mADone,\ \mADthree
+        \text{Knows}\ \langle\mPriv{I},\ \mPub{I}\rangle,\ \mIdcredi,\ \mIdcredr,\ 
+\mADone,\ \mADthree
     \end{array}
@@
     \end{array}$};
-    \action{3.7em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\ \mGx\rangle$};
-    \msg{7em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
+    \action{3.7em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\ 
+\mGx\rangle$};
+    \msg{7em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, 
+\mADone};
     \action{8em}{res}{$
@@
         \mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
-        \mMactwo = \mAead(\mKtwom; \langle \mIdcredr, \mTHtwo, \mCredr, \mADtwo \rangle; \textrm{``\phantom{}''}) \\
+        \mMactwo = \mAead(\mKtwom; \langle \mIdcredr, \mTHtwo, \mCredr, 
+\mADtwo \rangle; \textrm{``\phantom{}''}) \\
         \mKtwoe = \mHkdfExpand(\mPRKtwo, \mTHtwo)
       \end{array}$};
-    \msg{24em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\ \mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
+    \msg{24em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\ 
+\mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
     \action{25em}{ini}{$
@@
         \mKthreem = \mHkdfExpand(\mPRKfour, \mTHthree) \\
-        \mMacthree = \mAead(\mKthreem; \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle; \textrm{``\phantom{}''}) \\
-        \mSigthree = \mSign{I}(\langle \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle, \mMacthree \rangle)
+        \mMacthree = \mAead(\mKthreem; \langle \mIdcredi, \mTHthree, 
+\mCredi, \mADthree \rangle; \textrm{``\phantom{}''}) \\
+        \mSigthree = \mSign{I}(\langle \langle \mIdcredi, \mTHthree, \mCredi, 
+\mADthree \rangle, \mMacthree \rangle)
       \end{array}$};
-    \msg{38.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae; \mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
+    \msg{38.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae; 
+\mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
     \action{40em}{res}{$
@@
 }
-    \caption{The \mSigStat{} method for \mEdhoc{}. $\langle\cdot\rangle$ denotes a tuple, and the hash function \mHash{} is as established in the ciphersuite \mSuites{}.~\cite{Norr21}}
+    \caption{The \mSigStat{} method for \mEdhoc{}. $\langle\cdot\rangle$ 
+denotes a tuple, and the hash function \mHash{} is as established in the 
+ciphersuite \mSuites{}.~\cite{Norr21}}
 \label{fig:edhocsigstat}
@@
 Other, like the Universal Subscriber Identity Modules (USIM) used for
-authentication to 3GPP mobile networks, have application specific interfaces for
+authentication to 3GPP mobile networks, have application specific interfaces 
+for
 authentication and key agreement protocols etc.
 %
-A key aspect is how much of the application is placed in the TEE and how much is
+A key aspect is how much of the application is placed in the TEE and how 
+much is
 outside.
@@
 Because the security critical code runs in the area where the long-term keys
-reside, an implementation error here risks leaking information of the key to the
+reside, an implementation error here risks leaking information of the key to 
+the
 adversary.

### 2022-07-22 formalization.tex
[add] thread identifiers on figure 4
[fix] pfs -> pcs

@@
 \begin{align*}
-    \mPredPfs \triangleq\ & \forall I, R, \mSessKey, t_2, t_3\mLogicDot
+    \mPredPcs \triangleq\ & \forall \mTID, I, R, \mSessKey, t_2, t_3\mLogicDot
     \mK^{t_3}(\mSessKey)\  \land\ 
-    (\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(I, R, 
+    (\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(\mTID, I, R, 
 \mSessKey))
@@
     &(\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)
-    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
-    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
-    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))
-\end{align*}
-\begin{align*}
+    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
+    &\ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
+    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\[2em]
+%
     \mPredInjI \triangleq\ &
-    \forall I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, 
+    \forall \mTID_I, I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, 
 \mSessKey, S)
     \rightarrow\\
-    &(\exists t_1\mLogicDot \mRStart^{t_1}(R, \mSessKey, S) \land t_1 \lessdot 
-t_2)
-    \land (\forall I' R' t_1' \mLogicDot \mIComplete^{t_1'}(I' , R', \mSessKey, S)
-        \rightarrow t_1' \doteq t_1)
-    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
-\end{align*}
-\begin{align*}
+    &(\exists \mTID_R, t_1\mLogicDot \mRStart^{t_1}(\mTID_R, R, \mSessKey, S) \land t_1 \lessdot t_2
+    \land (\forall \mTID_I', I', R', t_1' \mLogicDot \mIComplete^{t_1'}(\mTID_I', I' , R', \mSessKey, S)
+        \rightarrow t_1' \doteq t_1))\\
+    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\[2em]
+%
     \mPredInjR \triangleq\ &
-    \forall I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(I, R, 
+    \forall \mTID_R, I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, I, R, 
 \mSessKey, S)
     \rightarrow\\
-    &(\exists t_1\mLogicDot \mIStart^{t_1}(I, R, \mSessKey, S) \land t_1 \lessdot 
-t_2)
-    \land (\forall I' R' t_1' \mLogicDot \mRComplete^{t_1'}(I' , R', \mSessKey, S)
-        \rightarrow t_1' \doteq t_1)
-    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2).
-\end{align*}
-\begin{align*}
+    &(\exists \mTID_I, t_1\mLogicDot \mIStart^{t_1}(\mTID_I, I, R, \mSessKey, S) \land t_1 \lessdot t_2
+    \land (\forall \mTID_R', I' R' t_1' \mLogicDot \mRComplete^{t_1'}(\mTID_R', I' , R', \mSessKey, S)
+        \rightarrow t_1' \doteq t_1))\\
+    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\[2em]
+%
     \mPredImpI \triangleq\ &
-    \forall I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(I, R, \mSessKey, 
+    \forall \mTID_I, I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(\mTID_I, I, R, \mSessKey, 
 S)
     \rightarrow\\
-      &(\forall I', R', S', t_2\mLogicDot \mRComplete^{t_2}(I', R', \mSessKey, S') 
-\rightarrow
-             (I=I' \land R=R' \land S=S'))\\
-      &\land (\forall I', R', S', t_1'\mLogicDot
-        (\mIComplete^{t_1'}(I', R', \mSessKey, S') \rightarrow t_1' \doteq t_1
-        )\\
-      &\ \ \ \lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
+      &(\forall \mTID_R, I', R', S', t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, I', R', \mSessKey, S') \rightarrow
+             (I=I' \land R=R' \land S=S')\\
+      &\ \ \land (\forall \mTID_I', I', R', S', t_1'\mLogicDot
+        \mIComplete^{t_1'}(\mTID_I', I', R', \mSessKey, S') \rightarrow t_1' \doteq t_1
+        ))\\
+    &\lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
     \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
-    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey)).
+    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
+    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))
 \end{align*}
@@
 %
-PFS for session key material \mSessKey{} is formalized as \mPredPfs{}
+PFS for session key material \mSessKey{} is formalized as \mPredPcs{}
 in Figure~\ref{fig:props}.
@@
 %
-We show the \mTamarin{} lemma which encodes the \mPredPfs{} property.
+We show the \mTamarin{} lemma which encodes the \mPredPcs{} property.
 %

### 2022-07-22 formalization.tex
[fix] moved rule 2 to one column

@@
 
-{
-\bgroup
-\parindent 0pt
-\begin{minipage}{0.49\textwidth}
-\bgroup
-\parindent 0pt
+\begin{minipage}{\textwidth}
 \begin{scriptsize}
@@
 15   assocData2 = <protected2, extAad2>
-16   MAC_2 = aead(`e', K_2m, assocData2,
-                  $cAEAD0)
+16   MAC_2 = aead(`e', K_2m, assocData2, $cAEAD0)
 17   authV = sign(<assocData2, MAC_2>, ~ltk)
 18   plainText2 = <$V, authV>
-\end{verbatim}
-\end{scriptsize}
-\egroup
-\end{minipage}
-\begin{minipage}{0.02\textwidth}
-\end{minipage}
-\begin{minipage}{0.48\textwidth}
-\bgroup
-\parindent 0pt
-\begin{scriptsize}
-\begin{verbatim}
-19   K_2e = expa(<$cAEAD0, TH_2,
-                 `K_2e'>, prk_2e)
-20   K_2e_1 = expa(<$cAEAD0, TH_2,
-                   `K_2e', `1'>, prk_2e)
-21   K_2e_2 = expa(<$cAEAD0, TH_2,
-                   `K_2e', `2'>, prk_2e)
-22   CIPHERTEXT_2 = <$V XOR K_2e_1,
-                     authV XOR K_2e_2>
+19   K_2e = expa(<$cAEAD0, TH_2, `K_2e'>, prk_2e)
+20   K_2e_1 = expa(<$cAEAD0, TH_2, `K_2e', `1'>, prk_2e)
+21   K_2e_2 = expa(<$cAEAD0, TH_2, `K_2e', `2'>, prk_2e)
+22   CIPHERTEXT_2 = <$V XOR K_2e_1, authV XOR K_2e_2>
 23   m2 = <data_2, CIPHERTEXT_2>
@@
 25 in
-26   [!LTK_SIG($V, ~ltk), !PK_SIG($V, pkV),
-      In(m1), Fr(~CR), Fr(~yy), Fr(~tid)]
-27   --[ExpRunningR(~tid, $V, exp_sk, agreed),
-        R2(~tid, $V, m1, m2)]->
-28   [StR2_STAT_SIG($V, ~ltk, ~yy, prk_3e2m,
-              TH_2, CIPHERTEXT_2, gx^~yy,
-              ~tid, m1, m2, agreed),
+26   [!LTK_SIG($V, ~ltk), !PK_SIG($V, pkV), In(m1), Fr(~CR), Fr(~yy), Fr(~tid)]
+27   --[ExpRunningR(~tid, $V, exp_sk, agreed), R2(~tid, $V, m1, m2)]->
+28   [StR2_STAT_SIG($V, ~ltk, ~yy, prk_3e2m, TH_2, CIPHERTEXT_2, gx^~yy, ~tid, m1, m2, agreed),
 29    Out(m2)]
@@
 \end{scriptsize}
-\egroup
 \end{minipage}
-\\
-\egroup
-%
-}
 \vspace{5mm}

### 2022-07-22 formalization.tex
Fixed quote marks in code

@@
 \begin{minipage}{0.48\textwidth}
-\begin{small}
+\begin{scriptsize}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{small}
+\end{scriptsize}
 \end{minipage}
@@
 3    agreed = <CS0, CI, ~CR>
-4    gx = `g'^xx
-5    data_2 = <`g'^~yy, CI, ~CR>
-6    m1 = <`STAT', `SIG', CS0, CI, gx>
+4    gx = 'g'^xx
+5    data_2 = <'g'^~yy, CI, ~CR>
+6    m1 = <'STAT', 'SIG', CS0, CI, gx>
 7    TH_2 = h(<$H0, m1, data_2>)
-8    prk_2e = extr(`e', gx^~yy)
+8    prk_2e = extr('e', gx^~yy)
 9    prk_3e2m = prk_2e
-10   K_2m = expa(<$cAEAD0, TH_2, `K_2m'>,
+10   K_2m = expa(<$cAEAD0, TH_2, 'K_2m'>,
 11               prk_3e2m)
@@
 15   assocData2 = <protected2, extAad2>
-16   MAC_2 = aead(`e', K_2m, assocData2, $cAEAD0)
+16   MAC_2 = aead('e', K_2m, assocData2,
+                  $cAEAD0)
 17   authV = sign(<assocData2, MAC_2>, ~ltk)
 18   plainText2 = <$V, authV>
-19   K_2e = expa(<$cAEAD0, TH_2, `K_2e'>, prk_2e)
-20   K_2e_1 = expa(<$cAEAD0, TH_2, `K_2e', `1'>, prk_2e)
-21   K_2e_2 = expa(<$cAEAD0, TH_2, `K_2e', `2'>, prk_2e)
-22   CIPHERTEXT_2 = <$V XOR K_2e_1, authV XOR K_2e_2>
+\end{verbatim}
+\end{scriptsize}
+\egroup
+\end{minipage}
+\begin{minipage}{0.02\textwidth}
+\end{minipage}
+\begin{minipage}{0.48\textwidth}
+\bgroup
+\parindent 0pt
+\begin{scriptsize}
+\begin{verbatim}
+19   K_2e = expa(<$cAEAD0, TH_2,
+                 'K_2e'>, prk_2e)
+20   K_2e_1 = expa(<$cAEAD0, TH_2,
+                   'K_2e', '1'>, prk_2e)
+21   K_2e_2 = expa(<$cAEAD0, TH_2,
+                   'K_2e', '2'>, prk_2e)
+22   CIPHERTEXT_2 = <$V XOR K_2e_1,
+                     authV XOR K_2e_2>
 23   m2 = <data_2, CIPHERTEXT_2>
@@
 2    all-traces
-3    ``All u v sk tid #t3 #t2.
+3    ''All u v sk tid #t3 #t2.
 4        (K(sk)@t3 & CompletedRun(u, v, sk, tid)@t2) ==>
@@
 9        | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & (t0 < k | k < t2))
-10       )"
+10       )''
 \end{verbatim}

### 2022-07-22 protocol.tex
Fixed quote marks in code

@@
 However, we will discuss the ramifications of such a renegotiation procedure
-and the error messages later, in Section~\ref{sec:}.
+and the error messages later, in Section~\ref{sec:errorHandling}.
 %
@@
 %
-This is similar to the 1-RTT semi-static pattern in \mOptls which computes 
-the
+This is similar to the 1-RTT semi-static pattern in \mOptls{} which computes the
 MAC key \textsf{sfk} for the message

### 2022-07-22 paper.tex
Fixed quote marks in code
[fix] title, authors, affiliations etc

@@
+% !TEX TS-program = pdflatexmk
+
 % This is samplepaper.tex, a sample chapter demonstrating the
@@
 \documentclass[runningheads]{llncs}
+\usepackage[T1]{fontenc}
 
 \usepackage{epsfig}
-\usepackage{subcaption}
+%\usepackage{subcaption}
 \usepackage{calc}
@@
 \usepackage{tikz}
-\usetikzlibrary{trees,snakes,arrows}
+%\usetikzlibrary{trees,snakes,arrows}
+\usetikzlibrary{trees,arrows}
 \usetikzlibrary{shapes,chains}
@@
 
-\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT 
-Devices}
+\begin{document}
+\title{Extended Formal Analysis of the EDHOC Protocol in Tamarin}
 \author{
-        Karl Norrman and 
-        Vaishnavi Sundararajan and
-        Alessandro Bruni
+    Karl Norrman\inst{1,2}\orcidID{0000-0003-0164-1478} \and
+    Vaishnavi Sundararajan\inst{3} \and
+    Alessandro Bruni\inst{4}
+}
+\authorrunning{Norrman, Sundararajan and Bruni}
+\institute{
+    KTH Royal Institute of Technology, Stockholm, Sweden \and
+    Ericsson Research, Security, Stockholm, Sweden\\
+    \email{karl.norrman@ericsson.com} \and
+    University of California Santa Cruz, USA\\
+    \email{vasundar@ucsc.edu} \and
+    IT University of Copenhagen, Copenhagen, Denmark\\
+    \email{brun@itu.dk}
 }
 
-\begin{document}
 \maketitle
 %
+
 \begin{abstract}
@@
 integrity of some additional data used as part of the protocol.
+\keywords{Formal Verification, Symbolic Dolev-Yao Model,
+           Authenticated Key Establishment, Protocols, IoT, Tamarin.}
+
 \end{abstract}

### 2022-07-22 formalization.tex
Fixed errors
[fix] improved section 4.1

@@
 operations are assumed to be perfect, e.g., encrypted messages can only be
-decrypted with the correct key, and there are no hash collisions exist.
+decrypted with the correct key, and there are no hash collisions.
 %
 
-The adversary is assumed to be the network, and thus controls the 
-communication
+The adversary is assumed to be in control of the communication
 channel and can see all messages being communicated as part of the protocol.
@@
 
-On top of the Dolev-Yao model,~\cite{Norr21} allow the adversary two extra
-capabilities with respect to getting access to the long-term and ephemeral 
-keys.
+On top of the standard Dolev-Yao model,~\cite{Norr21} allows the adversary to
+access the long-term and ephemeral keys.
 %
 Long-term key reveal, denoted $\mRevLTK^{t}(A)$, stands for the adversary
-gaining access to a party $A$'s long-term private key \mPriv{A} at a time point 
-$t$.
+gaining access to a party $A$'s long-term private key \mPriv{A} at a time point $t$.
 %
@@
 %
+Furthermore, we extend the results from~\cite{Norr21} by strengthening the
+adversary capabilities, as detailed next.
+%
 
@@
 %
-We extend the adversary model of~\cite{Norr21} by allowing the adversary
-access to a TEE interface on a party of their choosing before and after the
-\mEdhoc{} session.
+We extend the adversary model of~\cite{Norr21} by introducing an interface to
+allow the adversary to use the long term key material of any party, though
+without directly accessing such key material.
 %
-This allows us to distinguish between a full compromise and a weak one, 
-using
-the terminology of~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
+This allows us to model the scenario where the adversary has gained access to
+a device, but the long term key material is protected by a trusted execution
+environment (TEE), and thus only accessible through the TEE interface.
 %
-A weak compromise is when the adversary have temporary access to the TEE.
+According to the terminology of~\cite{DBLP:conf/csfw/Cohn-GordonCG16}, a
+protocol is considered to have \emph{weak post-compromise security}
+(Weak-PCS) if it achieves its security goals for a specific session
+even when the attacker has limited access to the long-term key material
+of the parties involved before the session starts -- through an interface 
+that maintains the key material secure but allows to run cryptographic 
+operations with it -- and full access to the key material of the parties
+involved after the end of the session, as well as access to the key material
+of all other parties.
 %
-Seen in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add 
-adversary
+Seen in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add adversary
 capabilities corresponding to a server adversary, i.e., an adversary that
@@
 %
-The model of~\cite{Norr21} only captures server adversary capabilities without
-access to the TEE, via the leakage of ephemeral keys, and local adversary
-capabilities, or more specifically, a secure-world adversary with read-write
-access to the
-secure world where the long-term keys reside via the leakage of long-term 
-keys.
-%
-Our extension allows analysis of so-called weak post-compromise security
-(PCS)~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
+Our extension to the model of~\cite{Norr21} thus allows us to verify all the
+previous security properties under a Weak-PCS model.
 %
@@
 More precisely, parties using the \mSig{} authentication method use a TEE 
-with
-an interface which accepts a message and returns the signature of that 
-message
-using the party's private long-term key.
+with an interface which accepts a message and returns the signature of that 
+message using the party's private long-term key.
 %
 A party $U$ using the \mStat{} authentication method use a TEE with an 
-interface
-accepting a point $P$ on the curve and returning $\mDH(P,\ \mGu{})$.
+interface accepting a point $P$ on the curve and returning $\mDH(P,\ \mGu{})$.
 %
@@
 This interface requires the least functionality from the TEE, reducing the 
-TEE's
-complexity and possibly cost of its implementation.
+TEE's complexity and possibly cost of its implementation.
 %
-The functional split is suitable even for the situation where a
-constrained device has implemented only the storage of the long-term key in 
-a
-special purpose circuit with minimal processing functionality.
+This functional split is suitable even when a constrained device has
+implemented only the storage of the long-term key in 
+a special purpose circuit with minimal processing functionality.
 %
@@
 %
-Care must be taken when specifying the start and the end.
-%
-Specifically, the start and end of the protocol run must be viewed w.r.t. the
+Care must be taken when specifying the start and the end: specifically,
+the start and end of the protocol run must be viewed w.r.t. the current
 role.
@@
 the \mSigSig{} method requires that the adversary does not have access to the
-TEE of the
-responder from the time of transmission of the first message until the second
-message is received by the initiator.
+TEE of the responder from the time of transmission of the first message until
+the second message is received by the initiator.
 %
@@
 
-We did not manage to get Tamarin to terminate for all lemmas and all 
-methods
-with the added TEE extension.
+With this additional attacker capability Tamarin manages to prove
+almost all lemmas and method combinations, except for the implicit
+authentication lemmas when the responder uses the \mStat{} authentication
+method: in these two cases Tamarin did not complete the within reasonable time.
 %
-When the responder uses the \mStat{} authentication method Tamarin did not
-complete the in a reasonable time for the authentication lemmas.
+We therefore simplified the modeling of XOR encryption of the second
+message to prove implicit authentication when the responder uses the 
+\mStat{} method.
 %
-To combat this, we simplified the modeling of XOR encryption of the second
-message for these lemmas when the responder uses \mStat{} authentication.
+More precisely, the second message encrypts two values: the responder's identity
+\verb|R| and the authentication information \verb|authR|.
 %
-More precisely, the second message contains the two values \verb|R| (the
-responder's identity) and
-\verb|authR| (the authentication information), and we generally model the
-XOR encryption by xoring each of them with their own key-stream terms.
+We model XOR encryption by xor-ing each term with their own key-stream term.
 %
-For the problematic lemmas however, we resorted to xoring the entire tuple
+However for the problematic cases we xor the entire tuple
 \verb|<R, authR>| with a single key-stream term.
 %
-Given that no attacks were identified using our original modeling, this
-simplification is intuitively not a severe restriction.
+This simplification might miss an attack on implicit authentication that is due
+to the combination of a malleable XOR encryption and access to the TEE interface.
 %
-However, it is possible that we now miss attacks on these methods
-combining properties of the two-term XOR encryption with TEE capabilities.
+However, given that no attacks were identified using our original modeling due
+to the use of XOR, nor in the current modeling for all other authentication 
+methods, we believe that this is not a severe restriction.
 %
 
-We now briefly describe \mTamarin{}, and how we model \mEdhoc{} in it 
-here.
+Next we describe \mTamarin{}, and our modeling of \mEdhoc{}.
 %
@@
 %
+{\footnotesize
 \begin{verbatim}
 [!KU(k), !KU(m), !KU(ad), !KU(ai)] --[]-> [!KU(aeadEncrypt(k, m, ad, ai))]
-\end{verbatim}
+\end{verbatim}}
 %
-We make used of \mTamarin{}'s builtin rules for other operations, like 
+We use \mTamarin{}'s builtin rules for other operations, like 
 \mXor{}.
@@
 
+{\parindent 0pt
 \begin{minipage}{\textwidth}
@@
 15   assocData2 = <protected2, extAad2>
-16   MAC_2 = aead('e', K_2m, assocData2,
-                  $cAEAD0)
+16   MAC_2 = aead('e', K_2m, assocData2, $cAEAD0)
 17   authV = sign(<assocData2, MAC_2>, ~ltk)
 18   plainText2 = <$V, authV>
-\end{verbatim}
-\end{scriptsize}
-\egroup
-\end{minipage}
-\begin{minipage}{0.02\textwidth}
-\end{minipage}
-\begin{minipage}{0.48\textwidth}
-\bgroup
-\parindent 0pt
-\begin{scriptsize}
-\begin{verbatim}
-19   K_2e = expa(<$cAEAD0, TH_2,
-                 'K_2e'>, prk_2e)
-20   K_2e_1 = expa(<$cAEAD0, TH_2,
-                   'K_2e', '1'>, prk_2e)
-21   K_2e_2 = expa(<$cAEAD0, TH_2,
-                   'K_2e', '2'>, prk_2e)
-22   CIPHERTEXT_2 = <$V XOR K_2e_1,
-                     authV XOR K_2e_2>
+19   K_2e = expa(<$cAEAD0, TH_2, 'K_2e'>, prk_2e)
+20   K_2e_1 = expa(<$cAEAD0, TH_2, 'K_2e', '1'>, prk_2e)
+21   K_2e_2 = expa(<$cAEAD0, TH_2, 'K_2e', '2'>, prk_2e)
+22   CIPHERTEXT_2 = <$V XOR K_2e_1, authV XOR K_2e_2>
 23   m2 = <data_2, CIPHERTEXT_2>
@@
 \end{scriptsize}
-\end{minipage}
+\end{minipage}}
 \vspace{5mm}

### 2022-07-22 formalization.tex
[fix] section 4.3

@@
 %
+Note that this attacker model is strictly more powerful than that of~\cite{Norr21},
+as it maintains all the previous attacker capabilities.
+%
 
@@
 authentication lemmas when the responder uses the \mStat{} authentication
-method: in these two cases Tamarin did not complete the within reasonable time.
+method: in these two cases Tamarin did not complete the within reasonable
+time\footnote{See Table~\ref{tab:props} for all verification results and
+computation times.}.
 %
@@
 %
-We use \mTamarin{}'s builtin rules for other operations, like 
-\mXor{}.
+Other than modeling authenticated encryption, we use \mTamarin{}'s builtin
+equational theories for signing, Diffie-Hellman, hashing and \mXor{}.
 %
 
-The rules modeling a Dolev-Yao adversary and the evolution of their 
-knowledge
-as the protocol executes are built into \mTamarin.
+Tamarin has also built-in rules for modeling a Dolev-Yao adversary and the
+evolution of their knowledge as the protocol executes.
 %
-We can extend and/or modify this behavior by adding other rules as well.
+We extend the Dolev-Yao adversary model by adding other rules that increase
+the capabilities of the attacker.
 %
@@
 %
-This 
-%stands for the fact that as the adversary interacts with the parties
+\anote{Removed:
+This stands for the fact that as the adversary interacts with the parties
 executing the protocol, at time point $t$, the adversary can derive the
 term $p$ using the Dolev-Yao deduction rules and any additional adversary
-capabilities specified in \mTamarin.
+capabilities specified in \mTamarin.}
 %
-As an example, consider the following implication
+As an example, the following implication
 \[
@@
 \]
-which represents the fact that if the adversary gets to know the pair of
+models that if the adversary gets to know the pair of
 keys $\langle k, k' \rangle$ at a time point $t$, then the adversary also
@@
 \label{sec:desired-properties}
-In this section, we describe how we model \mEdhoc{} and the various 
-properties
-we want it to satisfy.
+In this section we describe our modeling of \mEdhoc{} and its desired
+security properties.
 %
@@
 
-In this work, we consider the following properties:
-perfect forward secrecy, injective agreement for both the initiator and the
-responder, and implicit agreement.
+In this work we consider the following properties: secrecy of the session key,
+injective agreement and implicit agreement of the session key material for both
+initiator and responder, and secrecy and integrity of the application data sent
+on message 3 ($\mAD_3$)
 %
@@
     \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
 %
@@
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
 %
@@
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
 %
@@
     \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
-    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))
+    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))
 \end{align*}
@@
 %----------------------------------------------------------------------- PFS
-\subsubsection{Perfect Forward Secrecy (PFS).}
+\subsubsection{Secrecy of the session key material.}
 \label{sec:secrecy}
-Informally, by perfect forward secrecy, we mean that the adversary cannot
-gain access to the session key material \mSessKey{}, even if a long-term key is
-leaked after the establishment of \mSessKey{}.
+We show that the adversary cannot gain access to the session key material \mSessKey{},
+even under a weak post-compromise security model that allows the adversary limited access
+to the long-term keys of the parties involved through a secure interface before the session starts, 
+unrestricted access to the long-term keys of the involved parties after the session ends
+and of all other parties, as well as access to all other session keys.
 %
-PFS for session key material \mSessKey{} is formalized as \mPredPcs{}
+This property is formalized as \mPredPcs{}
 in Figure~\ref{fig:props}.
@@
 
-In the \mIComplete{} event, $I$ represents the identity of the initiator,
+In the \mIComplete{} event, \mTID{} represents a ``thread identifier'' for the
+current session, $I$ represents the identity of the initiator,
 and $R$ represents the identity of the party who the initiator believes is
@@
 %
-Intuitively, the PFS property states that an adversary obtains \mSessKey{}
-only if a party's long-term key is compromised before their run ends, if an
-ephemeral key is compromised at any time after a party's run began.
-%
-This is to hold even if the adversary has access to the TEE of either party
-before and after the protocol execution.
+Intuitively, the PCS property states that an adversary obtains \mSessKey{}
+only if one of the following conditions hold: either a party's long-term key
+is compromised before their run ends, or \mSessKey{} is directly revealed to
+the attacker, or the TEE interface for the responder is used during the current
+session for the initiator (i.e. after the initiator starts and before it
+commits), or, similarly, the TEE interface for the initiator is used during the
+current session for the responder.
 %
@@
 Therefore, in the \mTamarin{} code, instead of using the disjunction
-$\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, 
-\mSessKey)$
-to model the fact that either party can have completed their execution, we use
-a single action parametrized by the terms $I$, $R$, and \mSessKey.
+$\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\, 
+\mRComplete^{t_2}(\mTID, I, R, \mSessKey)$
+to model the fact that either party may have completed their execution, we use
+a single action parametrized by the terms $\mTID$, $I$, $R$, and \mSessKey.
 
@@
 %
-The property is to hold even if the adversary has access to the TEE of party $R$
+The property should hold even if the adversary has access to the TEE of party $R$
 before and after the protocol execution.
@@
 %%
-The analogous definition holds for the responder $R$.
+An analogous definition holds for the responder $R$.
 %
@@
 %------------------------------------------------------------- Implicit auth
-As part of our analysis, we found that all the \mEdhoc{} methods satisfy PFS.
+As part of our analysis, we found that all the \mEdhoc{} methods satisfy weak
+pot-compromise security.
 %

### 2022-07-22 formalization.tex
[fix] spec-citations and contributions

@@
 3        [!LTK_STAT($A, ~ltk),
-4         !PK_STAT($A, `g'^~ltk),
-5         Out(<$A, `g'^~ltk>)]
+4         !PK_STAT($A, 'g'^~ltk),
+5         Out(<$A, 'g'^~ltk>)]
 \end{verbatim}
@@
 2    all-traces
-3    ''All u v sk tid #t3 #t2.
+3    "All u v sk tid #t3 #t2.
 4        (K(sk)@t3 & CompletedRun(u, v, sk, tid)@t2) ==>
@@
 9        | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & (t0 < k | k < t2))
-10       )''
+10       )"
 \end{verbatim}

### 2022-07-22 intro.tex

@@
 %
-It allowed two different key establishment methods -- one involved a 
-pre-shared 
+It allowed two different key establishment methods -- one involved a pre-shared
 Diffie-Hellman (DH) \emph{cryptographic core}, and the other was a
@@
 Once these ingredients are added to a cryptographic core, we obtain a
-key-establishment 
-method.
+key-establishment method.
 %
@@
 %
-We extend the adversary model and \mTamarin{} models to capture weak
+We extend the adversary model and \mTamarin{} system models to capture weak
 post-compromise security (PCS) and model Trusted Execution Environments
@@
 %
-We also modify the model to consider a stronger version of the \mXor{} encryption whenever possible. 
-%
 We check for the following properties:
 \begin{itemize}
-\item Perfect forward secrecy for the session key material
 \item Injective agreement
 \item Implicit agreement
-\item Weak post-compromise security
+\item Weak post-compromise security for the session key material
+\item Perfect Forward Secrecy (PFS) for the session key material
 \item Secrecy and integrity of \mADthree
 \end{itemize}
-\vnote{Check this!}
+%
+We follow the definition of weak PCS by~\cite{DBLP:conf/csfw/Cohn-GordonCG16},
+which subsumes PFS.
 %
 We also discuss various issues arising in relation to the use of trusted
-execution environments (TEEs), denial of service (DoS) attacks, sources of 
-randomness, the negotiation of parameters (which the formal model abstracts 
+execution environments (TEEs), denial of service (DoS) attacks, error handling,
+the negotiation of parameters (which the formal model abstracts
 away) for establishing the protocol, and other potential attacks and concerns.
@@
 We have communicated these issues to the developers of the protocol.
+%
 

### 2022-07-22 kdfdiagram.tex

@@
 \draw [->, norm] (p2) -- (h2); 
-\draw [->, norm] (c1.east) -- ++(1.85,0) -- (h3.north) node[midway,above left] {Y};
+\draw [->, norm] (c1.east) -- ++(2.025,0) -- (h3.north) node[midway,above left] {Y};
 %\draw [->, norm] (h3.south) -- ++(0,-1) -- ++(-3,0);
@@
 \draw [->, norm] (p3) -- (h4); 
-\draw [->, norm] (c2.east) -- ++(1.91,0) -- (h5.north) node[midway,above left] {Y};
+\draw [->, norm] (c2.east) -- ++(2.07,0) -- (h5.north) node[midway,above left] {Y};
 %\draw [->, norm] (h5.south) -- ++(0,-1) -- ++(-3,0);
@@
 \draw [->] (h1.west) -- (p2.east);
-\draw [->] (u1.west) -- ++(-1.52,0) -- (h1.south);
+\draw [->] (u1.west) -- ++(-2.27,0) -- (h1.south);
 \draw [->] (p0.west) -- (h1.east);

### 2022-07-22 paper.tex

@@
 However, the responsibilities of verifying which identity was authenticated is
-not clearly split between the application and \mEdhoc{} and therefore an 
-attack
+not clearly split between the application and \mEdhoc{} and therefore an attack
 similar in effect to UKS is possible.
@@
 %
-However, the adversary has managed to gain control of one of these devices 
-(say
+However, the adversary has managed to gain control of one of these devices (say
 $A$).
@@
 %
-A valid \mEdhoc{} implementation may then look up the appropriate 
-credentials
+A valid \mEdhoc{} implementation may then look up the appropriate credentials
 for $A$, complete the execution and deliver the security context to the
@@
 %
-We communicated this to the authors of \mEdhoc{}, who in the latest version 
-of
+We communicated this to the authors of \mEdhoc{}, who in the latest version of
 the \mSpec{}, added that \mEdhoc{} makes \mIdcredr{} available to the
-application, and added an appendix D listing credential validations 
-that
+application, and added an appendix D listing credential validations that
 the application must take care of.
@@
 %
-The \mSpec{} does not include any specific measures for countering such 
-attacks,
-but assumes the application takes care of it.
+The \mSpec{} does not include any specific measures countering such
+attacks, but assumes the application takes care of it.
 
@@
 First, \mEdhoc{} relies on lower layers to mitigate DoS, and an example of DoS
-countermeasures is given: checking the return address upon receipt of the 
-first
+countermeasures is given: checking the return address upon receipt of the first
 message.
@@
 
-No recommendation is given on how to minimize DoS effects stemming from 
-the use
-of error messages, which can be used in a relatively arbitrary way in 
-\mEdhoc{}
+No recommendation is given on how to minimize DoS effects stemming from the use
+of error messages, which can be used in a relatively arbitrary way in \mEdhoc{}
 (see Section~\ref{sec:errorHandling}).
 %
-We now elaborate on some considerations that the application developers 
-need to
+We now elaborate on some considerations that the application developers need to
 take into account.
@@
 %
-For example, the content of the \mADone{} information element may need 
-the
+For example, the content of the \mADone{} information element may need the
 responder to take further action.
 %
-Since \mEdhoc{} is designed for constrained devices, this asymmetry needs to 
-be
+Since \mEdhoc{} is designed for constrained devices, this asymmetry needs to be
 considered even more carefully than for IKEv2~\cite{rfc8019}.
@@
 
-The asymmetry between the adversary's effort towards the first message and 
-that
+The asymmetry between the adversary's effort towards the first message and that
 of the responder may hence be significant and the possibility for applications
@@
 %
-The profiling locks the \mEdhoc{} initiator role to the \mCoap{} client role 
-and
-hence reduces the possibilities to change the \mEdhoc{} roles according to 
-DoS
+The profiling locks the \mEdhoc{} initiator role to the \mCoap{} client role and
+hence reduces the possibilities to change the \mEdhoc{} roles according to DoS
 vulnerability preferences.
@@
 These are attacks where an adversary spoofs the source address of a victim in a
-message and thereby tricks the responder into sending a message to the 
-victim.
+message and thereby tricks the responder into sending a message to the victim.
 %
@@
 In situations where the responder does not need to perform additional
-communications to verify certificates etc, a reachability mechanism may be 
-more
-expensive in terms of time and storage compared to continuing with the 
-protocol
+communications to verify certificates etc, a reachability mechanism may be more
+expensive in terms of time and storage compared to continuing with the protocol
 while keeping a half-open state for a period of time.
@@
 However, if round-trip times are large, and half-open states hence are kept by
-the receiver for a long time, a reachability mechanism can trade 
-communication
+the receiver for a long time, a reachability mechanism can trade communication
 for storage.
@@
 
-\mEdhoc{} responders querying external servers for certificate revocation 
-checks
+\mEdhoc{} responders querying external servers for certificate revocation checks
 or lookups may casue systems-level issues if many distributed \mEdhoc{}
@@
 %
-The \mEdhoc{} \mSpec{} recommends the initiator to not try that ciphersuite 
-with
+The \mEdhoc{} \mSpec{} recommends the initiator to not try that ciphersuite with
 the responder again in a new session.
@@
 In addition to unclear contexts in which
-\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail 
-about
+\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail about
 adversary capabilities.
@@
 %For example, \mOptls{}, which essentially has the same core as \mStat{}, is
-%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, 
-%which
-%explicitly separates long-term key storage from that of the ephemeral keys 
-%and
+%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, which
+%explicitly separates long-term key storage from that of the ephemeral keys and
 %session state.
@@
 %This fits in well with our use of TEEs.
-Since \mSigma{} cannot protect against compromised ephemeral keys, the 
-authors
-of \mEdhoc{} felt that considering the compromise of ephemeral keys 
-separately
+Since \mSigma{} cannot protect against compromised ephemeral keys, the authors
+of \mEdhoc{} felt that considering the compromise of ephemeral keys separately
 from that of long-term keys was not required~\cite{personalCommunication}.
@@
 %
-However, this only holds true for situations where one is only interested in 
-the
+However, this only holds true for situations where one is only interested in the
 session key confidentiality for an ongoing session.
 %
-Secure modules, in addition to being able to store long-term keys, are useful 
-in
+Secure modules, in addition to being able to store long-term keys, are useful in
 other ways too.
 %
-They also provide weak Post-Compromise Security (PCS) guarantees as 
-discussed
+They also provide weak Post-Compromise Security (PCS) guarantees as discussed
 above.
 %
-Upon discussion with the authors, they included recommendations on how to 
-store
+Upon discussion with the authors, they included recommendations on how to store
 long-term keys, and how to perform operations on these keys inside a secure
@@
 %
-As mentioned earlier, mutual explicit injective agreement cannot be obtained 
-for
+As mentioned earlier, mutual explicit injective agreement cannot be obtained for
 $P_{I}$.
 %
-This might prove to be a problem, and we proposed three potential 
-alternatives
+This might prove to be a problem, and we proposed three potential alternatives
 for addressing this to the \mEdhoc{} authors.
@@
 %
-While our formal analysis applies to at least the version from February 2021, there are newer versions.
+While our formal analysis applies to v5 of the \mSpec{} (February 2021),
+there are newer versions.
 %
-The latest version is from July 2022~\cite{draft-ietf-lake-edhoc-15}, which, among other things, changes the key derivation function.
+The latest version at the time of writing is
+v15 (July 2022)~\cite{draft-ietf-lake-edhoc-15}, which, among other things,
+differ from v5 in terms of some inputs to the key derivations.
 %
-However, we do refer to various aspects (error handling, denial of service etc) of this latest version, in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
-
+However, we do refer to various aspects
+(error handling, denial of service etc) of this latest version,
+in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
 %
-
-\knote{we 
-%are already citing too many different versions, so I don't think we
-    should add this to the bib. Potentially, we should tomorrow instead remove
-    and just mention some of the later versions.}
-\vnote{We do need to cite the latest one, just so people can look it up. It would 
-be more sensible to remove the citation for the one from February 2021, and 
-instead be more concrete about which version of EDHOC is the latest one that 
-our formal analysis applies to.}
+Due to lack of time we unfortunately could not bring the Tamarin model up to
+date with v15~\cite{draft-ietf-lake-edhoc-15}.
 %

### 2022-07-22 formalization.tex
[fix] remove editor's notes
[fix] grammatical and consistency fixes

@@
 in a session where they establish key material $k$.
-%
-% Ephemeral keys might be stored in a less secure part of a device,
-% but long-term keys are often stored and operate within a secure model,
-% like a trusted execution environment (TEE). \vnote{Do we want to say more 
-% about TEEs here?}
 %
@@
 %
-Other than modeling authenticated encryption, we use \mTamarin{}'s builtin
+Other than modeling authenticated encryption, we use \mTamarin{}'s built-in
 equational theories for signing, Diffie-Hellman, hashing and \mXor{}.
@@
 In this work we consider the following properties: secrecy of the session key,
-injective agreement and implicit agreement of the session key material for both
+injective agreement and implicit agreement of the session-key material for both
 initiator and responder, and secrecy and integrity of the application data sent
@@
 Agreement is considered on a set of parameters $S$ which also contains the
-session key material \mSessKey{}.
+session-key material \mSessKey{}.
 %
@@
 %----------------------------------------------------------------------- PFS
-\subsubsection{Secrecy of the session key material.}
+\subsubsection{Secrecy of the Session-key Material.}
 \label{sec:secrecy}
-We show that the adversary cannot gain access to the session key material \mSessKey{},
-even under a weak post-compromise security model that allows the adversary limited access
-to the long-term keys of the parties involved through a secure interface before the session starts, 
-unrestricted access to the long-term keys of the involved parties after the session ends
-and of all other parties, as well as access to all other session keys.
+We show that the adversary cannot gain access to the session-key material
+\mSessKey{}, even under a weak post-compromise security model that allows the
+adversary limited access to the long-term keys of the parties involved through a
+secure interface before the session starts, unrestricted access to the long-term
+keys of the involved parties after the session ends and of all other parties, as
+well as access to all other session keys.
 %
-This property is formalized as \mPredPcs{}
-in Figure~\ref{fig:props}.
+This property is formalized as \mPredPcs{} in Figure~\ref{fig:props}.
 %
@@
 playing the responder role, while \mSessKey{} stands for the established
-session key material.
+session-key material.
 %
@@
 the attacker, or the TEE interface for the responder is used during the current
-session for the initiator (i.e. after the initiator starts and before it
-commits), or, similarly, the TEE interface for the initiator is used during the
+session for the initiator, i.e., after the initiator starts and before it
+commits, or, similarly, the TEE interface for the initiator is used during the
 current session for the responder.
@@
 uniquely corresponds to one of $R$ where the set of parameters is $S$ and
-includes, in particular, the session key material \mSessKey{}.
+includes, in particular, the session-key material \mSessKey{}.
 %
@@
 The ``implicit'' in the name of the property stands for the fact that a party
-$A$ assumes that any party $B$ who has access to the session key material
+$A$ assumes that any party $B$ who has access to the session-key material
 \mSessKey{} must, in fact, be the intended party, and that if $B$ is honest,
@@
 be revealed since the property relies on the fact that the intended responder
-is the only one who knows the session key material.
+is the only one who knows the session-key material.
 %
@@
 with
-the public keys of $I$ and $R$ to compute the session key material.
+the public keys of $I$ and $R$ to compute the session-key material.
 %
@@
     \item responder identity,
-    \item session key material (which varies depending on the \mEdhoc{} 
+    \item session-key material (which varies depending on the \mEdhoc{} 
 method),
@@
 The full set of agreed parameters $S_F$ is $S_P \cup \{I, P_I\}$ when $P_I$
-is part of the session key material, and $S_P \cup \{I\}$ otherwise.
+is part of the session-key material, and $S_P \cup \{I\}$ otherwise.
 %
@@
 Our above notions of agreement ensure that both parties agree on each
-other's identity, role, and session key material.
+other's identity, role, and session-key material.
 %
@@
 As part of a UKS attack, a party $A$ can be made to believe that it finished
-an execution with a party $B$, but where the session key material is actually
+an execution with a party $B$, but where the session-key material is actually
 shared between $A$ and $C$ instead.
 %
-Again, due to the agreement on identities and session key material, any 
+Again, due to the agreement on identities and session-key material, any 
 method
@@
 \mRStart{} parametrized by the session id, the responder's identity, and the
-session key material \mT{exp\_sk}.
+session-key material \mT{exp\_sk}.
 %
-The \mT{exp} in the name of the variable for session key material represents
+The \mT{exp} in the name of the variable for session-key material represents
 the fact that the agreement property satisfied by this key is explicit, i.e.,
@@
 %
-We use \mT{imp\_sk} for the corresponding session key material which does 
-not
+We use \mT{imp\_sk} for the corresponding session-key material which does not
 include $P_{I}$.

### 2022-07-22 paper.tex
[fix] grammatical and consistency fixes

@@
 fails for some reason, then ``typically" an error message is sent, processing is
-terminated and protocol state is discarded. 
+terminated, and protocol state is discarded.
 %
@@
 \subsubsection{Non-Repudiation.}
-Access control mechanisms can vary depending on what the underlying 
-application
+Access control mechanisms can vary depending on what the underlying application
 is.
 %
-A nuclear power-plant might need to keep track of who enters and leaves, but 
-a
+A nuclear power-plant might need to keep track of who enters and leaves, but a
 coffee machine that logs every user and their coffee preference might lead to a
@@
 \mEdhoc{} responders querying external servers for certificate revocation checks
-or lookups may casue systems-level issues if many distributed \mEdhoc{}
+or lookups may cause systems-level issues if many distributed \mEdhoc{}
 responders frequently queries such servers when DoSed by clients.
@@
 As we have seen, \mEdhoc{} establishes some
-session key material, which can be fed into the \mEdhoc{}-Exporter to derive
+session-key material, which can be fed into the \mEdhoc{}-Exporter to derive
 session keys.
@@
 %
-    \item Derive session key material without using $P_{I}$:
+    \item Derive session-key material without using $P_{I}$:
 %
@@
 \label{sec:conclusion}
-%\label{sec:newdrafts}
 We formally modeled all four
@@
 Mutual injective agreement covers the parameters $S_P$:
-responder identity, roles, session key material (except for $P_I$ when
+responder identity, roles, session-key material (except for $P_I$ when
 initiator uses the \mStat{} authentication
@@
           \caption{Verified properties. $S_P$ contains roles,
-            responder identity, session key material (excluding
+            responder identity, session-key material (excluding
             $P_I$), \mCi, \mCr, and \mSuites. $S_F$ is $S_{P}$,
@@
         %         Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-        %         PFS for session key material & \cm & \cm & \cm & \cm\\
+        %         PFS for session-key material & \cm & \cm & \cm & \cm\\
         %         \hline

### 2022-07-22 intro.tex
[fix] grammatical and consistency fixes

@@
 In~\cite{Norr21}, the authors check for injective agreement, implicit
-agreement, and perfect forward secrecy for the session key material.
+agreement, and perfect forward secrecy for the session-key material.
 %
@@
 \item Implicit agreement
-\item Weak post-compromise security for the session key material
-\item Perfect Forward Secrecy (PFS) for the session key material
+\item Weak post-compromise security for the session-key material
+\item Perfect Forward Secrecy (PFS) for the session-key material
 \item Secrecy and integrity of \mADthree
@@
 %We give an explicit adversary model for the protocol and verify
-%properties such as session key material and entity authentication, and
+%properties such as session-key material and entity authentication, and
 %perfect
@@
 %definition
-%on the session key material, and none on the initiator's identity.
+%on the session-key material, and none on the initiator's identity.
 %%
 %However, we show that all methods fulfill an implicit agreement property
-%covering the session key material and the initiator's identity.
+%covering the session-key material and the initiator's identity.
 %%
@@
 %a
-%party establishes session key material with a trusted but compromised peer,
+%party establishes session-key material with a trusted but compromised peer,
 %even

### 2022-07-22 formalization.tex
[fix] heading consitency

@@
 
+%---------------------------------------------------------------------- subsub
 \subsection{Formal Model}\label{sec:threat-model}
@@
 %----------------------------------------------------------------------- PFS
-\subsubsection{Secrecy of the Session-key Material.}
+\subsubsection{Secrecy of the Session-Key Material.}
 \label{sec:secrecy}
@@
 %------------------------------------------------------- Agreed parameters
-\subsubsection{Parameters in set $S$.}
+\subsubsection{Set $S$ of Agreed Parameters.}
 \label{sec:agreedParams}
@@
 %---------------------------------------------------------------------------sub
-\subsection{Encoding \mEdhoc{} in \mTamarin}
+\subsection{Encoding \mEdhoc{} in \mTamarin{}}
 \label{sec:modeling}
@@
 Having described how we adapt the equational theory to model \mEdhoc,
-we now move on to the modelling of the adversary model and the 
+we now move on to the modeling of the adversary model and the 
 environment in

### 2022-07-22 paper.tex
[fix] heading consitency

@@
 that we analyze, and discussing some of the aspects of the latest \mEdhoc{}
-specification. In particular, we extend the modelling with trusted execution
+specification. In particular, we extend the modeling with trusted execution
 environments (TEEs), modify the way we model \mXor{} encryption, and
@@
 %-------------------------------------------------------------------------- sec
-\section{\uppercase{Error handling}}
+\section{\uppercase{Error Handling}}
 \label{sec:errorHandling}
@@
 
-\subsection{Lack of connection identifiers}
+\subsection{Lack of Connection Identifiers}
 Error messages do not include the optional connection identifiers \mCi{} or
@@
 
-\subsection{Improper handling}
+\subsection{Proper Reception Handling}
 \mEdhoc{} provides an error message mechanism, but little to no guidance on
@@
 %
-In such situations, implementors should do proper log-separation and log
+In such situations, implementers should do proper log-separation and log
 management.
 
-\subsection{Unspecified semantics for debugging messages}
+\subsection{Unspecified Semantics for Debugging Messages}
 Section 6.2 gives an example of sending a type 1 error message with the
@@
 %
-However, \mSuites is sent in the clear, and therefore can be modified by the
+However, \mSuites{} is sent in the clear, and therefore can be modified by the
 adversary.
@@
 %-------------------------------------------------------------------------- sec
-\section{\uppercase{Improvements to the specification}}
+\section{\uppercase{Improvements to the Specification}}
 \label{sec:discussion}

### 2022-07-22 intro.tex
[add] citation to correct spec in introduction

@@
 %
-We analyze the version of \mEdhoc{} as in~\cite{}.
+We formally analyze the version of \mEdhoc{} as in~\cite{}.
 %
+We formally analyze the \mEdhoc{} specification as of July
+2020~\cite{our-analysis-selander-lake-edhoc-00}.
+%
+Out formal analysis applies to v5 of the \mSpec{} (February 2021).
+%
+However, we do refer to various aspects
+(error handling, denial of service etc) of this latest version,
+in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
 We extend the adversary model and \mTamarin{} system models to capture weak
@@
 %
-We check for the following properties:
+We formally verify the following properties:
 \begin{itemize}

### 2022-07-22 formalization.tex
[fix] some fixes on the formalization section

--- 2022-07-22 formalization.tex
@@
     \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land \mTEE^{t_1}(R) \land t_0 \lessdot t_1)\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land t_0 \lessdot t_1)\\[2em]
 %
@@
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land t_0 \lessdot t_1)\\[2em]
 %
@@
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land t_0 \lessdot t_1)\\[2em]
 %
@@
     \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
-    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))
+    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land t_0 \lessdot t_1)
 \end{align*}
@@
 
-%----------------------------------------------------------------------- PFS
+%----------------------------------------------------------------------- PCS
 \subsubsection{Secrecy of the session key material.}
@@
 In the \mIComplete{} event, \mTID{} represents a ``thread identifier'' for the
-current session, $I$ represents the identity of the initiator,
+current session and is used to correlate multiple events for the same run of the 
+protocol for each party, $I$ represents the identity of the initiator,
 and $R$ represents the identity of the party who the initiator believes is
@@
 the attacker, or the TEE interface for the responder is used during the current
-session for the initiator (i.e. after the initiator starts and before it
-commits), or, similarly, the TEE interface for the initiator is used during the
+session for the initiator (i.e. after the initiator starts), or, similarly, 
+the TEE interface for the initiator is used during the
 current session for the responder.
+%
+We use the thread identifier \mTID for each party to relate the starting event
+to the corresponding commit event.
 %
@@
 \begin{verbatim}
-1 lemma secrecyPFS:
+1 lemma secrecyPCS:
 2    all-traces
@@
 7        | (Ex #t1. EphKeyRev(sk)@t1)
-8        | (Ex m1 #t0 #k. I1(tid, u, v, m1)@t0 & TEE(v)@k & (t0 < k | k < t2))
-9        | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & (t0 < k | k < t2))
+8        | (Ex m1 #t0 #k. I1(tid, u, v, m1)@t0 & TEE(v)@k & t0 < k)
+9        | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & t0 < k)
 10       )"
@@
 Lines 8 and 9 captures that the parties TEEs must be inaccessible to the
-adversary between the start and end of the execution as seen by each party
+adversary after the start of the session as seen by each party
 respectively.

### 2022-07-22 formalization.tex

--- 2022-07-22 formalization.tex
@@
 
-%----------------------------------------------------------------------- PFS
-\subsubsection{Secrecy of the Session-Key Material.}
+%----------------------------------------------------------------------- PCS
+\subsubsection{Secrecy of the session key material.}
 \label{sec:secrecy}
@@
 the attacker, or the TEE interface for the responder is used during the current
-session for the initiator, i.e., after the initiator starts and before it
-commits, or, similarly, the TEE interface for the initiator is used during the
+session for the initiator, i.e., after the initiator starts, or, similarly, the 
+TEE interface for the initiator is used during the
 current session for the responder.
+%
+We use the thread identifier \mTID for each party to relate the starting event
+to the corresponding commit event.
 %

### 2022-07-22 protocol.tex
[fix] remove editor's notes

@@
 %
-This context includes the session key material (we denote this by 
+This context includes the session-key material (we denote this by
 \mSessKey{}).
@@
 
-At the end of a successful run of the protocol, the session key material is
+At the end of a successful run of the protocol, the session-key material is
 established as \mSessKey{}, which we define as a set of various keys.
@@
 key derivations used to construct these messages.
-%\vnote{Do we still have the other method diagrams that we wrote out for the 
-%original paper but did not include?}
-%\knote{I don't think we have space for that.}
 %
@@
 \label{sec:TEE}
-Authentication of a specific IoT device assumes that this devices is the only
+Authentication of a specific IoT device assumes that this device is the only
 entity with access to the long-term key associated with the corresponding
@@
 %
-A state of the art approach is to use a Trusted Execution Environment (TEE),
+A state-of-the-art approach is to use a TEE,
 which holds the key and provides an API for operations using the key.
@@
 This is the approach taken by the
-Trust-Zone based \mMuEdhoc{}~\cite{DBLP:conf/codaspy/Hristozov0XFLS21}
+TrustZone based \mMuEdhoc{}~\cite{DBLP:conf/codaspy/Hristozov0XFLS21}
 implementation for example.
@@
 %
-Some, like ARM Trust Zone and Intel SGX are general purpose execution
+Some, like ARM TrustZone and Intel SGX are general purpose execution
 environments, which are flexibly programmable.
@@
 
-For larger devices that include general purpose processors with Trust Zone or
+For larger devices that include general purpose processors with TrustZone or
 SGX, entire \mEdhoc{} and \mOscore{} may reside inside the TEE.
@@
 From this perspective, it may be beneficial to follow the minimalistic approach
-even when having access to Trust Zone or SGX.
+even when having access to Trust Zoneor SGX.
 %
@@
 the TEE and extend the interface to accept messages and return the
-(en/de)crypted counter part.
+(en/de)crypted counterpart.
 %

### 2022-07-23 formalization.tex
Fixed typos and references and merged everything into paper.tex

--- 2022-07-22 formalization.tex
@@
 %
-Dolev-Yao is a well-established model for symbolic verification of security
+The Dolev-Yao model is well established for the symbolic verification of 
+security
 protocols.
@@
 operations are assumed to be perfect, e.g., encrypted messages can only be
-decrypted with the correct key, and there are no hash collisions.
+decrypted with the correct key, there are no hash collisions etc.
 %
@@
 In addition, they can interact with an unbounded number of protocol sessions,
-drop, inject and modify messages at will.
+and drop, inject and modify messages at will.
 %
 
-On top of the standard Dolev-Yao model,~\cite{Norr21} allows the adversary to
-access the long-term and ephemeral keys.
+On top of the standard Dolev-Yao model, \cite{Norr21} allows the adversary to
+access long-term and ephemeral keys via specific events.
 %
-Long-term key reveal, denoted $\mRevLTK^{t}(A)$, stands for the adversary
-gaining access to a party $A$'s long-term private key \mPriv{A} at a time point $t$.
+Long-term key reveal, denoted by $\mRevLTK^{t}(A)$, stands for the adversary
+gaining access to a party $A$'s long-term private key \mPriv{A} at a time point 
+$t$.
 %
-Ephemeral key reveal, denoted $\mRevEph^{t}(A, k)$, stands for the adversary
+Ephemeral key reveal, denoted by $\mRevEph^{t}(A, k)$, stands for the 
+adversary
 obtaining, at time $t$, the ephemeral private key \mPrivE{A} used by party $A$
-in a session where they establish key material $k$.
+in a session where they establish a session key $k$.
 %
@@
 %
-Furthermore, we extend the results from~\cite{Norr21} by strengthening the
+Furthermore, we modify the model from~\cite{Norr21} by strengthening the
 adversary capabilities, as detailed next.
@@
 %
-We extend the adversary model of~\cite{Norr21} by introducing an interface to
-allow the adversary to use the long term key material of any party, though
-without directly accessing such key material.
+We extend the adversary model of~\cite{Norr21} by allowing the adversary to 
+use the long term key of any party via an interface, without directly accessing 
+the key itself.
 %
 This allows us to model the scenario where the adversary has gained access to
-a device, but the long term key material is protected by a trusted execution
-environment (TEE), and thus only accessible through the TEE interface.
+a device, but the long-term key is protected by a trusted execution
+environment (TEE), and thus only accessible via the TEE interface.
 %
 According to the terminology of~\cite{DBLP:conf/csfw/Cohn-GordonCG16}, a
-protocol is considered to have \emph{weak post-compromise security}
+protocol is considered to enjoy \emph{weak post-compromise security}
 (Weak-PCS) if it achieves its security goals for a specific session
-even when the attacker has limited access to the long-term key material
-of the parties involved before the session starts -- through an interface 
-that maintains the key material secure but allows to run cryptographic 
-operations with it -- and full access to the key material of the parties
-involved after the end of the session, as well as access to the key material
-of all other parties.
+even when the following hold:
+\begin{itemize}
+\item Before the session starts, the adversary has limited access to the 
+long-term 
+keys of the parties involved, through an interface that securely maintains 
+the keys but allows principals to run cryptographic operations using them, and
+\item The adversary has full access to the keys of the parties involved after the 
+end of the session, as well as access to the keys of all other parties.
+\end{itemize}
 %
-Seen in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add adversary
-capabilities corresponding to a server adversary, i.e., an adversary that
-compromises a party, learns its ephemeral keys and may temporarily access its
-TEE, but does not learn the party's long-term key.
+As in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add 
+adversary
+capabilities corresponding to a server adversary, i.e., upon 
+compromising a party, the adversary learns their ephemeral keys and may temporarily access their TEE, but does not learn their long-term key.
 %
-Our extension to the model of~\cite{Norr21} thus allows us to verify all the
-previous security properties under a Weak-PCS model.
+This extension to the model of~\cite{Norr21} thus allows us to verify all the
+previous security properties under a ``Weak-PCS model''.
 %
-Note that this attacker model is strictly more powerful than that of~\cite{Norr21},
-as it maintains all the previous attacker capabilities.
+Note that this attacker model is strictly more powerful than that 
+of~\cite{Norr21}, as it subsumes the previous attacker capabilities.
 %
 
-We split the \mEdhoc{} functionality as follows.
+We split the functionality of \mEdhoc{} as follows.
 %
 The TEE contains the long-term key and allows the non-TEE parts of the
-application to perform the operations on it via its interface.
+application to perform the operations using it via an interface.
 %
 More precisely, parties using the \mSig{} authentication method use a TEE 
-with an interface which accepts a message and returns the signature of that 
+interface which accepts a message and returns the signature of that 
 message using the party's private long-term key.
 %
-A party $U$ using the \mStat{} authentication method use a TEE with an 
-interface accepting a point $P$ on the curve and returning $\mDH(P,\ \mGu{})$.
+A party $U$ using the \mStat{} authentication method uses a TEE with an 
+interface accepting a point $P$ on the curve and 
+%returning $\mDH(P,\ \mGu{})$.
+returning $\mDH(\mPriv{U},\ P)$.
 %
 
-The non-TEE parts implement \mEdhoc{} using this interface.
+%The non-TEE parts implement \mEdhoc{} using this interface.
 %
 This interface requires the least functionality from the TEE, reducing the 
-TEE's complexity and possibly cost of its implementation.
+TEE's complexity and (possibly) its cost.
 %
 This functional split is suitable even when a constrained device has
-implemented only the storage of the long-term key in 
+implemented the storage of only the long-term key in 
 a special purpose circuit with minimal processing functionality.
@@
 These rules allow the adversary to obtain terms representing signatures on a
-value of their choice to forge signatures (\verb|forge_SIG|), or to obtain terms
+value of their choice (\verb|forge_SIG|), or to obtain terms
 representing a curve point of their choice raised to the power of the
@@
 
-Because it is a trivial attack when the adversary access these rules with values
-from the test session, we must disqualify those rule applications.
+Because it is a trivial attack when the adversary accesses these rules with values
+from a session in progress, we must disqualify those rule applications.
 %
@@
 properties with a condition that no such action fact exists from the start of
-the protocol execution and its end.
+the protocol execution till its end.
 %
 Care must be taken when specifying the start and the end: specifically,
-the start and end of the protocol run must be viewed w.r.t. the current
+the start and end of the execution must be viewed with respect to the current
 role.
@@
 the \mSigSig{} method requires that the adversary does not have access to the
-TEE of the responder from the time of transmission of the first message until
-the second message is received by the initiator.
+TEE of the responder from the time that the first message is transmitted till
+the second message is received by the initiator (or equivalently, till the third message is transmitted by the initiator, since reception of a message and transmission of the next one by a party is one atomic operation).
 %
-Because reception of one message and transmission of the next one at a party
-is an atomic operation, these timepoints represent the start and end of the
+Thus, these timepoints represent the start and end of the
 protocol run from the perspective of the initiator.
@@
 
-With this additional attacker capability Tamarin manages to prove
-almost all lemmas and method combinations, except for the implicit
-authentication lemmas when the responder uses the \mStat{} authentication
-method: in these two cases Tamarin did not complete the within reasonable
-time\footnote{See Table~\ref{tab:props} for all verification results and
-computation times.}.
-%
-We therefore simplified the modeling of XOR encryption of the second
-message to prove implicit authentication when the responder uses the 
-\mStat{} method.
-%
-More precisely, the second message encrypts two values: the responder's identity
+The second message encrypts two values: the responder's 
+identity
 \verb|R| and the authentication information \verb|authR|.
 %
-We model XOR encryption by xor-ing each term with their own key-stream term.
+We model XOR encryption by \mXor-ing each term with their own key-stream 
+term.
 %
-However for the problematic cases we xor the entire tuple
+However for some problematic methods\footnote{See Table~\ref{tab:props} for all verification results and computation times.} we \mXor{} the entire tuple
 \verb|<R, authR>| with a single key-stream term.
 %
-This simplification might miss an attack on implicit authentication that is due
-to the combination of a malleable XOR encryption and access to the TEE interface.
+This simplification is likely to miss an attack on implicit authentication which occurs due to the combination of a malleable XOR encryption and access to the TEE 
+interface.
 %
-However, given that no attacks were identified using our original modeling due
-to the use of XOR, nor in the current modeling for all other authentication 
-methods, we believe that this is not a severe restriction.
+However, given that no attacks were identified due to the use of XOR in our original modeling, nor in the current modeling for all other authentication methods, we believe that this is not a severe restriction.
 %
 
-Next we describe \mTamarin{}, and our modeling of \mEdhoc{}.
+Next we provide a quick overview of the \mTamarin{} tool, and describe our modeling of \mEdhoc{} in it.
 %
@@
 \label{sec:tamarin}
-We extend the formal \mEdhoc{} model of~\cite{Norr21}, using the symbolic
-model tool \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an 
+We extend the formal \mEdhoc{} model of~\cite{Norr21}, using the tool \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an 
 interactive
-tool for the formal verification of security protocols.
+tool for the symbolic verification of security protocols.
 %
@@
 
-An annotated multiset rewrite-rule is represented as
+An annotated multiset-rewrite rule is represented as
 $l \ifarrow[e] r$, where $l$ and $r$ are multisets, and $e$ is
@@
 A sequence of actions yields a protocol execution trace.
+%
+Properties are defined as formulas in a fragment of temporal first order logic,
+and these formulas can be verified over execution traces.
 %
@@
 
-Properties are defined as formulas in a fragment of temporal first order logic,
-and these formulas can be verified over execution traces.
-%
-
 Protocol verification in \mTamarin{} happens under an equational theory $E$.
@@
 Users can extend the default term algebra and equational theory in
-\mTamarin{} with new function symbols and unification rules for these new 
+\mTamarin{} with new function symbols and unification rules (if any) for these new 
 symbols.
@@
 %
-\anote{Removed:
-This stands for the fact that as the adversary interacts with the parties
-executing the protocol, at time point $t$, the adversary can derive the
-term $p$ using the Dolev-Yao deduction rules and any additional adversary
-capabilities specified in \mTamarin.}
-%
 As an example, the following implication
@@
 \]
-models that if the adversary gets to know the pair of
-keys $\langle k, k' \rangle$ at a time point $t$, then the adversary also
+models the fact that if the adversary gets to know the pair of
+keys $\langle k, k' \rangle$ at a time point $t$, then the adversary
 knows each of those keys $k$ and $k'$ at time point $t$ as well.
@@
 \label{sec:desired-properties}
-In this section we describe our modeling of \mEdhoc{} and its desired
+In this section, we describe our modeling of \mEdhoc{} and its desired
 security properties.
@@
 In this work we consider the following properties: secrecy of the session key,
-injective agreement and implicit agreement of the session-key material for both
+injective agreement and implicit agreement on the session-key material for 
+both
 initiator and responder, and secrecy and integrity of the application data sent
-on message 3 ($\mAD_3$)
+as part of message 3 ($\mAD_3$).
 %
@@
 %
-We will first describe in detail both properties, and then describe the
+We will first describe in detail all these properties, and then describe the
 contents of the set $S$.
@@
 a modified version of a figure in~\cite{Norr21}.
+%
+%Note that the properties in Figure~\ref{fig:props} are unlike the actual \mTamarin{} lemma in one minor way -- in the \mTamarin{} code, we only check that the TEE is accessed after the start of the session, since the long-term key reveal event covers the reveal of all long-term and session keys after the end of the session.
+%\vnote{Check that this is right}
+
 \begin{figure*}[ht]
@@
     \mK^{t_3}(\mSessKey)\  \land\ 
-    (\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(\mTID, I, R, 
+    (\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\, 
+\mRComplete^{t_2}(\mTID, 
+I, R, 
 \mSessKey))
@@
     \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+	&\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land 
+\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, I, R) 
+\land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+%    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) 
+%\land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
 %
@@
     \rightarrow\\
-    &(\exists \mTID_R, t_1\mLogicDot \mRStart^{t_1}(\mTID_R, R, \mSessKey, S) \land t_1 \lessdot t_2
-    \land (\forall \mTID_I', I', R', t_1' \mLogicDot \mIComplete^{t_1'}(\mTID_I', I' , R', \mSessKey, S)
+    &(\exists \mTID_R, t_1\mLogicDot \mRStart^{t_1}(\mTID_R, R, \mSessKey, 
+S) \land t_1 \lessdot t_2
+    \land (\forall \mTID_I', I', R', t_1' \mLogicDot \mIComplete^{t_1'}(\mTID_I', 
+I' , R', \mSessKey, S)
         \rightarrow t_1' \doteq t_1))\\
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
+\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
 %
     \mPredInjR \triangleq\ &
-    \forall \mTID_R, I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, I, R, 
+    \forall \mTID_R, I, R, \mSessKey, S, t_2\mLogicDot 
+\mRComplete^{t_2}(\mTID_R, 
+I, R, 
 \mSessKey, S)
     \rightarrow\\
-    &(\exists \mTID_I, t_1\mLogicDot \mIStart^{t_1}(\mTID_I, I, R, \mSessKey, S) \land t_1 \lessdot t_2
-    \land (\forall \mTID_R', I' R' t_1' \mLogicDot \mRComplete^{t_1'}(\mTID_R', I' , R', \mSessKey, S)
+    &(\exists \mTID_I, t_1\mLogicDot \mIStart^{t_1}(\mTID_I, I, R, \mSessKey, 
+S) \land t_1 \lessdot t_2
+    \land (\forall \mTID_R', I' R' t_1' \mLogicDot 
+\mRComplete^{t_1'}(\mTID_R', 
+I' , R', \mSessKey, S)
         \rightarrow t_1' \doteq t_1))\\
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, 
+\mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot 
+t_2))\\[2em]
 %
     \mPredImpI \triangleq\ &
-    \forall \mTID_I, I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(\mTID_I, I, R, \mSessKey, 
+    \forall \mTID_I, I, R, \mSessKey, S, t_1\mLogicDot 
+\mIComplete^{t_1}(\mTID_I, 
+I, R, \mSessKey, 
 S)
     \rightarrow\\
-      &(\forall \mTID_R, I', R', S', t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, I', R', \mSessKey, S') \rightarrow
+      &(\forall \mTID_R, I', R', S', t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, 
+I', R', \mSessKey, S') \rightarrow
              (I=I' \land R=R' \land S=S')\\
       &\ \ \land (\forall \mTID_I', I', R', S', t_1'\mLogicDot
-        \mIComplete^{t_1'}(\mTID_I', I', R', \mSessKey, S') \rightarrow t_1' \doteq t_1
+        \mIComplete^{t_1'}(\mTID_I', I', R', \mSessKey, S') \rightarrow t_1' \doteq 
+t_1
         ))\\
@@
     \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
-    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))
+    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
+\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))
 \end{align*}
@@
 We show that the adversary cannot gain access to the session-key material
-\mSessKey{}, even under a weak post-compromise security model that allows the
-adversary limited access to the long-term keys of the parties involved through a
-secure interface before the session starts, unrestricted access to the long-term
-keys of the involved parties after the session ends and of all other parties, as
-well as access to all other session keys.
+\mSessKey{}, even under the weak post-compromise security model which allows 
+the
+adversary limited access to the long-term keys of the involved parties via
+a
+secure interface before the session starts, and unrestricted access to the long-term
+keys of all parties as well as access to all other session keys after the session ends.
 %
@@
 
-In the \mIComplete{} event, \mTID{} represents a ``thread identifier'' for the
+In the \mIComplete{} event, \mTID{} represents a ``thread identifier'' which uniquely identifies the
 current session, $I$ represents the identity of the initiator,
@@
 session for the initiator, i.e., after the initiator starts and before it
-commits, or, similarly, the TEE interface for the initiator is used during the
+completes, or, similarly, the TEE interface for the initiator is used during the
 current session for the responder.
 %
-The property in Figure~\ref{fig:props} is unlike the actual \mTamarin{} lemma
-in one minor way: \mTamarin's logic does not allow disjunctions to appear on
+Note that the properties in Figure~\ref{fig:props} are unlike the actual \mTamarin{} lemma in one minor way: \mTamarin's logic does not allow disjunctions to appear on
 the left-hand side of an implication inside a universally-quantified formula.
@@
 %
-The property should hold even if the adversary has access to the TEE of party $R$
+The property should hold even if the adversary has access to the TEE of party 
+$R$
 before and after the protocol execution.
-%
-%If the adversary has managed to compromise $R$'s long-term key, they can
-%generate any message of their liking and sign it using said key,
-%or compute a $\mathit{MAC}_{R}$ using the long-term keys \mPubE{I},
-%\mPriv{R}, and an ephemeral key pair $\langle\mPrivE{R},\ 
-%\mPubE{R}\rangle$
-%of 
-%their choice.
 %%
@@
 %
+%As part of the formalization, note that we parametrize the \mIStart{} and \mRStart{} events with some subset of \mTID, $I$, $R$, and $Z$, as well as the set $S$ of parameters upon which we check for agreement.
+%%
+%The set $S$ will often include the other party's identity, so, for example, \mRStart{} will not mention the term $I$ as a parameter outside of $S$, and instead just be parametrized by \mTID, $R$, $Z$, and $S$.
 
@@
 As part of our analysis, we found that all the \mEdhoc{} methods satisfy weak
-pot-compromise security.
+post-compromise security.
 %
@@
 the last message, $A$ concludes that if this message reaches $B$, then $A$
-and $B$ agree on each other's identities and roles, and the set $S$.
+and $B$ agree on each other's identities and roles, as well as the set $S$.
 %
@@
 
-In addition to these default operations, \mEdhoc{} is built over
-\mHkdfExpand, \mHkdfExtract, and the \mAead{} functions.
+In addition to these default operations, \mEdhoc{} is built over the following functions: \mHkdfExpand, \mHkdfExtract, and \mAead{}.
 %
@@
 %
-A term encrypted using \mAead{} is represented by \mT{aeadEncrypt(m, k, ad, 
+A term encrypted using the \mAead{} algorithm is represented by \mT{aeadEncrypt(m, k, ad, 
 ai)},
@@
 Note that we do not model any error message that $R$ might send in response
-to message \mMsgone rejecting $I$'s choice of ciphersuite and/or method.
+to message \mMsgone{} rejecting $I$'s choice of ciphersuite and/or method.
 %
@@
 %
-This can be seen in lines 19--22 in the code for \mT{R2\_STAT\_SIG}.
+This can be seen in lines 18--21 in the code for \mT{R2\_STAT\_SIG}.
 %
@@
 9    prk_3e2m = prk_2e
-10   K_2m = expa(<$cAEAD0, TH_2, 'K_2m'>,
-11               prk_3e2m)
+10   K_2m = expa(<$cAEAD0, TH_2, 'K_2m'>, prk_3e2m)
 12   protected2 = $V // ID_CRED_V
@@
 %
-We use \mT{imp\_sk} for the corresponding session-key material which does not
+We use \mT{imp\_sk} for the corresponding session-key material which does 
+not
 include $P_{I}$.
@@
 same.
-%
-
-%-------------------------------------------------------------------------- sub
-%\subsection{Encoding the Desired Properties in \mTamarin}
-%\label{sec:propertyFormalization}
-%The properties and adversary model we defined in
-%Section~\ref{sec:desired-properties} translate directly into \mTamarin's logic,
-%using the straightforward mapping of events to the actions emitted from the
-%model.
 %
@@
 
-\begin{small}
+\begin{scriptsize}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{small}
+\end{scriptsize}
 %

### 2022-07-23 intro.tex
Fixed typos and references and merged everything into paper.tex

--- 2022-07-22 intro.tex
@@
 %
-IETF is standardizing new protocols to secure communications between
-constrained devices.
+The Internet Engineering Task Force (IETF) is standardizing new protocols to secure communications between devices that operate under such restrictions.
 %
@@
 COSE
-(\mEdhoc{}) is under discussion in the IETF.
+(\mEdhoc{}) is being discussed in the IETF.
 %
@@
 While establishing security contexts for \mOscore{} is the primary goal for the
-\mEdhoc{} protocol, there might well be other use cases, which have not been
+\mEdhoc{} protocol, there might well be other use cases which have not been
 explored in depth yet.
 %
-It is therefore important to ensure that most of the fundamental properties
+It is therefore important to ensure that fundamental properties
 expected of key exchange protocols as established in the literature are satisfied
@@
 signatures were added as well~\cite{our-analysis-selander-lake-edhoc-00}.
+%
+While this is the version we base our formal analysis on, there have been later versions. 
+%
+See Section~\ref{sec:newdrafts} for a more detailed discussion.
 
@@
 %
-The authors check  various properties, namely secrecy, identity protection
-against an active adversary, strong authentication, perfect forward secrecy
+The authors check  the following properties: secrecy, identity protection, strong authentication, perfect forward secrecy
 (PFS), and integrity of application data.
 %
-Later work~\cite{Norr21} analyze the May 2020 version of \mEdhoc{} in the
+Later work~\cite{Norr21} analyzes the July 2020 version of \mEdhoc{} in the
 \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13}.
 %
-This version of the protocol has five key establishment methods.
+This version of the protocol has four key establishment methods.
 %
-In~\cite{Norr21}, the authors check for injective agreement, implicit
-agreement, and perfect forward secrecy for the session-key material.
+In~\cite{Norr21}, the properties checked for are injective agreement, implicit
+agreement, and perfect forward secrecy for the session key material.
 %
-The authors also discuss some fallouts of the various design choices made as
+That work also includes a discussion about the various design choices made as
 part of \mEdhoc{}, and the impact of \mEdhoc{} in multiple use-case scenarios.
@@
 %
-We analyze the version of \mEdhoc{} as in~\cite{}.
+We formally analyze the \mEdhoc{} specification as of July
+2020~\cite{our-analysis-selander-lake-edhoc-00}, but our formal analysis applies as far as the version of the \mSpec{} from February 2021.
 %
@@
 %
-We check for the following properties:
+We also alter the formal modelling used for encryption under \mXor{}.
+%
+We formally verify the following properties:
 \begin{itemize}
@@
 \item Implicit agreement
+\item Perfect Forward Secrecy (PFS) for the session-key material
 \item Weak post-compromise security for the session-key material
-\item Perfect Forward Secrecy (PFS) for the session-key material
 \item Secrecy and integrity of \mADthree
@@
 %
-We also discuss various issues arising in relation to the use of trusted
+Since \mEdhoc{} is still a protocol under development, the latest version of \mEdhoc{} is the one from July 2022~\cite{draft-ietf-lake-edhoc-15}, which contains details which are not covered by our formal model.
+%
+We do, however, refer to various aspects
+(error handling, denial of service etc) of the July 2022 version of the \mSpec{}, namely 
+in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
+%
+We also discuss various issues arising due to the use of trusted
 execution environments (TEEs), denial of service (DoS) attacks, error handling,
@@
 %
-
-%In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
-%establishment methods) using the \mTamarin{}
-%tool~\cite{DBLP:conf/cav/MeierSCB13}.
-%%
-%We
-%present a formal model we constructed of the protocol as given in the
-%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00}.
-%%
-%
-%We give an explicit adversary model for the protocol and verify
-%properties such as session-key material and entity authentication, and
-%perfect
-%forward secrecy, for all four methods.
-%%
-%
-%The model itself is valuable as a basis for verifying further updates in the
-%ongoing standardization.
-%%
-%It is publicly available~\cite{edhocTamarinRepo}.
-%%
-%It took several person-months to interpret the
-%specification and construct the model.
-%%
-%Termination requires a hand-crafted proof oracle to guide \mTamarin{}.
-%%
-%
-%We show that not all \mEdhoc{}'s key establishment
-%methods provide authentication according to the injective agreement
-%definition
-%on the session-key material, and none on the initiator's identity.
-%%
-%However, we show that all methods fulfill an implicit agreement property
-%covering the session-key material and the initiator's identity.
-%%
-%We identify a number of subtleties, ambiguities and weaknesses in the
-%specification.
-%%
-%For example, the authentication policy requirements allows situations where
-%a
-%party establishes session-key material with a trusted but compromised peer,
-%even
-%though the intention was to establish it with a different trusted party.
-%%
-%We provide remedies for the identified issues and have
-%communicated these to the IETF and the specification authors, who
-%incorporated
-%some of our suggestions and currently consider how to deal with the
-%remaining
-%ones.
-%
-
-%-------------------------------------------------------------------------- sub
-%\subsection{Comparison with Related Work}
-%The May 2018 version of \mEdhoc{} was formally analyzed by
-%\cite{DBLP:conf/secsr/BruniJPS18} using the \mProverif{}
-%tool~\cite{DBLP:conf/csfw/Blanchet01}.
-%%
-%Their analysis covered a pre-shared key authenticated core and one
-%based on \mSigma.
-%%
-%The properties checked for therein were secrecy, PFS and integrity of
-%application data, identity protection against an active adversary,
-%and strong authentication.
-%%
-%
-%In contrast to the key establishment methods analyzed by Bruni et~al.,
-%which
-%were based on the well-understood pre-shared key DH and \mSigma{}
-%protocols,
-%the three newly added
-%methods combine two unilateral authentication protocols with the goal to
-%constructing mutual authentication protocols.
-%%
-%Combining two protocols, which individually provide unilateral
-%authentication,
-%is not guaranteed to result in a secure mutual authentication
-%protocol~\cite{DBLP:conf/ccs/Krawczyk16}.
-%%
-%Consequently, even though the framework is similar to the one analyzed by
-%Bruni
-%et~al., the cryptographic underpinnings have significantly increased in
-%complexity, and is using mechanisms which have not previously been
-%formally analyzed.
-%%
-%The set of properties we check for is also different.
-%%
-%Our analysis is further carried out using a different tool,
-%namely \mTamarin; different kinds of strategies to formulate and
-%successfully analyze the protocol are required when working with this tool.

### 2022-07-23 paper.tex
[fix] some fixes on the formalization section
[fix] daggers
Fixed typos and references and merged everything into paper.tex

@@
 %
-%By now, the protocol has been fleshed out much more, and has been
-%augmented with multiple new key establishment methods.
 By 2021, the protocol had been fleshed out much more, with multiple new key
@@
 %
-Here, we build on this work, by modifying the model, adding new properties
-that we analyze, and discussing some of the aspects of the latest \mEdhoc{}
-specification. In particular, we extend the modeling with trusted execution
-environments (TEEs), modify the way we model \mXor{} encryption, and
-verify weak post-compromise security (PCS) as well as the secrecy and
-integrity of some additional data used as part of the protocol.
+In this paper, we build on that work, by modifying the model, analyzing some new properties, and discussing some aspects of the latest \mEdhoc{}
+specification.
+In particular, we extend the modeling in~\cite{Norr21} with trusted execution environments (TEEs), modify the way we model \mXor{} encryption, and in addition to the properties verified in~\cite{Norr21}, we verify weak post-compromise security (PCS) as well as the secrecy and integrity of some additional data used as part of the protocol.
+
 \keywords{Formal Verification, Symbolic Dolev-Yao Model,
@@
 \label{sec:introduction}
-\input{intro}
+IoT protocols are often run on devices which operate under severe restrictions
+on resources like bandwidth and energy consumption.
+%
+These constrained devices are often simple in their operation, but need to
+communicate and function without human interference or maintenance for
+extended periods of time.
+%
+The Internet Engineering Task Force (IETF) is standardizing new protocols to secure communications between devices that operate under such restrictions.
+%
+One such is the Object Security for Constrained RESTful Environments
+(\mOscore{}) protocol.
+%
+However, \mOscore{} requires the pre-establishment of a security context.
+%
+To this end, a key exchange protocol named Ephemeral Diffie-Hellman Over
+COSE
+(\mEdhoc{}) is being discussed in the IETF.
+%
+Since \mEdhoc{} will establish security contexts for \mOscore{}, the same
+resource constraints (especially those pertaining to message size) apply to the
+former as for the latter.
+%
+While establishing security contexts for \mOscore{} is the primary goal for the
+\mEdhoc{} protocol, there might well be other use cases which have not been
+explored in depth yet.
+%
+It is therefore important to ensure that fundamental properties
+expected of key exchange protocols as established in the literature are satisfied
+by \mEdhoc{} as well.
+%
+
+%-------------------------------------------------------------------------- sub
+\subsection{Evolution of \mEdhoc}
+\label{sec:edhocevol}
+The first \mEdhoc{} framework was introduced in March 2016.
+%
+It allowed two different key establishment methods -- one involved a pre-shared
+Diffie-Hellman (DH) \emph{cryptographic core}, and the other was a
+variation on challenge-response signatures, {\`a} la
+\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
+
+%
+A \emph{cryptographic core}, often just called a core, is an academic protocol,
+i.e., with no encodings or application-specific details as needed for an
+industrial protocol.
+%
+Once these ingredients are added to a cryptographic core, we obtain a
+key-establishment method.
+%
+Since then, the protocol has seen multiple changes.
+%
+In May 2018, the designers replaced the challenge-response signature core with
+one based on \mSigma{}
+(SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}, and
+in
+2020, three new cores, which mixed challenge-response signatures and regular
+signatures were added as well~\cite{our-analysis-selander-lake-edhoc-00}.
+%
+While this is the version we base our formal analysis on, there have been later versions. 
+%
+See Section~\ref{sec:newdrafts} for a more detailed discussion.
+
+%-------------------------------------------------------------------------- sub
+\subsection{Related Work and Contributions}
+\label{sub:related}
+The earliest related work is~\cite{DBLP:conf/secsr/BruniJPS18}, which formally
+analyzes the May 2018 version of \mEdhoc{} using the \mProverif{}
+tool~\cite{DBLP:conf/csfw/Blanchet01}.
+%
+In this paper, the authors analyze two key establishment methods -- one built
+on
+a pre-shared key authentication core, and one based on \mSigma{}.
+%
+The authors check  the following properties: secrecy, identity protection, strong authentication, perfect forward secrecy
+(PFS), and integrity of application data.
+%
+Later work~\cite{Norr21} analyzes the July 2020 version of \mEdhoc{} in the
+\mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13}.
+%
+This version of the protocol has four key establishment methods.
+%
+In~\cite{Norr21}, the properties checked for are injective agreement, implicit
+agreement, and perfect forward secrecy for the session key material.
+%
+That work also includes a discussion about the various design choices made as
+part of \mEdhoc{}, and the impact of \mEdhoc{} in multiple use-case scenarios.
+%%
+
+%-------------------------------------------------------------------------- sub
+\subsection{Contributions}
+\label{sec:contributions}
+In this paper, we extend the work presented in~\cite{Norr21}.
+%
+We formally analyze the \mEdhoc{} specification as of July
+2020~\cite{our-analysis-selander-lake-edhoc-00}, but our formal analysis applies as far as the version of the \mSpec{} from February 2021.
+%
+We extend the adversary model and \mTamarin{} system models to capture weak
+post-compromise security (PCS) and model Trusted Execution Environments
+(TEE).
+%
+We also alter the formal modelling used for encryption under \mXor{}.
+%
+We formally verify the following properties:
+\begin{itemize}
+\item Injective agreement
+\item Implicit agreement
+\item Perfect Forward Secrecy (PFS) for the session-key material
+\item Weak post-compromise security for the session-key material
+\item Secrecy and integrity of \mADthree
+\end{itemize}
+%
+We follow the definition of weak PCS by~\cite{DBLP:conf/csfw/Cohn-GordonCG16},
+which subsumes PFS.
+%
+Since \mEdhoc{} is still a protocol under development, the latest version of \mEdhoc{} is the one from July 2022~\cite{draft-ietf-lake-edhoc-15}, which contains details which are not covered by our formal model.
+%
+We do, however, refer to various aspects
+(error handling, denial of service etc) of the July 2022 version of the \mSpec{}, namely 
+in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.
+%
+We also discuss various issues arising due to the use of trusted
+execution environments (TEEs), denial of service (DoS) attacks, error handling,
+the negotiation of parameters (which the formal model abstracts
+away) for establishing the protocol, and other potential attacks and concerns.
+%
+We have communicated these issues to the developers of the protocol.
+%
 
@@
 \label{sec:edhoc}
-\input{protocol}
+% !TEX root = paper.tex
+In this section, we describe the various key establishment methods of the
+\mEdhoc{} protocol.
+%
+Following~\cite{Norr21}, we refer to the two roles executing the protocol as
+the initiator $I$ and the responder $R$.
+%
+We annotate values with $I$ and $R$ to make explicit which role they belong
+to.
+%
+
+%------------------------------------------------------------------------- sub
+\subsection{Notation}
+\label{sec:notation}
+We denote by $\langle d_{t, \mathit{id}}, Q_{t, \mathit{id}}\rangle$
+public-private key pairs, where $d$ is the private key, $Q$ is the public key,
+$t \in \{e, s\}$ indicates whether the key is ephemeral or static,
+and $\mathit{id}$ indicates the party to whom the key pair belongs.
+%
+When clear from context, we will often drop some (or all) of the subscripts.
+%
+Static key pairs (suitable for regular or challenge-response signatures)
+are long-term authentication credentials, whereas ephemeral key pairs are
+those generated afresh for each execution of the protocol.
+%
+
+Parties can authenticate using regular signatures or challenge-response
+signatures.
+%
+In the former case, we say that they use the
+\emph{signature-based authentication method} (\mSig{}).
+%
+In the latter case, we say, following the terminology in the \mSpec{},
+that they use the \emph{static key authentication method} (\mStat{}).
+%
+We adopt the challenge-response terminology
+from~\cite{DBLP:conf/crypto/Krawczyk05}.
+%
+
+\mEdhoc{} fundamentally uses elliptic curves and associated Diffie-Hellman
+operations.
+%
+Signatures using a party $A$'s keys are denoted by \mSign{A}$(\cdot)$, while
+the verification thereof is denoted by \mVf{A}$(\cdot)$.
+%
+A Diffie-Hellman operation which combines a private key $d$ and a point $P$
+on the elliptic curve is represented as $\mDH(d, P)$.
+%
+We will often overload notation to let $P$ stand for both the point on the
+elliptic curve as well as the corresponding bitstring encoding.
+
+%------------------------------------------------------------------------- sub
+\subsection{Overall Description}
+\label{sec:description}
+\mEdhoc{} is designed to establish a security context for the \mOscore{} protocol.
+%
+This context, in particular, includes the session-key material (we denote this by
+\mSessKey{}).
+%
+The generalized abstract protocol as given in the \mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} consists of three
+messages, and is shown in Figure~\ref{fig:edhocFramework}.
+%
+The abstract structure is the same across methods.
+%
+However, the authentication mechanisms and key derivation procedures differ between
+methods.
+%
+\mEdhoc{} may also transfer application data \mADone{}, \mADtwo{}, and \mADthree{} in
+addition to establishing the \mOscore{} security context.
+%
+
+%
+\begin{figure}[ht]
+\centering
+\tikzset{>=latex, every msg/.style={draw=thick}, every
+node/.style={fill=none,text=black}}
+\begin{tikzpicture}
+    \node (ini) at (0, 0) {Initiator};
+    \draw [very thick] (0, -0.25) -- (0,-2.3);
+    \draw [very thick] (5.75, -0.25) -- (5.75,-2.3);
+    \node (res) at (5.75,0) {Responder};
+    \msg{1em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi,
+\mADone};
+    \msg{3em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, \{\mIdcredr, \mAuthr,
+\mADtwo\}};
+    \msg{5em}{ini}{res}{\mMsgthree: \mCr, \{\mIdcredi, \mAuthi, \mADthree\}};
+    \draw [line width=1mm] (-0.75,-2.3) -- (0.75,-2.3);
+    \draw [line width=1mm] (5.75-0.75,-2.3) -- (5.75+0.75,-2.3);
+    \node (padding) at (0,-2.5) {};
+    \end{tikzpicture}
+    \caption{Structure of \mEdhoc{}: $\{t\}$ means $t$ is encrypted and integrity
+protected.~\cite{Norr21}}
+\label{fig:edhocFramework}
+\end{figure}
+%
+Of the three messages \mMsgone, \mMsgtwo, and \mMsgthree, the first two, among other things, establish a
+common authentication method \mMethod{} and ciphersuite \mSuites{}.
+%
+The party playing the initiator role uses \mMethod{} to propose which
+authentication methods the two parties shall use, and in \mSuites{}, proposes an ordered list
+of choices for the ciphersuite.
+%
+The chosen authentication methods may differ for the two roles,
+yielding four possible
+combinations: \mSigSig{}, \mSigStat{}, \mStatSig{}, and \mStatStat{},
+where the first authentication method in each combination is used by the
+initiator, and the second by the responder.
+\footnote{As in the \mSpec{}, we will from now on overload
+notation and refer to the combinations of authentication methods as methods
+as
+well.}
+%
+The party executing the responder role may choose to reject the method and/or
+ciphersuite chosen by the initiator by sending an error message.
+%
+This results in abandoning this session and renegotiating, as the initiator
+goes down their list of choices for ciphersuites, and picks the next option for
+a next execution of the protocol.
+%
+Our analysis does not cover such renegotiation which requires maintaining
+state
+between executions to remember the rejected ciphersuites.
+%
+However, we will discuss the ramifications of such a renegotiation procedure
+and the error messages later, in Section~\ref{sec:errorHandling}.
+%
+
+In addition to negotiating the method and ciphersuite, the first two messages are
+also instrumental for the exchange of public ephemeral keys \mGx{} and
+\mGy{},
+and connection identifiers \mCi{} and \mCr{}, for the initiator and responder
+roles respectively.
+%
+The \mSpec{} states that the connection identifiers serve only to route
+messages
+to the correct party executing \mEdhoc{}, but also claims that they may be
+used in turn by protocols (like \mOscore{}) which use the security context
+established by \mEdhoc{}.
+%
+While the \mSpec{} does not require any explicit security guarantees to be
+satisfied by these connection identifiers, it does, however, require that the
+identifiers be unique, i.e., in any session, $\mCi{} \neq \mCr{}$, and that the
+parties involved in the session can verify this uniqueness.
+%
+More precisely, the \mSpec{} states that \mOscore{} should be able to use these
+identifiers to retrieve any particular security context.
+%
+In this work, as in~\cite{Norr21}, we verify that the parties agree on the
+values of \mCi{} and \mCr{}.
+%
+The second and third messages also serve to identify and authenticate each
+party to the other.
+%
+These messages contain long-term key identifiers (\mIdcredi{} and
+\mIdcredr{}).
+%
+Additionally, the messages contain authenticating information
+(\mAuthi{} and \mAuthr{}), which lets each
+party know that the other party does indeed control the long-term key
+associated with these identifiers.
+%
+The authentication information is structured differently for each authentication method.
+%
+
+Consider the following scenario.
+%
+The initiator $I$ chooses the method $\mSigStat{}$, and sends this via
+$\mMethod{}$ to the responder $R$.
+%
+$R$ now has the option to reject this choice of method.
+%
+However, if $R$ chooses to accept this method, they need to provide an
+identifier \mIdcredr{} for a key pair which can be used with
+challenge-response
+signatures
+as well as authenticating information \mAuthr{}, as dictated by
+the static key authentication method \mStat{}.
+%
+$I$ will then respond with an identifier \mIdcredi{} corresponding to a key
+pair, which is suitable for regular signatures, and provide authentication
+\mAuthi{} as appropriate for the signature-based method \mSig{}.
+
+%-------------------------------------------------------------------------- sub
+\subsection{Key Schedule}
+\label{sec:keysched}
+The second and third messages of \mEdhoc{} contain
+authenticating information.
+%
+This information is either a signature or a message authentication code (MAC),
+as we will describe in more detail in Section~\ref{sec:edhocauth}.
+%
+The keys for these are generated using a key schedule which is intrinsic to the functioning of \mEdhoc{}.
+%
+The key schedule takes a DH key \mGxy{} as basic input and builds upon it, as shown in Figure~\ref{fig:kdfdiagram}.
+%
+\begin{figure}[!h]
+\scalebox{.785}{
+\begin{tikzpicture}[%
+    >=latex,              % Nice arrows; your taste may be different
+    start chain=going below,    % General flow is top-to-bottom
+    node distance=4mm and 60mm, % Global setup of box spacing
+    every join/.style={norm},   % Default linetype for connecting boxes
+    ]
+\tikzset{
+terminput/.style={rounded corners},
+term/.style={rounded corners},
+  base/.style={draw, thick, on chain, on grid, align=center, minimum height=4ex},
+  dhkbox/.style={draw=cbgreen, fill=cbgreen!25, rectangle},
+  dhk/.style={base, dhkbox},
+  prkbox/.style={draw=cborange, fill=cborange!25, rectangle},
+  prk/.style={base, prkbox},
+  hkdfext/.style={base, draw=black, fill=none, rectangle},
+  hkdfexp/.style={base, draw=black, fill=none, rectangle},
+  keybbox/.style={draw=cbnavy, fill=cbnavy!25, rectangle},
+  keyb/.style={base, keybbox, text width=4em},
+  norm/.style={->, draw, black},
+  cond/.style={base, draw=black, dashed, fill=none, rectangle},
+  txt/.style={base, draw=none, fill=none}
+  }
+\node [prk, join] (p2) {\mPRKtwo};
+\node [cond, join] (c1) {R uses \mStat?};
+
+\node [prk, below=6mm of c1.south] (p3) {\mPRKthree};
+\draw [->, norm] (c1.south) -- (p3.north) node[midway, right] {N};
+
+\node [cond, join, below=10mm of p3.south] (c2) {I uses \mStat?};
+\node [prk, below=8mm of c2.south] (p4) {\mPRKfour};
+\draw [->, norm] (c2.south) -- (p4.north) node[midway, right] {N};
+
+\node [hkdfext, right=3cm of p3] (h3) {\mHkdfExtract};
+\node [hkdfext, right=3cm of p4] (h5) {\mHkdfExtract};
+
+\node [hkdfexp, shape border rotate=180, left= 2.5cm of p4] (h6) {\mHkdfExpand};
+\node [keyb, join, left=3cm of h6] (k3) {\mKthreem};
+\node [hkdfexp, shape border rotate=180, below= 0.8cm of h6] (h9) {\mHkdfExpand};
+\node [txt, join, left=1cm of h9.west] (t4) {EDHOC-Exporter};
+
+\node [hkdfexp, shape border rotate=180, left= 2.5cm of p3] (h4) {\mHkdfExpand};
+\node [keyb, join, left=3cm of h4] (k2) {\mKtwom};
+
+\node [hkdfexp, shape border rotate=180, left= 2.5cm of p2] (h2) {\mHkdfExpand};
+\node [keyb, join, left=3cm of h2] (k1) {\mKtwoe};
+
+\node [hkdfexp, shape border rotate=180, below= 0.8cm of h4] (h8) {\mHkdfExpand};
+\node [keyb, below=0.8cm of k2] (k2b) {\mKthreeae};
+
+\node [txt, left=1cm of k1.west] (t1) {Enc (XOR) \\ in m2};
+\node [txt, left=1cm of k2.west] (t2) {\mMactwo~(signed if \\ R uses \mSig)};
+\node [txt, left=1cm of k2b.west] (t2b) {\mAead\ in m3};
+\node [txt, left=1cm of k3.west] (t3) {\mMacthree~(signed if \\ I uses \mSig)};
+
+\draw [->, norm] (k1.west) -- (t1.east);
+\draw [->, norm] (k2.west) -- (t2.east);
+\draw [->, norm] (k2b.west) -- (t2b.east);
+\draw [->, norm] (k3.west) -- (t3.east);
+
+\draw [->, norm] (p3.south) ++(0,-0.5) -- (h8);
+\draw [->, norm] (h8) -- (k2b);
+\draw [->, norm] (p2) -- (h2); 
+\draw [->, norm] (c1.east) -- ++(2.025,0) -- (h3.north) node[midway,above left] {Y};
+\draw [->, norm] (h3.west) -- (p3.east);
+\draw [->, norm] (p3) -- (h4); 
+\draw [->, norm] (c2.east) -- ++(2.07,0) -- (h5.north) node[midway,above left] {Y};
+\draw [->, norm] (h5.west) -- (p4.east);
+\draw [->, norm] (p4) -- (h6);
+\draw [->, norm] (p4.west) ++(-0.25,-0) -- ++(0,-0.8) -- (h9.east);
+
+\node [hkdfext, right=3cm of p2] (h1) {\mHkdfExtract};
+\node [dhk, right=2.7cm of h1] (p0) {$\mGxy$};
+\node [terminput, text width=2em, below = 0.2cm of p0] (u1) {Salt};
+\draw [->] (h1.west) -- (p2.east);
+\draw [->] (u1.west) -- ++(-2.27,0) -- (h1.south);
+\draw [->] (p0.west) -- (h1.east);
+
+\node [dhk, right = 2.7cm of h3] (u2) {$P_R$};
+\draw [->, norm] (u2.west) -- (h3.east);
+
+\node [dhk, right = 2.7cm of h5] (u3) {$P_I$};
+\draw [->, norm] (u3.west) -- (h5.east);
+
+\node [term, above = 0.55cm of h4] (u5) {\mTHtwo};
+\draw [->, dotted, shorten >=1mm] (u5) -- (h4);
+\draw [->, dotted, shorten >=1mm] (u5) -- (h2);
+
+\node [term, above = 0.5cm of h6] (u6) {\mTHthree};
+\draw [->, dotted, shorten >=1mm] (u6) -- (h6);
+\draw [->, dotted, shorten >=1mm] (u6) -- (h8);
+
+\node [term, below=0.4cm of p4] (u7) {\mTHfour};
+\draw [-> , dotted ] (u7.west) -- ([yshift=-0.4em] h9.east);
+\end{tikzpicture}
+}
+\caption{Key schedule for~\cite{our-analysis-selander-lake-edhoc-00}: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo,
+\mPRKthree, \mPRKfour{} are the intermediate key material, and
+\mKtwoe,\mKtwom,
+\mKthreeae, \mKthreem{} are the encryption keys for \mAead{} or
+\mXor{}. Dashed boxes are conditionals.~\cite{Norr21}}
+\label{fig:kdfdiagram}
+\end{figure}
+%
+
+To derive keys, \mEdhoc{} uses two functions from the \mHkdf{}
+interface~\cite{rfc5869}, \mHkdfExtract{} and \mHkdfExpand{}.
+%
+Both functions take as argument two values -- an input and a salt.
+%
+For \mHkdfExtract{}, the input is a DH key, while for \mHkdfExpand{},
+it is intermediate key material.
+%
+
+As mentioned earlier, the fundamental building block for the key schedule is
+the ephemeral DH key \mGxy{}, which is computed in two different ways by
+$I$ (as $\mDH(\mX, \mGy)$) and $R$ (as $\mDH(\mY, \mGx)$).
+%
+This key gives rise to intermediate keys \mPRKtwo{}, \mPRKthree{} and
+\mPRKfour{}, which can be derived as part of protocol execution.
+%
+Each intermediate key gives rise to encryption and integrity keys
+(\mKtwoe, \mKtwom{}, \mKthreeae, and \mKthreem)
+corresponding to each message in the protocol.
+%
+
+In order to generate the final keys, the two \mHkdf{} algorithms use various
+values for salt.
+%
+\mPRKtwo{} is generated by the \mHkdfExtract{} algorithm while using the
+empty
+string as the salt.
+%
+\mPRKthree{} and \mPRKfour{} are separately generated if $R$ or $I$ uses the
+\mStat{} method, using the corresponding DH key as input and the previous
+intermediate key as salt.
+%
+The key $P_{R}$, which is computed as $\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\
+\mGx)$, is used if the responder uses the \mStat{} authentication
+method.
+%
+Similarly, the key $P_{I}$ is used if the initiator uses \mStat{}, and is
+computed as $\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$.
+%
+
+These intermediate keys are fed into \mHkdfExpand{}, which uses as salt a value $\textit{th}$, which is a running
+hash of the information transmitted thus far as part of the protocol.
+%
+By $\textit{th}_{i}$, we denote the hash corresponding to the $i^{\rm{th}}$ message.
+%
+
+At the end of a successful run of the protocol, the session-key material is
+established as \mSessKey{}, which we define as a set of various keys.
+%
+This set always includes \mGxy{}, and if the initiator (resp. the responder) uses
+the \mStat{} authentication method, then it also includes $P_{I}$ (resp.
+$P_{R}$).
+%
+We
+discuss which material should be included in \mSessKey{} in more detail
+in Section~\ref{sec:formalization}, and the consequences of
+various choices in Section~\ref{sec:sessionKeyMaterial}.
+%
+Once \mSessKey{} has been established, an \mHkdf{}-based key exporter
+named
+\mEdhoc-Exporter extracts the keys required by the security protocol.
+%
+
+%-------------------------------------------------------------------------- sub
+\subsection{About Authentication in \mEdhoc{}}
+\label{sec:edhocauth}
+We now describe how the authentication information is constructed,
+depending on whether the \mSig{} or \mStat{} method is used.
+%
+For both methods, the following information is used to compute
+\mAuthr{}: \mIdcredr{}, \mCredr{}, a transcript hash of all the communicated
+information thus far in the protocol, and \mADtwo{}, if included.
+%
+\mAuthi{} uses the same pieces of information, but corresponding to the
+initiator role.
+%
+A MAC is obtained by feeding this material as additional data and the empty
+string as input to the Authenticated Encryption with Additional Data (AEAD) algorithm as indicated in the established ciphersuite \mSuites{}.
+%
+The encryption key for the AEAD algorithm is constructed, for both roles,
+using the ephemeral key material \mGx{}, \mGy{}, \mX{}, and \mY{}.
+%
+The initiator computes $\mDH(\mX, \mGy)$ while the responder computes
+$\mDH(\mY, \mGx)$, and DH operations give rise to the same key under both
+these computations.
+%
+
+When $I$ uses the \mSig{} authentication method, \mAuthi{} is $I$'s signature
+over the MAC computed as above, along with the data covered by the MAC.
+%
+However, when $I$ uses the \mStat{} method, \mAuthi{} is just the MAC, with
+one difference: the MAC key is derived using both the ephemeral key material
+\mGxy{} as well as the long-term key for the initiator
+$\langle\mPriv{I},\ \mPub{I}\rangle$.
+%
+This is similar to the 1-RTT semi-static pattern in \mOptls{} which computes the
+MAC key \textsf{sfk} for the message
+\textsf{sfin}~\cite{DBLP:conf/eurosp/KrawczykW16}.
+%
+The same procedures works for $R$ as well (with the values corresponding to
+$R$).
+%
+The abstract calculation for the authentication information values is as shown
+in Table~\ref{tab:authvalues}~\cite{Norr21}.
+%
+We use $\mathit{MAC}_{R}$ (resp. $\mathit{MAC}_{I}$) to denote a MAC
+which uses
+an encryption key constructed using
+$\langle\mPriv{R},\ \mPub{R}\rangle$ and $\langle\mX,\ \mGx\rangle$
+(resp. using $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\
+\mGy\rangle$).
+%
+\begin{table}[ht]
+\centering
+\begin{tabular}{|c|c|c|}
+        \hline
+        \mMethod & \mAuthi & \mAuthr\\
+        \hline
+        \mSigSig{} & $\mSign{I}(\cdot)$ & $\mSign{R}(\cdot)$ \\
+        \mSigStat{} & $\mSign{I}(\cdot)$ & $\textit{MAC}_R(\cdot)$\\
+        \mStatSig{} & $\textit{MAC}_I(\cdot)$ & $\mSign{R}(\cdot)$\\
+        \mStatStat{} & $\textit{MAC}_I(\cdot)$ & $\textit{MAC}_R(\cdot)$\\
+        \hline
+\end{tabular}
+\caption{The outer functions for each method \mMethod{}~\cite{Norr21}}
+\label{tab:authvalues}
+\end{table}
+%
+
+In addition to this, the second and third messages have
+provisions for transferring application messages which are encrypted and
+integrity protected, as can be seen in~\ref{fig:edhocFramework}.
+%
+The second message is encrypted by performing a bit-wise \mXor{} between
+the plaintext and the output of the key derivation function \mHkdf{}, as
+in Section~\ref{sec:keysched}.
+%
+For the third message, encryption and integrity protection is assured by the
+\mAead{} algorithm.
+%
+
+In Figure~\ref{fig:edhocsigstat} we show an example of protocol execution
+under
+the \mSigStat{} method.
+%
+The figure describes in detail the various message patterns, operations and
+key derivations used to construct these messages.
+%
+\begin{figure}[ht]
+\centering
+\scalebox{.7}{
+\tikzset{>=latex, every msg/.style={draw=thick}, every
+node/.style={fill=none,text=black}}
+\begin{tikzpicture}
+    \node (ini) at (0, 0) {Initiator};
+    \draw [very thick] (0, -0.5) -- (0,-14.8);
+    \draw [very thick] (9, -0.5) -- (9,-14.8);
+    \node[below=0.5em of ini,fill=white] {$
+    \begin{array}{c}
+        \text{Knows}\ \langle\mPriv{I},\ \mPub{I}\rangle,\ \mIdcredi,\ \mIdcredr,\
+\mADone,\ \mADthree
+    \end{array}
+    $};
+    \node (res) at (9,0) {Responder};
+    \node[below=0.5em of res,fill=white] {$
+    \begin{array}{c}
+        \text{Knows}\ \langle\mPriv{R},\ \mPub{R}\rangle, \ \mIdcredr,\ \mADtwo
+    \end{array}$};
+    \action{3.7em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\
+\mGx\rangle$};
+    \msg{7em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi,
+\mADone};
+    \action{8em}{res}{$
+      \begin{array}{c}
+          \text{Generates } \mCr,\ \langle\mY{},\ \mGy\rangle\\
+          \ \ P_e = \mDH(\mY,\ \mGx{})\\
+          \ \ P_R = \mDH(\mPriv{R},\ \mGx{})\\
+        \mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle)\\
+        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
+        \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
+        \mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
+        \mMactwo = \mAead(\mKtwom; \langle \mIdcredr, \mTHtwo, \mCredr,
+\mADtwo \rangle; \textrm{``\phantom{}''}) \\
+        \mKtwoe = \mHkdfExpand(\mPRKtwo, \mTHtwo)
+      \end{array}$};
+    \msg{24em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\
+\mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
+    \action{25em}{ini}{$
+      \begin{array}{c}
+        %\mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle) \
+        \ P_e = \mDH(\mX,\ \mGy{})\\
+        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
+        %\mKtwoe = \mHkdfExpand(\mPRKtwo,\mTHtwo)\\
+        %\mGrx = \mCredr^{x} \\
+        \ \ P_R = \mDH(\mX,\ \mPub{R})\\
+        \mPRKfour = \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
+        %\mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
+        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
+        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
+        \mKthreem = \mHkdfExpand(\mPRKfour, \mTHthree) \\
+        \mMacthree = \mAead(\mKthreem; \langle \mIdcredi, \mTHthree,
+\mCredi, \mADthree \rangle; \textrm{``\phantom{}''}) \\
+        \mSigthree = \mSign{I}(\langle \langle \mIdcredi, \mTHthree, \mCredi,
+\mADthree \rangle, \mMacthree \rangle)
+      \end{array}$};
+    \msg{38.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae;
+\mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
+    \action{40em}{res}{$
+    \begin{array}{c}
+        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
+        \mKthreem = \mHkdfExpand(\mPRKthree, \mTHthree) \\
+        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHthree)
+    \end{array}$};
+    \draw [line width=1mm] (-2,-14.8) -- (2,-14.8);
+    \draw [line width=1mm] (7,-14.8) -- (11,-14.8);
+    \end{tikzpicture}
+}
+    \caption{The \mSigStat{} method for \mEdhoc{}. $\langle\cdot\rangle$
+denotes a tuple, and the hash function \mHash{} is as established in the
+ciphersuite \mSuites{}.~\cite{Norr21}}
+\label{fig:edhocsigstat}
+\end{figure}
+
+%------------------------------------------------------------------------- sec
+\section{Implementation Aspects and Key Protection}
+\label{sec:TEE}
+Authentication of a specific IoT device assumes that the device is the only
+entity with access to the long-term key associated with its
+identity.
+%
+Since IoT devices may be accessible to adversaries, e.g., an insider cloning a
+key card, long-term keys must be appropriately protected.
+%
+A state-of-the-art approach is to use a trusted execution environment (TEE),
+which holds the key and provides an API for operations using the key.
+%
+This is the approach taken by the
+TrustZone-based \mMuEdhoc{}~\cite{DBLP:conf/codaspy/Hristozov0XFLS21}
+implementation, for example.
+%
+Typical operations include signatures using the long-term private key of a
+party.
+%
+
+TEEs can be of differing complexity.
+%
+Some, like ARM TrustZone and Intel SGX, are general-purpose execution
+environments, which can be programmed in many different ways.
+%
+Others, like the Universal Subscriber Identity Modules (USIM) used for
+authentication to 3GPP mobile networks, have application-specific interfaces
+for
+authentication, key agreement protocols etc.
+%
+
+A fundamental aspect of TEEs is how much of the application is placed in the TEE.
+%
+For larger devices that include general-purpose processors with TrustZone or
+SGX, entire \mEdhoc{} and \mOscore{} implementations may reside inside the TEE.
+%
+For constrained IoT devices on the lower end of the scale, a TEE might have to be
+implemented using a special-purpose integrated circuit.
+%
+In the latter case, it may be beneficial to follow a minimalistic approach, and
+only store in the TEE the long-term key and operations that need access to it, to reduce cost.
+%
+
+In general, it might appear more secure to implement as much as possible inside the TEE, but there is a security trade-off.
+%
+Because security-critical code runs in the same area as where long-term keys
+reside, an implementation error here risks leaking information of the key to
+the adversary.
+%
+From this perspective, it might be beneficial to follow the minimalistic approach
+even when having access to TrustZone or SGX.
+%
+
+A slightly more secure division of functionality is to keep both the long-term key and the session key inside the TEE and extend the interface to accept messages and return the (en/de)crypted counterpart, i.e.
+the interface exposes \mAead{} functions.
+%
 
@@
 \label{sec:formalization}
-\input{formalization}
+% !TEX root = paper.tex
+The \mEdhoc{} \mSpec{} \cite{our-analysis-selander-lake-edhoc-00} claims
+that \mEdhoc{} satisfies many security properties, but these are imprecisely
+expressed and motivated.
+%
+In particular, there is no coherent adversary model.
+%
+It is therefore not clear in which context properties should be verified.
+%
+We resolve this by clearly specifying an adversary model, in which we can 
+verify
+properties.
+%
+
+%---------------------------------------------------------------------- subsub
+\subsection{Formal Model}\label{sec:threat-model}
+As in~\cite{Norr21}, we verify \mEdhoc{} in an extended
+Dolev-Yao model~\cite{DY83}.
+%
+The Dolev-Yao model is well established for the symbolic verification of 
+security
+protocols.
+%
+Messages are modelled as terms in an algebra, and the various cryptographic
+operations are assumed to be perfect, e.g., encrypted messages can only be
+decrypted with the correct key, there are no hash collisions etc.
+%
+
+The adversary is assumed to be in control of the communication
+channel and can see all messages being communicated as part of the protocol.
+%
+In addition, they can interact with an unbounded number of protocol sessions,
+and drop, inject and modify messages at will.
+%
+
+On top of the standard Dolev-Yao model, \cite{Norr21} allows the adversary to
+access long-term and ephemeral keys via specific events.
+%
+Long-term key reveal, denoted by $\mRevLTK^{t}(A)$, stands for the adversary
+gaining access to a party $A$'s long-term private key \mPriv{A} at a time point 
+$t$.
+%
+Ephemeral key reveal, denoted by $\mRevEph^{t}(A, k)$, stands for the 
+adversary
+obtaining, at time $t$, the ephemeral private key \mPrivE{A} used by party $A$
+in a session where they establish a session key $k$.
+%
+Formalizing these two capabilities allows more fine-grained control
+over the access that an adversary has to these fundamentally different kinds of
+keys.
+%
+Furthermore, we modify the model from~\cite{Norr21} by strengthening the
+adversary capabilities, as detailed next.
+%
+
+%---------------------------------------------------------------------- subsub
+\subsubsection{System and Adversary Model Extensions Supporting TEE.}
+\label{sec:TEE:advModel}
+%
+We extend the adversary model of~\cite{Norr21} by allowing the adversary to 
+use the long term key of any party via an interface, without directly accessing 
+the key itself.
+%
+This allows us to model the scenario where the adversary has gained access to
+a device, but the long-term key is protected by a trusted execution
+environment (TEE), and thus only accessible via the TEE interface.
+%
+According to the terminology of~\cite{DBLP:conf/csfw/Cohn-GordonCG16}, a
+protocol is considered to enjoy \emph{weak post-compromise security}
+(Weak-PCS) if it achieves its security goals for a specific session
+even when the following hold:
+\begin{itemize}
+\item Before the session starts, the adversary has limited access to the 
+long-term 
+keys of the parties involved, through an interface that securely maintains 
+the keys but allows principals to run cryptographic operations using them, and
+\item The adversary has full access to the keys of the parties involved after the 
+end of the session, as well as access to the keys of all other parties.
+\end{itemize}
+%
+As in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add 
+adversary
+capabilities corresponding to a server adversary, i.e., upon 
+compromising a party, the adversary learns their ephemeral keys and may temporarily access their TEE, but does not learn their long-term key.
+%
+This extension to the model of~\cite{Norr21} thus allows us to verify all the
+previous security properties under a ``Weak-PCS model''.
+%
+Note that this attacker model is strictly more powerful than that 
+of~\cite{Norr21}, as it subsumes the previous attacker capabilities.
+%
+
+We split the functionality of \mEdhoc{} as follows.
+%
+The TEE contains the long-term key and allows the non-TEE parts of the
+application to perform the operations using it via an interface.
+%
+More precisely, parties using the \mSig{} authentication method use a TEE 
+interface which accepts a message and returns the signature of that 
+message using the party's private long-term key.
+%
+A party $U$ using the \mStat{} authentication method uses a TEE with an 
+interface accepting a point $P$ on the curve and 
+%returning $\mDH(P,\ \mGu{})$.
+returning $\mDH(\mPriv{U},\ P)$.
+%
+
+%The non-TEE parts implement \mEdhoc{} using this interface.
+%
+This interface requires the least functionality from the TEE, reducing the 
+TEE's complexity and (possibly) its cost.
+%
+This functional split is suitable even when a constrained device has
+implemented the storage of only the long-term key in 
+a special purpose circuit with minimal processing functionality.
+%
+Since \mEdhoc{} focuses on constrained IoT devices, it seems appropriate to
+cater for this setting.
+%
+
+%---------------------------------------------------------------------- subsub
+\subsubsection{Extended Formalism and Security Properties.}
+\label{sec:TEE:fmAndProps}
+Formally, we model the TEE interface by adding two new rewrite rules:
+%
+\begin{small}
+\begin{verbatim}
+rule forge_SIG:
+   [!LTK_SIG($A, ~ltk), In(xx)] --[TEE($A)]-> [Out(sign(xx, ~ltk))]
+
+rule exp_STAT:
+   [!LTK_STAT($A, ~ltk), In('g'^~xx)] --[TEE($A)]-> [Out(('g'^~xx)^~ltk)]
+\end{verbatim}
+\end{small}
+%
+These rules allow the adversary to obtain terms representing signatures on a
+value of their choice (\verb|forge_SIG|), or to obtain terms
+representing a curve point of their choice raised to the power of the
+long-term key (\verb|exp_STAT|).
+%
+
+Because it is a trivial attack when the adversary accesses these rules with values
+from a session in progress, we must disqualify those rule applications.
+%
+We do so by creating an action fact \verb|TEE($A)|, where \verb|$A| is the
+identity corresponding to the private key used, and then augmenting the
+properties with a condition that no such action fact exists from the start of
+the protocol execution till its end.
+%
+Care must be taken when specifying the start and the end: specifically,
+the start and end of the execution must be viewed with respect to the current
+role.
+%
+For example, the injective agreement property for the initiator in
+the \mSigSig{} method requires that the adversary does not have access to the
+TEE of the responder from the time that the first message is transmitted till
+the second message is received by the initiator (or equivalently, till the third message is transmitted by the initiator, since reception of a message and transmission of the next one by a party is one atomic operation).
+%
+Thus, these timepoints represent the start and end of the
+protocol run from the perspective of the initiator.
+%
+
+The second message encrypts two values: the responder's 
+identity
+\verb|R| and the authentication information \verb|authR|.
+%
+We model XOR encryption by \mXor-ing each term with their own key-stream 
+term.
+%
+However for some problematic methods\footnote{See Table~\ref{tab:props} for all verification results and computation times.} we \mXor{} the entire tuple
+\verb|<R, authR>| with a single key-stream term.
+%
+This simplification is likely to miss an attack on implicit authentication which occurs due to the combination of a malleable XOR encryption and access to the TEE 
+interface.
+%
+However, given that no attacks were identified due to the use of XOR in our original modeling, nor in the current modeling for all other authentication methods, we believe that this is not a severe restriction.
+%
+
+Next we provide a quick overview of the \mTamarin{} tool, and describe our modeling of \mEdhoc{} in it.
+%
+
+%-------------------------------------------------------------------------- sub
+\subsection{\mTamarin{}}
+\label{sec:tamarin}
+We extend the formal \mEdhoc{} model of~\cite{Norr21}, using the tool \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an 
+interactive
+tool for the symbolic verification of security protocols.
+%
+Protocols are modelled in \mTamarin{} as multiset rewrite rules which encode 
+a
+transition relation.
+%
+The elements of these multisets, called facts, contribute to the global
+system state.
+%
+For syntactic sugar, \mTamarin{} also allows the use of let-bindings and tuples.
+%
+For ease of presentation, we will present the model and properties in a
+slightly different syntax in this paper, but this syntax can be directly
+mapped to that of \mTamarin.
+%
+
+Rewrite rules can be annotated with events, called actions in \mTamarin{}.
+%
+Communicated messages in the protocol are modelled as terms in an algebra,
+which specifies sets of names, variables, and allowable function symbols.
+%
+Facts and actions are modelled as $n$-ary predicates in the term algebra,
+and actions can be parametrized using terms.
+%
+
+An annotated multiset-rewrite rule is represented as
+$l \ifarrow[e] r$, where $l$ and $r$ are multisets, and $e$ is
+a multiset of actions.
+%
+A sequence of actions yields a protocol execution trace.
+%
+Properties are defined as formulas in a fragment of temporal first order logic,
+and these formulas can be verified over execution traces.
+%
+Event types are predicates over global states generated during protocol 
+execution.
+%
+Consider an event type $E$ and a timestamp $t$ as part of a trace.
+%
+By $E^{t}(p_i)_{i\in\mathbb{N}}$, we represent an event of type $E$ occurring
+at time $t$ in a trace, parametrized by the sequence of values
+$(p_i)_{i\in\mathbb{N}}$
+(corresponding to the action fact $E(p_i)_{i\in\mathbb{N}}@t$ in \mTamarin).
+%
+Thus, the time points form a quasi order, and we denote the fact that
+$t_{1}$ comes before $t_{2}$ in a protocol trace by
+$t_{1} \lessdot t_{2}$, and that $t_{1}$ and $t_{2}$ stand for the same
+time point in a trace by $t_{1} \doteq t_{2}$.
+%
+\mTamarin{} allows events to occur at the same time point, with one
+restriction: multiple events of the same type cannot occur simultaneously,
+so if $t_{1} \doteq t_{2}$, then $E^{t_{1}} = E^{t_{2}}$.
+%
+
+Protocol verification in \mTamarin{} happens under an equational theory $E$.
+%
+For example, to represent the fact that $E$ satisfies the reversal of
+symmetric encryption by using a decryption operation with the key,
+one can write $\textit{dec}(\textit{enc}(x, y), y) =_{E} x$.
+%
+The equational theory $E$ is fixed upfront to handle the functions supported 
+by
+the term algebra, so we will omit the subscript for the rest of this paper.
+%
+
+Users can extend the default term algebra and equational theory in
+\mTamarin{} with new function symbols and unification rules (if any) for these new 
+symbols.
+%
+For example, \mEdhoc{} requires authenticated encryption, which we model 
+using
+the symbol \mT{aeadEncrypt}.
+%
+We augment \mTamarin{} with the following rule for this operation,
+which represents the fact that if the adversary knows a key \mT{k},
+a message \mT{m}, and authenticated data \mT{ad},
+and has access to an encryption algorithm \mT{ai},
+then they can obtain the message corresponding to the authenticated 
+encryption
+of \mT{m} with \mT{k}~\cite{Norr21}.
+%
+{\footnotesize
+\begin{verbatim}
+[!KU(k), !KU(m), !KU(ad), !KU(ai)] --[]-> [!KU(aeadEncrypt(k, m, ad, ai))]
+\end{verbatim}}
+%
+Other than modeling authenticated encryption, we use \mTamarin{}'s built-in
+equational theories for signing, Diffie-Hellman, hashing and \mXor{}.
+%
+
+Tamarin has also built-in rules for modeling a Dolev-Yao adversary and the
+evolution of their knowledge as the protocol executes.
+%
+We extend the Dolev-Yao adversary model by adding other rules that increase
+the capabilities of the attacker.
+%
+To denote that the adversary has access to a message $p$ at time $t$, we use 
+$\mK^{t}(p)$.
+%
+As an example, the following implication
+\[
+    \forall t, k, k'\mLogicDot \mK^{t}(\langle k, k'\rangle)\ \rightarrow\
+\mK^{t}(k) \land \mK^{t}(k'),
+\]
+models the fact that if the adversary gets to know the pair of
+keys $\langle k, k' \rangle$ at a time point $t$, then the adversary
+knows each of those keys $k$ and $k'$ at time point $t$ as well.
+%
+For more details about how \mTamarin{} manages adversary knowledge,
+see~\cite{DBLP:conf/cav/MeierSCB13}.
+%
+
+%-------------------------------------------------------------------------- sub
+\subsection{Model and Desired Properties}
+\label{sec:desired-properties}
+In this section, we describe our modeling of \mEdhoc{} and its desired
+security properties.
+%
+
+The party $I$ executing the initiator role considers a run of the protocol
+begun as soon as it sends the first message \mMsgone{} with event type
+\mIStart, and considers the run ended once it has sent the third message
+\mMsgthree{} with event type \mIComplete.
+%
+Similarly, the responder $R$ considers a run started upon receiving 
+\mMsgone{}
+with event type \mRStart, and finished upon receiving \mMsgthree{} with 
+type
+\mRComplete.
+%
+
+In this work we consider the following properties: secrecy of the session key,
+injective agreement and implicit agreement on the session-key material for 
+both
+initiator and responder, and secrecy and integrity of the application data sent
+as part of message 3 ($\mAD_3$).
+%
+Agreement is considered on a set of parameters $S$ which also contains the
+session-key material \mSessKey{}.
+%
+We will first describe in detail all these properties, and then describe the
+contents of the set $S$.
+%
+We formalize these properties as shown in Figure~\ref{fig:props}, which is
+a modified version of a figure in~\cite{Norr21}.
+%
+%Note that the properties in Figure~\ref{fig:props} are unlike the actual \mTamarin{} lemma in one minor way -- in the \mTamarin{} code, we only check that the TEE is accessed after the start of the session, since the long-term key reveal event covers the reveal of all long-term and session keys after the end of the session.
+%\vnote{Check that this is right}
+
+\begin{figure*}[ht]
+\begin{align*}
+    \mPredPcs \triangleq\ & \forall \mTID, I, R, \mSessKey, t_2, t_3\mLogicDot
+    \mK^{t_3}(\mSessKey)\  \land\ 
+    (\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\, 
+\mRComplete^{t_2}(\mTID, 
+I, R, 
+\mSessKey))
+    \rightarrow\\
+    &(\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)
+    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
+    &\ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
+    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
+	&\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land 
+\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, I, R) 
+\land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[1em]
+%    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) 
+%\land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+%
+    \mPredInjI \triangleq\ &
+    \forall \mTID_I, I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, 
+\mSessKey, S)
+    \rightarrow\\
+    &(\exists \mTID_R, t_1\mLogicDot \mRStart^{t_1}(\mTID_R, R, \mSessKey, 
+S) \land t_1 \lessdot t_2
+    \land (\forall \mTID_I', I', R', t_1' \mLogicDot \mIComplete^{t_1'}(\mTID_I', 
+I' , R', \mSessKey, S)
+        \rightarrow t_1' \doteq t_1))\\
+    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
+\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[1em]
+%
+    \mPredInjR \triangleq\ &
+    \forall \mTID_R, I, R, \mSessKey, S, t_2\mLogicDot 
+\mRComplete^{t_2}(\mTID_R, 
+I, R, 
+\mSessKey, S)
+    \rightarrow\\
+    &(\exists \mTID_I, t_1\mLogicDot \mIStart^{t_1}(\mTID_I, I, R, \mSessKey, 
+S) \land t_1 \lessdot t_2
+    \land (\forall \mTID_R', I' R' t_1' \mLogicDot 
+\mRComplete^{t_1'}(\mTID_R', 
+I' , R', \mSessKey, S)
+        \rightarrow t_1' \doteq t_1))\\
+    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, 
+\mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot 
+t_2))\\[1em]
+%
+    \mPredImpI \triangleq\ &
+    \forall \mTID_I, I, R, \mSessKey, S, t_1\mLogicDot 
+\mIComplete^{t_1}(\mTID_I, 
+I, R, \mSessKey, 
+S)
+    \rightarrow\\
+      &(\forall \mTID_R, I', R', S', t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, 
+I', R', \mSessKey, S') \rightarrow
+             (I=I' \land R=R' \land S=S')\\
+      &\ \land (\forall \mTID_I', I', R', S', t_1'\mLogicDot
+        \mIComplete^{t_1'}(\mTID_I', I', R', \mSessKey, S') \rightarrow t_1' \doteq 
+t_1
+        ))\\
+    &\lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
+    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
+    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
+    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
+\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))
+\end{align*}
+%
+\caption{Formalization of security properties and adversary model.}
+\label{fig:props}
+\end{figure*}
+
+%----------------------------------------------------------------------- PFS
+\subsubsection{Secrecy of the Session-Key Material.}
+\label{sec:secrecy}
+We show that the adversary cannot gain access to the session-key material
+\mSessKey{}, even under the weak post-compromise security model which allows 
+the
+adversary limited access to the long-term keys of the involved parties via
+a
+secure interface before the session starts, and unrestricted access to the long-term
+keys of all parties as well as access to all other session keys after the session ends.
+%
+This property is formalized as \mPredPcs{} in Figure~\ref{fig:props}.
+%
+
+In the \mIComplete{} event, \mTID{} represents a ``thread identifier'' which uniquely identifies the
+current session, $I$ represents the identity of the initiator,
+and $R$ represents the identity of the party who the initiator believes is
+playing the responder role, while \mSessKey{} stands for the established
+session-key material.
+%
+\mRComplete{} has analogous parameters; in particular, the responder $R$
+believes $I$ is the party playing the initiator role.
+%
+Intuitively, the PCS property states that an adversary obtains \mSessKey{}
+only if one of the following conditions hold: either a party's long-term key
+is compromised before their run ends, or \mSessKey{} is directly revealed to
+the attacker, or the TEE interface for the responder is used during the current
+session for the initiator, i.e., after the initiator starts and before it
+completes, or, similarly, the TEE interface for the initiator is used during the
+current session for the responder.
+%
+Note that the properties in Figure~\ref{fig:props} are unlike the actual \mTamarin{} lemma in one minor way: \mTamarin's logic does not allow disjunctions to appear on
+the left-hand side of an implication inside a universally-quantified formula.
+%
+Therefore, in the \mTamarin{} code, instead of using the disjunction
+$\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\,
+\mRComplete^{t_2}(\mTID, I, R, \mSessKey)$
+to model the fact that either party may have completed their execution, we use
+a single action parametrized by the terms $\mTID$, $I$, $R$, and \mSessKey.
+
+%-------------------------------------------------------------------- InjAgree
+\subsubsection{Authentication Properties.}
+\label{sec:authenticationDef}
+Following~\cite{Norr21}, we prove two different kinds of authentication
+properties, namely \emph{injective agreement} in the style
+of~\cite{DBLP:conf/csfw/Lowe97a}, and implicit agreement.
+%
+Injective agreement can be guaranteed to either party running the protocol.
+%
+For the initiator $I$, it guarantees to $I$ that whenever $I$ believes that
+they have completed a run with $R$ as responder, then the party $R$ has 
+indeed
+executed the protocol in the role of a responder, and that this run of $I$
+uniquely corresponds to one of $R$ where the set of parameters is $S$ and
+includes, in particular, the session-key material \mSessKey{}.
+%
+It can be defined for $R$ in a similar manner.
+
+%
+We formalize injective agreement for the initiator role as \mPredInjI{} and
+for the responder role as \mPredInjR{} in Figure~\ref{fig:props}.
+%
+For the initiator $I$, this property represents the fact that either
+injective agreement (as described above) holds, or the long-term key of
+the party $R$ assumed to be playing the responder role has been
+compromised before $I$'s role finished.
+%
+The property should hold even if the adversary has access to the TEE of party 
+$R$
+before and after the protocol execution.
+%%
+An analogous definition holds for the responder $R$.
+%
+%As part of the formalization, note that we parametrize the \mIStart{} and \mRStart{} events with some subset of \mTID, $I$, $R$, and $Z$, as well as the set $S$ of parameters upon which we check for agreement.
+%%
+%The set $S$ will often include the other party's identity, so, for example, \mRStart{} will not mention the term $I$ as a parameter outside of $S$, and instead just be parametrized by \mTID, $R$, $Z$, and $S$.
+
+%------------------------------------------------------------- Implicit auth
+As part of our analysis, we found that all the \mEdhoc{} methods satisfy weak
+post-compromise security.
+%
+However, this is not the case for the injective agreement property as stated 
+above.
+%
+Thus, we show a different property, a form of implicit agreement on the same
+set of parameters, which is guaranteed for all methods.
+%
+This modification is inspired by the definitions of implicit authentication in
+the computational model~\cite{DBLP:conf/csfw/GuilhemFW20}.
+%
+While that paper focuses on authenticating just a key and related identities,
+our definition encompasses a general set of parameters, as in the notion of
+injective agreement proposed by Lowe~\cite{DBLP:conf/csfw/Lowe97a}.
+
+The ``implicit'' in the name of the property stands for the fact that a party
+$A$ assumes that any party $B$ who has access to the session-key material
+\mSessKey{} must, in fact, be the intended party, and that if $B$ is honest,
+$B$ will agree on a set $S$ of parameters which includes \mSessKey.
+%
+Implicit agreement for both roles guarantees to $A$ that $A$ is or has been
+involved in exactly one protocol execution with $B$, and that $B$ agrees or
+will agree with $A$ on $S$.
+%
+This property diverges from injective agreement in that upon sending
+the last message, $A$ concludes that if this message reaches $B$, then $A$
+and $B$ agree on each other's identities and roles, as well as the set $S$.
+%
+
+Note that for implicit agreement to hold for $I$, the ephemeral keys must not
+be revealed since the property relies on the fact that the intended responder
+is the only one who knows the session-key material.
+%
+If the adversary has access to the ephemeral keys, they can use them along 
+with
+the public keys of $I$ and $R$ to compute the session-key material.
+%
+However, either party's long-term key can be revealed after the other
+party has finished their execution, since this still leaves the adversary
+unable to compute \mGxy{}.
+%
+
+Since \mTamarin{} runs out of memory to verify this property as is,
+we split it into two lemmas -- \mPredImpI{} for $I$ for one \mPredImpR{} for 
+$R$.
+%
+Figure~\ref{fig:props} contains the definition for \mPredImpI{}.
+%
+\mPredImpR{} is formalized similarly, so we omit it.
+%
+
+%------------------------------------------------------- Agreed parameters
+\subsubsection{Set $S$ of Agreed Parameters.}
+\label{sec:agreedParams}
+We now describe the set $S$ of parameters upon which the two parties obtain
+guarantees via the above properties.
+%
+The initiator $I$ gets injective and implicit agreement guarantees on the
+following partial set $S_P$ of parameters~\cite{Norr21}:
+\begin{itemize}
+    \item the roles played by itself and its peer,
+    \item responder identity,
+    \item session-key material (which varies depending on the \mEdhoc{} 
+method),
+    \item context identifiers \mCi{} and \mCr{}, and
+    \item cipher suites \mSuites{}.
+\end{itemize}
+%
+
+Due to the initiator being guaranteed identity protection under \mEdhoc{}, $I$
+cannot get explicit agreement with $R$ on the initiator's identity.
+%
+Similarly, when using the \mStat{} authentication method, $I$ does not get 
+any
+such guarantees about $P_{I}$.
+%
+However, $I$ does get implicit agreement with $R$ about $I$'s identity and the
+full set $S_{F}$ of agreed parameters.
+%
+In contrast, since $R$'s run finishes after that of $I$, $R$ can get explicit
+injective agreement assurances on the full set $S_{F}$ of agreed parameters.
+%
+The full set of agreed parameters $S_F$ is $S_P \cup \{I, P_I\}$ when $P_I$
+is part of the session-key material, and $S_P \cup \{I\}$ otherwise.
+%
+
+In addition to these properties, a couple of properties can be inferred
+without being explicitly modelled and verified.
+%
+One such property is Key-Compromise Impersonation
+(KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
+%
+A KCI attack occurs when an adversary who has access to $A$'s long-term 
+private
+key to make $A$ believe that they completed an execution with a peer $B$,
+while $B$ did not participate in said execution at all.
+%
+This is in particular relevant when \mStat{} authentication methods are used.
+%
+Our above notions of agreement ensure that both parties agree on each
+other's identity, role, and session-key material.
+%
+Therefore, all \mEdhoc{} methods that satisfy these agreement properties also
+avoid KCI attacks.
+%
+
+Another kind of attack is Unknown Key Share attacks
+(UKS)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
+%
+As part of a UKS attack, a party $A$ can be made to believe that it finished
+an execution with a party $B$, but where the session-key material is actually
+shared between $A$ and $C$ instead.
+%
+Again, due to the agreement on identities and session-key material, any 
+method
+that satisfies the above agreement properties also resists UKS attacks.
+%
+Overall, the agreement properties capture entity authentication,
+and satisfy any properties based on that notion.
+%
+However, see Section~\ref{sec:unintendedPeerAuth} for a discussion on the
+interaction between \mEdhoc{} and the application leading to similar issues.
+%
+
+%---------------------------------------------------------------------------sub
+\subsection{Encoding \mEdhoc{} in \mTamarin{}}
+\label{sec:modeling}
+%
+In this section, we describe how we model the \mSigSig, \mSigStat, \mStatSig,
+and \mStatStat{} methods of \mEdhoc{} in \mTamarin.
+%
+As in~\cite{Norr21}, we construct the \mTamarin{} model by utilizing the fact
+that all methods of \mEdhoc{} share a common underlying structure
+(as shown in Figure~\ref{fig:edhocFramework}).
+%
+We do so by using a single specification file in the M4 macro language,
+which generates all the methods.
+%
+As in~\cite{Norr21}, we only present the \mStatSig{} method which illustrates
+the use of two different authentication methods.
+%
+The full \mTamarin{} code for all models can be found 
+at~\cite{edhocTamarinRepo}.
+%
+
+As mentioned in Section~\ref{sec:tamarin}, we extend the default
+equational theory of \mTamarin{} to handle various operations used in 
+\mEdhoc.
+%
+Of the built-in theories, we use the ones for exclusive-or (\mXor),
+Diffie-Hellman operations, signatures (\mT{sign} and \mT{verify} operations),
+and 
+hashing~\cite{DBLP:conf/csfw/DreierHRS18,DBLP:conf/csfw/SchmidtMCB12}.
+%
+We 
+%augment the hashing function symbol to add an extra input, which we use 
+to
+model different hash functions.
+%
+
+In addition to these default operations, \mEdhoc{} is built over the following functions: \mHkdfExpand, \mHkdfExtract, and \mAead{}.
+%
+We represent \mHkdfExpand{} by \mT{expa} and \mHkdfExtract{} by 
+\mT{extr}.
+%
+\mAead{} operations need us to add extra equations to the underlying theory.
+%
+A term encrypted using the \mAead{} algorithm is represented by \mT{aeadEncrypt(m, k, ad, 
+ai)},
+where \mT{m} is the underlying message, \mT{k} is the encrypting key,
+\mT{ad} is the additional data, and \mT{ai} is the identifier for the
+encryption algorithm.
+%
+Decryption of such a term is defined via two equations that we add to
+\mTamarin's theory.
+%
+The following equation requires the decrypting party to know the additional
+data \mT{ad} to decrypt this encrypted term with verification of its integrity.
+\begin{small}
+\begin{verbatim}
+aeadDecrypt(aeadEncrypt(m, k, ad, ai), k, ad, ai) = m.
+\end{verbatim}
+\end{small}
+%
+Only the above equation is used by honest parties, but the adversary should
+also be able to decrypt without having to go through the additional step of
+identity verification.
+%
+To this end, we also add the following equation, where the adversary does not
+need access to the additional data \mT{ad} in order to decrypt.
+%
+\begin{small}\begin{verbatim}
+decrypt(aeadEncrypt(m, k, ad, ai), k, ai) = m.
+\end{verbatim}\end{small}
+%
+
+Having described how we adapt the equational theory to model \mEdhoc,
+we now move on to the modeling of the adversary model and the 
+environment in
+which the protocol is executed.
+%
+We extend the built-in Dolev-Yao adversary rules which are part of 
+\mTamarin.
+%
+We use the following rules to capture the link between a party's identity and
+their long-term key pairs, in the \mSig{}- and in the \mStat{}-based methods
+respectively.
+\begin{center}
+\begin{minipage}{0.48\textwidth}
+\begin{scriptsize}
+\begin{verbatim}
+1 rule registerLTK_SIG:
+2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
+3        [!LTK_SIG($A, ~ltk),
+4         !PK_SIG($A, pk(~ltk)),
+5         Out(<$A, pk(~ltk)>)]
+\end{verbatim}
+\end{scriptsize}
+\end{minipage}
+\hfill\vline\hfill
+\begin{minipage}{0.48\textwidth}
+\begin{scriptsize}
+\begin{verbatim}
+1 rule registerLTK_STAT:
+2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
+3        [!LTK_STAT($A, ~ltk),
+4         !PK_STAT($A, 'g'^~ltk),
+5         Out(<$A, 'g'^~ltk>)]
+\end{verbatim}
+\end{scriptsize}
+\end{minipage}
+\end{center}
+
+Using the fact \verb|Fr(~ltk)|, \mTamarin{} creates a new term \mT{ltk} and 
+uses
+it to represent a secret long-term key.
+%
+Via the \verb|Out(<$A, pk(~ltk)>)| fact, \mTamarin{} puts out onto the
+communication channel the identity of the party to whom this long-term
+key belongs, along with their public key.
+%
+Since the adversary has access to the communication channel, they can pick 
+up
+all of this information.
+%
+The event \mT{UniqLTK} parametrized by a party's identity and their 
+long-term
+key 
+models the unique correspondence between those two values.
+%
+As a result, this rules out a party owning multiple long-term keys
+-- in particular, it keeps the adversary from registering long-term keys in
+some honest party's name.
+%
+This aligns well with an external mechanism such as a certificate authority
+ensuring that long-term keys are uniquely assigned to the corresponding
+identities, which is ensured by the \mSpec.
+
+To model the reveal of long-term keys and ephemeral keys to an adversary,
+we use standard reveal rules and events of type \mRevLTK{} and \mRevEph,
+respectively.
+%
+It is also important to keep track of the time points at which these events 
+occur.
+%
+Long-term keys can be revealed on registration, even before the protocol 
+begins.
+%
+Ephemeral keys in our model can only be revealed when a party
+completes their role, i.e., simultaneously with events of type \mIComplete{}
+and \mRComplete.
+%
+Having set out the capabilities of the adversary, we now model the execution
+of the honest agents' roles.
+%
+
+For each protocol method, we use two rules apiece for the initiator and
+responder -- \mT{I1}, \mT{R2}, \mT{I3}, \mT{R4}.
+%
+Each of these stand for one step of the protocol as executed by either party.
+%
+To disambiguate, we will attach the method to the rule name.
+%
+These four rules directly map to the event types
+\mIStart, \mRStart, \mIComplete, and \mRComplete, respectively.
+%
+We show the \mT{R2\_STAT\_SIG} rule below to illustrate the various aspects
+of the \mTamarin{} modeling we are describing here.
+%
+
+In order to keep track of the initiator's state, we use facts prefixed with
+\mT{StI}, which carry information between the \mT{I1} and \mT{I3} rules.
+%
+Similarly, for the responder's state, we have \mT{StR} to carry state data
+between \mT{R2} and \mT{R4}.
+%
+In order to link two rules to a state fact, we use \mT{tid}, which
+is unique to the current session.
+%
+The use of these state facts can be seen in line 28 in the \mT{R2\_STAT\_SIG} 
+rule.
+%
+
+Note that we do not model any error message that $R$ might send in response
+to message \mMsgone{} rejecting $I$'s choice of ciphersuite and/or method.
+%
+
+As in~\cite{Norr21}, we model the \mXor{} encryption of 
+\mT{CIPHERTEXT\_2} with
+the key \mT{K\_2e} by allowing each part of the encrypted term to be
+separately attacked.
+%
+This means that we first expand \mT{K\_2e} to the same number of key terms 
+as
+subterms in the plaintext tuple.
+%
+This is done by applying \mHkdfExpand{} to unique inputs per subterm.
+%
+After this, we \mXor{} each subterm with its own key term.
+%
+This is more faithful to the \mSpec{} than \mXor-ing \mT{K\_2e} on its own
+with the plaintext tuple.
+%
+This can be seen in lines 18--21 in the code for \mT{R2\_STAT\_SIG}.
+%
+
+As we extended the model with TEEs and augmented the adversary's 
+capabilities
+with access to them, \mTamarin{} failed to complete in a reasonable time for
+some combination of authentication methods and security properties (see 
+Section~\ref{sec:conclusion} for a detailed discussion).
+%
+To circumvent the problem, we simplified the \mXor{} encryption to 
+\mXor-ing 
+a
+single term on the entire tuple for these cases. \\
+%
+
+{\parindent 0pt
+\begin{minipage}{\textwidth}
+\begin{scriptsize}
+\begin{verbatim}
+1 rule R2_STAT_SIG:
+2 let
+3    agreed = <CS0, CI, ~CR>
+4    gx = 'g'^xx
+5    data_2 = <'g'^~yy, CI, ~CR>
+6    m1 = <'STAT', 'SIG', CS0, CI, gx>
+7    TH_2 = h(<$H0, m1, data_2>)
+8    prk_2e = extr('e', gx^~yy)
+9    prk_3e2m = prk_2e
+10   K_2m = expa(<$cAEAD0, TH_2, 'K_2m'>, prk_3e2m)
+12   protected2 = $V // ID_CRED_V
+13   CRED_V = pkV
+14   extAad2 = <TH_2, CRED_V>
+15   assocData2 = <protected2, extAad2>
+16   MAC_2 = aead('e', K_2m, assocData2, $cAEAD0)
+17   authV = sign(<assocData2, MAC_2>, ~ltk)
+18   plainText2 = <$V, authV>
+19   K_2e = expa(<$cAEAD0, TH_2, 'K_2e'>, prk_2e)
+20   K_2e_1 = expa(<$cAEAD0, TH_2, 'K_2e', '1'>, prk_2e)
+21   K_2e_2 = expa(<$cAEAD0, TH_2, 'K_2e', '2'>, prk_2e)
+22   CIPHERTEXT_2 = <$V XOR K_2e_1, authV XOR K_2e_2>
+23   m2 = <data_2, CIPHERTEXT_2>
+24   exp_sk = <gx^~yy>
+25   in
+26   [!LTK_SIG($V, ~ltk), !PK_SIG($V, pkV), In(m1), Fr(~CR), Fr(~yy), Fr(~tid)]
+27   --[ExpRunningR(~tid, $V, exp_sk, agreed), R2(~tid, $V, m1, m2)]->
+28   [StR2_STAT_SIG($V, ~ltk, ~yy, prk_3e2m, TH_2, CIPHERTEXT_2, gx^~yy, ~tid, m1, m2, agreed),
+29    Out(m2)]
+\end{verbatim}
+\end{scriptsize}
+\end{minipage}}
+\vspace{5mm}
+
+As mentioned earlier, we use actions to represent parametrized events.
+%
+For example, in line 27 above, the action
+\verb|ExpRunningR(~tid, $V, exp_sk, agreed)| represents an event of type
+\mRStart{} parametrized by the session id, the responder's identity, and the
+session-key material \mT{exp\_sk}.
+%
+The \mT{exp} in the name of the variable for session-key material represents
+the fact that the agreement property satisfied by this key is explicit, i.e.,
+it includes $P_{I}$, as in Section~\ref{sec:agreedParams}.
+%
+We use \mT{imp\_sk} for the corresponding session-key material which does 
+not
+include $P_{I}$.
+%
+For the \mSigSig{} and \mSigStat{} methods, therefore, the two values are the
+same.
+%
+
+The properties we listed in Section~\ref{sec:desired-properties}
+translate directly into \mTamarin's logic.
+%
+We show the \mTamarin{} lemma which encodes the \mPredPcs{} property.
+%
+Other properties are formalized similarly. 
+%
+
+\begin{scriptsize}
+\begin{verbatim}
+1 lemma secrecyPFS:
+2    all-traces
+3    "All u v sk tid #t3 #t2.
+4       (K(sk)@t3 & CompletedRun(u, v, sk, tid)@t2) ==>
+5       ( (Ex #t1. LTKRev(u)@t1 & #t1 < #t2)
+6       | (Ex #t1. LTKRev(v)@t1 & #t1 < #t2)
+7       | (Ex #t1. EphKeyRev(sk)@t1)
+8       | (Ex m1 #t0 #k. I1(tid, u, v, m1)@t0 & TEE(v)@k & (t0 < k | k < t2))
+9       | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & (t0 < k | k < t2))
+10    )"
+\end{verbatim}
+\end{scriptsize}
+%
+
+In this formalization, we use the action \mT{CompletedRun(u, v, sk)}
+(in line 4) to represent the disjunction of the events $\mIComplete^{t_{2}}$
+and $\mRComplete^{t_{2}}$.
+%
+As expected, this action is emitted by both \mT{I3} and \mT{R4}.
+%
+Similarly, the action \mT{EphKeyRev(sk)} in line 7 stands for the reveal of
+the ephemeral key for $I$ or $R$ or both.
+%
+Lines 8 and 9 captures that the parties TEEs must be inaccessible to the
+adversary between the start and end of the execution as seen by each party
+respectively.
+%
 
@@
 %
-Upon discussion with the authors, they included recommendations on how to store
+Upon discussion with the authors, they included recommendations on storing
 long-term keys, and how to perform operations on these keys inside a secure
@@
 %
-    In those protocols, $P_{I}$'s presence in the computation of session key
+    In those protocols, including $P_{I}$ while computating session-key
     material is crucial to obtaining resistance against the compromise of the

### 2022-07-23 protocol.tex
Fixed typos and references and merged everything into paper.tex

@@
 %
-When clear from context, we will often drop part (or all) of the subscripts.
+When clear from context, we will often drop some (or all) of the subscripts.
 %
@@
 are long-term authentication credentials, whereas ephemeral key pairs are
-those
-which are generated afresh for each execution of the protocol.
+those generated afresh for each execution of the protocol.
 %
@@
 In the former case, we say that they use the
-\emph{signature based authentication method} (\mSig{}).
+\emph{signature-based authentication method} (\mSig{}).
 %
-In the latter, we say, following the terminology in the \mSpec{},
+In the latter case, we say, following the terminology in the \mSpec{},
 that they use the \emph{static key authentication method} (\mStat{}).
@@
 
-\mEdhoc{} fundamentally use elliptic curves and associated Diffie-Hellman
+\mEdhoc{} fundamentally uses elliptic curves and associated Diffie-Hellman
 operations.
@@
 elliptic curve as well as the corresponding bitstring encoding.
-%
-%\mEdhoc{} relies on \mCose{}~\cite{rfc8152} for elliptic curve operations and
-%transforming points into bitstrings, and we therefore abstract those as
-%follows.
-%
-%Signatures and verification thereof using party $A$'s key pair are
-%denoted \mSign{A}$(\cdot)$ and \mVf{A}$(\cdot)$ respectively.
-%%
-%The DH-primitive combining a private key $d$ and a point $P$ is denoted
-%$\mDH(d,P)$.
-%%
-%We abuse notation and let these function symbols denote operations on both
-%points and the corresponding bitstrings.
-%
 
@@
 \label{sec:description}
-\mEdhoc{} is designed to establish a security context for \mOscore{}.
+\mEdhoc{} is designed to establish a security context for the \mOscore{} protocol.
 %
-This context includes the session-key material (we denote this by
+This context, in particular, includes the session-key material (we denote this by
 \mSessKey{}).
 %
-The abstract protocol as given in the \mSpec{}~\cite{} consists of three
-messages, and is shown in Figure~\ref{fig:edhocFramework}~\cite{Norr21}.
+The generalized abstract protocol as given in the \mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} consists of three
+messages, and is shown in Figure~\ref{fig:edhocFramework}.
 %
-The generalized abstract protocol is the same across methods,
-and may transfer application data \mADone{}, \mADtwo{}, and \mADthree{} in
-addition to establishing the context.
+The abstract structure is the same across methods.
 %
-The authentication mechanisms and key derivation procedures differ between
+However, the authentication mechanisms and key derivation procedures differ between
 methods.
 %
-We will describe the various aspects of the protocol now.
+\mEdhoc{} may also transfer application data \mADone{}, \mADtwo{}, and \mADthree{} in
+addition to establishing the \mOscore{} security context.
+%
+
 %
@@
 %
-Of the three messages, the first two, among other things, serve to establish a
-common authentication method \mMethod{}, and a ciphersuite \mSuites{}.
+Of the three messages \mMsgone, \mMsgtwo, and \mMsgthree, the first two, among other things, establish a
+common authentication method \mMethod{} and ciphersuite \mSuites{}.
 %
-In \mMethod{}, the party playing the initiator role proposes which
-authentication methods the two parties shall use, and an ordered list
+The party playing the initiator role uses \mMethod{} to propose which
+authentication methods the two parties shall use, and in \mSuites{}, proposes an ordered list
 of choices for the ciphersuite.
 %
-As mentioned earlier, the authentication methods may differ for the two roles,
+The chosen authentication methods may differ for the two roles,
 yielding four possible
@@
 where the first authentication method in each combination is used by the
-initiator role, and the second by the responder role.
+initiator, and the second by the responder.
 \footnote{As in the \mSpec{}, we will from now on overload
@@
 %
-The party executing the responder role may choose to reject the method or
+The party executing the responder role may choose to reject the method and/or
 ciphersuite chosen by the initiator by sending an error message.
@@
 
-In addition to negotiating the ciphersuite, the first two messages are
+In addition to negotiating the method and ciphersuite, the first two messages are
 also instrumental for the exchange of public ephemeral keys \mGx{} and
@@
 to the correct party executing \mEdhoc{}, but also claims that they may be
-used in turn by protocols (like \mOscore{}) using the security context
+used in turn by protocols (like \mOscore{}) which use the security context
 established by \mEdhoc{}.
@@
 %
-The authentication information is different for each authentication method.
+The authentication information is structured differently for each authentication method.
 %
-Consider the following example scenario.
+
+Consider the following scenario.
 %
-The initiator $I$ chooses the method $\mSigStat{}$ and sends this via
+The initiator $I$ chooses the method $\mSigStat{}$, and sends this via
 $\mMethod{}$ to the responder $R$.
 %
-$R$ now has an option to either reject this choice of method, or to provide an
+$R$ now has the option to reject this choice of method.
+%
+However, if $R$ chooses to accept this method, they need to provide an
 identifier \mIdcredr{} for a key pair which can be used with
@@
 \label{sec:keysched}
-As mentioned above, the second and third messages of \mEdhoc{} contain
+The second and third messages of \mEdhoc{} contain
 authenticating information.
@@
 %
-The keys for these are generated by using a key schedule which takes at least
-one DH key \mGxy{} as basic input, as shown in Figure~\ref{fig:kdfdiagram},
-and is intrinsic to the functioning of \mEdhoc{}.
+The keys for these are generated using a key schedule which is intrinsic to the functioning of \mEdhoc{}.
+%
+The key schedule takes a DH key \mGxy{} as basic input and builds upon it, as shown in Figure~\ref{fig:kdfdiagram}.
 %
@@
 \scalebox{.785}{
-\input{kdfdiagram.tex}
+\begin{tikzpicture}[%
+    >=latex,              % Nice arrows; your taste may be different
+    start chain=going below,    % General flow is top-to-bottom
+    node distance=4mm and 60mm, % Global setup of box spacing
+    every join/.style={norm},   % Default linetype for connecting boxes
+    ]
+\tikzset{
+terminput/.style={rounded corners},
+term/.style={rounded corners},
+  base/.style={draw, thick, on chain, on grid, align=center, minimum height=4ex},
+  dhkbox/.style={draw=cbgreen, fill=cbgreen!25, rectangle},
+  dhk/.style={base, dhkbox},
+  prkbox/.style={draw=cborange, fill=cborange!25, rectangle},
+  prk/.style={base, prkbox},
+  hkdfext/.style={base, draw=black, fill=none, rectangle},
+  hkdfexp/.style={base, draw=black, fill=none, rectangle},
+  keybbox/.style={draw=cbnavy, fill=cbnavy!25, rectangle},
+  keyb/.style={base, keybbox, text width=4em},
+  norm/.style={->, draw, black},
+  cond/.style={base, draw=black, dashed, fill=none, rectangle},
+  txt/.style={base, draw=none, fill=none}
   }
-\caption{Key schedule: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo,
+\node [prk, join] (p2) {\mPRKtwo};
+\node [cond, join] (c1) {R uses \mStat?};
+
+\node [prk, below=6mm of c1.south] (p3) {\mPRKthree};
+\draw [->, norm] (c1.south) -- (p3.north) node[midway, right] {N};
+
+\node [cond, join, below=10mm of p3.south] (c2) {I uses \mStat?};
+\node [prk, below=8mm of c2.south] (p4) {\mPRKfour};
+\draw [->, norm] (c2.south) -- (p4.north) node[midway, right] {N};
+
+\node [hkdfext, right=3cm of p3] (h3) {\mHkdfExtract};
+\node [hkdfext, right=3cm of p4] (h5) {\mHkdfExtract};
+
+\node [hkdfexp, shape border rotate=180, left= 2.5cm of p4] (h6) {\mHkdfExpand};
+\node [keyb, join, left=3cm of h6] (k3) {\mKthreem};
+\node [hkdfexp, shape border rotate=180, below= 0.8cm of h6] (h9) {\mHkdfExpand};
+\node [txt, join, left=1cm of h9.west] (t4) {EDHOC-Exporter};
+
+\node [hkdfexp, shape border rotate=180, left= 2.5cm of p3] (h4) {\mHkdfExpand};
+\node [keyb, join, left=3cm of h4] (k2) {\mKtwom};
+
+\node [hkdfexp, shape border rotate=180, left= 2.5cm of p2] (h2) {\mHkdfExpand};
+\node [keyb, join, left=3cm of h2] (k1) {\mKtwoe};
+
+\node [hkdfexp, shape border rotate=180, below= 0.8cm of h4] (h8) {\mHkdfExpand};
+\node [keyb, below=0.8cm of k2] (k2b) {\mKthreeae};
+
+\node [txt, left=1cm of k1.west] (t1) {Enc (XOR) \\ in m2};
+\node [txt, left=1cm of k2.west] (t2) {\mMactwo~(signed if \\ R uses \mSig)};
+\node [txt, left=1cm of k2b.west] (t2b) {\mAead\ in m3};
+\node [txt, left=1cm of k3.west] (t3) {\mMacthree~(signed if \\ I uses \mSig)};
+
+\draw [->, norm] (k1.west) -- (t1.east);
+\draw [->, norm] (k2.west) -- (t2.east);
+\draw [->, norm] (k2b.west) -- (t2b.east);
+\draw [->, norm] (k3.west) -- (t3.east);
+
+\draw [->, norm] (p3.south) ++(0,-0.5) -- (h8);
+\draw [->, norm] (h8) -- (k2b);
+\draw [->, norm] (p2) -- (h2); 
+\draw [->, norm] (c1.east) -- ++(2.025,0) -- (h3.north) node[midway,above left] {Y};
+\draw [->, norm] (h3.west) -- (p3.east);
+\draw [->, norm] (p3) -- (h4); 
+\draw [->, norm] (c2.east) -- ++(2.07,0) -- (h5.north) node[midway,above left] {Y};
+\draw [->, norm] (h5.west) -- (p4.east);
+\draw [->, norm] (p4) -- (h6);
+\draw [->, norm] (p4.west) ++(-0.25,-0) -- ++(0,-0.8) -- (h9.east);
+
+\node [hkdfext, right=3cm of p2] (h1) {\mHkdfExtract};
+\node [dhk, right=2.7cm of h1] (p0) {$\mGxy$};
+\node [terminput, text width=2em, below = 0.2cm of p0] (u1) {Salt};
+\draw [->] (h1.west) -- (p2.east);
+\draw [->] (u1.west) -- ++(-2.27,0) -- (h1.south);
+\draw [->] (p0.west) -- (h1.east);
+
+\node [dhk, right = 2.7cm of h3] (u2) {$P_R$};
+\draw [->, norm] (u2.west) -- (h3.east);
+
+\node [dhk, right = 2.7cm of h5] (u3) {$P_I$};
+\draw [->, norm] (u3.west) -- (h5.east);
+
+\node [term, above = 0.55cm of h4] (u5) {\mTHtwo};
+\draw [->, dotted, shorten >=1mm] (u5) -- (h4);
+\draw [->, dotted, shorten >=1mm] (u5) -- (h2);
+
+\node [term, above = 0.5cm of h6] (u6) {\mTHthree};
+\draw [->, dotted, shorten >=1mm] (u6) -- (h6);
+\draw [->, dotted, shorten >=1mm] (u6) -- (h8);
+
+\node [term, below=0.4cm of p4] (u7) {\mTHfour};
+\draw [-> , dotted ] (u7.west) -- ([yshift=-0.4em] h9.east);
+\end{tikzpicture}
+}
+\caption{Key schedule for~\cite{our-analysis-selander-lake-edhoc-00}: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo,
 \mPRKthree, \mPRKfour{} are the intermediate key material, and
@@
 %
-Both functions take as argument two values -- a salt, and an input.
+Both functions take as argument two values -- an input and a salt.
 %
@@
 (\mKtwoe, \mKtwom{}, \mKthreeae, and \mKthreem)
-for a particular message in the protocol.
+corresponding to each message in the protocol.
 %
@@
 %
-The key $P_{R}$ is used if the responder uses the \mStat{} authentication
+The key $P_{R}$, which is computed as $\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\
+\mGx)$, is used if the responder uses the \mStat{} authentication
 method.
-%
-$P_{R}$ is the key computed as $\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\
-\mGx)$.
 %
@@
 
-These intermediate keys are fed into \mHkdfExpand{}, as shown in
-Figure~\ref{fig:kdfdiagram}.
-%
-\mHkdfExpand{} here uses as salt a value $\textit{th}$, which is a running
+These intermediate keys are fed into \mHkdfExpand{}, which uses as salt a value $\textit{th}$, which is a running
 hash of the information transmitted thus far as part of the protocol.
 %
-$\textit{th}_{i}$ denotes the hash corresponding to the $i^{\rm{th}}$ message.
+By $\textit{th}_{i}$, we denote the hash corresponding to the $i^{\rm{th}}$ message.
 %
@@
 We
-%discuss which material should be included in \mSessKey{} in more detail
-in
-the formal model in Section~\ref{sec:formalization} and the consequences of
+discuss which material should be included in \mSessKey{} in more detail
+in Section~\ref{sec:formalization}, and the consequences of
 various choices in Section~\ref{sec:sessionKeyMaterial}.
@@
 We now describe how the authentication information is constructed,
-depending on which method is used.
+depending on whether the \mSig{} or \mStat{} method is used.
 %
-For both \mStat{} and \mSig{}, the following information is used to compute
+For both methods, the following information is used to compute
 \mAuthr{}: \mIdcredr{}, \mCredr{}, a transcript hash of all the communicated
@@
 A MAC is obtained by feeding this material as additional data and the empty
-string as input to the Authenticated Encryption with Additional Data (AEAD)
-encryption algorithm as indicated in the established ciphersuite \mSuites{}.
+string as input to the Authenticated Encryption with Additional Data (AEAD) algorithm as indicated in the established ciphersuite \mSuites{}.
 %
@@
 When $I$ uses the \mSig{} authentication method, \mAuthi{} is $I$'s signature
-over the MAC along with the data covered by the MAC.
+over the MAC computed as above, along with the data covered by the MAC.
 %
@@
 %
-We use $\mathit{MAC}_{I}$ (resp. $\mathit{MAC}_{R}$) to denote a MAC
+We use $\mathit{MAC}_{R}$ (resp. $\mathit{MAC}_{I}$) to denote a MAC
 which uses
@@
 For the third message, encryption and integrity protection is assured by the
-\mAead{} algorithm which is part of the established ciphersuite \mSuites{}.
+\mAead{} algorithm.
 %
@@
 \label{sec:TEE}
-Authentication of a specific IoT device assumes that this device is the only
-entity with access to the long-term key associated with the corresponding
+Authentication of a specific IoT device assumes that the device is the only
+entity with access to the long-term key associated with its
 identity.
@@
 Since IoT devices may be accessible to adversaries, e.g., an insider cloning a
-key card, the long-term keys must be appropriately protected.
+key card, long-term keys must be appropriately protected.
 %
-A state-of-the-art approach is to use a TEE,
+A state-of-the-art approach is to use a trusted execution environment (TEE),
 which holds the key and provides an API for operations using the key.
@@
 This is the approach taken by the
-TrustZone based \mMuEdhoc{}~\cite{DBLP:conf/codaspy/Hristozov0XFLS21}
-implementation for example.
+TrustZone-based \mMuEdhoc{}~\cite{DBLP:conf/codaspy/Hristozov0XFLS21}
+implementation, for example.
 %
@@
 
-TEEs come in various forms of differing complexity.
+TEEs can be of differing complexity.
 %
-Some, like ARM TrustZone and Intel SGX are general purpose execution
-environments, which are flexibly programmable.
+Some, like ARM TrustZone and Intel SGX, are general-purpose execution
+environments, which can be programmed in many different ways.
 %
-Other, like the Universal Subscriber Identity Modules (USIM) used for
-authentication to 3GPP mobile networks, have application specific interfaces
+Others, like the Universal Subscriber Identity Modules (USIM) used for
+authentication to 3GPP mobile networks, have application-specific interfaces
 for
-authentication and key agreement protocols etc.
-%
-A key aspect is how much of the application is placed in the TEE and how
-much is
-outside.
+authentication, key agreement protocols etc.
 %
 
-For larger devices that include general purpose processors with TrustZone or
-SGX, entire \mEdhoc{} and \mOscore{} may reside inside the TEE.
+A fundamental aspect of TEEs is how much of the application is placed in the TEE.
 %
-For constrained IoT devices in the lower end of the scale, a TEE may have to be
+For larger devices that include general-purpose processors with TrustZone or
+SGX, entire \mEdhoc{} and \mOscore{} implementations may reside inside the TEE.
+%
+For constrained IoT devices on the lower end of the scale, a TEE might have to be
 implemented using a special-purpose integrated circuit.
@@
 In the latter case, it may be beneficial to follow a minimalistic approach, and
-contain the long-term key and only
-the operations that need access to it in the TEE for cost reasons.
+only store in the TEE the long-term key and operations that need access to it, to reduce cost.
 %
 
-It may at first appear more secure to implement as much as possible inside the
-TEE, but there is a security trade-off: the more code inside the TEE,
-the higher the risk of implementation errors in the security-critical code.
+In general, it might appear more secure to implement as much as possible inside the TEE, but there is a security trade-off.
 %
-Because the security critical code runs in the area where the long-term keys
+Because security-critical code runs in the same area as where long-term keys
 reside, an implementation error here risks leaking information of the key to
-the
-adversary.
+the adversary.
 %
-From this perspective, it may be beneficial to follow the minimalistic approach
-even when having access to Trust Zoneor SGX.
+From this perspective, it might be beneficial to follow the minimalistic approach
+even when having access to TrustZone or SGX.
 %
 
-A slightly more secure division of functionality is to also keep the
-session key inside
-the TEE and extend the interface to accept messages and return the
-(en/de)crypted counterpart.
+A slightly more secure division of functionality is to keep both the long-term key and the session key inside the TEE and extend the interface to accept messages and return the (en/de)crypted counterpart, i.e.
+the interface exposes \mAead{} functions.
 %
-That is, the interface exposes \mAead{} functions in the interface.
-%

### 2022-07-23 formalization.tex

@@
     \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
-	&\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land 
-\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, I, R) 
-\land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
-%    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) 
-%\land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land \mTEE^{t_1}(R) \land t_0 \lessdot t_1)\\
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land t_0 \lessdot t_1)\\[2em]
 %
@@
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
-\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land t_0 \lessdot t_1)\\[2em]
 %
@@
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, 
-\mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot 
-t_2))\\[2em]
+    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land t_0 \lessdot t_1)\\[2em]
 %
@@
     \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
-    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
-\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))
+    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land t_0 \lessdot t_1)
 \end{align*}
@@
 
-In the \mIComplete{} event, \mTID{} represents a ``thread identifier'' which uniquely identifies the
-current session, $I$ represents the identity of the initiator,
+In the \mIComplete{} event, \mTID{} represents a ``thread identifier'' for the
+current session and is used to correlate multiple events for the same run of the 
+protocol for each party, $I$ represents the identity of the initiator,
 and $R$ represents the identity of the party who the initiator believes is
@@
 the attacker, or the TEE interface for the responder is used during the current
-session for the initiator, i.e., after the initiator starts and before it
-completes, or, similarly, the TEE interface for the initiator is used during the
+session for the initiator, i.e., after the initiator starts, or, similarly, the 
+TEE interface for the initiator is used during the
 current session for the responder.
 %
-Note that the properties in Figure~\ref{fig:props} are unlike the actual \mTamarin{} lemma in one minor way: \mTamarin's logic does not allow disjunctions to appear on
+We use the thread identifier \mTID for each party to relate the starting event
+to the corresponding commit event.
+%
+The property in Figure~\ref{fig:props} is unlike the actual \mTamarin{} lemma
+in one minor way: \mTamarin's logic does not allow disjunctions to appear on
 the left-hand side of an implication inside a universally-quantified formula.

### 2022-07-23 paper.tex
Final version

@@
 
-\input{macros}
+%\input{macros}
+\usepackage{ifthen}
+\newcommand{\mPoint}[1]{{\color{red}\textbf{Point: }#1}}
+\newcommand{\mcneed}{\textbf{Citation needed}}
+\newcommand{\mcfix}{\textbf{Fix citation}}
+
+\definecolor{cbnavy}{RGB}{15, 32, 128}
+\definecolor{cborange}{RGB}{245, 121, 58}
+\definecolor{cbgreen}{RGB}{0, 150, 50}
+
+%\newcommand{\runhead}[1]{\noindent\textbf{#1. }}
+\newcommand{\runhead}[1]{\subsubsection{#1}}
+
+% -- Fonts and styles for names
+%\newcommand{\mItemStyle}[1]{\ensuremath{#1}}
+%\newcommand{\mSetStyle}[1]{\ensuremath{\mathrm{\mathbf{#1}}}}
+%\newcommand{\mFunStyle}[1]{\text{\textup{\textsf{#1}}}}
+%\newcommand{\mConstStyle}[1]{\text{\textup{\textsf{#1}}}}
+%\newcommand{\mVarStyle}[1]{\mathit{#1}}
+%\newcommand{\mFactStyle}[1]{\text{\textsf{#1}}}
+%\newcommand{\mMethodStyle}[1]{\mbox{\mConstStyle{#1}}}
+%\newcommand{\mProtocolStyle}[1]{\mbox{\textrm{#1}}}
+
+%\newcommand{\mItemStyle}[1]{\ensuremath{#1}}
+%\newcommand{\mSetStyle}[1]{\ensuremath{\mathrm{\mathbf{#1}}}}
+\newcommand{\mFunStyle}[1]{\textsf{#1}}
+\newcommand{\mConstStyle}[1]{\textsf{#1}}
+\newcommand{\mVarStyle}[1]{\mathit{#1}}
+\newcommand{\mFactStyle}[1]{\textsf{#1}}
+\newcommand{\mMethodStyle}[1]{\mConstStyle{#1}}
+\newcommand{\mProtocolStyle}[1]{\text{#1}}
+
+% -- Macros for consistent wording
+\newcommand{\mSpec}{specification}  % The EDHOC spec document we are analyzing
+\newcommand{\mDate}{2020-07-06}
+
+% -- Formalization notation
+\newcommand{\mRevLTK}{\ensuremath{\mathbf{A}_\mFunStyle{LTK}}}
+\newcommand{\mTEE}{\ensuremath{\mathbf{A}_\mFunStyle{TEE}}}
+\newcommand{\mRevEph}{\ensuremath{\mathbf{A}_\mFunStyle{Eph}}}
+\newcommand{\mIStart}{\ensuremath{\mathbf{I}_\mFunStyle{S}}}
+\newcommand{\mIComplete}{\ensuremath{\mathbf{I}_\mFunStyle{C}}}
+\newcommand{\mRStart}{\ensuremath{\mathbf{R}_\mFunStyle{S}}}
+\newcommand{\mRComplete}{\ensuremath{\mathbf{R}_\mFunStyle{C}}}
+\newcommand{\mISend}{\ensuremath{\mathbf{I}_{\mFunStyle{SEND}}}}
+\newcommand{\mRRecv}{\ensuremath{\mathbf{R}_{\mFunStyle{Recv}}}}
+
+\newcommand{\mPredPcs}{\ensuremath{\mathbf{PCS}}}
+\newcommand{\mPredInjI}{\ensuremath{\mathbf{InjAgree}_I}}
+\newcommand{\mPredInjR}{\ensuremath{\mathbf{InjAgree}_R}}
+\newcommand{\mPredImpI}{\ensuremath{\mathbf{ImpAgree}_I}}
+\newcommand{\mPredImpR}{\ensuremath{\mathbf{ImpAgree}_R}}
+\newcommand{\mSecADthree}{\ensuremath{\mathbf{{Sec}\mADthree}}}
+\newcommand{\mIntADthree}{\ensuremath{\mathbf{Integ\mADthree}}}
+\newcommand{\mK}{\ensuremath{\mathcal{K}}}
+\DeclareMathOperator{\mLogicDot}{.}
+
+% -- Domain specific macros
+\newcommand{\mArxiv}{\texttt{arXiv}}
+\newcommand{\mTamarin}{\mProtocolStyle{Tamarin}}
+\newcommand{\mProverif}{\mProtocolStyle{ProVerif}}
+\newcommand{\mEdhoc}{\mProtocolStyle{EDHOC}}
+\newcommand{\mMuEdhoc}{\mProtocolStyle{\ensuremath{\mu}EDHOC}}
+\newcommand{\mOscore}{\mProtocolStyle{OSCORE}}
+\newcommand{\mSigma}{\mProtocolStyle{SIGMA}}
+\newcommand{\mSigmaI}{\mProtocolStyle{SIGMA\nobreakdash-I}}
+\newcommand{\mCbor}{\mProtocolStyle{CBOR}}
+\newcommand{\mCose}{\mProtocolStyle{COSE}}
+\newcommand{\mCoseEncrypt}{\mProtocolStyle{COSE\_Encrypt0}}
+\newcommand{\mCoseSign}{\mProtocolStyle{Cose\_Sign1}}
+\newcommand{\mHkdf}{\mProtocolStyle{HKDF}}
+\newcommand{\mHkdfExtract}{\mProtocolStyle{HKDF\nobreakdash-extract}}
+\newcommand{\mHkdfExpand}{\mProtocolStyle{HKDF\nobreakdash-expand}}
+\newcommand{\mHmac}{\mProtocolStyle{HMAC}}
+\newcommand{\mAead}{\mProtocolStyle{AEAD}}
+\newcommand{\mAeadDecrypt}{\mProtocolStyle{AEAD\nobreakdash-decrypt}}
+\newcommand{\mDecrypt}{\mProtocolStyle{decrypt}}
+\newcommand{\mOptls}{\mProtocolStyle{OPTLS}}
+\newcommand{\mNoise}{\mProtocolStyle{Noise}}
+\newcommand{\mTls}{\mProtocolStyle{TLS}}
+\newcommand{\mDandTls}{\mProtocolStyle{(D)TLS}}
+\newcommand{\mCtls}{\mProtocolStyle{cTLS}}
+\newcommand{\mCoap}{\mProtocolStyle{CoAP}}
+
+\newcommand{\mStat}{\mMethodStyle{STAT}}
+\newcommand{\mSig}{\mMethodStyle{SIG}}
+\newcommand{\mPsk}{\mMethodStyle{PSK}}
+\newcommand{\mStatStat}{\mMethodStyle{STAT-STAT}}
+\newcommand{\mStatSig}{\mMethodStyle{STAT-SIG}}
+\newcommand{\mSigStat}{\mMethodStyle{SIG-STAT}}
+\newcommand{\mSigSig}{\mMethodStyle{SIG-SIG}}
+\newcommand{\mPskPsk}{\mMethodStyle{PSK-PSK}}
+
+\newcommand{\mSid}{\mConstStyle{sid}}    % session id = (u, v, s-key)
+
+\newcommand{\mXor}{\mConstStyle{XOR}}
+\newcommand{\mSuites}{\ensuremath{S_I}}
+\newcommand{\mMethod}{\ensuremath{M}}
+\newcommand{\mCi}{\ensuremath{C_I}}
+\newcommand{\mCr}{\ensuremath{C_R}}
+\newcommand{\mGu}{\ensuremath{Q_{s,U}}}
+\newcommand{\mGi}{\ensuremath{Q_{s,I}}}
+\newcommand{\mGr}{\ensuremath{Q_{s,R}}}
+\newcommand{\mPriv}[1]{\ensuremath{d_{s,#1}}}
+\newcommand{\mPub}[1]{\ensuremath{Q_{s,#1}}}
+\newcommand{\mPrivE}[1]{\ensuremath{d_{e,#1}}}
+\newcommand{\mPubE}[1]{\ensuremath{Q_{e,#1}}}
+\newcommand{\mX}{\ensuremath{d_{e,I}}}
+\newcommand{\mY}{\ensuremath{d_{e,R}}}
+\newcommand{\mGiy}{\ensuremath{\mY\cdot\mGi}}
+\newcommand{\mGrx}{\ensuremath{\mX\cdot\mGr}}
+\newcommand{\mGx}{\ensuremath{Q_{e,I}}}
+\newcommand{\mGy}{\ensuremath{Q_{e,R}}}
+\newcommand{\mGxy}{\ensuremath{P_e}}
+\newcommand{\mSessKey}{\ensuremath{Z}}
+\newcommand{\mIDPsk}{\mConstStyle{ID\_PSK}}
+\newcommand{\mSign}[1]{\ensuremath{\mathit{sign_{#1}}}}
+\newcommand{\mVf}[1]{\ensuremath{\mathit{vf_{#1}}}}
+\newcommand{\mDH}{\ensuremath{\mathit{dh}}}
+\newcommand{\mTH}{\ensuremath{\mathit{th}}}
+\newcommand{\mTHtwo}{\ensuremath{\mathit{th}_2}}
+\newcommand{\mKtwoe}{\ensuremath{K_\mathit{2e}}}
+\newcommand{\mKtwom}{\ensuremath{K_\mathit{2m}}}
+\newcommand{\mKtwoae}{\ensuremath{K_\mathit{2ae}}}
+\newcommand{\mTID}{\ensuremath{\mathit{tid}}}
+
+\newcommand{\mKthreeae}{\ensuremath{K_\mathit{3ae}}}
+\newcommand{\mKthreem}{\ensuremath{K_\mathit{3m}}}
+
+\newcommand{\mTHthree}{\ensuremath{\mathit{th}_3}}
+\newcommand{\mhplain}{\ensuremath{h''}}
+\newcommand{\mCredi}{\ensuremath{Q_{s,I}}}
+\newcommand{\mCredr}{\ensuremath{Q_{s,R}}}
+\newcommand{\mHash}{\ensuremath{h}}
+
+\newcommand{\mTHfour}{\ensuremath{\mathit{th}_4}}
+\newcommand{\mAuthi}{\ensuremath{\mathit{Auth}_I}}
+\newcommand{\mAuthr}{\ensuremath{\mathit{Auth}_R}}
+
+\newcommand{\mMactwo}{\ensuremath{\mathit{MAC}_2}}
+\newcommand{\mMacthree}{\ensuremath{\mathit{MAC}_3}}
+
+\newcommand{\mSigtwo}{\ensuremath{\mathit{Sig}_2}}
+\newcommand{\mSigthree}{\ensuremath{\mathit{Sig}_3}}
+
+% \newcommand{\mMsgone}{\mConstStyle{m1}}
+% \newcommand{\mMsgtwo}{\mConstStyle{m2}}
+% \newcommand{\mMsgthree}{\mConstStyle{m3}}
+\newcommand{\mMsgone}{\ensuremath{m_1}}
+\newcommand{\mMsgtwo}{\ensuremath{m_2}}
+\newcommand{\mMsgthree}{\ensuremath{m_3}}
+
+\newcommand{\mCipher}{\mConstStyle{CIPHERTEXT\_2}}
+
+\newcommand{\mAD}{\ensuremath{\mathit{ad}}}
+\newcommand{\mADone}{\ensuremath{\mathit{ad}_1}}
+\newcommand{\mADtwo}{\ensuremath{\mathit{ad}_2}}
+\newcommand{\mADthree}{\ensuremath{\mathit{ad}_3}}
+
+\newcommand{\mPRK}{\ensuremath{\mathit{PRK}}}
+\newcommand{\mPRKtwo}{\ensuremath{\mPRK_\mathit{2e}}}
+\newcommand{\mPRKthree}{\ensuremath{\mPRK_\mathit{3e2m}}}
+\newcommand{\mPRKfour}{\ensuremath{\mPRK_\mathit{4x3m}}}
+
+\newcommand{\mIdcredi}{\ensuremath{ID_I}}
+\newcommand{\mIdcredr}{\ensuremath{ID_R}}
+\newcommand{\mLtki}{\mConstStyle{ltk\_I}}
+\newcommand{\mLtkr}{\mConstStyle{ltk\_R}}
+\newcommand{\mLtk}{\mConstStyle{ltk}}
+
+\newcommand{\cm}{\checkmark}
+
+% TIKZ messages and actions
+\newcommand{\msg}[4]{\draw[->,thick] ([yshift=-#1]#2.south) coordinate (l1)--(l1-|#3) node[midway, above]{#4}}
+\newcommand{\action}[3]{\node[draw,thick,fill=white,align=center,below={#1} of {#2}]{#3}}
+
+% Tamarin symbols
+\newcommand{\ifarrow}[1][]{\ifthenelse{\equal{#1}{}}{\rightarrow}{-\hspace{-4.22pt}[{#1}]\hspace{-5.6pt}\rightarrow}}
+\newcommand{\semarrow}[1][]{\ifthenelse{\equal{#1}{}}{\Rightarrow}{=\hspace{-3.0pt}[{#1}]\hspace{-3pt}\Rightarrow}}
+\newcommand{\mIn}{\mathsf{In}}
+\newcommand{\mOut}{\mathsf{Out}}
+\newcommand{\mFr}{\mathsf{Fr}}
+\newcommand{\mKD}{\mathsf{KD}}
+\newcommand{\mKU}{\mathsf{KU}}
+\newcommand{\mT}[1]{\text{\texttt{#1}}}
 
@@
     Karl Norrman\inst{1,2}\orcidID{0000-0003-0164-1478} \and
-    Vaishnavi Sundararajan\inst{3} \and
+    Vaishnavi Sundararajan\inst{3}\orcidID{0000-0002-5945-5208} \and
     Alessandro Bruni\inst{4}
@@
     \email{karl.norrman@ericsson.com} \and
-    University of California Santa Cruz, USA\\
+    University of California, Santa Cruz, USA\\
     \email{vasundar@ucsc.edu} \and
@@
 \label{sec:tamarin}
-We extend the formal \mEdhoc{} model of~\cite{Norr21}, using the tool \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an 
+We extend the formal model of~\cite{Norr21}, using the tool \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an 
 interactive
@@
 
-\begin{figure*}[ht]
+\begin{figure*}[htp]
 \begin{align*}
@@
     \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
-	&\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land 
+	&\ \lor\ (\exists t_0, t_1, Z'\mLogicDot \mIStart^{t_0}(\mTID, I, R, Z') \land 
 \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, I, R) 
+    &\ \lor\ (\exists t_0, t_1, Z'\mLogicDot \mRStart^{t_0}(\mTID, I, R, Z') 
 \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[1em]
@@
     \mPredInjI \triangleq\ &
-    \forall \mTID_I, I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, 
+    \forall \mTID_I, I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(\mTID_{I}, I, R, 
 \mSessKey, S)
@@
 I' , R', \mSessKey, S)
-        \rightarrow t_1' \doteq t_1))\\
+        \rightarrow t_1' \doteq t_2))\\
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
+    &\ \lor\ (\exists t_0, t_1, Z'\mLogicDot \mIStart^{t_0}(\mTID_I, I, R, Z') \land 
 \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[1em]
@@
 I' , R', \mSessKey, S)
-        \rightarrow t_1' \doteq t_1))\\
+        \rightarrow t_1' \doteq t_2))\\
     &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
-    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, 
-\mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot 
+    &\ \lor\ (\exists t_0, t_1, Z', S'\mLogicDot \mRStart^{t_0}(\mTID_R, R, 
+Z', S') \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot 
 t_2))\\[1em]
@@
     \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
-    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
-\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))
+    &\lor\ (\exists t_0, t_1, Z'\mLogicDot \mIStart^{t_0}(\mTID_I, I, R, Z') \land 
+\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[1em]
+%
+	\mSecADthree \triangleq\ &
+	\forall \mADthree, I, R, \mSessKey, \mTID_{I}, t_{3}, t_{2}\mLogicDot  \mK^{t_{3}}(\mADthree) \land \mISend^{t_{2}}(I, R, \mSessKey, \mADthree, \mTID_{I}) \rightarrow \\
+	& ( (\exists t_{1}\mLogicDot \mRevLTK^{t_{1}}(I) \land t_{1} \lessdot t_{2})\lor(\exists t_{1}\mLogicDot \mRevLTK^{t_{1}}(R) \land t_{1} \lessdot t_{2}) \\
+	&\ \ \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
+     \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey)) \\
+    &\ \ \lor(\exists t_{0}, t_{1}, Z'\mLogicDot \mIStart^{t_{0}}(\mTID_{I}, I, R, Z') \land \mTEE^{t_{1}}(R) \land t_{0} \lessdot t_{1}) \\
+    &\ \ \lor(\exists t_{0}, t_{1}, t_{4}, \mTID_{R}, Z', S'\mLogicDot \mTEE^{t_{1}}(I) \land \mRRecv^{t_{4}}(I, R, Z, \mADthree, \mTID_{R}) \land \mRStart^{t_{0}}(\mTID_{R}, R, Z', S') \land t_{0} \lessdot t_{1}))\\[1em]
+%
+	\mIntADthree \triangleq\ &
+	\forall \mADthree, I, R, \mSessKey, \mTID_{R}, t_{3}\mLogicDot 
+	\mRRecv^{t_{3}}(I, R, \mSessKey, \mADthree, \mTID_{R}) \rightarrow \\
+	& ( (\exists \mTID_{I}, t_{2}\mLogicDot \mISend^{t_{2}}(I, R, \mSessKey, \mADthree, \mTID_{I}) \land t_{2} \lessdot t_{3}) \\
+	&\ \lor (\exists t_{2}\mLogicDot \mRevLTK^{t_{2}}(I) \land t_{2} \lessdot t_{3}) \lor(\exists t_{2}\mLogicDot \mRevLTK^{t_{2}}(R) \land t_{2} \lessdot t_{3}) \\
+	&\ \ \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
+     \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey)) \\
+     &\ \ \lor(\exists t_{0}, t_{1}, t_{2}, \mTID_{I}, Z'\mLogicDot \mTEE^{t_{1}}(R) \land \mISend^{t_{2}}(I, R, \mSessKey, \mADthree, \mTID_{I}) \land \mIStart^{t_{0}}(\mTID_{I}, I, R, Z') \land t_{0} \lessdot t_{1}) \\
+     &\ \ \lor(\exists t_{0}, t_{1}, Z', S'\mLogicDot \mTEE^{t_{1}}(I) \land \mRStart^{t_{0}}(\mTID_{R}, R, Z', S') \land t_{0} \lessdot t_{1})
+	)
 \end{align*}
-%
-\caption{Formalization of security properties and adversary model.}
+\caption{Formalization of security properties}
 \label{fig:props}
@@
 only if one of the following conditions hold: either a party's long-term key
-is compromised before their run ends, or \mSessKey{} is directly revealed to
-the attacker, or the TEE interface for the responder is used during the current
+is compromised before their run ends, or the TEE interface for the responder is used during the current
 session for the initiator, i.e., after the initiator starts and before it
@@
 The full \mTamarin{} code for all models can be found 
-at~\cite{edhocTamarinRepo}.
+at~\cite{dropboxrepo}.
 %
@@
 %
-
+The entire code can be found at~\cite{dropboxrepo}.
 %-------------------------------------------------------------------------- sec
@@
               \textbf{Injective agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-              \textbf{Time (real/user)} & 0m27s / 11m17s & 9m27s / 239m49s & 
-1m21s / 35m36s & 18m16s / 499m2s\\
+              \textbf{Time (real/user)} & 3m10.782s / 76m23.828s & 52m57.153s / 1909m1.455s$\dagger$ & 
+              5m32.432s / 139m36.018s & 30m54.891s / 848m36.833s \\
               \textbf{Implicit agreement / I} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
               \textbf{Implicit agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-              \textbf{Time (real/user)} & 2m37s / 65m40s & 5m28s / 131m12s 
-$\dagger$ 
-& 8m54s / 212m21s & 7m37s / 188m46s $\dagger$\\
+              \textbf{Time (real/user)} & 3m34.222s / 89m41.770s & 6m46.187s / 180m6.615s 
+& 10m34.228s / 253m22.845s & 1m32.187s / 67m2.723s $\dagger$\\
               \textbf{Session key secrecy} & \cm & \cm & \cm & \cm\\
-              \textbf{Time (real/user)} & 0m41s / 16m4s & 1m52s / 45m27s & 6m31s / 
-156m26s & 9m47s / 232m2s\\
+              \textbf{Time (real/user)} & 1m23.317s / 52m16.347s & 3m54.009s / 155m54.834s & 12m33.884s / 457m32.698s & 20m40.391s / 915m15.590s\\
               \textbf{Secrecy of AD3} & \cm & \cm & \cm & \cm\\
               \textbf{Integrity of AD3} & \cm & \cm & \cm & \cm\\
-              \textbf{Time (real/user)} & 1m40s / 38m57s & 5m53s / 138m43s & 
-3m35s / 85m12s & 45m6s / 1029m1s\\
+              \textbf{Time (real/user)} & 2m39.727s / 101m2.371s & 5m51.728s / 223m59.728s & 
+              5m51.728s / 223m59.728s & 46m0.643s / 1786m3.825s\\
 			\hline
@@
                     \label{tab:props}
-        % \begin{tabular}{|l|c|c|c|c|}
-        %         \hline
-        %         & \mSigSig & \mSigStat & \mStatSig & \mStatStat \\
-        %         \hline
-        %         Injective agreement for I & $S_F$ & $S_F$ & $S_P$ & $S_P$\\
-        %         Injective agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-        %         Implicit agreement for I & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-        %         Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
-        %         PFS for session-key material & \cm & \cm & \cm & \cm\\
-        %         \hline
-        % \end{tabular}
 \end{table*}
@@
 %
+Modelling the new key derivation function would have needed sweeping changes across our entire formal model, and due to paucity of time, we could not modify the Tamarin model to be up to date with the current version.
+%
 However, we do refer to various aspects
@@
 %
-Due to lack of time we unfortunately could not bring the Tamarin model up to
-date with v15~\cite{draft-ietf-lake-edhoc-15}.
+Modelling the latest version of the protocol is left as future work.
 %

### 2022-08-09 paper.tex
[add] vaishnavi's conclusion updates

@@
 \label{sec:conclusion}
-We formally modeled all four
-methods of the \mEdhoc{} \mSpec{} using \mTamarin.
-%
-We formulated several important security properties and identified precise
-adversary models in which we verified these.
-%
-The properties are shown in Table~\ref{tab:props}.
-%
-Mutual injective agreement covers the parameters $S_P$:
-responder identity, roles, session-key material (except for $P_I$ when
-initiator uses the \mStat{} authentication
-method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
-%
-The responder in addition is ensured agreement on the initiators identity and
-$P_I$, i.e., on the set $S_F$.
-%
-Implicit agreement covers all previously mentioned parameters for both 
-peers.
-%
-All measurements were taken on a server with 2 Intel(R) Xeon(R) Gold
-6242 CPU @ 2.80GHz with a total of 32 cores / 64 threads and reserved
-256 GiB of memory for the scheduled job (the machine has a total of
-384 GiB installed).
-%
-Time measurements are rounded to the nearest second and we report both
-the real time (time to complete the task) and user time (total CPU time
-spent for the task).
-%
+As part of this paper, we modeled all four authentication methods of the \mEdhoc{} protocol using the \mTamarin{} verification tool.
+%
+We used a server with 2 Intel(R) Xeon(R) Gold 6242 CPU @ 2.80GHz, which had a total of 32 cores / 64 threads, and of the total 384GiB, the server allotted 256GiB of memory for the \mTamarin{} job.
+%
+We report both the time taken to complete the task (``real time'') and the CPU time spent on the task (``user time'').
+%
+Time measurements are rounded to the nearest second.
+%
+We formulated and verified the following properties in a precise adversary model with controlled access to TEEs and a precise modelling of \mXor.
+%
+\begin{itemize}
+\item Injective agreement for both $I$ and $R$
+\item Implicit agreement for both $I$ and $R$
+\item Session key secrecy
+\item Secrecy and integrity of \mADthree
+\end{itemize}
+%
+We consider two separate sets of parameters, $S_P$ and $S_F$. 
+%
+The set $S_P$ consists of the following pieces of information: the identity of the responder, roles, context identifierss \mCi{} and \mCr, cipher suites \mSuites, and session-key material excluding $P_I$.
+%
+For injective agreement, when the initiator uses a \mStat{} method, they are ensured agreement only on this above set of parameters. 
+%
+The set $S_F$ is the set $S_P$ along with the initiator's identity and $P_I$. 
+%
+The responder is guaranteed injective agreement on the parameters $S_F$, irrespective of what method they run, while the initiator is guaranteed injective agreement on $S_F$ as long as they run a \mSig{} method.
+%
+Implicit agreement can be reached for both parties on the set $S_F$.
+%
+In addition, mutual entity authentication, UKS- and KCI-resistance can be inferred from the verified properties.
+%
+We present the results in Table~\ref{tab:props}.
+
+%We formally modeled all four
+%methods of the \mEdhoc{} \mSpec{} using \mTamarin.
+%
+% We formulated several important security properties and identified precise
+% adversary models in which we verified these.
+% %
+%The properties are shown in Table~\ref{tab:props}.
+%
+% Mutual injective agreement covers the parameters $S_P$:
+% responder identity, roles, session-key material (except for $P_I$ when
+% initiator uses the \mStat{} authentication
+% method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
+% %
+% The responder in addition is ensured agreement on the initiators identity and
+% $P_I$, i.e., on the set $S_F$.
+% %
+% Implicit agreement covers all previously mentioned parameters for both 
+% peers.
+% %
+% All measurements were taken on a server with 2 Intel(R) Xeon(R) Gold
+% 6242 CPU @ 2.80GHz with a total of 32 cores / 64 threads and reserved
+% 256 GiB of memory for the scheduled job (the machine has a total of
+% 384 GiB installed).
+% %
+% Time measurements are rounded to the nearest second and we report both
+% the real time (time to complete the task) and user time (total CPU time
+% spent for the task).
+% %
 Mutual entity authentication, UKS- and KCI-resistance can be inferred
@@
 enable easier verification.
+%
+We also discussed some aspects of the protocol with respect to error handling and denial of service attacks.
 
@@
 %
-Modelling the latest version of the protocol is left as future work.
+Modelling the latest version of the protocol is also left as future work.
 %

### 2022-08-09 paper.tex
[add] clarifications and spelling corrections

@@
 %
-We also alter the formal modelling used for encryption under \mXor{}.
+We also alter the formal modeling used for encryption under \mXor{}.
 %
@@
 %
-    In those protocols, including $P_{I}$ while computating session-key
+    In those protocols, including $P_{I}$ while computing session-key
     material is crucial to obtaining resistance against the compromise of the
@@
 \label{sec:conclusion}
-As part of this paper, we modeled all four authentication methods of the \mEdhoc{} protocol using the \mTamarin{} verification tool.
-%
-We used a server with 2 Intel(R) Xeon(R) Gold 6242 CPU @ 2.80GHz, which had a total of 32 cores / 64 threads, and of the total 384GiB, the server allotted 256GiB of memory for the \mTamarin{} job.
-%
-We report both the time taken to complete the task (``real time'') and the CPU time spent on the task (``user time'').
+As part of this paper, we modeled all four authentication methods of the
+\mEdhoc{} protocol using the \mTamarin{} verification tool.
+%
+We used a server with 2 Intel(R) Xeon(R) Gold 6242 CPU @ 2.80GHz, which had a
+total of 32 cores / 64 threads, and of the total 384GiB, the server allotted
+256GiB of memory for the \mTamarin{} job.
+%
+We report both the time taken to complete the task (``real time'') and the CPU
+time spent on the task (``user time'').
 %
@@
 %
-We formulated and verified the following properties in a precise adversary model with controlled access to TEEs and a precise modelling of \mXor.
+We formulated and verified the following properties in a precise adversary model
+with controlled access to TEEs and a precise modeling of \mXor.
 %
@@
 \item Implicit agreement for both $I$ and $R$
-\item Session key secrecy
-\item Secrecy and integrity of \mADthree
+\item Perfect Forward Secrecy (PFS) for the session-key material
+\item Weak post-compromise security for the session-key material
+\item Secrecy and integrity of \mADthree{}
 \end{itemize}
 %
-We consider two separate sets of parameters, $S_P$ and $S_F$. 
-%
-The set $S_P$ consists of the following pieces of information: the identity of the responder, roles, context identifierss \mCi{} and \mCr, cipher suites \mSuites, and session-key material excluding $P_I$.
-%
-For injective agreement, when the initiator uses a \mStat{} method, they are ensured agreement only on this above set of parameters. 
+We consider two separate sets of parameters, a partial set ($S_P$) and the full
+set ($S_F$).
+%
+The set $S_P$ consists of the following pieces of information: the identity of
+the responder, roles, context identifiers \mCi{} and \mCr, cipher suites
+\mSuites, and session-key material excluding $P_I$.
+%
+For injective agreement, when the initiator uses a \mStat{} method, they are
+ensured agreement only on this above set of parameters.
 %
@@
 %
-The responder is guaranteed injective agreement on the parameters $S_F$, irrespective of what method they run, while the initiator is guaranteed injective agreement on $S_F$ as long as they run a \mSig{} method.
+The responder is guaranteed injective agreement on the parameters $S_F$,
+irrespective of what method they run, while the initiator is guaranteed
+injective agreement on $S_F$ as long as they run a \mSig{} method.
 %
@@
 %
-In addition, mutual entity authentication, UKS- and KCI-resistance can be inferred from the verified properties.
+In addition, mutual entity authentication, UKS- and KCI-resistance can be
+inferred from the verified properties.
 %
 We present the results in Table~\ref{tab:props}.
-
-%We formally modeled all four
-%methods of the \mEdhoc{} \mSpec{} using \mTamarin.
-%
-% We formulated several important security properties and identified precise
-% adversary models in which we verified these.
-% %
-%The properties are shown in Table~\ref{tab:props}.
-%
-% Mutual injective agreement covers the parameters $S_P$:
-% responder identity, roles, session-key material (except for $P_I$ when
-% initiator uses the \mStat{} authentication
-% method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
-% %
-% The responder in addition is ensured agreement on the initiators identity and
-% $P_I$, i.e., on the set $S_F$.
-% %
-% Implicit agreement covers all previously mentioned parameters for both 
-% peers.
-% %
-% All measurements were taken on a server with 2 Intel(R) Xeon(R) Gold
-% 6242 CPU @ 2.80GHz with a total of 32 cores / 64 threads and reserved
-% 256 GiB of memory for the scheduled job (the machine has a total of
-% 384 GiB installed).
-% %
-% Time measurements are rounded to the nearest second and we report both
-% the real time (time to complete the task) and user time (total CPU time
-% spent for the task).
-% %
+%
+
 Mutual entity authentication, UKS- and KCI-resistance can be inferred
@@
 
-Further, we identified a situation where initiators may establish an 
-\mOscore{}
+Further, we identified a situation where initiators may establish an \mOscore{}
 security context with a different party than the application using \mEdhoc{}
@@
 We discussed how the IETF may extract and better define security properties
-to
-enable easier verification.
-%
-We also discussed some aspects of the protocol with respect to error handling and denial of service attacks.
-
-We verified each method in isolation, and leave as future work to verify 
-whether
+to enable easier verification.
+%
+We also discussed some aspects of the protocol with respect to error handling
+and denial of service attacks.
+%
+
+We verified each method in isolation, and leave as future work to verify whether
 the methods are secure under composition.
@@
 %
-Modelling the new key derivation function would have needed sweeping changes across our entire formal model, and due to paucity of time, we could not modify the Tamarin model to be up to date with the current version.
+Modeling the new key derivation function would have needed sweeping changes
+across our entire formal model, and due to paucity of time, we could not modify
+the \mTamarin{} model to be up to date with the current version.
 %
@@
 %
-Modelling the latest version of the protocol is also left as future work.
+Modeling the latest version of the protocol is also left as future work.
 %

### 2022-08-09 paper.tex
[add] minor fixes and dead-code removal

@@
 %
-%Note that the properties in Figure~\ref{fig:props} are unlike the actual \mTamarin{} lemma in one minor way -- in the \mTamarin{} code, we only check that the TEE is accessed after the start of the session, since the long-term key reveal event covers the reveal of all long-term and session keys after the end of the session.
-%\vnote{Check that this is right}
 
@@
 %
-%As part of the formalization, note that we parametrize the \mIStart{} and \mRStart{} events with some subset of \mTID, $I$, $R$, and $Z$, as well as the set $S$ of parameters upon which we check for agreement.
-%%
-%The set $S$ will often include the other party's identity, so, for example, \mRStart{} will not mention the term $I$ as a parameter outside of $S$, and instead just be parametrized by \mTID, $R$, $Z$, and $S$.
 
@@
 %
-We 
-%augment the hashing function symbol to add an extra input, which we use 
-to
-model different hash functions.
-%
-
-In addition to these default operations, \mEdhoc{} is built over the following functions: \mHkdfExpand, \mHkdfExtract, and \mAead{}.
+
+In addition to these default operations, \mEdhoc{} is built over the following
+functions: \mHkdfExpand, \mHkdfExtract, and \mAead{}.
 %
@@
 %
-A term encrypted using the \mAead{} algorithm is represented by \mT{aeadEncrypt(m, k, ad, 
-ai)},
+A term encrypted using the \mAead{} algorithm is represented by
+\mT{aeadEncrypt(m, k, ad, ai)},
 where \mT{m} is the underlying message, \mT{k} is the encrypting key,
@@
 
-Using the fact \verb|Fr(~ltk)|, \mTamarin{} creates a new term \mT{ltk} and 
-uses
+Using the fact \verb|Fr(~ltk)|, \mTamarin{} creates a new term \mT{ltk} and uses
 it to represent a secret long-term key.
@@
 %
-Since the adversary has access to the communication channel, they can pick 
-up
+Since the adversary has access to the communication channel, they can pick up
 all of this information.
@@
 The event \mT{UniqLTK} parametrized by a party's identity and their
-long-term
-key 
+long-term key
 models the unique correspondence between those two values.
@@
 %
-Long-term keys can be revealed on registration, even before the protocol 
-begins.
+Long-term keys can be revealed on registration, even before the protocol begins.
 %
@@
 To circumvent the problem, we simplified the \mXor{} encryption to
-\mXor-ing 
-a
-single term on the entire tuple for these cases. \\
+\mXor-ing a single term on the entire tuple for these cases. \\
 %
@@
 %
-%For example, \mOptls{}, which essentially has the same core as \mStat{}, is
-%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, which
-%explicitly separates long-term key storage from that of the ephemeral keys and
-%session state.
-%%
-%This fits in well with our use of TEEs.
 Since \mSigma{} cannot protect against compromised ephemeral keys, the authors
@@
 preferable to obtain a homogeneous security level among the \mEdhoc{}
-methods.~\cite{Norr21}
+methods~\cite{Norr21}.
 %
